### 2024-03-13 analyses/simulate-sce/notebooks/splatter-ScPCA.Rmd
add dependencies and initial nb

---
title: "Simulating Single Cell experiments from ScPCA data"
output: html_notebook
---

```{r setup}
suppressMessages({
  library(SingleCellExperiment)
})
```

```{r}
sample_id <- "SCPCS000430"
library_id <- "SCPCL000750"
suffixes <- c( "_processed.rds", "_filtered.rds", "_unfiltered.rds")

infiles <- file.path("data", sample_id, paste0(library_id, suffixes))

sce <- readRDS(infiles[1])
```

```{r}
ncells <- 100
sce_subset <- sce[, sample.int(ncol(sce), ncells)]
sce_sim <- sce_subset
# splatter doesn't know new matrix types?
sim_params <- splatter::splatEstimate(as.matrix(counts(sce_subset)))
counts(sce_sim, withDimnames = FALSE) <- counts(splatter::splatSimulate(sim_params))

sce_sim <- sce_sim |>
  scater::logNormCounts()|>
  scater::runPCA(ncomponents = 10) |> # we don't need all the PCA components
  scater::runUMAP()
```

Are the same genes zero counted as before? That would be ideal, but maybe isn't the case.

```{r}
sce_sim
```

```{r}
metadata(sce_sim)$miQC_model <- NULL
readr::write_rds(sce_sim, paste0(infiles[1], "_sim_gz.rds"), compress = "gz")
readr::write_rds(sce_sim, paste0(infiles[1], "_sim_bz.rds"), compress = "bz2")
```

```{r}
sce2 <- readRDS(infiles[2])
ncells <- 100
sce2_subset <- sce2[, sample.int(ncol(sce), ncells)]
sce2_sim <- sce2_subset
# splatter doesn't know new matrix types?
sim_params <- splatter::splatEstimate(as.matrix(counts(sce_subset)))
counts(sce2_sim, withDimnames = FALSE) <- counts(splatter::splatSimulate(sim_params))
```

```{r}
metadata(sce2_sim)$miQC_model <- NULL
readr::write_rds(sce2_sim, paste0(infiles[2], "_sim_gz.rds"), compress = "gz")
readr::write_rds(sce2_sim, paste0(infiles[2], "_sim_bz.rds"), compress = "bz2")
```