### 2024-03-13 analyses/simulate-sce/notebooks/splatter-ScPCA.Rmd
Notebook version of sim complete

---
title: "Simulating Single Cell experiments from ScPCA data"
output: html_notebook
---

The data used in this notebook was previously downloaded with the following command:

```
./download-data.py --include "filtered,unfiltered,processed" --format "sce,anndata"
```

Note that we will only use the SCE data, but having the AnnData is useful for comparison in the final output.

```{r setup}
suppressMessages({
  library(SingleCellExperiment)
})

# get the current data directory (data/current), tracing from to the repo root
data_dir <- file.path(rprojroot::find_root(rprojroot::is_git_root), "data", "current")
results_dir <- here::here("results", "simulated") # starts from this modules root, since it is an r project.
```

For now, we will pick a single library to work with.
```{r}
project_id <- "SCPCP000015"
sample_id <- "SCPCS000490"
library_id <- "SCPCL000822"

suffixes <- c( "_processed.rds", "_filtered.rds", "_unfiltered.rds")

infiles <- file.path(data_dir, project_id, sample_id, paste0(library_id, suffixes))
outfiles <- file.path(results_dir, project_id, sample_id, paste0(library_id, suffixes))

sce <- readRDS(infiles[1]) # start with the processed file
```

Start by creating a subset of the original data set and adjusting the metadata.

```{r}
ncells <- 100

# reduce the data set to a smaller number of cells
cell_subset <- sample.int(ncol(sce), ncells)
sce_sim <- sce[, cell_subset]

# remove participant id, but leave it as a string
metadata(sce_sim)$sample_metadata$participant_id <- ""

# remove any miQC model
metadata(sce_sim)$miQC_model <- NULL

# reduce the cell type data matrices, if present
if (!is.null(metadata(sce_sim)$singler_results)) {
  metadata(sce_sim)$singler_results <- metadata(sce_sim)$singler_results[cell_subset,]
}
if (!is.null(metadata(sce_sim)$cellassign_predictions)) {
  metadata(sce_sim)$cellassign_predictions <- metadata(sce_sim)$cellassign_predictions[cell_subset,]
}
```

Create a function to randomly select labels from a set, ensuring each label is included at least once.

```{r}
random_label<- function(label_set, n){
  # randomly select n labels from the label set, ensuring each label is included at least once
  # ensure labels are unique
  label_set <- unique(label_set)
  if (length(label_set) > n) {
    stop("The number of labels must not be greater than the number requested")
  }
  r_labels <- sample(label_set, n, replace = TRUE)
  # add each label at least once
  idx <- sample.int(n, length(label_set))
  r_labels[idx] <- label_set
  return(r_labels)
}

```

Modify the cluster and cell type labels in the subest SCE to ensure they are all represented

```{r}
# Adjust cluster/cell type labels, using original label sets

# get label sets
cluster_set <- unique(colData(sce)$cluster)
cellassign_set <- unique(colData(sce)$cellassign_celltype_annotation)

# create a mapping for singler ontology and annotation
singler_df <- colData(sce)|>
  data.frame()|>
  dplyr::select(singler_celltype_ontology, singler_celltype_annotation) |>
  dplyr::distinct()

singler_dict <- setNames(
  singler_df$singler_celltype_annotation,
  singler_df$singler_celltype_ontology
)

colData(sce_sim)$cluster <- random_label(cluster_set, ncells)
colData(sce_sim)$singler_celltype_ontology <- random_label(names(singler_dict), ncells)
colData(sce_sim)$singler_celltype_annotation <- unname(
  singler_dict[colData(sce_sim)$singler_celltype_ontology]
)
colData(sce_sim)$cellassign_celltype_annotation <- random_label(cellassign_set, ncells)
```

Now we can do the simulation.
We will use the `splatter` package to simulate the data.
Not sure whether we want to do simulations with groups, but we probably don't need to.

```{r}
# for simulation, get probabilities of each cluster
# cluster_probs <- c(table(colData(sce)$cluster)/ncol(sce))
sim_params <- splatter::splatEstimate(as.matrix(counts(sce_sim)))

# sim_params <- splatter::setParams(sim_params, group.prob = cluster_probs)

counts(sce_sim, withDimnames = FALSE) <- splatter::splatSimulate(sim_params)|>
  counts()
# recalculate dimension reduction
sce_sim <- sce_sim |>
  scater::logNormCounts()|>
  scater::runPCA(10) |> # we don't need all the PCA components
  scater::runUMAP(dimred = "PCA")
```

Replace and update column stats

```{r}
# store coldata column names to restore order later
colData_names <- names(colData(sce_sim))

mito_detected_factor <- colData(sce_sim)$subsets_mito_detected / colData(sce_sim)$detected

# remove stats that will be recalculated
remove_stats <- c(
  "sum",
  "detected",
  "total",
  "subsets_mito_sum",
  "subsets_mito_detected",
  names(colData(sce_sim))[grep("altExp_.*_(sum|detected|percent)", names(colData(sce_sim)))]
)
colData(sce_sim)[,remove_stats] <- NULL

sce_sim <- scuttle::addPerCellQC(sce_sim)
colData(sce_sim)$subsets_mito_sum <- colData(sce_sim)$sum * colData(sce_sim)$subsets_mito_percent / 100
colData(sce_sim)$subsets_mito_detected <- round(colData(sce_sim)$detected * mito_detected_factor)

# restore column order
colData(sce_sim) <- colData(sce_sim)[,colData_names]

```

```{r}
fs::dir_create(dirname(outfiles[1]))
readr::write_rds(sce_sim, paste0(outfiles[1]), compress = "bz2")
```