### 2024-03-16 analyses/hello-R/notebooks/hello.Rmd
start to respond to review

---
title: Hello OpenScPCA
author:
  - Joshua A. Shapiro, Childhood Cancer Data Lab, ALSF
date: "`r Sys.Date()`"
output: html_notebook
---

## Introduction

This notebook presents a simple example of an analysis in the OpenScPCA project, using R/Bioconductor.

We will go through the processed SCE objects in the project `data` library and count the number of cells in each object, plotting the results.

## Setup

Load libraries and set initial paths.

```{r libraries}
# load required libraries
suppressPackageStartupMessages({
  library("SingleCellExperiment")
  library("ggplot2")
})
```

Set the paths to the data, results, and plots directories.

```{r}
# Find the repository and module root directories
repo_root <- rprojroot::find_root(rprojroot::is_git_root)
module_root <- rprojroot::find_root(rprojroot::is_renv_project)

# set current data dir
data_dir <- file.path(repo_root, "data", "current")

# set results and plots directories (using the analysis project file to find root)
results_dir <- file.path(module_root, "results")
plots_dir <- file.path(module_root, "plots")
```

We will be creating two output files: a table and a plot.
Defining their locations in the `Setup` section will make it easy for future notebook users to know what to expect.

```{r}
count_file <- file.path(results_dir, "cell_counts.csv")
plot_file <- file.path(plots_dir, "cell_counts.pdf")
```

## Functions

First, we will define a function to load and count the cells in a SCE object, returning a one line data frame with the project ID, sample ID, library ID, and number of cells.

```{r}
# function to count cells in an sce file
# returns a data frame with ids and cell count
count_sce <- function(sce_file) {
  # load the SCE object
  sce <- readRDS(sce_file)

  # get stats and count the cells
  result_df <- data.frame(
    project_id = metadata(sce)$project_id,
    sample_id = metadata(sce)$sample_id,
    library_id = metadata(sce)$library_id,
    n_cells = ncol(sce)
  )
  return(result_df)
}
```

I think that is it! Now we can get the list of SCE objects and count the cells.

## Count cells and plot

First we will list all of the processed RDS files in the data directory.
We will use `list.files()` to do this all in one shot, using the option to search recursively.

```{r}
# get the list of processed SCE files
sce_files <- list.files(
  data_dir,
  pattern = "_processed.rds$",
  full.names = TRUE,
  recursive = TRUE
)
```

Now we will use the `purrr` package to apply our counting function to each of the files, and then combine all of the results into a single data frame.

```{r}
# count the cells in each file
cell_counts <- sce_files |>
  purrr::map(count_sce) |>
  purrr::list_rbind() # combine results into a single data frame
```

Let's look at the table of results.

```{r}
cell_counts
```

And save it to our results directory.

```{r}
readr::write_csv(cell_counts, count_file)
```

Finally, we will plot the results as a faceted histogram, colored by project ID.

```{r}
count_plot <- ggplot(cell_counts, aes(x = n_cells, fill = project_id)) +
  geom_histogram(bins = 20) +
  labs(x = "Number of cells") +
  facet_wrap(facets = vars(project_id)) +
  guides(fill = "none") + # don't show the legend
  theme_bw()
count_plot # show the plot
```

Excellent!

Now let's save the plot.

```{r}
ggsave(plot_file, count_plot, width = 8, height = 6)
```

## Session Info

```{r}
sessioninfo::session_info()
```