### 2024-03-18 templates/analysis-module-R/notebook-template.Rmd
Add R notebook template

---
title: Analysis Title
author:
  - Your name here
date: "`r Sys.Date()`"
output: html_notebook
---

## Introduction

This a template for an analysis notebook using R Markdown.

It is a good idea to start with a brief introduction to the analysis, including the purpose of the analysis, the data used, and the methods applied.
Replace this text with your own introduction, and be sure to update the `title` and `author` fields at the top of the document.

## Setup

### Packages

Load required packages in the following chunk, if needed.
Do not install packages here, only load them with the `library()` function.

```{r packages}
# load required packages
# e.g. library("tidyverse")
```

You may also want to use the `package::function()` syntax within your code for better tracking of which functions come from which packages, and to avoid conflicts between functions with the same name in different packages.

### Paths

Setting paths to directories and files for input and output at the start of a notebook can be helpful for organization and reproducibility.
You may not know all of the paths you need when you start a notebook, but once you do, keeping them all in one place can be helpful, both for yourself in the future and for other users of the notebook.

### Base directories

In the chunk below, we included some common directories that you might need, finding their locations using the very useful `rprojroot` package.

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")

# This module's base path, found by the presence of an `renv` directory
module_base <- rprojroot::find_root(rprojroot::is_renv_project)
# If you are not using `renv`, you might use something like the following instead:
# module_base <- file.path(repository_base, "analyses", "{module_name}")
```

### Input and output files

Set paths to input and output directories and files in the chunk below

```{r paths}
# Input files

# Output files
```

## Analysis content

Organize the remainder of your content into sections and subsections as appropriate for your analysis.

Add a new code chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I* on Windows or Linux, *Cmd+Option+I* on a Mac.
Be sure to comment your code and explain your steps as you go.

When you save the notebook, an HTML file containing the code and output will be saved alongside it.
To preview the HTML file, click the *Preview* button or press *Ctrl+Shift+K* on Windows or Linux, *Cmd+Shift+K* on a Mac.

The preview shows you a rendered HTML copy of the contents of the editor.
Consequently, *Preview* does not run any R code chunks.
Instead, the output of the chunk when it was last run in the editor is displayed.

It is usually a good idea to rerun the entire notebook before sharing any output, to ensure that the notebook and output files are fully up to date.
You can do this by clicking on the *Run* menu in the toolbar and clicking on *Restart R and Run All Chunks*.

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```