### 2024-04-05 analyses/simulate-sce/notebooks/splatter-ScPCA.Rmd

---
title: "Simulating Single Cell experiments from ScPCA data"
output: html_notebook
---

This notebook was used as exploration for the `simulate-sce.R` script.
The code here is not fully current with that script, but has been left here for reference about the development process and for future use.

The data used in this notebook was previously downloaded with the following command:

```
./download-data.py --include "filtered,unfiltered,processed" --format "sce,anndata"
```

Note that we will only use the SCE data, but having the AnnData is useful for comparison in the final output.

```{r setup}
suppressMessages({
  library(SingleCellExperiment)
})

# get the current data directory (data/current), tracing from to the repo root
data_dir <- file.path(rprojroot::find_root(rprojroot::is_git_root), "data", "current")
results_dir <- here::here("results", "simulated") # starts from this modules root, since it is an r project.
```

For now, we will pick a single library to work with.
```{r}
project_id <- "SCPCP000015"
sample_id <- "SCPCS000490"
library_id <- "SCPCL000822"

suffixes <- c("_processed.rds", "_filtered.rds", "_unfiltered.rds")

infiles <- file.path(data_dir, project_id, sample_id, paste0(library_id, suffixes))
outfiles <- file.path(results_dir, project_id, sample_id, paste0(library_id, suffixes))

sce <- readRDS(infiles[1]) # start with the processed file
```

Start by creating a subset of the original data set and adjusting the metadata.

```{r}
ncells <- 100

# reduce the data set to a smaller number of cells
cell_subset <- sample.int(ncol(sce), ncells)
sce_sim <- sce[, cell_subset]

# remove participant id, but leave it as a string
metadata(sce_sim)$sample_metadata$participant_id <- ""

# remove any miQC model
metadata(sce_sim)$miQC_model <- NULL

# reduce the cell type data matrices, if present
if (!is.null(metadata(sce_sim)$singler_results)) {
  metadata(sce_sim)$singler_results <- metadata(sce_sim)$singler_results[cell_subset, ]
}
if (!is.null(metadata(sce_sim)$cellassign_predictions)) {
  metadata(sce_sim)$cellassign_predictions <- metadata(sce_sim)$cellassign_predictions[cell_subset, ]
}
```

Create a function to randomly select labels from a set, ensuring each label is included at least once.

```{r}
random_label <- function(label_set, n) {
  # randomly select n labels from the label set, ensuring each label is included at least once
  # ensure labels are unique
  label_set <- unique(label_set)
  if (length(label_set) > n) {
    stop("The number of labels must not be greater than the number requested")
  }
  r_labels <- sample(label_set, n, replace = TRUE)
  # add each label at least once
  idx <- sample.int(n, length(label_set))
  r_labels[idx] <- label_set
  return(r_labels)
}
```

Modify the cluster and cell type labels in the subset SCE to ensure they are all represented

```{r}
# Adjust cluster/cell type labels, using original label sets

# get label sets
cluster_set <- unique(colData(sce)$cluster)
cellassign_set <- unique(colData(sce)$cellassign_celltype_annotation)

# create a mapping for singler ontology and annotation
singler_df <- colData(sce) |>
  data.frame() |>
  dplyr::select(singler_celltype_ontology, singler_celltype_annotation) |>
  dplyr::distinct()

singler_dict <- setNames(
  singler_df$singler_celltype_annotation,
  singler_df$singler_celltype_ontology
)

colData(sce_sim)$cluster <- random_label(cluster_set, ncells)
colData(sce_sim)$singler_celltype_ontology <- random_label(names(singler_dict), ncells)
colData(sce_sim)$singler_celltype_annotation <- unname(
  singler_dict[colData(sce_sim)$singler_celltype_ontology]
)
colData(sce_sim)$cellassign_celltype_annotation <- random_label(cellassign_set, ncells)
```

Now we can do the simulation.
We will use the `splatter` package to simulate the data.
Not sure whether we want to do simulations with groups, but we probably don't need to.

```{r}
# for simulation, get probabilities of each cluster
# cluster_probs <- c(table(colData(sce)$cluster)/ncol(sce))
sim_params <- splatter::splatEstimate(as.matrix(counts(sce_sim)))

# sim_params <- splatter::setParams(sim_params, group.prob = cluster_probs)

counts(sce_sim, withDimnames = FALSE) <- splatter::splatSimulate(sim_params) |>
  counts()
# recalculate dimension reduction
sce_sim <- sce_sim |>
  scater::logNormCounts() |>
  scater::runPCA(10) |> # we don't need all the PCA components
  scater::runUMAP(dimred = "PCA")
```

Replace and update column stats

```{r}
# store coldata column names to restore order later
colData_names <- names(colData(sce_sim))

mito_detected_factor <- colData(sce_sim)$subsets_mito_detected / colData(sce_sim)$detected

# remove stats that will be recalculated
remove_stats <- c(
  "sum",
  "detected",
  "total",
  "subsets_mito_sum",
  "subsets_mito_detected",
  names(colData(sce_sim))[grep("altExp_.*_(sum|detected|percent)", names(colData(sce_sim)))]
)
colData(sce_sim)[, remove_stats] <- NULL

sce_sim <- scuttle::addPerCellQC(sce_sim)
colData(sce_sim)$subsets_mito_sum <- colData(sce_sim)$sum * colData(sce_sim)$subsets_mito_percent / 100
colData(sce_sim)$subsets_mito_detected <- round(colData(sce_sim)$detected * mito_detected_factor)

# restore column order
colData(sce_sim) <- colData(sce_sim)[, colData_names]
```

```{r}
fs::dir_create(dirname(outfiles[1]))
readr::write_rds(sce_sim, paste0(outfiles[1]), compress = "bz2")
```