### 2024-04-25 analyses/ewings-cell-type/01-tumor-cell-classification.Rmd
initial exploration notebook for marker genes

---
title: "Explor tumor cell classification - SCPCS000490"
author: Ally Hawkins
date: "`r Sys.Date()`"
output: html_notebook
params:
  sample_id: SCPCS000490
  library_id: SCPCL000822
---

## Introduction

This notebook looks at expression of marker genes for tumor cells in a Ewing sarcoma sample, `r {params$sample_id}`. 
Additionally, we see if we can use those marker genes to classify tumor and normal cells using `CellAssign`.

- First we look at expression of each of the marker genes across all cells. 
- Then we plot the mean expression of all marker genes across all cells. 
- Next, we perform clustering and see if we can assign each cluster as tumor or normal cells based on mean expression of marker genes. 
- Finally, we use `CellAssign` with both the full list of marker genes and filtered list of marker genes. 
- The results from `CellAssign` are then compared to manual annotation of cells.

## Setup

```{r packages}
suppressMessages(
  # load required packages
  library(SingleCellExperiment),
  library(ggplot2)
)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")
sample_dir <- file.path(data_dir, "SCPCP000015", params$sample_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "ewings-cell-type")
```

```{r paths}
# Input files
sce_filename <- glue::glue("{params$library_id}_processed.rds")
sce_file <- file.path(sample_dir, sce_filename)

anndata_filename <- glue::glue("{params$library_id}_processed_rna.hdf5")
anndata_file <- file.path(sample_dir, anndata_filename)

marker_genes <- file.path(module_base, "references", "tumor_marker_genes.tsv")

# Output files
tumor_markers_mtx_file <- file.path(module_base, "references", "tumor-marker-cellassign.tsv")
filtered_markers_mtx_file <- file.path(module_base, "references", "filtered-tumor-marker-cellassign.tsv")
cellassign_predictions_file <- file.path(module_base, "results", "cellassign-tumor-predictions.tsv")
filtered_predictions_file <- file.path(module_base, "results", "filtered-cellassign-tumor-predictions.tsv")
```

```{r}
# read in processed sce
sce <- readr::read_rds(sce_file)

# read in marker genes table 
marker_genes_df <- readr::read_tsv(marker_genes) |> 
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |> 
  dplyr::distinct()

marker_genes <- marker_genes_df |> 
  dplyr::filter(cell_type == "tumor") |> 
  dplyr::pull(ensembl_gene_id)
```

## Analysis content

### Explore marker gene expression

The first thing we do here is just create a faceted UMAP showing the expression of each marker gene for tumor cells.

```{r}
# get the gene expression counts for all marker genes
marker_gene_exp <- logcounts(sce[marker_genes, ]) |>
    as.matrix() |>
    t() |>
    as.data.frame() |>
    tibble::rownames_to_column("barcodes")

# pull out the UMAP coordinates and make a data frame to use for plotting
umap_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
        \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP"),
        starts_with("UMAP")
      )

marker_gene_plot_df <- umap_df |> 
  # add in marker gene expression to dataframe 
  dplyr::left_join(marker_gene_exp, by = "barcodes") |> 
  # combine all genes into a single column for easy faceting
  tidyr::pivot_longer(
    cols = starts_with("ENSG"),
    names_to = "ensembl_gene_id",
    values_to = "gene_expression"
  ) |> 
  # join with marker gene df to get gene symbols for plotting 
  dplyr::left_join(marker_genes_df, by = c("ensembl_gene_id")) |>
  dplyr::select(barcodes, UMAP1, UMAP2, gene_symbol, ensembl_gene_id, gene_expression, clusters)
```

```{r}
# faceted umap showing a umap panel for each marker gene 
ggplot(marker_gene_plot_df, aes(x = UMAP1, y = UMAP2, color = gene_expression)) +
    geom_point(alpha = 0.1, size = 0.4) +
    facet_wrap(vars(gene_symbol)) +
    scale_color_viridis_c() +
    labs(
      color = "Log-normalized gene expression"
    ) +
    # remove axis numbers and background grid
    scale_x_continuous(labels = NULL, breaks = NULL) +
    scale_y_continuous(labels = NULL, breaks = NULL) +
    theme(
      aspect.ratio = 1,
      legend.position = "bottom",
      axis.title = element_text(size = 9, color = "black"),
      strip.text = element_text(size = 8),
      legend.title = element_text(size = 9),
      legend.text = element_text(size = 8)
    ) +
    guides(colour = guide_colorbar(title.position = "bottom", title.hjust = 0.5)) +
  theme_bw()
```

Looking at this plot, there is definitely some variation in the expression of marker genes. 
It looks like the highest gene expression is seen in the larger group of cells on the bottom right of the UMAP.

Below, we can get the mean gene expression of all marker genes and plot in a single UMAP.

```{r}
# calculate mean gene expression across all marker genes in list 
marker_mean_exp <- umap_df |> 
  dplyr::group_by(barcodes) |> 
  dplyr::mutate(mean_exp = mean(gene_expression)) |> 
  dplyr::select(barcodes, UMAP1, UMAP2, mean_exp) |> 
  dplyr::distinct()

# plot mean gene expression 
ggplot(marker_mean_exp, aes(x = UMAP1, y = UMAP2, color = mean_exp)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_color_viridis_c() +
  theme_bw()
```

Similar to the individual plots, it looks like there is one group of cells on the bottom right that has the highest mean gene expression. 
We would anticipate that these are most likely to be the tumor cells.

### Clustering

```{r}
ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = clusters)) +
  geom_point(size = 0.5, alpha = 0.5) +
  theme_bw()
```

```{r}
# classify cells as tumor/normal 
classification_df <- marker_mean_exp |> 
  dplyr::mutate(tumor_normal_classification = dplyr::if_else(mean_exp > 0.5, "Tumor", "Normal"))

ggplot(classification_df, aes(x = UMAP1, y = UMAP2, color = tumor_normal_classification)) +
  geom_point(size = 0.5, alpha = 0.5) +
  theme_bw()
```

### Run CellAssign with marker genes

Create marker genes reference

```{r}
binary_mtx <- marker_genes_df |>
  dplyr::select(cell_type, ensembl_gene_id) |> 
  unique() |> 
  tidyr::pivot_wider(
    id_cols = ensembl_gene_id,
    names_from = cell_type,
    values_from = cell_type,
    values_fn = length,
    values_fill = 0
  ) |>
  # add a column with no marker genes
  # cell assign will assign cells to "other" when no other cell types are appropriate
  dplyr::mutate(other = 0)

# export matrix 
readr::write_tsv(binary_mtx, tumor_markers_mtx_file)
```

Add script for running CellAssign 
```{r, eval=FALSE}
cellassign_call <- glue::glue("
python scripts/run-cellassign.py \
  --anndata_file '{anndata_file}' \
  --output_predictions '{cellassign_predictions_file}' \
  --reference '{tumor_markers_mtx}'
")

system(cellassign_call)
```

Read in CellAssign marker genes
```{r}
predictions <- readr::read_tsv(cellassign_predictions_file)

celltype_assignments <- predictions |>
      tidyr::pivot_longer(
        !barcode,
        names_to = "celltype",
        values_to = "prediction"
      ) |>
      dplyr::group_by(barcode) |>
      dplyr::slice_max(prediction, n = 1) |>
      dplyr::ungroup()

table(celltype_assignments$celltype)
```

What if we restrict our list of marker genes based on our dataset? So we want something that has a mean gene expression of 1.

```{r}
filtered_markers <- rowData(sce) |> 
  as.data.frame() |> 
  dplyr::filter(gene_symbol %in% marker_genes_df$gene_symbol,
                mean > 1) |> 
  dplyr::pull(gene_ids)

filtered_binary_mtx <- binary_mtx |> 
  dplyr::filter(ensembl_gene_id %in% filtered_markers)

readr::write_tsv(filtered_binary_mtx, filtered_markers_mtx_file)
```

```{r, eval=FALSE}
cellassign_call <- glue::glue("
python scripts/run-cellassign.py --anndata_file '{anndata_file}' --output_predictions '{filtered_predictions_file}' --reference '{filtered_markers_mtx_file}'
")

system(cellassign_call)
```

```{r}
predictions <- readr::read_tsv(filtered_predictions_file)

celltype_assignments <- predictions |>
      tidyr::pivot_longer(
        !barcode,
        names_to = "celltype",
        values_to = "prediction"
      ) |>
      dplyr::group_by(barcode) |>
      dplyr::slice_max(prediction, n = 1) |>
      dplyr::ungroup()

table(celltype_assignments$celltype)
```

```{r}
celltype_df <- marker_mean_exp |> 
  dplyr::left_join(celltype_assignments, by = c("barcodes" = "barcode"))

ggplot(celltype_df, aes(x = UMAP1, UMAP2, color = celltype)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  theme_bw()
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```