### 2024-05-01 analyses/ewings-cell-type/02-cellassign-tumor-cell-classification.Rmd
initiate cellassign notebook

---
title: "Explore tumor cell classification with CellAssign"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  sample_id: SCPCS000490
  library_id: SCPCL000822
---

## Introduction

This notebook looks at classifying tumor cells with `CellAssign` in a Ewing sarcoma sample, `r {params$sample_id}`.

- We use `CellAssign` with both a full list of marker genes and filtered list of marker genes. 
- We then compare the average gene expression of all marker genes between tumor and other cells. 
- Finally we use `CellAssign` with marker genes from [Visser et al.](https://doi.org/10.1158/2767-9764.CRC-23-0027) specified in the supplemental methods.

## Setup

```{r packages}
suppressMessages(
  # load required packages
  library(SingleCellExperiment),
  library(ggplot2)
)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")
sample_dir <- file.path(data_dir, "SCPCP000015", params$sample_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "ewings-cell-type")
```

```{r paths}
# Input files
sce_filename <- glue::glue("{params$library_id}_processed.rds")
sce_file <- file.path(sample_dir, sce_filename)

anndata_filename <- glue::glue("{params$library_id}_processed_rna.hdf5")
anndata_file <- file.path(sample_dir, anndata_filename)

# tumor marker genes
tumor_markers_file <- file.path(module_base, "references", "tumor-marker-genes.tsv")
# all markers 
all_markers_file <- file.path(module_base, "references", "all-marker-genes.tsv")

# Output reference matrix files
ref_dir <- file.path(module_base, "references", "cellassign_refs")

tumor_markers_mtx_file <- file.path(ref_dir, "tumor-marker-cellassign.tsv")
filtered_markers_mtx_file <- file.path(ref_dir, "filtered-tumor-marker-cellassign.tsv")
visser_markers_mtx_file <- file.path(ref_dir, "visser-all-marker-cellassign.tsv")

# output predictions from cellassign 
results_dir <- file.path(module_base, "results")
cellassign_predictions_file <- file.path(results_dir, "cellassign-tumor-predictions.tsv")
filtered_predictions_file <- file.path(results_dir, "filtered-cellassign-tumor-predictions.tsv")
visser_predictions_file <- file.path(results_dir, "visser-cellassign-predictions.tsv")

```

```{r}
# read in processed sce
sce <- readr::read_rds(sce_file)

# read in tumor marker genes table 
tumor_markers_df <- readr::read_tsv(tumor_markers_file) |> 
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |> 
  dplyr::distinct()

# read in all maker genes 
all_markers_df <- readr::read_tsv(all_markers_file) |> 
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |> 
  dplyr::distinct()
```

Define some helper functions for creating a reference matrix and obtaining cell type annotations from `CellAssign` predictions.

```{r}
# function for creating a binary matrix with cell types as columns and marker genes as rows 
build_binary_mtx <- function(marker_genes_df,
                             gene_col, 
                             cell_type_col) {

  binary_mtx <- marker_genes_df |>
  dplyr::select({{cell_type_col}}, {{gene_col}}) |> 
  unique() |> 
  tidyr::pivot_wider(
    id_cols = {{gene_col}},
    names_from = {{cell_type_col}},
    values_from = {{cell_type_col}},
    values_fn = length,
    values_fill = 0
  ) |>
  # add a column with no marker genes
  # cell assign will assign cells to "other" when no other cell types are appropriate
  dplyr::mutate(other = 0)

  return(binary_mtx)
}

# function to get assigned cell type based on max prediction returned by CellAssign 
get_celltype_assignment <- function(predictions){
  # get individual cell type assignments
  # those with the max prediction 
  celltype_assignments <- predictions |>
    tidyr::pivot_longer(
      !barcode,
      names_to = "celltype",
      values_to = "prediction"
    ) |>
    dplyr::group_by(barcode) |>
    dplyr::slice_max(prediction, n = 1) |>
    dplyr::ungroup()

  return(celltype_assignments)
}
```

## Analysis content

### Run CellAssign with only tumor marker genes

First let's see if we can use `CellAssign` to classify cells as tumor or normal using just the list of tumor marker genes.

First, we need to create a binary matrix with two columns, "tumor" and "other". 
The "other" will refer to any normal cells.

```{r}
tumor_ref_mtx <- build_binary_mtx(tumor_markers_df, 
                                  gene_col = "ensembl_gene_id",
                                  cell_type_col = "cell_type")

# export matrix 
readr::write_tsv(tumor_ref_mtx, tumor_markers_mtx_file)
```

Now run `CellAssign` with the full marker gene list as the reference.

```{r, eval=FALSE}
cellassign_call <- glue::glue(
  "python scripts/run-cellassign.py",
  "--anndata_file '{anndata_file}'",
  "--output_predictions '{cellassign_predictions_file}'",
  "--reference '{tumor_markers_mtx_file}'"
)

system(cellassign_call)
```

We can then read in the predictions file and find the predicted cell type for each cell, either "tumor" or "other".

```{r}
# read in predictions file
tumor_predictions <- readr::read_tsv(cellassign_predictions_file)

# get cell type assignments 
tumor_celltype_assignments <- get_celltype_assignment(tumor_predictions)

# create a table of assignments 
table(tumor_celltype_assignments$celltype)
```

Looking at this, it appears that `CellAssign` is classifying all cells as "other"? 
I find this hard to believe since this sample was taken directly from a Ewing sarcoma tumor, so there are bound to be tumor cells.

### Run CellAssign with a filtered set of tumor marker genes

In looking at the original UMAP there are quite a few genes that have very low expression across all cells, so perhaps they are adding noise to the model. 
Although this is a bit circular, we could restrict our list of marker genes to be only those that have a high mean gene expression in our dataset? 
I would expect that when we do that, we are able to identify more tumor cells that line up with what we would expect based on gene expression patterns.

```{r}
# grab the mean gene expression for all marker genes 
rowdata_df <- rowData(sce) |> 
  as.data.frame() |> 
  dplyr::filter(gene_symbol %in% marker_genes_df$gene_symbol)
rowdata_df
```

```{r}
# what's the mean gene expression of all genes in this sample for reference
mean(rowData(sce)$mean)
```

There are a few genes that have low gene expression < 1. 
Let's try and filter out those genes and then use `CellAssign`.

```{r}
# get list of marker genes to keep
filtered_markers <- rowdata_df |> 
  dplyr::filter(mean > 1) |> 
  dplyr::pull(gene_ids)

# filter binary mtx 
filtered_ref_mtx <- tumor_ref_mtx |> 
  dplyr::filter(ensembl_gene_id %in% filtered_markers)

readr::write_tsv(filtered_ref_mtx, filtered_markers_mtx_file)
```

```{r, eval=FALSE}
cellassign_call <- glue::glue(
  "python scripts/run-cellassign.py",
  "--anndata_file '{anndata_file}'",
  "--output_predictions '{filtered_predictions_file}'",
  "--reference '{filtered_markers_mtx_file}'"
)

system(cellassign_call)
```

Now we can read in the predictions for the filtered marker gene list.

```{r}
filtered_predictions <- readr::read_tsv(filtered_predictions_file)

filtered_celltype_assignments <- get_celltype_assignment(filtered_predictions)

table(filtered_celltype_assignments$celltype)
```

Now we are seeing a few more tumor cells, let's see where they are?

```{r}
celltype_df <- marker_mean_exp |> 
  dplyr::left_join(celltype_assignments, by = c("barcodes" = "barcode"))

ggplot(celltype_df, aes(x = UMAP1, UMAP2, color = celltype)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  theme_bw()
```

Not quite where I would expect to see tumor cells. 
Let's look at the mean expression of marker genes between tumor and other cells as classified by `CellAssign`.

```{r}
ggplot(celltype_df, aes(x = mean_exp, fill = celltype)) +
  geom_density() +
  facet_grid(rows = vars(celltype)) +
  theme_bw()
```

Unlike what I would expect, the "other" cells appear to have higher expression of marker genes? 
Maybe using `CellAssign` with only two options and a small set of marker genes doesn't work in the way we would expect.

### Run CellAssign with tumor markers and all marker genes from Visser et al.

```{r}
visser_ref_mtx <- build_binary_mtx(all_markers_df, 
                                   gene_col = "ensembl_gene_id",
                                   cell_type_col = "cell_type")

# export matrix 
readr::write_tsv(visser_ref_mtx, visser_markers_mtx_file)
```

Now run `CellAssign` with the full marker gene list as the reference.

```{r, eval=FALSE}
cellassign_call <- glue::glue(
  "python scripts/run-cellassign.py ",
  "--anndata_file '{anndata_file}' ",
  "--output_predictions '{visser_predictions_file}' ",
  "--reference '{visser_markers_mtx_file}'"
)

system(cellassign_call)
```

```{r}
visser_predictions <- readr::read_tsv(visser_predictions_file)

visser_celltype_assignments <- get_celltype_assignment(visser_predictions)

table(visser_celltype_assignments$celltype)
```

## Conclusions

- Using `CellAssign` with the same list of marker genes doesn't identify any tumor cells. 
- Using `CellAssign` with a restricted list of marker genes identfies tumor cells that have lower marker gene expression than other cells. 
This is opposite of the assignment we obtained using manual annotation.

```{r}

```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```