### 2024-05-01 analyses/ewings-cell-type/02-cellassign-tumor-cell-classification.Rmd
add visser marker gene analysis

---
title: "Explore tumor cell classification with CellAssign"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  sample_id: SCPCS000490
  library_id: SCPCL000822
---

## Introduction

This notebook looks at classifying tumor cells with `CellAssign` in a Ewing sarcoma sample, `r {params$sample_id}`.

- We use `CellAssign` with both a full list of marker genes and filtered list of marker genes. 
- We then compare the average gene expression of all marker genes between tumor and other cells. 
- Finally we use `CellAssign` with marker genes from [Visser et al.](https://doi.org/10.1158/2767-9764.CRC-23-0027) specified in the supplemental methods.

## Setup

```{r packages}
suppressMessages(
  # load required packages
  library(SingleCellExperiment),
  library(ggplot2)
)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")
sample_dir <- file.path(data_dir, "SCPCP000015", params$sample_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "ewings-cell-type")

# source in helper functions
jaccard_functions <- file.path(module_base, "scripts", "utils", "jaccard-functions.R")
source(jaccard_functions)
```

```{r paths}
# Input files
sce_filename <- glue::glue("{params$library_id}_processed.rds")
sce_file <- file.path(sample_dir, sce_filename)

anndata_filename <- glue::glue("{params$library_id}_processed_rna.hdf5")
anndata_file <- file.path(sample_dir, anndata_filename)

# tumor normal classifications 
classifications_results_dir <- file.path(module_base, "results", "marker_gene_analysis")
classifications_filename <- glue::glue("{params$library_id}_tumor_normal_classifications.tsv")
classifications_file <- file.path(classifications_results_dir, classifications_filename)

# tumor marker genes
tumor_markers_file <- file.path(module_base, "references", "tumor-marker-genes.tsv")
# all markers 
all_markers_file <- file.path(module_base, "references", "all-marker-genes.tsv")
# panglao markers
panglao_markers_file <- file.path(module_base, "references", "PanglaoDB_markers_2020-03-27.tsv")

# Output reference matrix files
ref_dir <- file.path(module_base, "references", "cellassign_refs")

tumor_markers_mtx_file <- file.path(ref_dir, "tumor-marker-cellassign.tsv")
filtered_markers_mtx_file <- file.path(ref_dir, "filtered-tumor-marker-cellassign.tsv")
visser_markers_mtx_file <- file.path(ref_dir, "visser-all-marker-cellassign.tsv")
panglao_markers_mtx_file <- file.path(ref_dir, "panglao-endo-fibro-cellassign.tsv")

# output predictions from cellassign 
preds_results_dir <- file.path(module_base, "results", "cellassign_predictions")
fs::dir_create(preds_results_dir)

cellassign_predictions_file <- file.path(preds_results_dir, "cellassign-tumor-predictions.tsv")
filtered_predictions_file <- file.path(preds_results_dir, "filtered-cellassign-tumor-predictions.tsv")
visser_predictions_file <- file.path(preds_results_dir, "visser-cellassign-predictions.tsv")
panglao_predictions_file <- file.path(preds_results_dir, "panglao-endo-fibro-predictions.tsv")
```

```{r}
# read in processed sce
sce <- readr::read_rds(sce_file)

# read in tumor marker genes table 
tumor_markers_df <- readr::read_tsv(tumor_markers_file) |> 
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |> 
  dplyr::distinct()

# read in all maker genes 
all_markers_df <- readr::read_tsv(all_markers_file) |> 
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |> 
  dplyr::distinct()
```

Define some helper functions for creating a reference matrix and obtaining cell type annotations from `CellAssign` predictions.

```{r}
# function for creating a binary matrix with cell types as columns and marker genes as rows 
build_binary_mtx <- function(marker_genes_df,
                             gene_col, 
                             cell_type_col) {

  binary_mtx <- marker_genes_df |>
  dplyr::select({{cell_type_col}}, {{gene_col}}) |> 
  unique() |> 
  tidyr::pivot_wider(
    id_cols = {{gene_col}},
    names_from = {{cell_type_col}},
    values_from = {{cell_type_col}},
    values_fn = length,
    values_fill = 0
  ) |>
  # add a column with no marker genes
  # cell assign will assign cells to "other" when no other cell types are appropriate
  dplyr::mutate(other = 0)

  return(binary_mtx)
}

# function to get assigned cell type based on max prediction returned by CellAssign 
get_celltype_assignment <- function(predictions){
  # get individual cell type assignments
  # those with the max prediction 
  celltype_assignments <- predictions |>
    tidyr::pivot_longer(
      !barcode,
      names_to = "celltype",
      values_to = "prediction"
    ) |>
    dplyr::group_by(barcode) |>
    dplyr::slice_max(prediction, n = 1) |>
    dplyr::ungroup()

  return(celltype_assignments)
}
```

## Analysis content

### Run CellAssign with only tumor marker genes

First let's see if we can use `CellAssign` to classify cells as tumor or normal using just the list of tumor marker genes.

First, we need to create a binary matrix with two columns, "tumor" and "other". 
The "other" will refer to any normal cells.

```{r}
tumor_ref_mtx <- build_binary_mtx(tumor_markers_df, 
                                  gene_col = "ensembl_gene_id",
                                  cell_type_col = "cell_type")

# export matrix 
readr::write_tsv(tumor_ref_mtx, tumor_markers_mtx_file)
```

Now run `CellAssign` with the full marker gene list as the reference.

```{r, eval=FALSE}
cellassign_call <- glue::glue(
  "python scripts/run-cellassign.py",
  "--anndata_file '{anndata_file}'",
  "--output_predictions '{cellassign_predictions_file}'",
  "--reference '{tumor_markers_mtx_file}'"
)

system(cellassign_call)
```

We can then read in the predictions file and find the predicted cell type for each cell, either "tumor" or "other".

```{r}
# read in predictions file
tumor_predictions <- readr::read_tsv(cellassign_predictions_file)

# get cell type assignments 
tumor_celltype_assignments <- get_celltype_assignment(tumor_predictions)

# create a table of assignments 
table(tumor_celltype_assignments$celltype)
```

Looking at this, it appears that `CellAssign` is classifying all cells as "other"? 
I find this hard to believe since this sample was taken directly from a Ewing sarcoma tumor, so there are bound to be tumor cells.

### Run CellAssign with a filtered set of tumor marker genes

In looking at the UMAPs of all marker genes (see `01-marker-gene-tumor-classification.Rmd`) there are quite a few genes that have very low expression across all cells, so perhaps they are adding noise to the model. 
Although this is a bit circular, we could restrict our list of marker genes to be only those that have a high mean gene expression in our dataset? 
I would expect that when we do that, we are able to identify more tumor cells that line up with what we would expect based on gene expression patterns.

```{r}
# grab the mean gene expression for all marker genes 
rowdata_df <- rowData(sce) |> 
  as.data.frame() |> 
  dplyr::filter(gene_symbol %in% tumor_markers_df$gene_symbol)
rowdata_df
```

```{r}
# what's the mean gene expression of all genes in this sample for reference
mean(rowData(sce)$mean)
```

There are a few genes that have low gene expression < 1. 
Let's try and filter out those genes and then use `CellAssign`.

```{r}
# get list of marker genes to keep
filtered_markers <- rowdata_df |> 
  dplyr::filter(mean > 1) |> 
  dplyr::pull(gene_ids)

# filter binary mtx 
filtered_ref_mtx <- tumor_ref_mtx |> 
  dplyr::filter(ensembl_gene_id %in% filtered_markers)

readr::write_tsv(filtered_ref_mtx, filtered_markers_mtx_file)
```

```{r, eval=FALSE}
cellassign_call <- glue::glue(
  "python scripts/run-cellassign.py",
  "--anndata_file '{anndata_file}'",
  "--output_predictions '{filtered_predictions_file}'",
  "--reference '{filtered_markers_mtx_file}'"
)

system(cellassign_call)
```

Now we can read in the predictions for the filtered marker gene list.

```{r}
filtered_predictions <- readr::read_tsv(filtered_predictions_file)

filtered_celltype_assignments <- get_celltype_assignment(filtered_predictions)

table(filtered_celltype_assignments$celltype)
```

Now we are seeing a few more tumor cells, let's see where they are on the UMAP. 
We can also compare with the tumor normal classifications we made using marker genes alone.

```{r}
# first read in tumor normal classifications 
classifications_df <- readr::read_tsv(classifications_file)

filtered_celltype_assignments <- filtered_celltype_assignments |> 
  dplyr::rename(filtered_cellassign_annotations = celltype)

celltype_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
        \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP"),
        starts_with("UMAP")
      ) |> 
  # join with tumor normal
  dplyr::left_join(classifications_df) |> 
  # join with filtered cell type assignments from cellassign 
  dplyr::left_join(filtered_celltype_assignments, by = c("barcodes" = "barcode"))

ggplot(celltype_df, aes(x = UMAP1, UMAP2, color = filtered_cellassign_annotations)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  theme_bw()
```

Not quite where I would expect to see tumor cells. 
Let's see how many tumor cells overlap with our manual classifications using the Jaccard similarity index.

```{r}
jaccard_mtx <- make_jaccard_matrix(celltype_df, "filtered_cellassign_annotations", "marker_gene_classification")

ComplexHeatmap::Heatmap(
  t(jaccard_mtx), # transpose because matrix rows are in common & we want a vertical arrangement
  col = circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue")),
  border = TRUE,
  ## Row parameters
  cluster_rows = FALSE,
  row_title = "Marker gene",
  row_title_gp = grid::gpar(fontsize = 12),
  row_title_side = "left",
  row_names_side = "left",
  row_dend_side = "right",
  row_names_gp = grid::gpar(fontsize = 10),
  ## Column parameters
  cluster_columns = FALSE,
  column_title = "CellAssign",
  column_title_gp = grid::gpar(fontsize = 12),
  column_names_side = "bottom",
  column_names_gp = grid::gpar(fontsize = 10),
  column_names_rot = 90,
  ## Legend parameters
  heatmap_legend_param = list(
    title = "Jaccard index",
    direction = "vertical",
    legend_width = unit(1.5, "in")
  ),
  show_heatmap_legend = TRUE,
)
```
As expected, it looks like CellAssign is assigning very few cells to tumor that overlap with the tumor cells annotated with marker gene analysis. 
It looks like most tumor cells are being classified as normal with the marker gene analysis.

Let's look at the mean expression of marker genes between tumor and other cells as classified by `CellAssign`.

```{r}
# get marker gene expression 
marker_gene_exp <- logcounts(sce[filtered_markers, ]) |>
  as.matrix() |>
  t() |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes")

marker_gene_exp_df <- celltype_df |> 
  dplyr::left_join(marker_gene_exp, by = "barcodes") |> 
  tidyr::pivot_longer(
    cols = starts_with("ENSG"),
    names_to = "ensembl_gene_id",
    values_to = "transformed_gene_expression"
  ) |> 
  dplyr::group_by(barcodes) |> 
  dplyr::mutate(sum_exp = sum(transformed_gene_expression))

ggplot(marker_gene_exp_df, aes(x = sum_exp, fill = filtered_cellassign_annotations)) +
  geom_density() +
  facet_grid(rows = vars(filtered_cellassign_annotations)) +
  theme_bw()
```

Unlike what I would expect, the "other" cells appear to have higher expression of marker genes? 
Maybe using `CellAssign` with only two options and a small set of marker genes doesn't work in the way we would expect.

### Run CellAssign with tumor markers and all marker genes from Visser et al.

Perhaps we need to add a few more cell types, rather than just have the two options. 
Below we will use CellAssign but with all tumor marker genes and all marker genes in Visser et al. for non-tumor cells.

```{r}
visser_ref_mtx <- build_binary_mtx(all_markers_df, 
                                   gene_col = "ensembl_gene_id",
                                   cell_type_col = "cell_type")

# export matrix 
readr::write_tsv(visser_ref_mtx, visser_markers_mtx_file)
```

Now run `CellAssign` with the full marker gene list as the reference.

```{r, eval=FALSE}
cellassign_call <- glue::glue(
  "python scripts/run-cellassign.py ",
  "--anndata_file '{anndata_file}' ",
  "--output_predictions '{visser_predictions_file}' ",
  "--reference '{visser_markers_mtx_file}'"
)

system(cellassign_call)
```

```{r}
visser_predictions <- readr::read_tsv(visser_predictions_file)

visser_celltype_assignments <- get_celltype_assignment(visser_predictions)

table(visser_celltype_assignments$celltype)
```
Here we see that there are still only a few hundred tumor cells and most cells are endothelial cells. 
Generally, I expect that there to be a large portion of tumor cells with much fewer cells from normal cell types.

Below we will compare assignments between CellAssign and manual marker gene assignment.

```{r}
# first rename column to combine with cell type df 
visser_celltype_assignments <- visser_celltype_assignments |> 
  dplyr::rename(visser_cellassign_annotations = celltype)

celltype_df <- celltype_df |> 
  dplyr::left_join(visser_celltype_assignments, by = c("barcodes" = "barcode"))

ggplot(celltype_df, aes(x = UMAP1, UMAP2, color = visser_cellassign_annotations)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  theme_bw()
```

```{r}
jaccard_mtx <- make_jaccard_matrix(celltype_df, "visser_cellassign_annotations", "marker_gene_classification")

ComplexHeatmap::Heatmap(
  t(jaccard_mtx), # transpose because matrix rows are in common & we want a vertical arrangement
  col = circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue")),
  border = TRUE,
  ## Row parameters
  cluster_rows = FALSE,
  row_title = "Marker gene",
  row_title_gp = grid::gpar(fontsize = 12),
  row_title_side = "left",
  row_names_side = "left",
  row_dend_side = "right",
  row_names_gp = grid::gpar(fontsize = 10),
  ## Column parameters
  cluster_columns = FALSE,
  column_title = "CellAssign - Visser",
  column_title_gp = grid::gpar(fontsize = 12),
  column_names_side = "bottom",
  column_names_gp = grid::gpar(fontsize = 10),
  column_names_rot = 90,
  ## Legend parameters
  heatmap_legend_param = list(
    title = "Jaccard index",
    direction = "vertical",
    legend_width = unit(1.5, "in")
  ),
  show_heatmap_legend = TRUE,
)
```

```{r}
all_markers <- all_markers_df |> 
  dplyr::pull(ensembl_gene_id)

marker_gene_exp <- logcounts(sce[all_markers, ]) |>
  as.matrix() |>
  t() |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes")

marker_gene_exp_df <- celltype_df |> 
  dplyr::left_join(marker_gene_exp, by = "barcodes") |> 
  tidyr::pivot_longer(
    cols = starts_with("ENSG"),
    names_to = "ensembl_gene_id",
    values_to = "gene_expression"
  ) |> 
  dplyr::left_join(all_markers_df)

celltypes <- unique(all_markers_df$cell_type)

celltypes |> 
  purrr::map(\(celltype){
    plot_df <- marker_gene_exp_df |> 
      dplyr::filter(cell_type == celltype)

    # faceted umap showing a umap panel for each marker gene 
    ggplot(plot_df, aes(x = UMAP1, y = UMAP2, color = gene_expression)) +
      geom_point(alpha = 0.1, size = 0.2) +
      facet_wrap(vars(gene_symbol)) +
      scale_color_viridis_c() +
      labs(
        title = celltype,
        color = "Log-normalized gene expression"
      ) +
      # remove axis numbers and background grid
      scale_x_continuous(labels = NULL, breaks = NULL) +
      scale_y_continuous(labels = NULL, breaks = NULL) +
      theme(
        aspect.ratio = 1,
        legend.position = "bottom",
        axis.title = element_text(size = 9, color = "black"),
        strip.text = element_text(size = 8),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8)
      ) +
      guides(colour = guide_colorbar(title.position = "bottom", title.hjust = 0.5)) +
      theme_bw()

  })
```

Then we will look at the gene expression by cell type for each marker gene. 
We would expect that the specified marker genes for each cell type are highly expressed in those cells compared to other cells.

```{r}
transformed_exp_df <- marker_gene_exp_df |> 
  dplyr::select(barcodes, gene_symbol, gene_expression) |> 
  tidyr::pivot_wider(
    names_from = "gene_symbol",
    values_from = "gene_expression") |> 
  dplyr::mutate(across(!barcodes, ~ scale(.x))) |> 
  tidyr::pivot_longer(
    cols = - barcodes,
    names_to = "gene_symbol",
    values_to = "transformed_gene_expression"
  )

# add the transformed gene expression back to the main df used for plotting
grouped_celltype_df <- marker_gene_exp_df |> 
  dplyr::left_join(transformed_exp_df, by = c("barcodes", "gene_symbol")) |> 
  dplyr::group_by(barcodes, cell_type) |> 
  dplyr::mutate(sum_exp = sum(transformed_gene_expression))
```

```{r}
ggplot(grouped_celltype_df, aes(x = sum_exp, color = cell_type)) +
  geom_density() +
  facet_wrap(vars(visser_cellassign_annotations)) +
  theme_bw() +
  labs(
    color = "Marker gene group"
  )
```

### Use markers from PanglaoDB

```{r}
panglao_markers_df <- readr::read_tsv(panglao_markers_file) |> 
  dplyr::filter(`cell type` %in% c("Endothelial cells", "Fibroblasts")) |> 
  dplyr::select(`official gene symbol`, `cell type`)

panglao_tumor_markers_df <- rowData(sce) |> 
  as.data.frame() |> 
  dplyr::filter(gene_symbol %in% panglao_markers_df$`official gene symbol`) |> 
  dplyr::left_join(panglao_markers_df, by = c("gene_symbol" = "official gene symbol")) |> 
  dplyr::select(gene_symbol, ensembl_gene_id = gene_ids, cell_type = `cell type`) |> 
  dplyr::bind_rows(tumor_markers_df)

```

```{r}
panglao_tumor_ref_mtx <- build_binary_mtx(panglao_tumor_markers_df, 
                                          gene_col = "ensembl_gene_id",
                                          cell_type_col = "cell_type")

# export matrix 
readr::write_tsv(panglao_tumor_ref_mtx, panglao_markers_mtx_file)
```

Now run `CellAssign` with the full marker gene list as the reference.

```{r, eval=FALSE}
cellassign_call <- glue::glue(
  "python scripts/run-cellassign.py ",
  "--anndata_file '{anndata_file}' ",
  "--output_predictions '{panglao_predictions_file}' ",
  "--reference '{panglao_markers_mtx_file}' "
)

system(cellassign_call)
```

We can then read in the predictions file and find the predicted cell type for each cell, either "tumor" or "other".

```{r}
# read in predictions file
panglao_predictions <- readr::read_tsv(panglao_markers_mtx_file)

# get cell type assignments 
panglao_celltype_assignments <- get_celltype_assignment(panglao_predictions)

# create a table of assignments 
table(panglao_celltype_assignments$celltype)
```

## Conclusions

- Using `CellAssign` with the same list of marker genes doesn't identify any tumor cells. 
- Using `CellAssign` with a restricted list of marker genes identfies tumor cells that have lower marker gene expression than other cells. 
This is opposite of the assignment we obtained using manual annotation.

```{r}

```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```