### 2024-05-08 analyses/ewings-cell-type/04-infercnv.Rmd
initial notebook

---
title: "Copy number inference with InferCNV"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  sample_id: SCPCS000490
  library_id: SCPCL000822
---

## Introduction

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")
sample_dir <- file.path(data_dir, "SCPCP000015", params$sample_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "ewings-cell-type")
```

```{r base paths}
# source in helper functions
jaccard_functions <- file.path(module_base, "scripts", "utils", "jaccard-functions.R")
source(jaccard_functions)
```

```{r paths}
# Input files
sce_filename <- glue::glue("{params$library_id}_processed.rds")
sce_file <- file.path(sample_dir, sce_filename)

# tumor normal classifications 
classifications_results_dir <- file.path(module_base, "results", "marker_gene_analysis")
classifications_filename <- glue::glue("{params$library_id}_tumor_normal_classifications.tsv")
classifications_file <- file.path(classifications_results_dir, classifications_filename)

# output from running infercnv 
```

```{r}
sce <- readr::read_rds(sce_file)
```

```{r}
# create the infercnv object
infercnv_obj <- infercnv::CreateInfercnvObject(raw_counts_matrix=counts(sce),
                                               annotations_file=annotations_file,
                                               delim="\t",
                                               gene_order_file=gene_order_file,
                                               ref_group_names=c("endothelial"))

```
```{r}
# perform infercnv operations to reveal cnv signal
infercnv_obj <- infercnv::run(infercnv_obj,
                              cutoff=0.1,  # use 1 for smart-seq, 0.1 for 10x-genomics
                              out_dir=infer_results_dir,  # dir is auto-created for storing outputs
                              denoise=T,
                              HMM=T ,
                              report_by = "cell"
)
```

```{r}
infercnv_obj@observation_grouped_cell_indices

infercnv_obj@expr.data[1:5, 1:5]
```