### 2024-05-09 analyses/ewings-cell-type/04-infercnv.Rmd
look at cnvs for each cell

---
title: "Copy number inference with InferCNV"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  sample_id: SCPCS000490
  library_id: SCPCL000822
---

## Introduction

Here we look at using [`InferCNV`](https://github.com/broadinstitute/infercnv) to identify tumor and normal cells in `r params$sample_id`.

Before running this notebook, `InferCNV` is run using the `run-infercnv.R` script. 
This script runs `InferCNV` once with Endothelial cells as the reference cells and once without any cells as the reference.

Here, we look the copy number variations found reported by `InferCNV` and use that copy number information to call cells as tumor or normal cells. 
We then compare these classifications to manual classifications and annotations from `SingleR` and `CellAssign`.

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
  library(dplyr)
})
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")
sample_dir <- file.path(data_dir, "SCPCP000015", params$sample_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "ewings-cell-type")
```

```{r}
# source in helper functions
jaccard_functions <- file.path(module_base, "scripts", "utils", "jaccard-functions.R")
source(jaccard_functions)
```

```{r paths}
# Input files
sce_filename <- glue::glue("{params$library_id}_processed.rds")
sce_file <- file.path(sample_dir, sce_filename)

# tumor normal classifications 
classifications_results_dir <- file.path(module_base, "results", "marker_gene_analysis")
classifications_filename <- glue::glue("{params$library_id}_tumor_normal_classifications.tsv")
classifications_file <- file.path(classifications_results_dir, classifications_filename)

# output from running infercnv 
infer_cnv_dir <- file.path(module_base, "results", "infercnv", "SCPCL000822")

# add code to read in both ref and no ref? 
```

```{r}
# read in sce object
sce <- readr::read_rds(sce_file)

# read in tumor normal classifications 
manual_classifications_df <- readr::read_tsv(classifications_file)
```

## InferCNV Heatmap

Below is the heatmap produced by `InferCNV` showing the CNV's for each cell. 
Each row is a cell and each column is a gene. 
The top group contains all cells that were used as the reference group and the bottom is all other cells. 
The color indicates the relative expression intensity of each gene with the darker colors either indicating a gain of expression or loss of expression compared to the reference cells.

![InferCNV Heatmap - Endothelial cells as reference](results/infercnv/SCPCL000822/infercnv.png)
From this heatmap we see that there is a distinct group of cells within the bottom portion of the plot that appear to have a lot more CNVs than the rest of the cells. 
It's probable that these correspond to the tumor cells, but `InferCNV` does not provide any classification of cells as tumor or normal, it just provides information on copy number variations.

## Using CNVs to annotate tumor cells

`InferCNV` uses an HMM-based model to predict whether or not a copy number variation (either loss or gain) is present for each chromosome in each cell. 
For each predicted CNV region, a posterior probability is calculated for that CNV for each cell.

From the [`InferCNV` documentation](https://github.com/broadinstitute/inferCNV/wiki/inferCNV-HMM-based-CNV-Prediction-Methods):

> CNV regions with mean posterior probabilities of being normal (no CNV) that are above a maximum threshold are removed as likely false positive predictions.

The predictions information lies in a few files that live in the output directory. 
We want to know which cells had CNVs and how many CNVs were found. 
To map the information about CNVs to individual cells, they recommend using the `infercnv::add_to_seurat()` function which takes the output directory from running `InferCNV` and reads in all necessary files and adds the per cell CNV information to the `meta.data`. 
So we will start by converting our processed SCE object to a Seurat object, incorporate the CNV information, and then add the CNV information to the SCE object.

```{r}
# first create a Seurat object from our SCE object
seurat_obj <- scpcaTools::sce_to_seurat(sce)

# add in the CNV information 
# this adds new columns to the meta.data indicating presence of CNVs for each chromosome
seurat_obj <- infercnv::add_to_seurat(seurat_obj = seurat_obj,
                                      infercnv_output_path = infer_cnv_dir)

# grab metadata from Seurat and add back to SCE 
seurat_metadata <- seurat_obj@meta.data |> 
  dplyr::select(-c(orig.ident, nCount_RNA, nFeature_RNA))

colData(sce) <- DataFrame(seurat_metadata, row.names = seurat_metadata$barcodes)
```

The metadata that we added includes the following columns for each chromosome:

- `has_cnv_chr1`
- `has_loss_chr1`
- `has_dupli_chr1`
- `proportion_cnv_chr1`
- `proportion_loss_chr1`
- `proportion_dupli_chr1`
- `proportion_scaled_cnv_chr1`
- `proportion_scaled_loss_chr1`
- `proportion_scaled_dupli_chr1`

The `has` columns are boolean columns indicating if that cell has presence of a CNV, loss, or duplication event. 
If `has_cnv` is TRUE if either `has_loss` or `has_dupli` is TRUE.

See [the comments in this issue](https://github.com/broadinstitute/infercnv/issues/208) explaining what the contents of the columns are.

```{r}
coldata_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
        \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")
      )

# remove MT and GL chr 
chr_columns_to_keep <- which(!stringr::str_detect(names(coldata_df), "chrMT|chrGL000219.1"))
cnv_df <- coldata_df[, chr_columns_to_keep]
```

One approach to classifying tumor cells is to look for cells that have a lot of CNVs. 
Let's start by just summing up the total CNVs/cell and plotting that.

```{r}
# grab any columns that have has_cnv 
has_cnv_cols <- names(cnv_df)[startsWith(colnames(cnv_df), "has_cnv")]

cnv_df <- cnv_df |>
  # get total number of cnvs per cell and then scale that number 
  dplyr::mutate(has_cnv = rowSums(across(has_cnv_cols)),
                scaled_cnv = scale(has_cnv))
```

```{r}
# plot the total number of CNVs
ggplot(cnv_df, aes(x = UMAP1, y= UMAP2, color = has_cnv)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() +
  scale_color_viridis_c()
```
This shows that the total CNVs are highest in the group of cells that correspond to the same cells that are manually annotated as tumor cells, which is promising. 
If we were to use total CNVs we would need some sort of cut off for the number. 
Let's look at the distribution of total CNVs and also the scaled total CNVs to see if we can identify a cut off.

```{r}
ggplot(cnv_df, aes(x = has_cnv)) +
  geom_density() +
  theme_bw() +
  labs(x = "Total number of CNVs")

ggplot(cnv_df, aes(x= scaled_cnv)) + 
  geom_density() +
  theme_bw() +
  labs(x = "Scaled number of CNVS")
```

Looking at the scaled number of CNVs it does look like we could use 0 as the cutoff and call any cell that has > 0 CNVs (after scaling) as tumor and all other cells as normal.

```{r}
# add cnv classification 
cnv_df <- cnv_df |> 
  dplyr::mutate(cnv_classification = dplyr::if_else(scaled_cnv > 0, "Tumor", "Normal"))

ggplot(cnv_df, aes(x = UMAP1, y= UMAP2, color = cnv_classification)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() 
```

Another option is to use the scaled proportion of CNVs which tells us the proportion of genes on that chromosome that are part of the CNV.

```{r}
proportion_cols <- names(cnv_df)[startsWith(colnames(cnv_df), "proportion_scaled_cnv")]

# get mean proportion of cnvs per cell and then scale that number 
cnv_df <- cnv_df |> 
  dplyr::mutate(mean_proportion = rowMeans(across(proportion_cols)),
                scaled_mean_proportion = scale(mean_proportion))
```

```{r}
ggplot(cnv_df, aes(x = UMAP1, y= UMAP2, color = mean_proportion)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() +
  scale_color_viridis_c()

ggplot(cnv_df, aes(x = UMAP1, y= UMAP2, color = scaled_mean_proportion)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() +
  scale_color_viridis_c()
```

```{r}
ggplot(cnv_df, aes(x = scaled_mean_proportion)) +
  geom_density() +
  theme_bw()
```
It looks like maybe we could again use 0 as the cutoff here and anything above 0 is a tumor cell. 
Let's see how this matches up with the tumor/normal classifications we defined just using total CNVs.

```{r}
ggplot(cnv_df, aes(x = scaled_mean_proportion, color = cnv_classification)) +
  geom_density() +
  theme_bw()
```

Interestingly, the cells we labeled as tumor have a higher mean proportion value then those that we labeled as normal. 
I'm not really sure which is "better" to use the mean proportion or just the detection of CNVs to call tumor cells.

## Compare InferCNV annotations to manual annotations

```{r}
all_classifications_df <- cnv_df |> 
  dplyr::left_join(manual_classifications_df) |>
  # relevel classifications for confusion matrix 
  dplyr::mutate(
    cnv_classification = forcats::fct_relevel(cnv_classification, "Tumor"),
    marker_gene_classification = forcats::fct_relevel(marker_gene_classification, "Tumor")
  )

caret::confusionMatrix(
  table(
    all_classifications_df$cnv_classification,
    all_classifications_df$marker_gene_classification
  )
)
```

```{r}
celltype_columns <- c(
  "singler_celltype_annotation",
  "cellassign_celltype_annotation"
)

# create jaccard matrices for SingleR and CellAssign compared to aneuploid/diploid 
jaccard_matrices <- celltype_columns |>
  purrr::map(\(name) {
    make_jaccard_matrix(
      cnv_df,
      "cnv_classification",
      name
    )
  }) |> 
  purrr::set_names("SingleR", "CellAssign")
```

```{r, fig.height=10, fig.width=7}
# Set heatmap padding option
heatmap_padding <- 0.2
ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(heatmap_padding, "in"))
# list of heatmaps looking at SingleR/ CellAssign vs tumor/normal 
heatmap <- jaccard_matrices |>
  purrr::imap(
    \(celltype_mat, celltype_method) {
      ComplexHeatmap::Heatmap(
        t(celltype_mat), # transpose because matrix rows are in common & we want a vertical arrangement
        col = circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue")),
        border = TRUE,
        ## Row parameters
        cluster_rows = TRUE,
        row_title = celltype_method,
        row_title_gp = grid::gpar(fontsize = 12),
        row_title_side = "left",
        row_names_side = "left",
        row_dend_side = "right",
        row_names_gp = grid::gpar(fontsize = 10),
        ## Column parameters
        cluster_columns = FALSE,
        column_title = "",
        column_title_gp = grid::gpar(fontsize = 12),
        column_names_side = "bottom",
        column_names_gp = grid::gpar(fontsize = 10),
        column_names_rot = 90,
        ## Legend parameters
        heatmap_legend_param = list(
          title = "Jaccard index",
          direction = "vertical",
          legend_width = unit(1.5, "in")
        ),
        show_heatmap_legend = celltype_method == "SingleR",
      )
    }) |>
  # concatenate vertically into HeatmapList object
  purrr::reduce(ComplexHeatmap::`%v%`) |>
  ComplexHeatmap::draw(
    heatmap_legend_side = "right",
    # add a margin to the heatmap so labels don't get cut off
    padding = unit(c(2, 20, 2, 2), "mm")
  )
```

## Check for known CNVs in Ewing's

There are a few known copy number variations in Ewing's sarcoma:

- Gain of Chr8
- Gain of Chr12
- Gain of Chr1p
- Loss of Chr16q

Although these are the most frequent, there are patients who do not have any of these alterations and patients that only have some of these alterations. 
See [Tirode et al.,](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4264969/) and [Crompton et al.](https://doi.org/10.1158/2159-8290.CD-13-1037).

```{r}
# umaps showing where cells are with gains/losses 
ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_dupli_chr8)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr8 gain") +
  theme_bw()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_dupli_chr12)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr12 gain") +
  theme_bw()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_dupli_chr1)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr1 gain") +
  theme_bw()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_loss_chr16)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr16 loss") +
  theme_bw()
```

From this we see that the group of tumor cells at the bottom have chr8 and chr1 gains with some having gains of chr12. 
It doesn't look like any cells have loss of chr16. 
The other interesting thing is that there is a group of cells that have gains of chr8, chr1, and chr12 that we did not classify as tumor cells. Should they be...

```{r}
ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = proportion_scaled_dupli_chr8)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr8 gain") +
  theme_bw() +
  scale_color_viridis_c()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = proportion_scaled_dupli_chr12)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr12 gain") +
  theme_bw() +
  scale_color_viridis_c()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = proportion_scaled_dupli_chr1)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr1 gain") +
  theme_bw() +
  scale_color_viridis_c()
```
Now when we look at the proportion, we see that only the group of tumor cells on the bottom seem to have a higher proportion of genes in chr8 that have a gain in expression. 
Based on the papers, we should see a gain in the whole chr8 and chr12, not just one arm or specific genes, so we expect a high value for the proportion. 
We see this for chr8 but not in chr12.

I think with the combination of marker gene expression and CNVs, the bottom group of cells represent the tumor cells.

## Do the same thing but without a reference set of cells?

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```