### 2024-05-13 analyses/doublet-detection/01_compare-doublet-methods.Rmd

---
title: "Compare doublet detection methods"
author:
  - Stephanie J. Spielman
date: "`r Sys.Date()`"
output: html_notebook
---

## Introduction

This notebook will compare doublet detection results from several methods across a set of ScPCA libraries (_not multiplexed ones_) with varying size and diagnoses. 
The methods include [`scDblFinder`](https://bioconductor.org/packages/3.18/scDblFinder) and three methods from [`scds`](https://bioconductor.org/packages/devel/bioc/vignettes/scds/inst/doc/scds.html): `cxds`, `bsds`, and `hybrid` (a combination of the previous two methods).

## Setup

### Packages

```{r packages}
# load required packages
library(SingleCellExperiment)
library(scDblFinder)
library(scds)
```

### Paths

#### Base directories

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")

# The path to this module
module_base <- file.path(repository_base, "analyses", "doublet-detection")
```

#### Input and output files

Set paths to input and output directories and files in the chunk below.

```{r paths}
# Input files

# Output files
```

## Analysis content

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```