### 2024-05-14 analyses/cell-type-ewings/template_notebooks/01-marker-genes.Rmd
add template notebooks

---
title: "Manual classification of tumor cells"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
params:
  sample_id: SCPCS000490
  library_id: SCPCL000822
  results_dir: "../results/annotate_tumor_cells_output"
  reference_cell_file: "../references/cell_lists/SCPCS000490/SCPCL000822_reference-cells.tsv"
---

## Introduction

This notebook looks at expression of marker genes for tumor cells in a Ewing sarcoma sample, `r {params$sample_id}`. 
We then see if we can use expression of marker genes to manually classify tumor and normal cells.
The main goal of this notebook is only to identify tumor cells, identification and labelling of the other cells is a separate question that we do not answer here.

- First we look at expression of each of the marker genes across all cells. 
- Then we use a z-transform prior to summing expression of all marker genes. 
Cells with a z-score for any marker gene > 0 are classified as tumor cells. 
- We look at marker genes for these tumor cells using `scran::findMarkers()` and compare to the tumor markers.
- Marker gene expression is examined across tumor and normal cells. 
- Finally, we compare the cells classified as tumor/normal manually to assignments from `SingleR` and `CellAssign`.

## Setup

```{r}
project_root <- here::here()
renv::load(project_root)
```

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_bw()
)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")
sample_dir <- file.path(data_dir, "SCPCP000015", params$sample_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-ewings")
```

```{r}
# source in helper functions: make_jaccard_matrix() and jaccard() 
jaccard_functions <- file.path(module_base, "scripts", "utils", "jaccard-functions.R")
source(jaccard_functions)
```

```{r paths}
# Input files
sce_filename <- glue::glue("{params$library_id}_processed.rds")
sce_file <- file.path(sample_dir, sce_filename)

marker_genes <- file.path(module_base, "references", "tumor-marker-genes.tsv")

# output tumor/normal classifications
fs::dir_create(params$results_dir)

classifications_filename <- glue::glue("{params$library_id}_tumor_normal_classifications.tsv")
output_classifications_file <- file.path(params$results_dir, classifications_filename)
```

```{r}
# read in processed sce
sce <- readr::read_rds(sce_file)

# read in marker genes table 
marker_genes_df <- readr::read_tsv(marker_genes) |> 
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |> 
  dplyr::distinct()

marker_genes <- marker_genes_df |> 
  dplyr::filter(cell_type == "tumor") |> 
  dplyr::pull(ensembl_gene_id)
```

```{r}
# read in normal and tumor cells for later comparisons 
if(file.exists(params$reference_cell_file)){
  ref_cells_df <- readr::read_tsv(params$reference_cell_file)
} else {
  ref_cells_df <- NULL 
}
```

## Analysis content

### Explore marker gene expression

The first thing we do here is just create a faceted UMAP showing the expression of each marker gene for tumor cells.

```{r}
# get the gene expression counts for all marker genes
marker_gene_exp <- logcounts(sce[marker_genes, ]) |>
  as.matrix() |>
  t() |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes")

# pull out the UMAP coordinates and make a data frame to use for plotting
umap_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
        \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")
      ) |> 
  # add in marker gene expression to dataframe 
  dplyr::left_join(marker_gene_exp, by = "barcodes") |> 
  # combine all genes into a single column for easy faceting
  tidyr::pivot_longer(
    cols = starts_with("ENSG"),
    names_to = "ensembl_gene_id",
    values_to = "gene_expression"
  ) |> 
  # join with marker gene df to get gene symbols for plotting 
  dplyr::left_join(marker_genes_df, by = c("ensembl_gene_id")) |>
  dplyr::select(barcodes, UMAP1, UMAP2, gene_symbol, ensembl_gene_id, gene_expression, cluster)
```

```{r}
# faceted umap showing a umap panel for each marker gene 
ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = gene_expression)) +
    geom_point(alpha = 0.1, size = 0.2) +
    facet_wrap(vars(gene_symbol)) +
    scale_color_viridis_c() +
    labs(
      color = "Log-normalized gene expression"
    ) +
    # remove axis numbers and background grid
    scale_x_continuous(labels = NULL, breaks = NULL) +
    scale_y_continuous(labels = NULL, breaks = NULL) +
    theme(
      aspect.ratio = 1,
      legend.position = "bottom",
      axis.title = element_text(size = 9, color = "black"),
      strip.text = element_text(size = 8),
      legend.title = element_text(size = 9),
      legend.text = element_text(size = 8)
    ) +
    guides(colour = guide_colorbar(title.position = "bottom", title.hjust = 0.5)) 
```

We can also look at the distributions for each marker gene. 
We would expect to see some sort of bimodal distribution separating cells that do and do not have expression of the marker gene.

```{r}
ggplot(umap_df, aes(x = gene_expression, fill = gene_symbol)) +
  geom_density() +
  facet_wrap(vars(gene_symbol)) +
  theme(legend.position = "none")
```

Now we will transform each of the gene expression vectors by generating z-scores and use a cutoff of z-score > 0 to call tumor cells.

```{r}
umap_df <- umap_df |> 
  dplyr::group_by(gene_symbol) |>
  # get z-scores for each gene
  dplyr::mutate(transformed_gene_expression = scale(gene_expression)[, 1]) |>
  dplyr::ungroup()
```

```{r}
ggplot(umap_df, aes(x = transformed_gene_expression, fill = gene_symbol)) +
  geom_density() +
  facet_wrap(vars(gene_symbol)) +
  theme(legend.position = "none")
```

### Classify tumor cells using marker genes only

To classify tumor cells, we use a cutoff of z-score > 0 to count that cell as a tumor cell. 
We will call cells as tumor if any marker gene is present (z-score > 0). 
This also means the sum of all z-scores > 0.

Below, we can get the sum of the transformed gene expression of all marker genes and plot in a single UMAP.

```{r}
# calculate sum gene expression across all marker genes in list 
marker_sum_exp <- umap_df |> 
  dplyr::group_by(barcodes) |> 
  dplyr::mutate(sum_exp = sum(transformed_gene_expression)) |> 
  dplyr::select(barcodes, UMAP1, UMAP2, sum_exp, cluster) |> 
  dplyr::distinct()

# plot mean gene expression 
ggplot(marker_sum_exp, aes(x = UMAP1, y = UMAP2, color = sum_exp)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_color_viridis_c()
```

Now let's classify any cell that has a sum of marker genes > 0 (after z-transformation) as tumor cells.

```{r}
# classify tumor cells based on presence of any marker genes 
marker_sum_exp <- marker_sum_exp |> 
  dplyr::mutate(sum_classification = dplyr::if_else(sum_exp > 0, "Tumor", "Normal"))

ggplot(marker_sum_exp, aes(x = UMAP1, y = UMAP2, color = sum_classification)) +
  geom_point(size = 0.5, alpha = 0.5)
```

This gives us a rough idea of cells that may be classified as tumor cells. 
Let's look at the expression of each marker gene in tumor vs. normal cells and see if we see distinct separation between all of them or only a subset.

```{r}
# add cell types from sum classification to plotting df 
celltype_df <- umap_df |> 
  dplyr::left_join(marker_sum_exp)

# make plot with gene on y-axis, expression on x-axis and then color density by tumor/normal 
# do all marker genes so differences between tumor/normal or only a subset? 
ggplot(celltype_df, aes(x=transformed_gene_expression, color = sum_classification)) +
  geom_density() +
  facet_wrap(vars(gene_symbol)) 
```

### Marker genes using `findMarkers`

Below, we will use `scran::findMarkers()` to find any markers between tumor and normal cells. 
Since we used marker genes to classify tumor cells, we expect that at least some of the top markers would be from our list of tumor markers.

```{r}
# add in tumor/normal classification to colData of SCE 
celltypes_only_df <- marker_sum_exp |> 
  dplyr::select(barcodes, sum_classification)

coldata_df <- as.data.frame(colData(sce)) |> 
  dplyr::left_join(celltypes_only_df)

colData(sce) <- DataFrame(coldata_df, row.names = rownames(coldata_df))

# find markers between tumor and normal 
markers <- scran::findMarkers(sce,
                              groups = sce$sum_classification,
                              pval.type = "all") 
```

```{r}
# pull out top markers for tumor cells 
tumor_markers <- markers$Tumor |> 
  as.data.frame() |> 
  tibble::rownames_to_column("ensembl_gene_id") |> 
  # join to get gene symbols to print out later
  dplyr::left_join(marker_genes_df) |>
  # just take those with FDR < 0.05 and logFC > 1.5
  dplyr::filter(FDR < 0.05, summary.logFC > 1.5) |> 
  dplyr::arrange(FDR)

# get any marker genes that are in both findMarkers and our curated list
marker_gene_intersect <- tumor_markers |> 
  dplyr::filter(cell_type == "tumor")

# print out genes that are marker genes for both 
marker_gene_intersect
```

### Looking at marker genes across tumor cells

We expect that cells we are labeling as tumor cells have expression of the tumor marker genes. 
Cells labeled as normal should not have expression of the markers.

Below we visualize the total number of cells that are classified as tumor or normal using the sum of marker gene expression. 
Then for each set of cells, how many of those actually express each marker gene.

We expect that good marker genes will be expressed in tumor cells but not normal cells.

```{r, fig.height=10, fig.width=10}
marker_plot_df <- celltype_df |> 
  dplyr::mutate(gene_expresed = dplyr::if_else(transformed_gene_expression > 0, TRUE, FALSE)) |> 
  dplyr::select(gene_symbol, gene_expresed, sum_classification)

ggplot(marker_plot_df, aes(x = sum_classification, fill = gene_expresed)) +
  geom_bar(stat = "count") +
  facet_grid(rows = vars(gene_symbol)) +
  scale_fill_manual(values = c("grey", "blue")) +
  theme_classic()

```

Looking at this might help give us an idea of genes that are specifically expressed in only tumor cells.

### Compare with SingleR and CellAssign classifictions

The last thing we will do here is compare the manual annotations with the annotations from `SingleR` and `CellAssign`.

Let's create a heatmap looking at the Jaccard similarity index between each of the assigned cell types and the tumor/normal classification.

```{r}
# get all assigned cell types in one dataframe 
celltypes_df <- as.data.frame(colData(sce)) |> 
  dplyr::select(barcodes, singler_celltype_annotation, cellassign_celltype_annotation) |> 
  dplyr::left_join(marker_sum_exp)

celltype_columns <- c(
  "singler_celltype_annotation",
  "cellassign_celltype_annotation"
)

# create jaccard matrices for SingleR and CellAssign compared to tumor/normal 
jaccard_matrices <- celltype_columns |>
  purrr::map(\(name) {
    make_jaccard_matrix(
      celltypes_df,
      "sum_classification",
      name
    )
  }) |> 
  purrr::set_names("SingleR", "CellAssign")
```

```{r, fig.height=10, fig.width=7}
# Set heatmap padding option
heatmap_padding <- 0.2
ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(heatmap_padding, "in"))

# list of heatmaps looking at SingleR/ CellAssign vs tumor/normal 
heatmap <- jaccard_matrices |>
  purrr::imap(
    \(celltype_mat, celltype_method) {
      ComplexHeatmap::Heatmap(
        t(celltype_mat), # transpose because matrix rows are in common & we want a vertical arrangement
        col = circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue")),
        border = TRUE,
        ## Row parameters
        cluster_rows = TRUE,
        row_title = celltype_method,
        row_title_gp = grid::gpar(fontsize = 12),
        row_title_side = "left",
        row_names_side = "left",
        row_dend_side = "right",
        row_names_gp = grid::gpar(fontsize = 10),
        ## Column parameters
        cluster_columns = FALSE,
        column_title = "",
        column_title_gp = grid::gpar(fontsize = 12),
        column_names_side = "bottom",
        column_names_gp = grid::gpar(fontsize = 10),
        column_names_rot = 90,
        ## Legend parameters
        heatmap_legend_param = list(
          title = "Jaccard index",
          direction = "vertical",
          legend_width = unit(1.5, "in")
        ),
        show_heatmap_legend = celltype_method == "SingleR",
      )
    }) |>
  # concatenate vertically into HeatmapList object
  purrr::reduce(ComplexHeatmap::`%v%`) |>
  ComplexHeatmap::draw(
    heatmap_legend_side = "right",
    # add a margin to the heatmap so labels don't get cut off
    padding = unit(c(2, 20, 2, 2), "mm")
  )
```

```{r}
# plot comparing marker gene expression in cells annotated as "normal" and "tumor" by SingleR and CellAssign 
# use already identified normal/tumor cell types 
if(!is.null(ref_cells_df)){

  # get list of normal cells 
  normal_cells <- ref_cells_df |> 
    dplyr::filter(reference_cell_class == "Normal") |> 
    dplyr::pull(barcodes) |> 
    unique()

  # get list of tumor cells 
  tumor_cells <- ref_cells_df |> 
    dplyr::filter(reference_cell_class == "Tumor") |> 
    dplyr::pull(barcodes) |> 
    unique()

  # only create this plot if normal and tumor cells are provided 
  if(length(normal_cells) > 0 && length(tumor_cells) > 0){

    # indicate cells that are part of normal and tumor lists
    select_celltype_df <- celltypes_df |> 
      dplyr::mutate(
        automated_annotation = dplyr::case_when(
          barcodes %in% normal_cells ~ "Normal cell",
          barcodes %in% tumor_cells ~ "Tumor cell"
        )
      ) |> 
      dplyr::select(barcodes, automated_annotation)

    # join with umap df
    annotation_plot_df <- umap_df |> 
      dplyr::left_join(select_celltype_df) |>
      dplyr::filter(!is.na(automated_annotation))

   ggplot(annotation_plot_df, aes(x = transformed_gene_expression, color = automated_annotation)) +
      geom_density() +
      facet_wrap(vars(gene_symbol)) 
  } 
} 
```

## Save outputs

```{r}
# get a dataframe with barcodes, sum_classification and cluster_classification 
celltypes_df <- celltypes_df |> 
  dplyr::select(
    barcodes,
    marker_gene_classification = sum_classification
  )

readr::write_tsv(celltypes_df, output_classifications_file)
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```