### 2024-05-20 analyses/cell-type-ewings/exploratory_analysis/04-infercnv.Rmd
scale proportion by number of genes

---
title: "Copy number inference with InferCNV"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  sample_id: SCPCS000490
  library_id: SCPCL000822
---

## Introduction

Here we look at using [`InferCNV`](https://github.com/broadinstitute/infercnv) to identify tumor and normal cells in `r params$sample_id`.

Before running this notebook, `InferCNV` is run using the `run-infercnv.R` script. 
This script runs `InferCNV` once with Endothelial cells as the reference cells and once without any cells as the reference.

Here, we look the copy number variations found reported by `InferCNV` and use that copy number information to call cells as tumor or normal cells. 
We then compare these classifications to manual classifications and annotations from `SingleR` and `CellAssign`.

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
  library(dplyr)
})
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")
sample_dir <- file.path(data_dir, "SCPCP000015", params$sample_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-ewings")
```

```{r}
# source in helper functions
jaccard_functions <- file.path(module_base, "scripts", "utils", "jaccard-functions.R")
source(jaccard_functions)
```

```{r paths}
# Input files
sce_filename <- glue::glue("{params$library_id}_processed.rds")
sce_file <- file.path(sample_dir, sce_filename)

# tumor normal classifications 
classifications_results_dir <- file.path(module_base, "results", "marker_gene_analysis")
classifications_filename <- glue::glue("{params$library_id}_tumor_normal_classifications.tsv")
classifications_file <- file.path(classifications_results_dir, classifications_filename)

# output from running infercnv 
results_dir <- file.path(module_base, "results", "infercnv", params$library_id)

all_results_dir_list <- list(
  no_ref = file.path(results_dir, "no_reference"),
  with_ref = file.path(results_dir, "endothelial_reference")
)

# png ref files 
png_files <- all_results_dir_list |>
  purrr::map_chr(\(dir) file.path(dir, glue::glue("{params$library_id}_infercnv.png"))) |> 
  # make sure png files exists
  purrr::map(\(file)
    stopifnot("Missing PNG file" = file.exists(file))
  )

# with ref cnv metadata file 
cnv_metadata_file <- file.path(results_dir, "endothelial_reference", glue::glue("{params$library_id}_cnv-metadata.tsv"))

# full infercnv object with information on genes/ chr to use later 
final_cnv_obj_file <- file.path(results_dir, "endothelial_reference", glue::glue("{params$library_id}_cnv-obj.rds"))
```

```{r}
# read in sce object
sce <- readr::read_rds(sce_file)

# read in cnv metadata 
# use read.delim instead of readr because we want the rownames to be read in properly 
cnv_metadata_df <- read.delim(cnv_metadata_file) |> 
  tibble::rownames_to_column("barcodes")

# read in tumor normal classifications 
manual_classifications_df <- readr::read_tsv(classifications_file)

# read in cnv object 
final_cnv_obj <- readr::read_rds(final_cnv_obj_file)
```

## InferCNV Heatmap

### Endothelial cells as reference

Below is the heatmap produced by `InferCNV` showing the CNV's for each cell using endothelial cells as a reference. 
Each row is a cell and each column is a gene. 
The top group contains all cells that were used as the reference group and the bottom is all other cells. 
The color indicates the relative expression intensity of each gene with the darker colors either indicating a gain of expression or loss of expression compared to the reference cells.

From the [`InferCNV` documentation](https://github.com/broadinstitute/inferCNV/wiki/Interpreting-the-figure):

> The normal cells in the top heatmap define baseline expression for genes in normal cells. This baseline distribution of normal gene expression is subtracted from both the normal cells from which was defined as well as the tumor cells. Afterwards, the normal cell expression heatmap should be largely devoid of signal, with the exception of certain outlier gene expression values in certain cells. Removing this baseline of normal expression signal from the tumor cells should reveal those chromosomal regions that have significantly more or less expression than the normal cells, highlighting likely amplified or deleted whole chromosomes or large chromosomal regions.

The dendogram on the left side is hierarchical clustering of all non-reference cells. 
The color annotation next to it indicates subdivisions of the dendogram by cutting it in `k_obs_groups` groups, with the default `k_obs_groups` = 1.

![InferCNV Heatmap - Endothelial cells as reference](`r png_files["with_ref"]`)

From this heatmap we see that there is a distinct group of cells within the bottom portion of the plot that appear to have a lot more CNVs than the rest of the cells. 
It's probable that these correspond to the tumor cells, but `InferCNV` does not provide any classification of cells as tumor or normal, it just provides information on copy number variations.

### No reference

Below is the heatmap produced by `InferCNV` showing the CNV's for each cell without using a reference.

From the [`InferCNV` documentation](https://github.com/broadinstitute/inferCNV/wiki/Running-InferCNV):

> Note, if you do not have reference cells, you can set ref_group_names=NULL, in which case the average signal across all cells will be used to define the baseline. This can work well when there are sufficient differences among the cells included (ie. they do not all show a chromosomal deletion at the same place).

Given this information, I wouldn't expect the CNV inference to work well because we see that all cells that do have prominent CNVs have the same CNVs. 
There is not a lot of variation.

![InferCNV Heatmap - No reference](`r png_files["no_ref"]`)

Looking at this we see a small group of cells that may have some variations, but they are different from the variations seen above. 
I don't know that using this without a reference is the best call since it's unlikely that it will identify distinct groups of cells from tumors with quiet genomes with no baseline. 
The rest of the analysis will use the results from running `InferCNV` with endothelial cells as the reference.

## Using CNVs to annotate tumor cells

`InferCNV` uses an HMM-based model to predict whether or not a copy number variation (either loss or gain) is present for each chromosome in each cell. 
For each predicted CNV region, a posterior probability is calculated for that CNV for each cell.

From the [`InferCNV` documentation](https://github.com/broadinstitute/inferCNV/wiki/inferCNV-HMM-based-CNV-Prediction-Methods):

> CNV regions with mean posterior probabilities of being normal (no CNV) that are above a maximum threshold are removed as likely false positive predictions.

Before we can make any plots we need to add some metadata that contains CNV information on a per cell level to the cell metadata.

```{r}
# make df for plotting
coldata_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
        \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")
      ) |> 
  # add in per cell cnv information 
  dplyr::left_join(cnv_metadata_df, by = c("barcodes"))

# remove MT and GL chr, comments in docs recommend that these are not useful and can be inaccurate 
chr_columns_to_keep <- which(!stringr::str_detect(names(coldata_df), "chrMT|chrGL000219.1|chrGL000194.1"))
all_cnv_df <- coldata_df[, chr_columns_to_keep]
```

The metadata that we added includes the following columns for each chromosome:

- `has_cnv_chr1`
- `has_loss_chr1`
- `has_dupli_chr1`
- `proportion_cnv_chr1`
- `proportion_loss_chr1`
- `proportion_dupli_chr1`
- `proportion_scaled_cnv_chr1`
- `proportion_scaled_loss_chr1`
- `proportion_scaled_dupli_chr1`

The `has` columns are boolean columns indicating if that cell has presence of a CNV, loss, or duplication event. 
If `has_cnv` is TRUE if either `has_loss` or `has_dupli` is TRUE.

See [the comments in this issue](https://github.com/broadinstitute/infercnv/issues/208) explaining what the contents of the columns are.

One approach to classifying tumor cells is to look for cells that have a lot of CNVs. 
Let's start by just summing up the total number of chromosomes with CNVs per cell and plotting that.

```{r}
# grab any columns that have has_cnv 
has_cnv_cols <- names(all_cnv_df)[startsWith(colnames(all_cnv_df), "has_cnv")]

reference_cells_only <- all_cnv_df |> 
  dplyr::filter(stringr::str_detect(subcluster, "reference")) |> 
  dplyr::mutate(has_cnv = NA,
                z_scaled_cnv = NA)

unknown_cells_only <- all_cnv_df |>
  dplyr::filter(stringr::str_detect(subcluster, "unknown")) |> 
  # get total number of chromosomes with cnvs per cell and then scale that number 
  dplyr::mutate(has_cnv = rowSums(across(has_cnv_cols)),
                z_scaled_cnv = scale(has_cnv))

# combine back together for plotting 
cnv_df <- bind_rows(reference_cells_only, unknown_cells_only)
```

```{r}
# plot the total number of CNVs
ggplot(cnv_df, aes(x = UMAP1, y= UMAP2, color = has_cnv)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() +
  scale_color_viridis_c()
```
This shows that the total CNVs are highest in the group of cells that correspond to the same cells that are manually annotated as tumor cells, which is promising. 
Those cells in grey (bottom left) are cells used as a reference. 
If we were to use total CNVs we would need some sort of cut off for the number. 
Let's look at the distribution of total CNVs and also the scaled total CNVs to see if we can identify a cut off.

```{r}
ggplot(cnv_df, aes(x = has_cnv)) +
  geom_density() +
  theme_bw() +
  labs(x = "Total number of CNVs")

ggplot(cnv_df, aes(x= z_scaled_cnv)) + 
  geom_density() +
  theme_bw() +
  labs(x = "z-scaled chromosome CNV count")
```

Looking at the scaled number of CNVs it does look like we could use 0 as the cutoff and call any cell that has > 0 CNVs (after scaling) as tumor and all other cells as normal.

```{r}
# add cnv classification 
cnv_df <- cnv_df |> 
  dplyr::mutate(cnv_classification = dplyr::if_else(z_scaled_cnv > 0, "Tumor", "Normal"))

ggplot(cnv_df, aes(x = UMAP1, y= UMAP2, color = cnv_classification)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() 
```

Another option is to use the scaled proportion of CNVs which tells us the proportion of genes on that chromosome that are part of the CNV.

```{r}
proportion_cols <- names(cnv_df)[startsWith(colnames(cnv_df), "proportion_scaled_cnv")]

# get total number of genes per chromosome to use in weighted mean 
chr_weights <- final_cnv_obj@gene_order |> 
  as.data.frame() |> 
  dplyr::count(chr) |> 
  # only keep chr 1-22, drops MT and GL chr
  dplyr::filter(chr %in% glue::glue("chr{seq(1,22)}")) |> 
  dplyr::pull(n)

# get mean proportion of cnvs per cell weighted by chromsome size (number of genes)
cnv_df <- cnv_df |> 
  dplyr::rowwise() |> 
  # calculate weighted mean, excluding normal cells 
  dplyr::mutate(mean_proportion = ifelse(!is.na(has_cnv), weighted.mean(across(proportion_cols), num_genes_df$n), NA)) |> 
  dplyr::ungroup()
```

```{r}
ggplot(cnv_df, aes(x = UMAP1, y= UMAP2, color = mean_proportion)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() +
  scale_color_viridis_c()
```

```{r}
ggplot(cnv_df, aes(x = mean_proportion)) +
  geom_density() +
  theme_bw()
```
It looks like we could use a cutoff of 0.1 here to classify tumor/normal cells. 
Let's see how this matches up with the tumor/normal classifications we defined just using total CNVs.

```{r}
ggplot(cnv_df, aes(x = mean_proportion, color = cnv_classification)) +
  geom_density() +
  theme_bw()
```

As expected, the cells we labeled as tumor have a higher mean proportion value then those that we labeled as normal. 
I'm not really sure which is "better" to use the mean proportion or just the detection of CNVs to call tumor cells.

```{r}
# add cnv classification 
cnv_df <- cnv_df |> 
  dplyr::mutate(cnv_mean_classification = dplyr::if_else(mean_proportion > 0.1, "Tumor", "Normal"))

ggplot(cnv_df, aes(x = UMAP1, y= UMAP2, color = cnv_mean_classification)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() 
```

For the rest of the notebook, we will use the classifications of tumor cells based on the mean proportion of total CNV occurrences.

## Compare InferCNV annotations to manual annotations

Below we can compare these annotations from InferCNV to classifying tumor cells manually with marker genes.

```{r}
all_classifications_df <- cnv_df |> 
  dplyr::left_join(manual_classifications_df) |>
  # relevel classifications for confusion matrix 
  dplyr::mutate(
    cnv_classification = forcats::fct_relevel(cnv_mean_classification, "Tumor"),
    marker_gene_classification = forcats::fct_relevel(marker_gene_classification, "Tumor")
  )

caret::confusionMatrix(
  table(
    all_classifications_df$cnv_classification,
    all_classifications_df$marker_gene_classification
  )
)
```

This looks pretty consistent with what we see when we annotate cells manually.

We can also compare to the existing annotations obtained using `SingleR` and `CellAssign` as part of `scpca-nf`.

```{r}
celltype_columns <- c(
  "singler_celltype_annotation",
  "cellassign_celltype_annotation"
)

# create jaccard matrices for SingleR and CellAssign compared to aneuploid/diploid 
jaccard_matrices <- celltype_columns |>
  purrr::map(\(name) {
    make_jaccard_matrix(
      cnv_df,
      "cnv_mean_classification",
      name
    )
  }) |> 
  purrr::set_names("SingleR", "CellAssign")
```

```{r, fig.height=10, fig.width=7}
# Set heatmap padding option
heatmap_padding <- 0.2
ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(heatmap_padding, "in"))
# list of heatmaps looking at SingleR/ CellAssign vs tumor/normal 
heatmap <- jaccard_matrices |>
  purrr::imap(
    \(celltype_mat, celltype_method) {
      ComplexHeatmap::Heatmap(
        t(celltype_mat), # transpose because matrix rows are in common & we want a vertical arrangement
        col = circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue")),
        border = TRUE,
        ## Row parameters
        cluster_rows = TRUE,
        row_title = celltype_method,
        row_title_gp = grid::gpar(fontsize = 12),
        row_title_side = "left",
        row_names_side = "left",
        row_dend_side = "right",
        row_names_gp = grid::gpar(fontsize = 10),
        ## Column parameters
        cluster_columns = FALSE,
        column_title = "",
        column_title_gp = grid::gpar(fontsize = 12),
        column_names_side = "bottom",
        column_names_gp = grid::gpar(fontsize = 10),
        column_names_rot = 90,
        ## Legend parameters
        heatmap_legend_param = list(
          title = "Jaccard index",
          direction = "vertical",
          legend_width = unit(1.5, "in")
        ),
        show_heatmap_legend = celltype_method == "SingleR",
      )
    }) |>
  # concatenate vertically into HeatmapList object
  purrr::reduce(ComplexHeatmap::`%v%`) |>
  ComplexHeatmap::draw(
    heatmap_legend_side = "right",
    # add a margin to the heatmap so labels don't get cut off
    padding = unit(c(2, 20, 2, 2), "mm")
  )
```

This looks as we expect with mostly muscle cells lining up with tumor cells.

## Check for known CNVs in Ewing's

There are a few known copy number variations in Ewing's sarcoma:

- Gain of Chr8
- Gain of Chr12
- Gain of Chr1p
- Loss of Chr16q

Although these are the most frequent, there are patients who do not have any of these alterations and patients that only have some of these alterations. 
See [Tirode et al.,](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4264969/) and [Crompton et al.](https://doi.org/10.1158/2159-8290.CD-13-1037).

Below we will plot whether or not the gain or loss is present in each cell according to `InferCNV`.

```{r}
# umaps showing where cells are with gains/losses 
ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_dupli_chr8)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr8 gain") +
  theme_bw()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_dupli_chr12)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr12 gain") +
  theme_bw()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_dupli_chr1)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr1 gain") +
  theme_bw()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_loss_chr16)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr16 loss") +
  theme_bw()
```

From this we see that the group of tumor cells at the bottom have chr8 and chr1 gains with some having gains of chr12. 
It doesn't look like any cells have loss of chr16. 
The other interesting thing is that there is a group of cells that have gains of chr8, chr1, and chr12 that we did not classify as tumor cells. Should they be...

We can also look at the proportion of the chromosome with the gain.

```{r}
ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = proportion_scaled_dupli_chr8)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr8 gain") +
  theme_bw() +
  scale_color_viridis_c()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = proportion_scaled_dupli_chr12)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr12 gain") +
  theme_bw() +
  scale_color_viridis_c()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = proportion_scaled_dupli_chr1)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr1 gain") +
  theme_bw() +
  scale_color_viridis_c()
```

Now when we look at the proportion, we see that only the group of tumor cells on the bottom seem to have a higher proportion of genes in chr8 that have a gain in expression. 
Based on the papers, we should see a gain in the whole chr8 and chr12, not just one arm or specific genes, so we expect a high value for the proportion. 
We see this for chr8 but not in chr12. 
However, when you look at the heatmap you see that most of Chr8 is included in the gain, but only a small part of Chr12 has a gain, so these plots are not surprising.

I think with using the combination of marker gene expression and CNVs, the bottom group of cells represent the tumor cells.

## Conclusions

- If we have a known group of normal cells, `InferCNV` can be helpful in identifying cells that could be tumor cells, e.g., cells that have a high number of CNVs. 
- Using `InferCNV` without a group of normal cells does not seem feasible for this particular sample since all tumor cells appear to have the same set of CNVs with little variation. 
- We can define tumor cells by setting a cutoff for the number of CNVs that must be present in a cell, but I'm not sure how transferrable this is to other samples. 
- `InferCNV` can be helpful in looking at specific CNVs that are expected in a tumor type, something that is harder to look at in a tool like `CopyKAT`.

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```