### 2024-05-28 analyses/doublet-detection/exploratory-analyses/02_explore-benchmark-results.Rmd
notebook to explore individual benchmarking results on each benchmark dataset, and associated environment updates

---
params: 
  dataset: "hm-6k"
title: "`r glue::glue('Doublet benchmarking results: {params$dataset}')`"
author: Stephanie J. Spielman
date: "`r Sys.Date()`"
output: html_notebook
---

The goal of this notebook is to explore doublet inferences run on ground-truth benchmarking data. 
These three doublet detection methods are explored:

- `scDblFinder`
- `scrublet`
- A version of `cxds` calculated by `scDblFinder`.
Unlike the other two methods, this one reports only a score and does not attempt to make single/doublet calls, so we use a `0.5` threshold to make `cxds` calls.

This notebook explores inferences made on a single dataset at a time (here, **`r params$dataset`**) with the goal of examining the relationship between scores and accuracy of calls (if made) compared to the ground-truth values.
Note that ground-truth calls are available only for heterotypic doublets, which these methods are more tuned for in the first place (vs homotypic).

## Setup

### Packages

```{r packages}
library(SingleCellExperiment)
library(readr)
library(ggplot2)
library(caret)

theme_set(theme_bw())
```

### Paths

#### Base directories

```{r base paths}
module_base <- rprojroot::find_root(rprojroot::is_renv_project)
data_dir <- file.path(module_base, "scratch", "benchmark_datasets", params$dataset)
result_dir <- file.path(module_base, "results", "benchmark_results")
```

#### Input and output files

Set paths to input and output directories and files in the chunk below.

```{r paths}
# Input files
scdbl_tsv <- file.path(
  result_dir, 
  glue::glue("{params$dataset}_scdblfinder.tsv")
)
scrub_tsv <- file.path(
  result_dir, 
  glue::glue("{params$dataset}_scrublet.tsv")
)
truth_tsv <- file.path(
  data_dir, 
  glue::glue("{params$dataset}_ground-truth.tsv")
)

# Output files
```

### Read and prepare input data

First, we'll read in and combine TSV files with doublet results.

```{r}
scdbl_df <- read_tsv(scdbl_tsv) |>
  dplyr::select(
    barcodes,
    cxds_score, 
    scdbl_score = score, 
    scdbl_prediction  = class
  ) |>
  # add cxds calls at 0.5 threshold
  dplyr::mutate(
    cxds_prediction = dplyr::if_else(
      cxds_score >= 0.5,
      "doublet",
      "singlet"
    )
  )

scrub_df <- read_tsv(scrub_tsv)

true_df <- read_tsv(truth_tsv)

doublet_df <- dplyr::inner_join(
  scdbl_df,
  scrub_df, 
  by = "barcodes"
) |>
  dplyr::inner_join(
    true_df, 
    by = "barcodes"
  )
```

## Explore results: Score distributions

Both `scDblFinder` and `scrublet` make doublet calls based on an internally-determined threshold, which may not be optimal.

These sections explore the score distributions and how they relate to calls, including a confusion matrix and associated performance metrics. 
We're particularly interested in "Balanced accuracy," as there is a much higher proportion of singlets vs doublets.

### scDblFinder results

```{r}
ggplot(doublet_df) +
  aes(x = ground_truth,
      y = scdbl_score, 
      color = scdbl_prediction) + 
  geom_jitter(
    alpha = 0.2, 
    width = 0.1
  ) + 
  labs(
    x = "Ground truth call",
    y = "scDblFinder score",
    color = "scDblFinder prediction"
  )

confusion <-  caret::confusionMatrix(
    # truth should be first
    table(
      "Truth" = doublet_df$ground_truth,
      "Prediction" = doublet_df$scdbl_prediction
    ), 
    positive = "doublet"
)

print(confusion)
```

### scrublet results

```{r}
ggplot(doublet_df) +
  aes(x = ground_truth,
      y = scrublet_score, 
      color = scrublet_prediction) + 
  geom_jitter(
    alpha = 0.2, 
    width = 0.1
  ) + 
  labs(
    x = "Ground truth call",
    y = "scrublet score",
    color = "scrublet prediction"
  )

confusion <-  caret::confusionMatrix(
    # truth should be first
    table(
      "Truth" = doublet_df$ground_truth,
      "Prediction" = doublet_df$scrublet_prediction
    ), 
    positive = "doublet"
)

print(confusion)
```

### cxds results

> This method does not call droplets as singlet/doublet.
> Instead, performance is assessed below at a threshold of 0.5.

```{r}
ggplot(doublet_df) +
  aes(x = ground_truth,
      y = cxds_score, 
      color = cxds_prediction) + 
  geom_jitter(
    alpha = 0.2, 
    width = 0.1
  ) + 
  labs(
    x = "Ground truth call",
    y = "cxds score",
    color = "cxds prediction"
  )

confusion <-  caret::confusionMatrix(
    # truth should be first
    table(
      "Truth" = doublet_df$ground_truth,
      "Prediction" = doublet_df$cxds_prediction
    ), 
    positive = "doublet"
)

print(confusion)
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```