### 2024-06-03 analyses/doublet-detection/template-notebooks/03_compare-benchmark-results.Rmd
PCA section for comparison notebook

---
title: "Comparison among doublet predictions on benchmarking datasets"
author: Stephanie J. Spielman
date: "`r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_depth: 4
    code_folding: hide
---

## Setup

### Packages

```{r packages}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(ggplot2)
  library(patchwork)
  library(caret)
  library(UpSetR)
})

theme_set(theme_bw())
```

### Paths

```{r base paths}
module_base <- rprojroot::find_root(rprojroot::is_renv_project)
data_dir <- file.path(module_base, "scratch", "benchmark-datasets")
result_dir <- file.path(module_base, "results", "benchmark-results")
```

### Functions

```{r}
plot_pca_calls <- function(df, 
                           color_column, 
                           pred_column,
                           color_lab) {
  # Plot PCs colored by singlet/doublet, showing doublets on top
  # df is expected to contain columns PC1, PC2, `color_column`, and `pred_column`. These should _not_ be provided as strings
  ggplot(df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(
    size = 0.75, 
    alpha = 0.6
  ) +
  scale_color_manual(name = color_lab, values = c("black", "lightblue")) + 
  geom_point(
    data = dplyr::filter(df, {{color_column}} == "doublet"), 
    color = "black",
    size = 0.75
  ) +
  theme(
    legend.title.position = "top",
    legend.position = "bottom"
  )
}

plot_pca_metrics <- function(df, color_column, false_colors) {
  # Plot PCs colored by performance metric, showing false calls on top
  # false_colors is a vector of colors used for fn and fp points
  # df is expected to contain columns PC1, PC2, and `color_column`. This should _not_ be provided as a string.
  ggplot(df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(
    size = 0.75, 
    alpha = 0.6
  ) + 
  geom_point(
    data = dplyr::filter(df, {{color_column}} %in% false_colors), 
    size = 0.75
  ) +
  scale_color_identity() +
  theme(legend.position = "none")
}
```

### Read and prepare input data

First, we'll read in and combine doublet results into a list of data frames for each dataset.
We'll also two columns for each dataset:

- `consensus_call`, which will be "doublet" if _all_ methods predict "doublet," and "singlet" otherwise
- `call_type`, which will classify the consensus call as one of "tp", "tn", "fp", or "fn" (true/false positive/negative)

```{r paths}
# find all dataset names to process:
dataset_names <- list.files(result_dir, pattern = "*_scrublet.tsv") |>
  stringr::str_remove("_scrublet.tsv")
```

```{r read_data}
confusion_colors <- c(
  "tp" = "lightblue",
  "tn" = "pink",
  "fp" = "blue",
  "fn" = "firebrick2"
)

# Read in and data for analysis
doublet_df_list <- dataset_names |>
  purrr::map(
    \(dataset) {

      scdbl_tsv <- file.path(result_dir, glue::glue("{dataset}_scdblfinder.tsv"))
      scrub_tsv <- file.path(result_dir, glue::glue("{dataset}_scrublet.tsv"))
      sce_file <- file.path(data_dir, dataset, glue::glue("{dataset}_sce.rds"))

      scdbl_df <- scdbl_tsv |>
        readr::read_tsv(show_col_types = FALSE) |>
        dplyr::select(
          barcodes,
          cxds_score, 
          scdbl_score = score, 
          scdbl_prediction  = class
        ) |>
        # add cxds calls at 0.75 threshold
        dplyr::mutate(
          cxds_prediction = dplyr::if_else(
            cxds_score >= 0.75,
            "doublet",
            "singlet"
          )
        )

      scrub_df <- readr::read_tsv(scrub_tsv, show_col_types = FALSE)

      # grab ground truth and PCA coordinates
      sce <- readr::read_rds(sce_file)
      sce_df <- scuttle::makePerCellDF(sce, use.dimred = "PCA") |>
        tibble::rownames_to_column(var = "barcodes") |>
        dplyr::select(barcodes,
                      ground_truth = ground_truth_doublets, 
                      PC1 = PCA.1, 
                      PC2 = PCA.2)
      rm(sce)

      dataset_df <- scdbl_df |>
        dplyr::left_join(
          scrub_df, 
          by = "barcodes"
        ) |>
        dplyr::left_join(
          sce_df, 
          by = "barcodes"
        )

      # Add a consensus call column
      dataset_df <- dataset_df |>
        dplyr::rowwise() |>
        dplyr::mutate(consensus_call = dplyr::if_else(
          all(
            c(scdbl_prediction, scrublet_prediction, cxds_prediction) == "doublet"
          ),
          "doublet", 
          "singlet"
        )) |>
        dplyr::mutate(
          call_type = dplyr::case_when(
            consensus_call == "doublet" && ground_truth == "doublet" ~ "tp",
            consensus_call == "singlet" && ground_truth == "singlet" ~ "tn",
            consensus_call == "doublet" && ground_truth == "singlet" ~ "fp",
            consensus_call == "singlet" && ground_truth == "doublet" ~ "fn"
          ), 
          # set associated plotting colors
          call_type_color = dplyr::case_when(
            call_type == "tp" ~ unname(confusion_colors["tp"]),
            call_type == "tn" ~ unname(confusion_colors["tn"]),
            call_type == "fp" ~ unname(confusion_colors["fp"]),
            call_type == "fn" ~ unname(confusion_colors["fn"])
          )
          )

      return(dataset_df)
    }
  )
names(doublet_df_list) <- dataset_names
```

## Visualizations

### PCA

This section plots the PCA for each dataset, with three color schemes from left to right:
- Ground truth doublets are shown in black
- Consensus doublets are shown in black
- Points are colored as tp, tn, fp, or fn based on comparing the consensus call to the ground truth

```{r, fig.width = 12, fig.height = 6}
# Make a legend for the confusion-colored PCA
legend_plot <- data.frame(
  x = factor(names(confusion_colors), levels = names(confusion_colors)), y = 1:4
) |>
 ggplot(aes(x = x, y = y, color = x)) + 
  geom_point(size = 3) + 
  scale_color_manual(name = "Metric", values = confusion_colors) 
confusion_legend <- ggpubr::get_legend(legend_plot) |> ggpubr::as_ggplot()

doublet_df_list |>
  purrr::iwalk(
    \(df, dataset) {

      # First, ground truth
      p1 <- plot_pca_calls(
        df, 
        color_column = ground_truth, 
        color_lab = "Ground truth"
      )

      # Second, consensus call
      p2 <- plot_pca_calls(
        df, 
        color_column = consensus_call, 
        color_lab = "Consensus call"
      )

      # Third, call type
      p3 <- plot_pca_metrics(
        df,
        call_type_color,
        false_colors = unname(c(confusion_colors["fn"], confusion_colors["fp"]))
      )

      # combine and plot
      plot( p1 + p2 + p3 + confusion_legend + plot_annotation(glue::glue("PCA for {dataset}")) + plot_layout(ncol=4, widths = c(1,1,1,0.25)) )
    }
  )
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```