### 2024-06-05 analyses/cell-type-ewings/cell_type_annotations/SCPCS000490/SCPCL000822_tumor-cell-validation.Rmd
Apply suggestions from code review
Co-authored-by: Joshua Shapiro <josh.shapiro@ccdatalab.org>

---
title: "SCPCL000822 - Validation of tumor cells"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

## Introduction

In this notebook we compare tumor cell annotations obtained from marker genes, `CopyKAT`, and `InferCNV` to identify a group of tumor cells we are confident in.

- First we look at the cells that are annotated as tumor cells by each method. 
Using this information, we call any cell that is annotated as a tumor cell by both `CopyKAT` and `InferCNV` as a tumor cell. 
- We then validate that these are in fact tumor cells by doing the following: 
  - Looking at marker gene expression. 
  - Looking at expression of EWS-FLI1 target gene sets. 
  - Confirming presence of known CNVs in Ewing sarcoma

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_bw()
)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "2024-05-01")
sample_dir <- file.path(data_dir, "SCPCP000015", "SCPCS000490")

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-ewings")
```

```{r}
# Input files
sce_file <- file.path(sample_dir, "SCPCL000822_processed.rds")
marker_genes <- file.path(module_base, "references", "tumor-marker-genes.tsv")

# results from annotation workflow 
results_dir <- file.path(module_base, "results", "annotate_tumor_cells_output", "SCPCS000490")

marker_gene_results_file <- file.path(results_dir, "SCPCL000822_tumor-normal-classifications.tsv")
copykat_predictions_file <- file.path(results_dir, "SCPCL000822_copykat-classifications.tsv")

infercnv_predictions_file <- file.path(results_dir, "SCPCL000822_infercnv-classifications.tsv")
infercnv_metadata_file <- file.path(results_dir, "infercnv", "SCPCL000822_cnv-metadata.tsv")

geneset_scores_file <- file.path(results_dir, "SCPCL000822_gene-set-scores.tsv")

# output files 
final_annotations_dir <- file.path(module_base, "cell_type_annotations", "SCPCS000490")
fs::dir_create(final_annotations_dir)
final_annotations_file <- file.path(final_annotations_dir, "SCPCL000822_tumor-classifications.tsv")
```

```{r}
# read in all files 
sce <- readr::read_rds(sce_file)

# read in marker genes table 
marker_genes_df <- readr::read_tsv(marker_genes) |> 
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |> 
  dplyr::distinct()

marker_gene_results_df <- readr::read_tsv(marker_gene_results_file)

# copykat output
copykat_results_df <- readr::read_tsv(copykat_predictions_file) |> 
  # only select no_ref classification and rename to be consistent with naming of other classifications
  dplyr::select(barcodes, copykat_classification = no_ref) |> 
  dplyr::mutate(
    copykat_classification = dplyr::case_when(
      copykat_classification == "aneuploid" ~ "Tumor",
      copykat_classification == "diploid" ~ "Normal",
      .default = copykat_classification
    )
  )

# infercnv output 
infercnv_results_df <- readr::read_tsv(infercnv_predictions_file)
infercnv_metadata_df <- read.delim(infercnv_metadata_file) |> 
  tibble::rownames_to_column("barcodes")

# remove MT and GL chr, comments in docs recommend that these are not useful and can be inaccurate 
chr_columns_to_keep <- which(!stringr::str_detect(names(infercnv_metadata_df), "chrMT|chrGL"))
infercnv_metadata_df <- infercnv_metadata_df[, chr_columns_to_keep]

# gene set scores 
geneset_scores_df <- readr::read_tsv(geneset_scores_file)
```

```{r}
# grab all coldata and add in all classifications and results to use for plotting throughout notebook 
# get list of marker genes to use 
marker_genes <- marker_genes_df |> 
  dplyr::filter(cell_type == "tumor") |> 
  dplyr::pull(ensembl_gene_id)

# get the gene expression counts for all marker genes
marker_gene_exp <- logcounts(sce[marker_genes, ]) |>
  as.matrix() |>
  t() |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes")

# pull out the UMAP coordinates and make a data frame to use for plotting
coldata_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
        \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")
      ) |> 
  # add in marker gene expression to dataframe 
  dplyr::left_join(marker_gene_exp, by = "barcodes") |> 
  # combine all genes into a single column for easy faceting
  tidyr::pivot_longer(
    cols = starts_with("ENSG"),
    names_to = "ensembl_gene_id",
    values_to = "gene_expression"
  ) |> 
  # join with marker gene df to get gene symbols for plotting 
  dplyr::left_join(marker_genes_df, by = "ensembl_gene_id") |>
  # get rid of excess columns
  dplyr::select(barcodes, UMAP1, UMAP2, gene_symbol, ensembl_gene_id, gene_expression) |> 
  # add in marker gene classifications
  dplyr::left_join(marker_gene_results_df, by = "barcodes") |> 
  # add in copykat results 
  dplyr::left_join(copykat_results_df, by = "barcodes") |> 
  # add in infercnv results and metadata
  dplyr::left_join(infercnv_results_df, by = "barcodes") |> 
  dplyr::left_join(infercnv_metadata_df, by = "barcodes") |> 
  # gene set scores 
  dplyr::left_join(geneset_scores_df, by = "barcodes")

```

## Identify tumor cell population

Below is a UMAP showing which cells are labeled as tumor cells using each of the various classification methods.

```{r}
# plot tumor cells for each classification method 
tumor_umap_df <- coldata_df |> 
  dplyr::select(barcodes, UMAP1, UMAP2, marker_gene_classification, copykat_classification, cnv_proportion_classification) |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  )

ggplot(tumor_umap_df, aes(x = UMAP1, UMAP2, color = classification)) +
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(vars(method))
```

This plot shows us that there is a pretty clear consensus that most of the cells in the bottom left group are tumor cells. 
From the original marker gene analysis we know that not all marker genes are expressed and that there is quite a bit of variance in expression. 
Because of that and the fact that both CNV inference methods show similar annotations, we will classify any cells that are labeled as tumor by both `CopyKAT` and `InferCNV` as tumor cells.

```{r}
# create a combined classification between inferCNV and copyKAT
# cells that are classified as tumor by both are kept as tumor, all other are normal 
coldata_df <- coldata_df |> 
  dplyr::mutate(
    combined_cnv_classification = dplyr::if_else(
      cnv_proportion_classification == "Tumor" & copykat_classification == "Tumor",
      "Tumor",
      "Normal"
    ),
    # there are cells classified as NA by copykat, we want those to be normal 
    combined_cnv_classification = dplyr::if_else(is.na(copykat_classification) & is.na(combined_cnv_classification), "Normal", combined_cnv_classification)
  )
```

```{r}
# look at which cells are tumor/normal
# be sure to remove any duplicates 
coldata_df |> 
  dplyr::select(UMAP1, UMAP2, barcodes, combined_cnv_classification) |>
  dplyr::distinct() |> 
  ggplot(aes(x = UMAP1, y = UMAP2, color = combined_cnv_classification)) +
  geom_point(alpha = 0.5, size = 0.5)
```

The remainder of the notebook will focus on validating that these in fact are tumor cells and all other cells are classified as normal.

## Validate tumor cells

### Marker gene expression

First, we will look at the expression of the tumor marker genes. 
We expect that these should be expressed higher in the tumor cells than in the normal cells.

In this first plot, we show a density plot of marker gene expression coloring by the cell classification (tumor or normal). 
For this plot, we will include all classifications, including the new combined CNV classification.

```{r}
# calculate sum of the scaled gene expression values for plotting 
marker_density_df <- coldata_df |> 
  dplyr::select(barcodes, gene_symbol, gene_expression, ends_with("classification")) |> 
  unique() |> 
  dplyr::group_by(gene_symbol) |>
  # get z-scores for each gene
  dplyr::mutate(transformed_gene_expression = scale(gene_expression)[, 1]) |> 
  dplyr::ungroup() |> 
  dplyr::group_by(barcodes) |> 
  # add all the scores together 
  dplyr::mutate(sum_exp = sum(transformed_gene_expression)) |>
  # expand classifications column for easy faceting 
  dplyr::ungroup() |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  )

ggplot(marker_density_df, aes(x = sum_exp, color = classification)) +
  geom_density() +
  facet_wrap(vars(method))
```

This plot shows us that overall the tumor cells in each method have higher marker gene expression. 
Although the `InferCNV` alone and combined CNV classifications appear to have the most distinct separation in marker gene expression.

Next, we will plot the expression of each gene in each cell using a heatmap. 
Here the rows are the marker genes and the columns are the cells with the color indicating the expression of that gene in that cell.

```{r}
# create matrix with marker genes as rows and barcodes as columns
marker_gene_heatmap <- coldata_df |> 
  dplyr::select(gene_expression, gene_symbol, barcodes) |> 
  tidyr::pivot_wider(values_from = gene_expression, 
                     names_from = barcodes) |> 
  tibble::column_to_rownames("gene_symbol") |> 
  as.matrix()

# create annotation df, make sure that dimensions match up by only getting unique barcodes 
annotation_df <- coldata_df |> 
  dplyr::select(barcodes, combined_cnv_classification) |> 
  unique()

annotation <- ComplexHeatmap::columnAnnotation(
  classification = annotation_df$combined_cnv_classification,
  col = list(classification = c("Tumor" = "#00274C", "Normal" = "#FFCB05")))
```

```{r}
# plot heatmap of marker genes 
ComplexHeatmap::Heatmap(
  marker_gene_heatmap,
  col = circlize::colorRamp2(c(0, 1), colors = c("white", "#00274C")),
  border = TRUE,
  ## Row parameters
  cluster_rows = TRUE,
  row_title = "Marker gene symbol",
  row_title_gp = grid::gpar(fontsize = 12),
  row_title_side = "left",
  row_names_side = "left",
  row_dend_side = "right",
  row_names_gp = grid::gpar(fontsize = 10),
  ## Column parameters
  cluster_columns = TRUE,
  show_column_names = FALSE,
  bottom_annotation = annotation
)
```

As expected, the tumor cells appear to have more marker genes that are expressed than normal cells.

### Gene set scores

Next, we will look at the gene set scores for three different EWS-FLI1 gene sets. 
The below plots will all use the `mean-GENESET` columns.

First we will just look at the gene set scores for each cell on the UMAP.

```{r}
# plot tumor cells for each classification method 
geneset_plot_df <- coldata_df |> 
  dplyr::select(barcodes, UMAP1, UMAP2, ends_with("classification"), starts_with("mean")) |> 
  tidyr::pivot_longer(
    cols = starts_with("mean"),
    names_to = "geneset",
    values_to = "mean_score"
  ) |> 
  dplyr::mutate(
    geneset = stringr::word(geneset, -1, sep = "-")
  )

ggplot(geneset_plot_df, aes(x = UMAP1, UMAP2, color = mean_score)) +
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(vars(geneset)) +
  scale_color_viridis_c()
```

Generally, the higher gene set scores appear to be in the group of cells that are tumor cells with some slight variations between gene sets.

Below we look at gene set scores for cells classified as tumor cells vs normal cells for all classification methods.

```{r, fig.height=10}
geneset_plot_df |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  ggplot(aes(x = mean_score, color = classification)) +
  geom_density(bw = 0.05) +
  facet_grid(rows = vars(method), 
             cols = vars(geneset))
```

For the most part the gene set scores are higher in tumor cells regardless of classification.

Below we make a heatmap of the gene set scores for just the combined CNV classification. 
Each row is a gene set and each column is a barcode. 
The color indicates the gene set score for that cell.

```{r}
# make a matrix of gene set by barcode 
geneset_heatmap <- geneset_plot_df |> 
  dplyr::select(mean_score, geneset, barcodes) |> 
  unique() |> 
  tidyr::pivot_wider(values_from = mean_score, 
                     names_from = barcodes) |> 
  tibble::column_to_rownames("geneset") |> 
  as.matrix()

ComplexHeatmap::Heatmap(
  geneset_heatmap,
  col = circlize::colorRamp2(c(0, 1), colors = c("white", "#00274C")),
  border = TRUE,
  ## Row parameters
  cluster_rows = TRUE,
  row_title = "",
  row_title_gp = grid::gpar(fontsize = 12),
  row_title_side = "left",
  row_names_side = "left",
  row_dend_side = "right",
  row_names_gp = grid::gpar(fontsize = 10),
  ## Column parameters
  cluster_columns = TRUE,
  show_column_names = FALSE,
  bottom_annotation = annotation
)
```

For the most part we see that cells with higher gene set scores are classified as tumor cells. 
One thing to note is that the `SILIGAN` gene set appears to have some expression in normal cells. 
However, if you look at the spots where all three gene sets have higher scores, these apply to tumor cells.

### Validate based on CNAs

The last thing we will do here is look to see if tumor cells have any of the expected CNAs found in Ewing sarcoma:

- Gain of Chr8
- Gain of Chr12
- Gain of Chr1p
- Loss of Chr16q

Although these are the most frequent, there are patients who do not have any of these alterations and patients that only have some of these alterations. 
See [Tirode et al.,](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4264969/) and [Crompton et al.](https://doi.org/10.1158/2159-8290.CD-13-1037).

We will look specifically at gains and losses in each chromosome across tumor and normal cells. 
To do this, we will use the output from `InferCNV` since that output separates out gains and losses in a way that makes it easy for plotting.

The first heatmap looks at the proportion of each chromosome with a CNV that corresponds to an amplification or gain. 
Here each row is a cell and each column is a chromosome with the color corresponding to the proportion of the chromosome with a gain detected.

```{r}
# get matrix of barcodes by chromosome
gain_cnv_df <- coldata_df |> 
  dplyr::select(barcodes, starts_with("proportion_scaled_dupli")) |> 
  unique() |> 
  tibble::column_to_rownames("barcodes") |> 
  as.matrix()

# remove extra characters in colnames, proportion_scaled_dupli
colnames(gain_cnv_df) <- stringr::word(colnames(gain_cnv_df), -1, sep = "_")

# need to make a new annotation, since we want to annotate the rows here 
annotation <- ComplexHeatmap::rowAnnotation(classification = annotation_df$combined_cnv_classification,
                                            col = list(classification = c("Tumor" = "#00274C", "Normal" = "#FFCB05")))
```

```{r}
ComplexHeatmap::Heatmap(
  gain_cnv_df,
  col = circlize::colorRamp2(c(0, 1), colors = c("white", "#00274C")),
  border = TRUE,
  ## Row parameters
  cluster_rows = TRUE,
  row_title = "",
  row_title_gp = grid::gpar(fontsize = 12),
  row_title_side = "left",
  show_row_names = FALSE,
  row_dend_side = "right",
  row_names_gp = grid::gpar(fontsize = 10),
  ## Column parameters
  cluster_columns = FALSE,
  show_column_names = TRUE,
  right_annotation = annotation
)
```

Here we see a gain in Chr8 and Chr1 in tumor cells compared to normal cells. 
We also see other chromosomes with amplifications in tumor cells compared to normal cells and that the tumor and normal cells are nicely separated in the heirarchical clustering.

Below we make the same plot but with only the losses or deletions.

```{r}
# barcode by chromosome mtx with losses 
loss_cnv_df <- coldata_df |> 
  dplyr::select(barcodes, starts_with("proportion_scaled_loss")) |> 
  unique() |> 
  tibble::column_to_rownames("barcodes") |> 
  as.matrix()

colnames(loss_cnv_df) <- stringr::word(colnames(loss_cnv_df), -1, sep = "_")
```

```{r}
ComplexHeatmap::Heatmap(
  loss_cnv_df,
  col = circlize::colorRamp2(c(0, 1), colors = c("white", "#00274C")),
  border = TRUE,
  ## Row parameters
  cluster_rows = TRUE,
  row_title = "",
  row_title_gp = grid::gpar(fontsize = 12),
  row_title_side = "left",
  show_row_names = FALSE,
  row_dend_side = "right",
  row_names_gp = grid::gpar(fontsize = 10),
  ## Column parameters
  cluster_columns = FALSE,
  show_column_names = TRUE,
  right_annotation = annotation
)
```

We do not see the expected loss in Chr16, but this is not present in every patient. 
Additionally, we had already noted the gain in Chr8 in the tumor cells, so there is still at least one expected CNV present. 
Again, there is a clear separation of tumor and normal cells in terms of presence of CNVs.

## Save final tumor annotations

After validating that the cells classified as tumor by both `CopyKAT` and `InferCNV` show the expected expression patterns and CNV patterns as tumor cells, we can confidently label those cells as tumor cells. 
We will export these labels to the final annotations file for later use.

```{r}
# export final TSV with annotations 
final_classifications <- coldata_df |> 
  dplyr::select(
    cell_barcode = barcodes,
    tumor_cell_classification = combined_cnv_classification
  ) |> 
  unique()

readr::write_tsv(final_classifications, final_annotations_file)
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```