### 2024-06-14 analyses/cell-type-ewings/cell_type_annotations/SCPCS000492/SCPCL000824_tumor-cell-validation.Rmd
some more exploratory plots

---
title: "SCPCL000824 - Validation of tumor cells"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

## Introduction

In this notebook we compare tumor cell annotations obtained from marker genes, `CopyKAT`, and `InferCNV` to identify a group of tumor cells we are confident in.

- First we look at the cells that are annotated as tumor cells by each method. 
Using this information, we call any cell that is annotated as a tumor cell by both `CopyKAT` and `InferCNV` as a tumor cell. 
- We then validate that these are in fact tumor cells by doing the following: 
  - Looking at marker gene expression. 
  - Looking at expression of EWS-FLI1 target gene sets. 
  - Confirming presence of known CNVs in Ewing sarcoma

## Define functions

```{r}
# create a heatmap where rows are marker genes or gene sets and columns are cells 
# this function adds a column annotation since cells are columns 
plot_gene_heatmap <- function(df, 
                              row_title = "",
                              legend_title = "", 
                              annotation = NULL){

  # plot heatmap of marker genes 
  heatmap <- ComplexHeatmap::Heatmap(
    df,
    # set the color scale based on min and max values
    col = circlize::colorRamp2(seq(min(df), max(df), length = 2), colors = c("white", "#00274C")),
    border = TRUE,
    ## Row parameters
    cluster_rows = TRUE,
    row_title = row_title,
    row_title_gp = grid::gpar(fontsize = 12),
    row_title_side = "left",
    row_names_side = "left",
    row_dend_side = "right",
    row_names_gp = grid::gpar(fontsize = 10),
    ## Column parameters
    cluster_columns = TRUE,
    show_column_names = FALSE,
    bottom_annotation = annotation,
    heatmap_legend_param = list(
      title = legend_title
    )
  )

  return(heatmap)
}

# create a heatmap where rows are cells and columns are chromosomes 
# this function adds a row annotation since cells are rows 
plot_cnv_heatmap <- function(df,
                             cnv_col, 
                             annotation = NULL,
                             legend_title = ""){

  # barcode by chromosome df 
  cnv_df <- df |> 
    dplyr::select(barcodes, starts_with(cnv_col)) |> 
    unique() |> 
    tibble::column_to_rownames("barcodes")

  # just get the name of the chromosome as column name 
  colnames(cnv_df) <- stringr::word(colnames(cnv_df), -1, sep = "_")

  # do a little bit of column reordering before converting to a matrix
  cnv_df <- cnv_df |> 
    dplyr::relocate(chr10, .before = chr11) |> 
    dplyr::relocate(chr19, .before = chr20) |> 
    dplyr::relocate(chr21, .before = chr22) |> 
    as.matrix()

  heatmap <- ComplexHeatmap::Heatmap(
    cnv_df,
    col = circlize::colorRamp2(seq(min(cnv_df), max(cnv_df), length = 2), colors = c("white", "#00274C")),
    border = TRUE,
    ## Row parameters
    cluster_rows = TRUE,
    row_title = "",
    row_title_gp = grid::gpar(fontsize = 12),
    row_title_side = "left",
    show_row_names = FALSE,
    row_dend_side = "right",
    row_names_gp = grid::gpar(fontsize = 10),
    ## Column parameters
    cluster_columns = FALSE,
    show_column_names = TRUE,
    right_annotation = annotation,
    heatmap_legend_param = list(
      title = legend_title
    )
  )
  return(heatmap)
}

```

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_bw()
)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "2024-05-01")
sample_dir <- file.path(data_dir, "SCPCP000015", "SCPCS000492")

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-ewings")
```

```{r}
# Input files
sce_file <- file.path(sample_dir, "SCPCL000824_processed.rds")
marker_genes <- file.path(module_base, "references", "tumor-marker-genes.tsv")

# results from annotation workflow 
results_dir <- file.path(module_base, "results", "annotate_tumor_cells_output", "SCPCS000492")

marker_gene_results_file <- file.path(results_dir, "SCPCL000824_tumor-normal-classifications.tsv")
copykat_predictions_file <- file.path(results_dir, "SCPCL000824_copykat-classifications.tsv")

infercnv_predictions_file <- file.path(results_dir, "SCPCL000824_infercnv-classifications.tsv")
infercnv_metadata_file <- file.path(results_dir, "infercnv", "SCPCL000824_cnv-metadata.tsv")

geneset_scores_file <- file.path(results_dir, "SCPCL000824_gene-set-scores.tsv")

# output files 
final_annotations_dir <- file.path(module_base, "cell_type_annotations", "SCPCS000492")
fs::dir_create(final_annotations_dir)
final_annotations_file <- file.path(final_annotations_dir, "SCPCL000824_tumor-classifications.tsv")
```

```{r}
# read in all files 
sce <- readr::read_rds(sce_file)

# read in marker genes table 
marker_genes_df <- readr::read_tsv(marker_genes) |> 
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |> 
  dplyr::distinct()

marker_gene_results_df <- readr::read_tsv(marker_gene_results_file)

# copykat output
copykat_results_df <- readr::read_tsv(copykat_predictions_file) |> 
  # only select no_ref classification and rename to be consistent with naming of other classifications
  dplyr::select(barcodes, copykat_classification = no_ref, mean_cnv_detection) |> 
  dplyr::distinct() |> 
  dplyr::mutate(
    copykat_classification = dplyr::case_when(
      copykat_classification == "aneuploid" ~ "Tumor",
      copykat_classification == "diploid" ~ "Normal",
      TRUE ~ copykat_classification
    )
  )

# infercnv output 
infercnv_results_df <- readr::read_tsv(infercnv_predictions_file)
infercnv_metadata_df <- read.delim(infercnv_metadata_file) |> 
  tibble::rownames_to_column("barcodes")

# remove MT and GL chr, comments in docs recommend that these are not useful and can be inaccurate 
chr_columns_to_keep <- which(!stringr::str_detect(names(infercnv_metadata_df), "chrMT|chrGL"))
infercnv_metadata_df <- infercnv_metadata_df[, chr_columns_to_keep]

# gene set scores 
geneset_scores_df <- readr::read_tsv(geneset_scores_file)
```

```{r}
# grab all coldata and add in all classifications and results to use for plotting throughout notebook 
# get list of marker genes to use 
marker_genes <- marker_genes_df |> 
  dplyr::filter(cell_type == "tumor") |> 
  dplyr::pull(ensembl_gene_id)

# get the gene expression counts for all marker genes
marker_gene_exp <- logcounts(sce[marker_genes, ]) |>
  as.matrix() |>
  t() |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes")

# pull out the UMAP coordinates and make a data frame to use for plotting
classification_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
        \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")
      ) |>
  # get rid of excess columns
  dplyr::select(barcodes, UMAP1, UMAP2) |> 
  # add in marker gene classifications
  dplyr::left_join(marker_gene_results_df) |> 
  # add in copykat results 
  dplyr::left_join(copykat_results_df) |> 
  # add in infercnv results and metadata
  dplyr::left_join(infercnv_results_df) |> 
  dplyr::left_join(infercnv_metadata_df) |> 
  # gene set scores 
  dplyr::left_join(geneset_scores_df)
```

## Identify tumor cell population

Below is a UMAP showing which cells are labeled as tumor cells using each of the various classification methods.

```{r}
# plot tumor cells for each classification method 
tumor_umap_df <- classification_df |> 
  dplyr::select(barcodes, UMAP1, UMAP2, marker_gene_classification, copykat_classification, cnv_proportion_classification) |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  dplyr::mutate(
    method = dplyr::case_when(
      method == "marker_gene_classification" ~ "Marker genes only",
      method == "copykat_classification" ~ "CopyKAT - no reference",
      method == "cnv_proportion_classification" ~ "InferCNV proportion"
    )
  )

ggplot(tumor_umap_df, aes(x = UMAP1, UMAP2, color = classification)) +
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(vars(method))
```

```{r}
classification_df <- classification_df |>
  dplyr::mutate(
    geneset_classification = dplyr::if_else(
      `scaled_mean-RIGGI` > 0 | `scaled_mean-ZHANG` > 0 | `scaled_mean-SILIGAN` > 0,
      "Tumor",
      "Normal"
    ),
    combined_classification = dplyr::case_when(
      marker_gene_classification == "Normal" & geneset_classification == "Normal" ~ "Normal",
      marker_gene_classification == "Tumor" & geneset_classification == "Tumor" ~ "Tumor",
      .default = "Unknown"
    )
    # combined_classification = dplyr::case_when(
    #   copykat_classification == "Normal" & cnv_proportion_classification == "Normal" & marker_gene_classification == "Normal" ~ "Normal",
    #   copykat_classification == "Normal" & cnv_proportion_classification == "Reference" & marker_gene_classification == "Normal" ~ "Normal",
    #   copykat_classification == "Tumor" & cnv_proportion_classification == "Tumor" & marker_gene_classification == "Tumor" ~ "Tumor",
    #   .default = "Unknown"
    # )
  )
```

```{r}
ggplot(classification_df, aes(x = UMAP1, y = UMAP2, color = geneset_classification)) +
  geom_point(size = 0.5, alpha = 0.5)
```

```{r}
# calculate sum of the scaled gene expression values for plotting 
marker_density_df <- classification_df |> 
  # add in marker gene expression to dataframe 
  dplyr::left_join(marker_gene_exp, by = "barcodes") |> 
  # combine all genes into a single column for easy faceting
  tidyr::pivot_longer(
    cols = starts_with("ENSG"),
    names_to = "ensembl_gene_id",
    values_to = "gene_expression"
  ) |> 
  # join with marker gene df to get gene symbols for plotting 
  dplyr::left_join(marker_genes_df, by = c("ensembl_gene_id")) |> 
  dplyr::select(barcodes, gene_symbol, gene_expression, copykat_classification, cnv_proportion_classification, marker_gene_classification, combined_classification, geneset_classification) |> 
  unique() |> 
  dplyr::group_by(gene_symbol) |>
  # get z-scores for each gene
  dplyr::mutate(transformed_gene_expression = scale(gene_expression)[, 1]) |> 
  dplyr::ungroup() |> 
  dplyr::group_by(barcodes) |> 
  # add all the scores together 
  dplyr::mutate(sum_exp = sum(transformed_gene_expression)) |>
  # expand classifications column for easy faceting 
  dplyr::ungroup() |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  dplyr::mutate(
    method = dplyr::case_when(
      method == "marker_gene_classification" ~ "Marker genes only",
      method == "copykat_classification" ~ "CopyKAT - no reference",
      method == "cnv_proportion_classification" ~ "InferCNV proportion",
      method == "combined_classification" ~ "Combined",
      method == "geneset_classification" ~ "EWS-FLI1 Gene set"
    )
  )

ggplot(marker_density_df, aes(x = sum_exp, color = classification)) +
  geom_density() +
  facet_wrap(vars(method))
```

```{r}
# plot tumor cells for each classification method 
geneset_plot_df <- classification_df |> 
  dplyr::select(barcodes, UMAP1, UMAP2, ends_with("classification"), starts_with("scaled_mean")) |> 
  tidyr::pivot_longer(
    cols = starts_with("scaled_mean"),
    names_to = "geneset",
    values_to = "mean_score"
  ) |> 
  dplyr::mutate(
    geneset = stringr::word(geneset, -1, sep = "-")
  )

ggplot(geneset_plot_df, aes(x = UMAP1, UMAP2, color = mean_score)) +
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(vars(geneset)) +
  scale_color_viridis_c()
```

```{r}
geneset_plot_df |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  ggplot(aes(x = mean_score, color = classification)) +
  geom_density(bw = 0.05) +
  facet_grid(rows = vars(method), 
             cols = vars(geneset))
```

CNV heatmap annotated by copykat and marker gene prediction, two annotation columns

```{r}
annotation_df <- classification_df |> 
  dplyr::select(barcodes, ends_with("classification")) |> 
  unique()

# need to make a new annotation, since we want to annotate the rows here 
annotation <- ComplexHeatmap::rowAnnotation(classification = annotation_df$combined_classification,
                                            col = list(classification = c("Tumor" = "#00274C", "Normal" = "#FFCB05", "Unknown" = "grey")))

# plot gains in cnv
plot_cnv_heatmap(classification_df,
                 cnv_col = "proportion_scaled_dupli",
                 annotation = annotation,
                 legend_title = "Copy number \ngain")
```

```{r}
# plot gains in cnv
plot_cnv_heatmap(classification_df,
                 cnv_col = "proportion_scaled_loss",
                 annotation = annotation,
                 legend_title = "Copy number \nloss")
```

Density plot with mean proportion CNV from InferCNV on x-axis, colored by prediction, faceted by classification method

```{r}
classification_df |> 
  dplyr::select(barcodes, UMAP1, UMAP2, ends_with("classification"), mean_cnv_proportion) |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  ggplot(aes(x = mean_cnv_proportion, color = classification)) +
  geom_density(bw = 0.05) +
  facet_grid(rows = vars(method))
```

Mean InferCNV proportion on x-axis, mean copyKAT proportion on y-axis, color by scaled sum of marker gene expression

```{r}
cnv_prop_df <- classification_df |> 
  dplyr::select(barcodes, ends_with("classification"), starts_with("scaled_mean"), mean_cnv_proportion, mean_cnv_detection) |> 
  tidyr::pivot_longer(
    cols = starts_with("scaled_mean"),
    names_to = "geneset",
    values_to = "mean_score"
  ) |> 
  dplyr::mutate(
    geneset = stringr::word(geneset, -1, sep = "-")
  )

ggplot(cnv_prop_df, aes(x = mean_cnv_proportion, y = abs(mean_cnv_detection), color = mean_score)) +
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(vars(geneset)) +
  scale_color_viridis_c()
```

```{r}
ggplot(cnv_prop_df, aes(x = mean_score, y = abs(mean_cnv_detection))) +
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(vars(geneset)) +
  scale_color_viridis_c()
```

faceted density plot with marker gene expression on x axis, colored by copykat and infercnv prediction and faceted by marker gene

```{r}
geneset_scores_df |> 
  dplyr::pivot

ggplot(geneset_scores_df, aes(x = mean_score, color = ))
```

use SingleR to classify as tumor/normal using SCPCS000490

correlation of geneset score and marker gene score 
```{r}
geneset_plot_df |>

  dplyr::left_join(marker_density_df) |> 
  ggplot(aes(x = sum_exp, y = mean_score)) +
    facet_wrap()
```