### 2024-06-26 analyses/cell-type-ewings/cell_type_annotations/SCPCS000492/SCPCL000824_tumor-gene-expression-exploration.Rmd
update notebook to use UCell and AUCell

---
title: "`SCPCL000824` - Exploration of gene set based methods for tumor cell assignment"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: hide
---

## Introduction

In this notebook we attempt to identify tumor cells in `SCPCL000824`. 
We first compare tumor cell annotations obtained from marker genes, `CopyKAT`, and `InferCNV` to identify a group of tumor cells we are confident in. 
However, there is a lack of consensus between tumor cells identified by marker gene expression and CNV inference.

In an attempt to identify a group of cells that we are confident are tumor cells, here we test using the following orthogonal marker gene based methods:

- Marker gene expression from `SCPCL000822` to inform an appropriate cut off for marker gene expression in `SCPCL000824`. 
- [`AUCell`](https://www.bioconductor.org/packages/release/bioc/vignettes/AUCell/inst/doc/AUCell.html) with marker genes and three EWS-FLI1 gene sets in both samples. 
- [`UCell`](https://carmonalab.github.io/UCell_demo/UCell_matrix_vignette.html#content) with marker genes and three EWS-FLI1 gene sets in both samples.

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_bw()
)

# quiet messages
options(readr.show_col_types = FALSE)
ComplexHeatmap::ht_opt(message = FALSE)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "2024-05-01")
sample_dir <- file.path(data_dir, "SCPCP000015", "SCPCS000492")

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-ewings")
```

```{r}
# source in helper functions: plot_gene_heatmap() and plot_cnv_heatmap() 
# create_classification_df() and create_marker_gene_df()
validation_functions <- file.path(module_base, "scripts", "utils", "tumor-validation-helpers.R")
source(validation_functions)
```

```{r}
# Input files
sce_file <- file.path(sample_dir, "SCPCL000824_processed.rds")
marker_genes_file <- file.path(module_base, "references", "tumor-marker-genes.tsv")

# results from annotation workflow 
results_dir <- file.path(module_base, "results", "annotate_tumor_cells_output", "SCPCS000492")

marker_gene_results_file <- file.path(results_dir, "SCPCL000824_tumor-normal-classifications.tsv")
copykat_predictions_file <- file.path(results_dir, "SCPCL000824_copykat-classifications.tsv")

infercnv_predictions_file <- file.path(results_dir, "SCPCL000824_infercnv-classifications.tsv")
infercnv_metadata_file <- file.path(results_dir, "infercnv", "SCPCL000824_cnv-metadata.tsv")

geneset_scores_file <- file.path(results_dir, "SCPCL000824_gene-set-scores.tsv")

# output files 
final_annotations_dir <- file.path(module_base, "cell_type_annotations", "SCPCS000492")
fs::dir_create(final_annotations_dir)
final_annotations_file <- file.path(final_annotations_dir, "SCPCL000824_tumor-classifications.tsv")

# reference files to use with SingleR
ref_sce_file <- file.path(data_dir, "SCPCP000015", "SCPCS000490", "SCPCL000822_processed.rds")
ref_labels_file <- file.path(module_base, "cell_type_annotations", "SCPCS000490", "SCPCL000822_tumor-classifications.tsv")
ref_geneset_scores_file <- file.path(module_base, "results", "annotate_tumor_cells_output", "SCPCS000490", "SCPCL000822_gene-set-scores.tsv")
```

```{r}
# read in sce file 
sce <- readr::read_rds(sce_file)

# read in ref sce and ref annotations for comparing between samples 
# ref is SCPCL000822
ref_sce <- readr::read_rds(ref_sce_file)
ref_labels_df <- readr::read_tsv(ref_labels_file)
ref_geneset_df <- readr::read_tsv(ref_geneset_scores_file)
```

```{r}
# generate classification df 
classification_df <- create_classification_df(
  sce,
  marker_gene_results_file,
  copykat_predictions_file, 
  infercnv_predictions_file,
  infercnv_metadata_file,
  geneset_scores_file
) |> 
  # remove extra infercnv classification column we won't use 
  dplyr::select(- cnv_sum_classification)
```

## Evaluate tumor cell classifications

Below is a UMAP showing which cells are labeled as tumor cells using each of the various classification methods (marker genes, `CopyKAT`, and `InferCNV`).

```{r}
# plot tumor cells for each classification method 
tumor_umap_df <- classification_df |> 
  dplyr::select(barcodes, UMAP1, UMAP2, ends_with("classification")) |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  dplyr::mutate(
    method = dplyr::case_when(
      method == "marker_gene_classification" ~ "Marker genes only",
      method == "copykat_classification" ~ "CopyKAT - no reference",
      method == "cnv_proportion_classification" ~ "InferCNV proportion"
    )
  )

ggplot(tumor_umap_df, aes(x = UMAP1, UMAP2, color = classification)) +
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(vars(method))
```

It looks like we don't have a lot of agreement between the marker gene based classification and copy number inference classification. 
Although we do see some agreement between the two CNV methods.

### Marker gene and gene set expression

We can confirm this by looking at marker gene expression in cells classified as tumor or normal by each method.

```{r}
# generate marker genes df
plot_markers_df <- create_marker_gene_df(
  sce,
  classification_df, 
  marker_genes_file
)

# create a density plot showing the distribution of marker gene expression across classification methods 
marker_density_df <- plot_markers_df |>
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  dplyr::mutate(
    method = dplyr::case_when(
      method == "marker_gene_classification" ~ "Marker genes only",
      method == "copykat_classification" ~ "CopyKAT - no reference",
      method == "cnv_proportion_classification" ~ "InferCNV proportion"
    )
  )

ggplot(marker_density_df, aes(x = sum_transformed_exp, color = classification)) +
  geom_density() +
  facet_wrap(vars(method))
```

We see a few things from this plot:

- As expected given that marker gene expression was used to classify cells using the "Marker genes only" method, we see an increase in marker gene expression in tumor cells. 
However we don't see a clear bimodal distribution, which is something we did see in `SCPCL000822`. 
- For `InferCNV`, both tumor and normal cells have higher expression of marker genes than the reference. 
- For `CopyKAT`, the normal cells appear to have higher expression of marker genes.

Let's also look at gene set scores and see if we can use those to help us determine which cells are most likely to be tumor cells.

```{r}
# plot gene set scores for each cell  
geneset_plot_df <- classification_df |> 
  dplyr::select(barcodes, UMAP1, UMAP2, ends_with("classification"), starts_with("mean-")) |> 
  tidyr::pivot_longer(
    cols = starts_with("mean"),
    names_to = "geneset",
    values_to = "mean_score"
  ) |> 
  dplyr::mutate(
    geneset = stringr::word(geneset, -1, sep = "-")
  )

ggplot(geneset_plot_df, aes(x = UMAP1, UMAP2, color = mean_score)) +
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(vars(geneset)) +
  scale_color_viridis_c()
```

We do see that the scores across all three genesets seem to be pretty consistent. 
Cells that have higher scores for `RIGGI`, also have higher scores for `SILIGAN` and `ZHANG`.

Let's look at how the distribution of gene set scores coordinates with assignments from the CNV and marker gene methods. 
Below, we color the distribution based on the classification method used.

```{r}
geneset_plot_df |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  dplyr::mutate(
    method = dplyr::case_when(
      method == "marker_gene_classification" ~ "Marker genes only",
      method == "copykat_classification" ~ "CopyKAT - no reference",
      method == "cnv_proportion_classification" ~ "InferCNV proportion"
    )) |> 
  ggplot(aes(x = mean_score, color = classification)) +
  geom_density(bw = 0.05) +
  facet_grid(rows = vars(method), 
             cols = vars(geneset))
```

We see a few things from this plot:

- The marker genes classification has the most consistent increase in gene set scores in tumor cells compared to normal cells, although the separation is not super clear.  
- For the `RIGGI` and `ZHANG` gene sets, the scores are higher in both tumor and normal cells compared to reference cells as classified by `InferCNV`. 
- For all gene sets, the normal cells have a slightly higher gene set score as classified by `CopyKAT`.

## Using gene expression to classify tumor cells

Unlike with `SCPCL000822`, there is no clear consensus between CNV inference methods and marker gene based classification. 
Because these genomes are quiet, we expect CNV inference to be more difficult and less likely to accurately predict the tumor cells. 
Therefore, we should use methods that rely on gene expression.

The rest of this notebook will explore using different gene set based methods to classify tumor cells to see if we can more accurately identify the tumor cells in this sample.

### Comparing marker gene expression in `SCPCL000822` and `SCPCL000824`

With `SCPCL000822` we had a clear separation between marker gene expression in tumor cells and normal cells because there was a bimodal distribution. 
It does not appear that is the case here so it's going to be more difficult to pull out tumor cells.

First we will just compare the distribution of the raw marker gene expression in `SCPCL000822` and `SCPCL000824`. 
To do this, we will get the total marker gene expression by summing all marker genes in a cell and then plot the distribution. 
The first plot will look at the distribution across all cells and the second will show the classification of tumor vs. normal cells.

```{r}
# look at the raw sum of marker gene expression in both 822 and 824
# first create marker genes df for ref sce 
ref_classification_df <- ref_labels_df |> 
  dplyr::rename("barcodes" = "cell_barcode")

ref_markers_df <- create_marker_gene_df(sce = ref_sce, 
                                        classification_df = ref_classification_df,
                                        marker_genes_file) |> 
  dplyr::mutate(sample = "SCPCL000822")

# combine all marker gene data for both samples into one df 
combined_markers_df <- plot_markers_df |> 
  dplyr::select(barcodes, tumor_cell_classification = marker_gene_classification, gene_symbol, gene_expression, transformed_gene_expression, sum_raw_exp, sum_transformed_exp) |> 
  dplyr::mutate(sample = "SCPCL000824") |> 
  dplyr::bind_rows(ref_markers_df)
```

```{r}
# total distribution
ggplot(combined_markers_df, aes(x = sum_raw_exp)) +
  geom_density() +
  facet_grid(rows = vars(sample))

# color by tumor cell classification 
ggplot(combined_markers_df, aes(x = sum_raw_exp, color = tumor_cell_classification)) +
  geom_density() +
  facet_grid(rows = vars(sample))
```

Looking at this, we see that `SCPCL000822` has a bimodal distribution and shows a clear cutoff where cells are tumor vs. normal, but this is not the case for `SCPCL000824`. 
Let's find the local minima in the bimodal distribution for `SCPCL000822` and then use that to classify tumor cells in `SCPCL000824`.

```{r}
# create distribution 
density_data <- density(combined_markers_df$sum_raw_exp)
# find the local minima in the distribution
exp_cutoff <- optimize(approxfun(density_data$x, density_data$y), interval = c(1,10))$minimum
```

```{r}
# add new column with updated marker gene classification 
# use local minima from 822 to define 824
new_classification <- combined_markers_df |> 
  dplyr::filter(sample == "SCPCL000824") |> 
  dplyr::mutate(updated_marker_gene_classification = dplyr::if_else(sum_raw_exp >= exp_cutoff, "Tumor", "Normal")) |> 
  dplyr::select(barcodes, updated_marker_gene_classification)

# add new column to existing classification
classification_df <- classification_df |> 
  dplyr::left_join(new_classification)

# label cells based on new classifiation 
ggplot(classification_df, aes(x = UMAP1, y = UMAP2, color = updated_marker_gene_classification)) +
  geom_point(alpha = 0.5, size = 0.5)
```

This looks closer to what I would expect with most cells in the large group of cells corresponding to tumor cells. 
This also captures cells that are labeled as tumor cells by both marker genes and CNV inference.

### Comparing gene set expression in `SCPCL000822` and `SCPCL000824`

Now we will look at the gene set scores for each gene set in `SCPCL000822` and `SCPCL000824`.  
We did not use these to classify `SCPCL000822`, so we won't actually do any classification, but this will show us if the scores for tumor cells are similar to each other across samples.

```{r}
# get geneset score from reference sce 
ref_geneset_df <- ref_geneset_df |> 
  dplyr::select(barcodes, starts_with("mean")) |> 
  tidyr::pivot_longer(
    cols = starts_with("mean"),
    names_to = "geneset",
    values_to = "mean_score"
  ) |> 
  dplyr::mutate(
    geneset = stringr::word(geneset, -1, sep = "-"),
    sample = "SCPCL000822 - reference"
  )

# join with gene set scores from 824 and plot distribution 
geneset_plot_df |> 
  dplyr::select(barcodes, geneset, mean_score) |> 
  dplyr::mutate(sample = "SCPCL000824") |> 
  dplyr::bind_rows(ref_geneset_df) |> 
  ggplot(aes(x = mean_score, color = sample)) +
  geom_density() +
  facet_grid(rows = vars(geneset))
```

Again we see that the gene set scores are bimodal for `SCPCL000822`, at least for `RIGGI` and `ZHANG`. 
However, we don't see that for `SCPCL000824`, but we do see a very similar ranger of values. 
This makes me think that the majority of the cells are in fact tumor cells.

## Classifying tumor cells with AUCell

The next thing we will do is see if we can use a more data driven approach to identify cells with higher marker gene or gene set expression. 
We will use [`AUCell`](https://www.bioconductor.org/packages/release/bioc/vignettes/AUCell/inst/doc/AUCell.html) to do this and will run it on both samples. 
However, `AUCell` relies on having a bimodal distribution in gene set expression, so I expect we may need to do a similar thing where we use cutoffs determined in `SCPCL000822` to call cells in `SCPCL000824`.

```{r}
# names of gene sets to grab
ews_gene_sets <- c(
  "ZHANG_TARGETS_OF_EWSR1_FLI1_FUSION",
  "RIGGI_EWING_SARCOMA_PROGENITOR_UP",
  "SILIGAN_TARGETS_OF_EWS_FLI1_FUSION_DN"
)

# pull gene sets from msigbdr 
# all gene sets are part of C2, CGP
genes_df <- msigdbr::msigdbr(species= "Homo sapiens", 
                             category = "C2",
                             subcategory = "CGP") |> 
  # only keep relevant gene sets 
  dplyr::filter(gs_name %in% ews_gene_sets)
```

```{r}
# first create named list of genes
# we will need this to run UCell later too 
genes_list <- ews_gene_sets |> 
  purrr::map(\(name){
    genes <- genes_df |> 
      dplyr::filter(gs_name == name) |> 
      dplyr::pull(ensembl_gene)
  }) |> 
  purrr::set_names(ews_gene_sets)

# build GeneSetCollection for AUCell 
msig_gene_sets <- genes_list |> 
  purrr::imap(\(genes, name) GSEABase::GeneSet(genes, setName = name))

# get list of marker genes to add to GeneSets 
marker_genes <- readr::read_tsv(marker_genes_file, show_col_types = FALSE) |> 
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |> 
  dplyr::distinct() |> 
  dplyr::filter(cell_type == "tumor") |> 
  dplyr::pull(ensembl_gene_id)

# turn it into a gene set 
marker_gene_set <- marker_genes |> 
  GSEABase::GeneSet(setName = "markers")

# join msig and marker genes into a collection to input with AUCell 
collection <- GSEABase::GeneSetCollection(c(msig_gene_sets, marker_gene_set))
```

```{r}
# run AUCell
ref_auc <- AUCell::AUCell_run(counts(ref_sce), collection)

# assign cells based on threshold and output plots 
ref_assignment <- AUCell::AUCell_exploreThresholds(ref_auc, assign = TRUE)
```

Here we can see a bimodal distribution of all gene sets except `SILIGAN`, which is to be expected based on the original distribution plots. 
We also see that `AUCell` identifies a threshold based on the local minima of the AUC distribution where all cells above the threshold are classified as expressing the gene set.
Let's compare the assignments from our validated classifications to using `AUCell` with `SCPCL000822`.

```{r}
# compare AUCell classification vs cnv classification in ref
# pull out tumor cells found by AUCell
tumor_cells <- ref_assignment$markers$assignment

ref_classification_df <- ref_classification_df |> 
  # add new column with auc classification 
  dplyr::mutate(auc_classification = dplyr::if_else(barcodes %in% tumor_cells, "Tumor", "Normal"))

# filter out any ambiguous cells 
caret_df <- ref_classification_df |> 
  dplyr::filter(tumor_cell_classification != "Ambiguous")

# compare using a confusion matrix
caret::confusionMatrix(
    table(
      caret_df$tumor_cell_classification, 
      caret_df$auc_classification)
  ) 
```

It looks like we get pretty good agreement between our original classification and the updated `AUCell` classification using the marker gene list.

Now we repeat on `SCPCL000824`.

```{r}
# get new auc and plot thresholds 
new_auc <- AUCell::AUCell_run(counts(sce), collection)
new_assignment <- AUCell::AUCell_exploreThresholds(new_auc, assign = TRUE)
```

Here we do not see any bimodal distributions and that most cells are considered to not be expressing the specified gene sets, which doesn't fit with our previous findings. 
I expect that without a bimodal distribution this model does not do a good job of calculating a threshold.

Instead, we will use the threshold identified by `AUCell` for `SCPCL000822` to identify tumor cells in `SCPCL000824`.  
Here, we label any cells that have passed the threshold set for the marker gene sets as tumor cells.

```{r}
# pull out threshold used for assigning cells from ref sample 
ref_threshold <- ref_assignment$markers$aucThr$selected

# create a vector of tumor cells in 824 using the cutoff from 822
new_tumor_cells <- new_auc@assays@data$AUC |> 
  as.data.frame() |> 
  tibble::rownames_to_column("gene_set") |> 
  tidyr::pivot_longer(!gene_set, 
                      names_to = "barcodes", 
                      values_to = "auc") |> 
  dplyr::filter(gene_set == "markers", auc >= ref_threshold) |>
  dplyr::pull(barcodes)

# add to original classification df 
classification_df <- classification_df |> 
  dplyr::mutate(auc_classification = dplyr::if_else(barcodes %in% new_tumor_cells, "Tumor", "Normal"))
```

```{r}
# visualize which cells are classified as tumor 
ggplot(classification_df, aes(x = UMAP1, y = UMAP2, color = auc_classification)) +
  geom_point(alpha = 0.5, size = 0.5)
```

This looks similar to using the updated marker gene cutoff to classify cells where most cells in that bottom group are tumor cells.

## Classifying tumor cells with UCell

Here we will look at using [`UCell`](https://carmonalab.github.io/UCell_demo/UCell_matrix_vignette.html#content) to calculate a gene set score for each cell and then attempt to classify cells based on the distribution of those gene set scores. 
Again, we will look at both `SCPCL000822` and `SCPCL000824`.

```{r}
# create list to use for ucell of gene sets and marker genes
ucell_gene_sets <- c(genes_list, markers = list(marker_genes))

# run ucell on both ref (822) and 824
ref_ucell <- UCell::ScoreSignatures_UCell(counts(ref_sce), features = ucell_gene_sets)
new_ucell <- UCell::ScoreSignatures_UCell(counts(sce), features = ucell_gene_sets)
```

`UCell` returns a signature score for each gene set and does not calculate any thresholds for classification on it's own. 
We will look at the distribution of scores across both samples below and see if we can identify a good cut off to use for classification.

```{r}
# plot distribution of ucell scores for both samples 
ref_ucell <- ref_ucell |> 
  as.data.frame() |> 
  tibble::rownames_to_column("barcodes") |> 
  dplyr::mutate(sample = "SCPCL000822")

ucell_df <- new_ucell |> 
  as.data.frame() |> 
  tibble::rownames_to_column("barcodes") |> 
  dplyr::mutate(sample = "SCPCL000824") |> 
  dplyr::bind_rows(ref_ucell) |> 
  tidyr::pivot_longer(ends_with("UCell"), 
                      names_to = "gene_list", 
                      values_to = "signature_score") |> 
  dplyr::mutate(
    gene_list = stringr::word(gene_list, 1, sep = "_")
  )
```

```{r}
ggplot(ucell_df, aes(x = signature_score, colour = sample)) +
  geom_density() +
  facet_grid(rows = vars(gene_list))
```

Here we see that the marker gene set maybe has a small bimodal distribution in `SCPCL000822`, but it is very small and not quite as striking as looking at the scores from `AUCell` or raw gene expression. 
`RIGGI`, `ZHANG`, and `SILIGAN` both show very similar distributions to the gene set scores distributions from our manual calculations of gene set scores (mean of all genes in the gene set). 
Again, there is a bimodal distribution for `RIGGI` and `ZHANG`, but not for `SILIGAN` and it is only present in `SCPCL000822`.

Let's use the local minima from the `RIGGI` gene set scores for `SCPCL000822` to identify tumor cells. 
First we will look at how using `UCell` compares to our previous classifications for `SCPCL000822`.

```{r}
# create distribution 
density_data <- density(ref_ucell$RIGGI_EWING_SARCOMA_PROGENITOR_UP_UCell)
# find the local minima in the distribution
geneset_cutoff <- optimize(approxfun(density_data$x, density_data$y), interval = c(0.01, 0.2))$minimum

# get a vector of tumor cells from ref 
ref_ucell_tumor_cells <- ucell_df |> 
  dplyr::filter(gene_list == "RIGGI",
                signature_score >= geneset_cutoff,
                sample == "SCPCL000822") |> 
  dplyr::pull(barcodes) |> 
  unique()

# add classification column for comparison to original classification
caret_df <- caret_df |> 
  dplyr::mutate(
    ucell_classification = dplyr::if_else(barcodes %in% ref_ucell_tumor_cells, "Tumor", "Normal")
  )

# confusion matrix between original and ucell 
caret::confusionMatrix(
  table(
    caret_df$tumor_cell_classification, 
    caret_df$ucell_classification 
  )
)
```

It looks like we get pretty good agreement between using `UCell` with the `RIGGI` gene set to the original classification. 
Now we will use the cutoff from `SCPCL000822` to classify tumor cells in `SCPCL000824`.

```{r}
# get list of tumor cells using ucell 
new_ucell_tumor_cells <- ucell_df |> 
  dplyr::filter(gene_list == "RIGGI",
                signature_score >= geneset_cutoff,
                sample == "SCPCL000824") |> 
  dplyr::pull(barcodes) |> 
  unique()

# add ucell classification
classification_df <- classification_df |> 
  dplyr::mutate(
    ucell_classification = dplyr::if_else(barcodes %in% new_ucell_tumor_cells, "Tumor", "Normal")
  )

ggplot(classification_df, aes(x = UMAP1, y = UMAP2, color = ucell_classification)) +
  geom_point(alpha = 0.5, size = 0.5)
```

Here we see less cells classified as tumor cells than we saw with the `AUCell` classification. 
This also looks very similar to the original UMAP when we classified cells using marker gene expression.

## Validate annotations

Finally, we will look at the marker gene and gene set scores for every cell across annotation methods.

```{r}
# create annotation df, keeping all classification methods 
annotation_df <- classification_df |> 
  dplyr::select(barcodes, marker_gene_classification, updated_marker_gene_classification, auc_classification, ucell_classification) |> 
  unique()

# create matrix with marker genes as rows and barcodes as columns
marker_gene_heatmap <- plot_markers_df |> 
  dplyr::select(gene_expression, gene_symbol, barcodes) |> 
  tidyr::pivot_wider(values_from = gene_expression, 
                     names_from = barcodes) |> 
  tibble::column_to_rownames("gene_symbol") |> 
  as.matrix()

annotation <- ComplexHeatmap::columnAnnotation(
  marker_genes = annotation_df$marker_gene_classification,
  updated_marker_genes = annotation_df$updated_marker_gene_classification,
  AUCell = annotation_df$auc_classification, 
  UCell = annotation_df$ucell_classification,
  col = list(marker_genes = c("Tumor" = "#00274C", "Normal" = "#FFCB05", "Ambiguous" = "grey"),
             updated_marker_genes = c("Tumor" = "#00274C", "Normal" = "#FFCB05", "Ambiguous" = "grey"), 
             AUCell = c("Tumor" = "#00274C", "Normal" = "#FFCB05", "Ambiguous" = "grey"),
             UCell = c("Tumor" = "#00274C", "Normal" = "#FFCB05", "Ambiguous" = "grey")))
```

```{r}
# plot heatmap of marker genes 
plot_gene_heatmap(marker_gene_heatmap, 
                  row_title = "Marker gene symbol",
                  legend_title = "Marker gene \nexpression",
                  annotation = annotation)
```

Here we see that the updated marker gene classification (using raw gene expression cut offs determined from `SCPCL000822`) and `AUCell` tend to have the clearest sepration between tumor cells and normal cells that line up with expression of individual marker genes.

Now we will create the same plot but with gene set scores.

```{r}
# make a matrix of gene set by barcode 
geneset_heatmap <- geneset_plot_df |> 
  dplyr::select(mean_score, geneset, barcodes) |> 
  unique() |> 
  tidyr::pivot_wider(values_from = mean_score, 
                     names_from = barcodes) |> 
  tibble::column_to_rownames("geneset") |> 
  as.matrix()

# plot heatmap of gene set score 
plot_gene_heatmap(geneset_heatmap,
                  annotation = annotation,
                  legend_title = "Gene set \nscore")
```

Again, we see that the updated marker genes and `AUCell` have the most similar assignments and these tend to group by tumor/normal cells. 
Although we do see a similar separation of cells in `UCell`, but there are many more cells that are classified as normal.

## Save annotations

We will go ahead and save the annotations from all gene expression based methods for future use.

```{r}
# export final TSV with annotations 
classifications_output <- classification_df |> 
  dplyr::select(
    cell_barcode = barcodes,
    marker_gene_classification,
    updated_marker_gene_classification, 
    auc_classification,
    ucell_classification
  ) |> 
  unique()

readr::write_tsv(classifications_output, final_annotations_file)
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```