### 2024-06-27 analyses/cell-type-ewings/cell_type_annotations/SCPCS000492/SCPCL000824_singler-classification.Rmd
notebook for classifying with SIngleR

---
title: "`SCPCL000824` - Using `SingleR` for classification of tumor cells"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: hide
---

## Introduction

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_bw()
)

# quiet messages
options(readr.show_col_types = FALSE)
ComplexHeatmap::ht_opt(message = FALSE)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "2024-05-01")
sample_dir <- file.path(data_dir, "SCPCP000015", "SCPCS000492")

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-ewings")
```

```{r}
# source in helper functions: plot_gene_heatmap() and plot_cnv_heatmap() 
# create_classification_df() and create_marker_gene_df()
validation_functions <- file.path(module_base, "scripts", "utils", "tumor-validation-helpers.R")
source(validation_functions)

# source in helper functions for make_jaccard_matrix() and jaccard()
jaccard_functions <- file.path(module_base, "scripts", "utils", "jaccard-functions.R")
source(jaccard_functions)
```

```{r}
# Input files
sce_file <- file.path(sample_dir, "SCPCL000824_processed.rds")
marker_genes_file <- file.path(module_base, "references", "tumor-marker-genes.tsv")

# annotations from SCPCL000824_tumor-gene-expression-exploration.Rmd
annotations_dir <- file.path(module_base, "cell_type_annotations", "SCPCS000492")
annotations_file <- file.path(annotations_dir, "SCPCL000824_tumor-classifications.tsv.gz")

results_dir <- file.path(module_base, "results", "annotate_tumor_cells_output", "SCPCS000492")
geneset_scores_file <- file.path(results_dir, "SCPCL000824_gene-set-scores.tsv")

# singler results 
full_singler_results <- file.path(module_base, "scratch", "SingleR", "SCPCL000824_singler-results.rds")
singler_annotations_file <- file.path(annotations_dir, "SCPCL000824_singler-classifications.tsv")
```

```{r}
# read in sce file 
sce <- readr::read_rds(sce_file)

# read in annotations
annotations_df <- readr::read_tsv(annotations_file)
singler_annotations_df <- readr::read_tsv(singler_annotations_file)

# read in full singler results 
singler_results_list <- readr::read_rds(full_singler_results)

# read in gene set scores
geneset_scores_df <- readr::read_tsv(geneset_scores_file)
```

```{r}
# generate classification df
classification_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
    \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")
  ) |>
  # get rid of excess columns
  dplyr::select(barcodes, UMAP1, UMAP2, singler_celltype_annotation) |> 
  dplyr::left_join(annotations_df, by = c("barcodes" = "cell_barcode")) |> 
  dplyr::left_join(singler_annotations_df, by = "barcodes") |> 
  dplyr::left_join(geneset_scores_df, by = "barcodes") |>
  # get the top cell types for each classification method for plotting later 
  dplyr::mutate(
    across(
      ends_with("_classification"),
      \(x) {
        x |>
          forcats::fct_lump_n(7, other_level = "All remaining cell types", ties.method = "first") |>
          forcats::fct_infreq() |>
          forcats::fct_relevel("All remaining cell types", after = Inf)
      },
      .names = "{.col}_lumped"
    )
  )
```

## SingleR

```{r}
annotations_to_plot <- c(
  "blueprint_tumor_classification_lumped",
  "updated_singler_classification_lumped",
  "updated_singler_blueprint_classification_lumped"
)

annotations_to_plot |> 
  purrr::map(\(annotation_column){
    ggplot(
    classification_df,
    aes(x = UMAP1, y = UMAP2, color = .data[[annotation_column]])
  ) +
    # set points for all "other" points
    geom_point(
      data = dplyr::select(
        classification_df, -.data[[annotation_column]]
      ),
      color = "gray80",
      alpha = 0.5,
      size = 0.5
    ) +
    # set points for desired cell type
    geom_point(size = 0.5, alpha = 0.5) +
    facet_wrap(
      vars(.data[[annotation_column]]),
      ncol = 3
    ) +
    scale_color_brewer(palette = "Dark2") +
    # remove axis numbers and background grid
    scale_x_continuous(labels = NULL, breaks = NULL) +
    scale_y_continuous(labels = NULL, breaks = NULL) +
    guides(
      color = guide_legend(
        title = "Cell types",
        # more visible points in legend
        override.aes = list(
          alpha = 1,
          size = 1.5
        )
      )
    )
  })
```

Compare tumor vs. other in all classification methods vs. updated

```{r}
plot_jaccard <- function(classification_df,
                         singler_results_column,
                         methods_to_compare
                         ){

  # create jaccard matrices for SingleR and CellAssign compared to aneuploid/diploid 
  jaccard_matrices <- methods_to_compare |>
    purrr::map(\(name) {
      make_jaccard_matrix(
        classification_df,
        singler_results_column,
        name
      )
    })

  # Set heatmap padding option
  heatmap_padding <- 0.2
  ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(heatmap_padding, "in"))
  # list of heatmaps looking at SingleR/ CellAssign vs tumor/normal 
  heatmap <- jaccard_matrices |>
    purrr::imap(
      \(celltype_mat, celltype_method) {
        ComplexHeatmap::Heatmap(
          t(celltype_mat), # transpose because matrix rows are in common & we want a vertical arrangement
          col = circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue")),
          border = TRUE,
          ## Row parameters
          cluster_rows = TRUE,
          row_title = celltype_method,
          row_title_gp = grid::gpar(fontsize = 12),
          row_title_side = "left",
          row_names_side = "left",
          row_dend_side = "right",
          row_names_gp = grid::gpar(fontsize = 10),
          ## Column parameters
          cluster_columns = TRUE,
          column_title = singler_results_column,
          column_title_gp = grid::gpar(fontsize = 12),
          column_names_side = "bottom",
          column_names_gp = grid::gpar(fontsize = 10),
          column_names_rot = 90,
          ## Legend parameters
          heatmap_legend_param = list(
            title = "Jaccard index",
            direction = "vertical",
            legend_width = unit(1.5, "in")
          ),
          show_heatmap_legend = celltype_method == "UCell",
        )
      }) |>
    # concatenate vertically into HeatmapList object
    purrr::reduce(ComplexHeatmap::`%v%`) |>
    ComplexHeatmap::draw(
      heatmap_legend_side = "right",
      # add a margin to the heatmap so labels don't get cut off
      padding = unit(c(2, 20, 2, 2), "mm")
    )

  return(heatmap)

}
```

```{r}
 original_methods <- c(
    "updated_marker_gene_classification",
    "auc_classification",
    "ucell_classification"
  ) |> 
  purrr::set_names("Marker genes", "AUCell", "UCell")

heatmap_list <- annotations_to_plot |> 
  purrr::walk(\(annotation_column) {plot_jaccard(classification_df, annotation_column, original_methods)})
```

```{r}
methods <- c(
    "blueprint_tumor_classification_lumped",
    "updated_singler_classification_lumped",
    "updated_singler_blueprint_classification_lumped"
  ) |> 
  purrr::set_names("Blueprint", "Original", "Original and Blueprint")

heatmap_list <- plot_jaccard(classification_df, "singler_celltype_annotation", methods)
```

Validate marker gene and gene set expression across all three methods

```{r}
plot_markers_df <- create_marker_gene_df(
  sce,
  classification_df, 
  marker_genes_file
)
```

```{r}
# create a density plot showing the distribution of marker gene expression across classification methods 
marker_density_df <- plot_markers_df |>
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  dplyr::mutate(
    method = dplyr::case_when(
      method == "marker_gene_classification" ~ "Marker genes only",
      method == "updated_marker_gene_classification" ~ "Marker gene cutoff from 822",
      method == "auc_classification" ~ "AUCell",
      method == "ucell_classification" ~ "UCell",
      method == "blueprint_tumor_classification" ~ "Blueprint",
      method == "updated_singler_classification" ~ "Original SingleR with tumor",
      method == "updated_singler_blueprint_classification" ~ "Tumor + SingleR + Blueprint"
    ),
    classification = dplyr::if_else(classification == "Tumor", "Tumor", "Other")
  )

ggplot(marker_density_df, aes(x = sum_transformed_exp, color = classification)) +
  geom_density() +
  facet_wrap(vars(method))
```

```{r}
# plot gene set scores for each cell  
geneset_plot_df <- classification_df |> 
  dplyr::select(barcodes, UMAP1, UMAP2, ends_with("classification"), starts_with("mean-")) |> 
  tidyr::pivot_longer(
    cols = starts_with("mean"),
    names_to = "geneset",
    values_to = "mean_score"
  ) |> 
  dplyr::mutate(
    geneset = stringr::word(geneset, -1, sep = "-")
  ) |> 
  tidyr::pivot_longer(
    cols = ends_with("classification"),
    names_to = "method",
    values_to = "classification"
  ) |> 
  dplyr::mutate(
    method = dplyr::case_when(
      method == "marker_gene_classification" ~ "Marker genes only",
      method == "updated_marker_gene_classification" ~ "Marker gene cutoff from 822",
      method == "auc_classification" ~ "AUCell",
      method == "ucell_classification" ~ "UCell",
      method == "blueprint_tumor_classification" ~ "Blueprint",
      method == "updated_singler_classification" ~ "Original SingleR with tumor",
      method == "updated_singler_blueprint_classification" ~ "Tumor + SingleR + Blueprint"
    ),
    classification = dplyr::if_else(classification == "Tumor", "Tumor", "Other")
    )

ggplot(geneset_plot_df, aes(x = mean_score, color = classification)) +
  geom_density(bw = 0.05) +
  facet_grid(rows = vars(method), 
             cols = vars(geneset))

```

Look at consensus across all three references. 
For normal cells, identify some top markers and see if they are expressed in that group of cells compared to all other cells.