### 2024-07-10 analyses/cell-type-ewings/template_notebooks/auc-workflow/01-auc-results.Rmd

---
title: "Summary of tumor cell classification with `AUCell` for `r params$library_id`"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
params:
  library_id: NULL
  sce_file: NULL
  aucell_results_file: NULL
  ref_auc_results: NULL
  geneset_scores_file: NULL
  marker_genes_file: NULL
  auc_threshold: NULL
---

This report summarizes the classification of `r params$library_id` using [`AUCell`](https://www.bioconductor.org/packages/release/bioc/html/AUCell.html). 
Cells were classified as tumor if they had an AUC value greater than or equal to `r params$auc_threshold`.

## Setup

```{r}
# check that sce and results files exist 
stopifnot(
  "sce file does not exist" = file.exists(params$sce_file),
  "AUCell results file does not exist" = file.exists(params$aucell_results_file), 
  "reference results from AUCell does not exist" = file.exists(params$ref_auc_results), 
  "Marker genes file does not exist" = file.exists(params$marker_genes_file), 
  "Gene set scores file does not exist" = file.exists(params$geneset_scores_file)
)
```

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)

# quiet messages
options(readr.show_col_types = FALSE)
ComplexHeatmap::ht_opt(message = FALSE)

# set colors for tumor and normal cells throughout report 
plot_colors <- c("Tumor" = "#00274C", "Normal" = "#FFCB05")
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-ewings")

# source in helper functions: plot_gene_heatmap() and plot_cnv_heatmap() 
# create_classification_df() and create_marker_gene_df()
validation_functions <- file.path(module_base, "scripts", "utils", "tumor-validation-helpers.R")
source(validation_functions)
```

```{r}
# Read in files
sce <- readr::read_rds(params$sce_file)
auc_results_df <- readr::read_tsv(params$aucell_results_file)
geneset_scores_df <- readr::read_tsv(params$geneset_scores_file)

# reference auc results file 
ref_auc_df <- readr::read_tsv(params$ref_auc_results)
```

```{r}
# generate classification df to use for plots 
classification_df <- sce |> 
  scuttle::makePerCellDF(use.dimred = "UMAP") |> 
  # replace UMAP.1 with UMAP1
  dplyr::rename_with(
    \(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")
  ) |>
  # get rid of excess columns
  dplyr::select(barcodes, UMAP1, UMAP2, singler_celltype_annotation) |> 
  # join with previous annotations, singler results, and gene set scores 
  dplyr::left_join(auc_results_df, by = "barcodes") |> 
  dplyr::left_join(geneset_scores_df, by = "barcodes") |> 
  dplyr::mutate(
    # make a factor so tumor always appears first 
    auc_classification = forcats::fct_relevel(auc_classification, "Tumor")
  )

# get marker gene expression 
markers_df <- create_marker_gene_df(
  sce,
  classification_df, 
  params$marker_genes_file
)
```

## AUCell classification UMAP

Below is a UMAP showing cells that are labeled as either `Tumor` or `Normal` using `AUCell`.

```{r}
ggplot(classification_df, aes(x = UMAP1, y = UMAP2, color = auc_classification)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_color_manual(values = plot_colors) +
  labs(
    color = "AUCell classification"
  )
```

## Comparison of AUC distribution

The below plot compares the distribution of AUC values in the query library as compared to the reference library used to determine the AUC threshold.

```{r}
all_auc_df <- dplyr::bind_rows(
  "reference" = ref_auc_df, 
  "query" = auc_results_df,
  .id = "sample"
) |>
  dplyr::mutate(
    sample = ifelse(sample == "query", paste("query:", params$library_id), sample)
  )

ggplot(all_auc_df, aes(x = auc, color = sample)) +
  geom_density() + 
  geom_vline(xintercept = params$auc_threshold, linetype = "dashed") +
  labs(
    x = "AUC",
    color = ""
  )
```

## Expression of marker genes

The below plots summarize marker gene expression in tumor and normal cells as classified by `AUCell`.

### Density plot

The `logcounts` for each gene in the marker gene list is summed for each cell and shown on the x-axis.
We would expect tumor cells to have higher expression of marker genes than normal cells.

```{r}
# create a density plot showing the distribution of marker gene expression across classification methods 
markers_df |> 
  dplyr::select(barcodes, sum_raw_exp, auc_classification) |> 
  dplyr::distinct() |> 
  ggplot(aes(x = sum_raw_exp, color = auc_classification)) +
  geom_density() +
  scale_color_manual(values = plot_colors) +
  labs(
    x = "Raw sum of marker gene logcounts",
    color = "AUCell classification"
  )
```

### Heatmap

Here every row is a marker gene and every column is a cell with colors corresponding to the expression of a specified marker gene.  
The annotation bar indicates which cells are tumor or normal cells as classified by `AUCell`. 
We expect to see higher expression of individual marker genes in tumor cells compared to normal cells.

```{r}
# create matrix with marker genes as rows and barcodes as columns
marker_gene_matrix <- markers_df |> 
  dplyr::select(gene_expression, gene_symbol, barcodes) |> 
  tidyr::pivot_wider(values_from = gene_expression, 
                     names_from = barcodes) |> 
  tibble::column_to_rownames("gene_symbol") |> 
  as.matrix()

annotation <- ComplexHeatmap::columnAnnotation(
  AUCell = classification_df$auc_classification, 
  col = list(AUCell = plot_colors)
)
```

```{r}
# plot heatmap of marker genes 
plot_gene_heatmap(marker_gene_matrix, 
                  row_title = "Marker gene symbol",
                  legend_title = "Marker gene \nexpression",
                  annotation = annotation)
```

## Gene set scores

The below plots summarize the gene set scores in tumor and normal cells as classified by `AUCell`.

Gene set scores are determined by calculating the mean expression of all genes in a given gene set. 
Scores are calculated for the following gene sets from `MsigDb`:

  - [`ZHANG_TARGETS_OF_EWSR1_FLI1_FUSION`](https://www.gsea-msigdb.org/gsea/msigdb/human/geneset/ZHANG_TARGETS_OF_EWSR1_FLI1_FUSION.html)
  - [`RIGGI_EWING_SARCOMA_PROGENITOR_UP`](https://www.gsea-msigdb.org/gsea/msigdb/human/geneset/RIGGI_EWING_SARCOMA_PROGENITOR_UP.html?ex=1)
  - [`SILIGAN_TARGETS_OF_EWS_FLI1_FUSION_DN`](https://www.gsea-msigdb.org/gsea/msigdb/cards/SILIGAN_TARGETS_OF_EWS_FLI1_FUSION_DN)

### Density plot

```{r}
# prep data frame for plotting gene set scores 
geneset_plot_df <- classification_df |> 
  dplyr::select(barcodes, auc_classification, starts_with("mean-")) |> 
  # make one column containing all gene set scores 
  tidyr::pivot_longer(
    cols = starts_with("mean"),
    names_to = "geneset",
    values_to = "mean_score"
  ) |> 
  # simplify gene set name for labels
  dplyr::mutate(
    geneset = stringr::word(geneset, -1, sep = "-")
  )

# make density plot showing distribution of gene set scores for tumor/normal 
ggplot(geneset_plot_df, aes(x = mean_score, color = auc_classification)) +
  geom_density(bw = 0.05) +
  facet_grid(rows = vars(geneset),
             scales = "free_y") +
  scale_color_manual(values = plot_colors) +
  labs(
    x = "Gene set score",
    color = "AUCell classification"
  )
```

### Heatmap

Here every column is a cell and every row is a gene set with colors corresponding to the gene set score. 
The annotation bar indicates which cells are tumor or normal cells as classified by `AUCell`. 
We expect to see higher gene set scores in tumor cells compared to normal cells.

```{r}
# make a matrix of gene set by barcode 
geneset_matrix <- geneset_plot_df |> 
  dplyr::select(mean_score, geneset, barcodes) |> 
  unique() |> 
  tidyr::pivot_wider(values_from = mean_score, 
                     names_from = barcodes) |> 
  tibble::column_to_rownames("geneset") |> 
  as.matrix()

# plot heatmap of gene set score 
plot_gene_heatmap(geneset_matrix,
                  annotation = annotation,
                  legend_title = "Gene set \nscore")
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```