### 2024-07-23 analyses/doublet-detection/exploratory-notebooks/03_compare-benchmark-results.Rmd
spacing to trigger styling

---
params: 
  datasets: "hm-6k pbmc-1B-dm pdx-MULTI HMEC-orig-MULTI"
title: "Comparison among doublet predictions on benchmarking datasets"
author: Stephanie J. Spielman
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 4
    code_folding: hide
---

This notebook compares doublet results calculated on benchmarking datasets to one another, with the primary goal of addressing these questions:

1. Do methods tend to predict overlapping or distinct sets of doublets on the same dataset?
2. Is the consensus doublet call across methods predictive of the true doublet status?
Here, the "consensus doublets" are those droplets which all methods identify as doublets.

There are several items to bear in mind when interpreting these results:

- The ground truth calls themselves to which we are comparing consensus calls may not be entirely accurate, since they were also computationally identified generally with demultiplexing algorithms.
- `cxds`, as we are using it, does not have a specific threshold for calling droplets.
By contrast, both `scrublet` and `scDblFinder` identify a threshold based on the given dataset they are processing.
This notebook uses a doublet threshold of `>=0.5` for `cxds`, which may not be universally suitable (though choosing a universally suitable threshold is not easy either!).

## Setup

### Packages

```{r packages}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(ggplot2)
  library(patchwork)
  library(caret)
  library(UpSetR)
})

theme_set(
  theme_bw() +
    theme(
      legend.title = element_text(size = rel(1.1)),
      legend.text = element_text(size = rel(1.1))
    )
)

# define threshold used to call cxds
cxds_threshold <- 0.5
```

### Paths

```{r base paths}
module_base <- rprojroot::find_root(rprojroot::is_renv_project)
data_dir <- file.path(module_base, "scratch", "benchmark-datasets")
result_dir <- file.path(module_base, "results", "benchmark-results")
```

### Functions

```{r}
plot_pca_calls <- function(df,
                           color_column,
                           plot_type = "calls",
                           title) {
  # Plot PCs colored by either:
  # if plot_type == "calls", color by singlet or doublet, showing doublets on top
  # if plot_type == "class", color by singlet, doublet, or ambiguous, showing doublets & ambiguous on top
  # df is expected to contain columns PC1, PC2, `color_column`, which should _not_ be provided as a string
  plot <- ggplot(df) +
    aes(
      x = PC1,
      y = PC2,
      color = {{ color_column }}
    ) +
    geom_point(
      size = 0.75,
      alpha = 0.6
    ) +
    ggtitle(title) +
    guides(color = guide_legend(override.aes = list(size = 2)))

  if (plot_type == "calls") {
    plot <- plot +
      scale_color_manual(
        name = "",
        values = c("doublet" = "black", "singlet" = "gray90")
      ) +
      geom_point(
        data = dplyr::filter(df, {{ color_column }} == "doublet"),
        color = "black",
        size = 0.75
      )
  } else if (plot_type == "class") {
    plot <- plot +
      scale_color_manual(
        name = "Consensus",
        values = c(
          "ambiguous" = "yellow",
          "doublet" = "black",
          "singlet" = "gray90"
        )
      ) +
      geom_point(
        data = dplyr::filter(df, consensus_class != "singlet"),
        size = 0.75,
        alpha = 0.7,
        aes(color = consensus_class),
      )
  } else {
    stop("plot_type must be 'calls' or 'class'")
  }

  return(plot)
}

plot_pca_metrics <- function(df, color_column) {
  # Plot PCs colored by performance metric, showing false calls on top
  # metric_colors is a named vector of colors used for coloring tp/tn/fp/fn

  # used in PCA plots
  metric_colors <- c(
    "tp" = "lightblue",
    "tn" = "pink",
    "fp" = "blue",
    "fn" = "firebrick2"
  )

  ggplot(df) +
    aes(
      x = PC1,
      y = PC2,
      color = {{ color_column }}
    ) +
    geom_point(
      size = 0.75,
      alpha = 0.6
    ) +
    geom_point(
      data = dplyr::filter(df, {{ color_column }} %in% c("fp", "fn")),
      size = 0.75
    ) +
    scale_color_manual(name = "Call type", values = metric_colors) +
    ggtitle("Consensus call metrics") +
    guides(color = guide_legend(override.aes = list(size = 2)))
}
```

## Read and prepare input data

First, we'll read in and combine doublet results, as well as PCA coordinates, into a list of data frames for each dataset.

```{r dataset names}
# find all dataset names to process:
dataset_names <- stringr::str_split_1(params$datasets, pattern = " ")
```

```{r read_data}
# Read in data for analysis
doublet_df_list <- dataset_names |>
  purrr::map(
    \(dataset) {
      scdbl_tsv <- file.path(result_dir, glue::glue("{dataset}_scdblfinder.tsv"))
      scrub_tsv <- file.path(result_dir, glue::glue("{dataset}_scrublet.tsv"))
      sce_file <- file.path(data_dir, dataset, glue::glue("{dataset}.rds"))

      scdbl_df <- scdbl_tsv |>
        readr::read_tsv(show_col_types = FALSE) |>
        dplyr::select(
          barcodes,
          cxds_score,
          scdbl_score = score,
          scdbl_prediction = class
        ) |>
        # add cxds calls at `cxds_threshold` threshold
        dplyr::mutate(
          cxds_prediction = dplyr::if_else(
            cxds_score >= cxds_threshold,
            "doublet",
            "singlet"
          )
        )

      scrub_df <- readr::read_tsv(scrub_tsv, show_col_types = FALSE)

      # grab ground truth and PCA coordinates
      sce <- readr::read_rds(sce_file)

      scuttle::makePerCellDF(sce, use.dimred = "PCA") |>
        tibble::rownames_to_column(var = "barcodes") |>
        dplyr::select(barcodes,
          ground_truth = ground_truth_doublets,
          PC1 = PCA.1,
          PC2 = PCA.2
        ) |>
        dplyr::left_join(
          scrub_df,
          by = "barcodes"
        ) |>
        dplyr::left_join(
          scdbl_df,
          by = "barcodes"
        )
    }
  ) |>
  purrr::set_names(dataset_names)
```

## Upset plots

This section contains upset plots that show overlap across doublet calls from each method, displayed for each dataset.

```{r}
doublet_df_list |>
  purrr::iwalk(
    \(df, dataset) {
      doublet_barcodes <- list(
        "scDblFinder" = df$barcodes[df$scdbl_prediction == "doublet"],
        "scrublet"    = df$barcodes[df$scrublet_prediction == "doublet"],
        "cxds"        = df$barcodes[df$cxds_prediction == "doublet"]
      )

      UpSetR::upset(fromList(doublet_barcodes), order.by = "freq") |> print()
      grid::grid.text( # plot title
        dataset,
        x = 0.65,
        y = 0.95,
        gp = grid::gpar(fontsize = 16)
      )
    }
  )
```

## Explore consensus across methods

Next, we will explore the consensus of doublet predictions across methods.
To this end, we'll create some new columns for each dataset:

- `consensus_class`, which will be "doublet" if all three methods are doublet, "singlet" if all three methods are singlet, and "ambiguous" if there are any disagreements
- `consensus_call`, which will be "doublet" if _all_ methods predict "doublet," and "singlet" otherwise
- `confusion_call`, which will classify the consensus call as one of `tp`, `tn`, `fp`, or `fn` (true/false positive/negative) based on `consensus_call`

```{r}
doublet_df_list <- doublet_df_list |>
  purrr::map(
    \(dataset_df) {
      dataset_df |>
        dplyr::rowwise() |>
        dplyr::mutate(
          # Add column `consensus_call`
          # "doublet" if all three methods are doublet, and "singlet" otherwise
          consensus_call = dplyr::if_else(
            all(c(scdbl_prediction, scrublet_prediction, cxds_prediction) == "doublet"),
            "doublet",
            "singlet"
          ),
          # Add column `consensus_class`
          #  "doublet" if all three methods are doublet, "singlet" if all three methods are singlet,
          #  and "ambiguous" if there are any disagreements
          consensus_class = dplyr::case_when(
            consensus_call == "doublet" ~ "doublet",
            all(c(scdbl_prediction, scrublet_prediction, cxds_prediction) == "singlet") ~ "singlet",
            .default = "ambiguous"
          ),
          # Add column `confusion_call`
          #  tp/tn/fp/fn based on the `consensus_call` column
          confusion_call = dplyr::case_when(
            consensus_call == "doublet" && ground_truth == "doublet" ~ "tp",
            consensus_call == "singlet" && ground_truth == "singlet" ~ "tn",
            consensus_call == "doublet" && ground_truth == "singlet" ~ "fp",
            consensus_call == "singlet" && ground_truth == "doublet" ~ "fn"
          )
        )
    }
  )
```

### PCA

This section plots the PCA for each dataset, clockwise from the top left:

1. `scDblFinder` singlet/doublet calls
2. `scrublet` singlet/doublet calls
3. `cxds` singlet/doublet calls
4. Ground truth single/doublets
5. Consensus _call_ droplets are either "doublet" or "singlet"
6. Consensus _class_: droplets are either "doublet", "singlet", or "ambiguous"
7. Metrics based on the consensus call

- For the first five panels, doublets are shown in black and singlets in light gray.
- In the sixth panel showing the consensus class, doublets are shown in black, singlets in light gray, and ambiguous are yellow
- In the seventh panel showing metrics, colors are:
  - True positive: light blue
  - True negative: pink
  - False positive: blue
  - False negative: red

```{r, fig.width = 12, fig.height = 9}
pc_color_columns <- c(
  "scDblFinder prediction" = "scdbl_prediction",
  "scrublet prediction" = "scrublet_prediction",
  "cxds prediction" = "cxds_prediction",
  "Ground truth" = "ground_truth",
  "Consensus call" = "consensus_call"
)
doublet_df_list |>
  purrr::imap(
    \(df, dataset) {
      pc_plot_list <- pc_color_columns |>
        purrr::imap(
          \(color_column, plot_title) {
            plot_pca_calls(
              df,
              color_column = !!sym(color_column),
              plot_type = "calls",
              title = plot_title
            ) + theme(legend.position = "none")
          }
        )

      # plot for consensus_class, which needs three colors
      # note that the list name isn't actually used here
      pc_plot_list$consensus_class <- plot_pca_calls(
        df,
        color_column = consensus_class,
        plot_type = "class",
        title = "Consensus class"
      )

      # plot metrics
      pc_plot_list$metrics <- plot_pca_metrics(
        df,
        color_column = confusion_call
      )

      patchwork::wrap_plots(pc_plot_list) +
        plot_annotation(
          glue::glue("PCA for {dataset}"),
          theme = theme(plot.title = element_text(size = 16))
        ) +
        plot_layout(guides = "collect")
    }
  )
```

### Performance metrics

This section shows a table of the consensus class counts and calculates a confusion matrix with associated statistics from the consensus calls:

```{r}
metric_df <- doublet_df_list |>
  purrr::imap(
    \(df, dataset) {
      print(glue::glue("======================== {dataset} ========================"))

      cat("Table of consensus class counts:")
      print(table(df$consensus_class))

      cat("\n\n")

      # run confusion matrix only if there are both singlets and doublets
      if ("singlet" %in% df$consensus_call & "doublet" %in% df$consensus_call) {
        confusion_result <- caret::confusionMatrix(
          # truth should be first
          table(
            "Truth" = df$ground_truth,
            "Consensus prediction" = df$consensus_call
          ),
          positive = "doublet"
        )

        print(confusion_result)

        # Extract information we want to present later in a table
        return(
          tibble::tibble(
            "Dataset name" = dataset,
            "Kappa" = round(confusion_result$overall["Kappa"], 3),
            "Balanced accuracy" = round(confusion_result$byClass["Balanced Accuracy"], 3)
          )
        )
      }
    }
  ) |>
  dplyr::bind_rows()
```

### Conclusions

Overall, methods do not have substantial overlap with each other.
They each tend to detect different sets of doublets, leading to fairly small sets of consensus doublets.
Further, the consensus doublets called by all three methods have some, but not substantial, overlap with the ground truth.

For two out of four datasets, `scDblFinder` predicts a much larger number of doublets compared to other methods.

Based on the PCAs, it additionally looks like there are many more false negatives in the consensus prediction that `cxds` appears to capture but other methods do not.

The table below summarizes performance of the "consensus caller".
Note that, in the [benchmarking paper these datasets were originally analyzed in](https://doi.org/10.1016/j.cels.2020.11.008), `hm-6k` was observed to be one of the "easiest" datasets to classify across methods.
Consistent with that observation, it has the highest `kappa` value here.

```{r}
metric_df
```

## Explore consensus of `scDblFinder` and `cxds`

The above analysis suggests that `scrublet` may be more conservative than the other two methods, so including it in a consensus may be increasing the number of false positives.
Here, we'll repeat the same consensus analyses, but considering _only_ `scDblFinder` and `cxds`.
Since there are only two methods here, we will create two columns for this data:

### Prepare data

```{r}
# modify existing consensus columns to only consider 2 methods
doublet_df_list <- doublet_df_list |>
  purrr::map(
    \(dataset_df) {
      dataset_df <- dataset_df |>
        dplyr::rowwise() |>
        # Update column `consensus_call`
        dplyr::mutate(
          consensus_call = dplyr::if_else(
            all(c(scdbl_prediction, cxds_prediction) == "doublet"),
            "doublet",
            "singlet"
          )
        ) |>
        # Add Update `consensus_class`
        dplyr::mutate(
          consensus_class = dplyr::case_when(
            consensus_call == "doublet" ~ "doublet",
            all(c(scdbl_prediction, cxds_prediction) == "singlet") ~ "singlet",
            .default = "ambiguous"
          )
        ) |>
        # Add Update `confusion_call`
        dplyr::mutate(
          confusion_call = dplyr::case_when(
            consensus_call == "doublet" && ground_truth == "doublet" ~ "tp",
            consensus_call == "singlet" && ground_truth == "singlet" ~ "tn",
            consensus_call == "doublet" && ground_truth == "singlet" ~ "fp",
            consensus_call == "singlet" && ground_truth == "doublet" ~ "fn"
          )
        )

      return(dataset_df)
    }
  ) |>
  purrr::set_names(dataset_names)
```

### PCA

This section plots the PCA for each dataset, clockwise from the top left:

1. `scDblFinder` singlet/doublet calls
2. `cxds` singlet/doublet calls
3. Ground truth single/doublets
4. Consensus _call_ droplets are either "doublet" or "singlet"
5. Consensus _class_: droplets are either "doublet", "singlet", or "ambiguous"
6. Metrics based on the consensus call

```{r, fig.width = 12, fig.height = 6}
pc_color_columns <- c(
  "scDblFinder prediction" = "scdbl_prediction",
  "cxds prediction" = "cxds_prediction",
  "Ground truth" = "ground_truth",
  "Consensus call" = "consensus_call"
)

doublet_df_list |>
  purrr::imap(
    \(df, dataset) {
      pc_plot_list <- pc_color_columns |>
        purrr::imap(
          \(color_column, plot_title) {
            plot_pca_calls(
              df,
              color_column = !!sym(color_column),
              plot_type = "calls",
              title = plot_title
            ) + theme(legend.position = "none")
          }
        )
      # plot for consensus_class, which needs three colors
      # note that the list name isn't actually used here
      pc_plot_list$consensus_class <- plot_pca_calls(
        df,
        color_column = consensus_class,
        plot_type = "class",
        title = "Consensus class"
      )

      # plot metrics
      pc_plot_list$metrics <- plot_pca_metrics(
        df,
        color_column = confusion_call
      )

      patchwork::wrap_plots(pc_plot_list) +
        plot_annotation(
          glue::glue("PCA for {dataset}"),
          theme = theme(plot.title = element_text(size = 16))
        ) +
        plot_layout(guides = "collect")
    }
  )
```

### Performance metrics

This section shows a table of the consensus class counts and calculates a confusion matrix with associated statistics from the consensus calls:

```{r}
metric_df <- doublet_df_list |>
  purrr::imap(
    \(df, dataset) {
      print(glue::glue("======================== {dataset} ========================"))

      cat("Table of consensus class counts:")
      print(table(df$consensus_class))

      cat("\n\n")

      # run confusion matrix only if there are both singlets and doublets
      if ("singlet" %in% df$consensus_call & "doublet" %in% df$consensus_call) {
        confusion_result <- caret::confusionMatrix(
          # truth should be first
          table(
            "Truth" = df$ground_truth,
            "Consensus prediction" = df$consensus_call
          ),
          positive = "doublet"
        )

        print(confusion_result)

        # Extract information we want to present later in a table
        return(
          tibble::tibble(
            "Dataset name" = dataset,
            "Kappa" = round(confusion_result$overall["Kappa"], 3),
            "Balanced accuracy" = round(confusion_result$byClass["Balanced Accuracy"], 3)
          )
        )
      }
    }
  ) |>
  dplyr::bind_rows()
```

### Conclusions

Considering only `scDblFinder` and `cxds`, we see several particular differences in statistics:

- Balanced accuracy for `hm-6k` and `HMEC-orig-MULTI` are about the same here as for the consensus among all methods, but accuracy has substantially _decreased_ for `pbmc-1B-dm` and `pdx-MULTI1`
- By contrast, kappa values have increased for all datasets, with the most marked increase for `pdx-MULTI`.

Even so, only `hm-6k` (the "easy" dataset) has both a high kappa and high balanced accuracy.
While `HMEC-orig-MULTI`'s balanced accuracy is fairly high, its kappa value is not.

```{r}
metric_df
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```