### 2024-08-05 analyses/cell-type-ewings/template_notebooks/auc-singler-workflow/03-clustering.Rmd
clustering template

---
title: "Clustering exploration for `r params$library_id`"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
params:
  library_id: "SCPCL000822"
  sce_file: "../../../../data/current/SCPCP000015/SCPCS000490/SCPCL000822_processed.rds"
  singler_results_file: "../../results/aucell_singler_annotation/SCPCS000490/SCPCL000822_singler-classifications.tsv"
  cluster_results_file: "../../results/aucell_singler_annotation/SCPCS000490/SCPCL000822_clusters.tsv" 
  marker_genes_file: "../../references/visser-all-marker-genes.tsv"
---

This notebook summarizes clustering for `r params$library_id`.

- Louvain clusters are calculated using a range of parameters specifying nearest neighbors (5, 10, 15, 20, 25, 30, 35, and 40). 
- Metrics are then calculated to evaluate clustering results across all parameters: 
  - Silhouette width 
  - Cluster purity 
  - Cluster stability 
- The clusters are then compared to the results from running `SingleR` in the `aucell-singler-annotation.sh` workflow.
- We then look at marker gene expression across the clusters and assign each cluster to a cell type.

## Setup

```{r}
# check that sce and results files exist
stopifnot(
  "sce file does not exist" = file.exists(params$sce_file),
  "SingleR results file does not exist" = file.exists(params$singler_results_file),
  "Marker genes file does not exist" = file.exists(params$marker_genes_file)
)
```

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)

# quiet messages
options(readr.show_col_types = FALSE)
ComplexHeatmap::ht_opt(message = FALSE)
```

```{r base paths}
# The path to this module
module_base <- rprojroot::find_root(rprojroot::is_renv_project)

# source in helper functions: calculate_sum_markers()
validation_functions <- file.path(module_base, "scripts", "utils", "tumor-validation-helpers.R")
source(validation_functions)

# source in clustering functions
clustering_functions <- file.path(module_base, "scripts", "utils", "clustering-functions.R")
source(clustering_functions)
```

```{r}
# read in input sce files
sce <- readr::read_rds(params$sce_file)

# read in singler results
singler_classification_df <- readr::read_tsv(params$singler_results_file)
```

## Clustering

Below we perform Louvain, Jaccard clustering, varying `k`. 
The minimum `k` is 5 and the maximum `k` is 40 with a step size of 5.

```{r}
# perform clustering across k = 5-40 with increments of 5
all_cluster_results <- cluster_sweep(sce) |>
  dplyr::mutate(
    nn_param = forcats::fct_relevel(nn_param, "k_5", after = 0)
  )
```

```{r}
# get umap embeddings and combine into a data frame with cluster assignments
umap_df <- sce |>
  scuttle::makePerCellDF(use.dimred = "UMAP") |>
  # replace UMAP.1 with UMAP1 and get rid of excess columns
  dplyr::select(barcodes, UMAP1 = UMAP.1, UMAP2 = UMAP.2) |>
  dplyr::left_join(all_cluster_results, by = "barcodes")
```

Below we visualize the cluster assignments using each parameter on a UMAP. 
This can be helpful to see any values of K that may have obvious over or under clustering.

```{r, fig.height=15, fig.width=10}
# look at clustering results for each library across all params
ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = cluster)) +
  geom_point(alpha = 0.5, size = 0.1) +
  facet_wrap(vars(nn_param)) +
  theme(
    aspect.ratio = 1
  ) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 1.5)))
```

## Clustering statistics

Below we calculate a series of statistics:

- Average silhouette width: This metric evaluates cluster separation. 
Cells with large positive silhouette widths are closer to other cells in the same cluster than to cells in different clusters. 
- Average cluster purity: This metric also evaluates cluster separation and tells us the proportion of neighboring cells that are assigned to the same cluster. 
Higher purity values indicate that clusters are well separated. 
- Cluster stability: This evaluates how stable the clustering is to input data. 
Higher values of cluster stability indicate more reproducible clusters.

```{r, warning=FALSE}
# get a combined stats dataframe with purity and width for all clusters
all_stats_df <- get_cluster_stats(sce, all_cluster_results) |>
  dplyr::mutate(
    # make sure the order is correct
    nn_param = forcats::fct_relevel(nn_param, "k_5", after = 0)
  )
```

### Silhouette width

```{r}
# silhouette width for different params
plot_cluster_stats(all_stats_df, width)
```

### Cluster purity

```{r}
# cluster purity for different params
plot_cluster_stats(all_stats_df, purity)
```

### Cluster stability

```{r}
# calculate cluster stability
stability_df <- get_cluster_stability(sce, all_cluster_results) |>
  dplyr::mutate(
    k_value = as.factor(k_value),
    # make sure that k = 5 comes first
    k_value = forcats::fct_relevel(k_value, "5", after = 0)
  )

# plot stability across all values of k
ggplot(stability_df, aes(x = k_value, y = ari)) +
  geom_jitter(width = 0.1) +
  labs(title = "Cluster stability") +
  stat_summary(
    aes(group = k_value),
    color = "red",
    # median and quartiles for point range
    fun = "median",
    fun.min = function(x) {
      quantile(x, 0.25)
    },
    fun.max = function(x) {
      quantile(x, 0.75)
    }
  )
```

## Compare clusters to cell types from `SingleR`

Now we will compare the clustering results to the cell type assignments obtained from `SingleR` in the `aucell-singler-annotation.sh` workflow. 
To compare results we will calculate the Jaccard similarity index between clusters and cell types. 
We expect that good clustering will line up with the cell type assignments so that there is ~ 1 cell type per cluster.

For the plots, we will only display the top 7 cell types and all other cells will be lumped together into `All remaining cell types`.

```{r}
cluster_classification_df <- all_cluster_results |>
  # add in classifications from singler
  dplyr::left_join(singler_classification_df, by = "barcodes") |>
  dplyr::mutate(
    # first grab anything that is tumor and label it tumor
    # NA should be unknown
    singler_annotation = dplyr::case_when(
      stringr::str_detect(singler_annotation, "tumor") ~ "tumor",
      is.na(singler_annotation) ~ "unknown", # make sure to separate out unknown labels
      .default = singler_annotation
    ) |>
      forcats::fct_relevel("tumor", after = 0),
    # get the top cell types for plotting later
    singler_lumped = singler_annotation |>
      forcats::fct_lump_n(7, other_level = "All remaining cell types", ties.method = "first") |>
      forcats::fct_infreq() |>
      forcats::fct_relevel("All remaining cell types", after = Inf),
    nn_param = forcats::fct_relevel(nn_param, "k_5", after = 0)
  )
```

```{r, fig.height=10, fig.width=7}
# get heatmap for showing cluster assignments vs. singler assignments
cluster_celltype_heatmap(cluster_classification_df)
```

## Marker gene expression across clusters

We will also look at the expression of marker genes across each cluster. 
In these plots, each row shows the distribution of the specified marker genes in that cluster. 
Each panel is labeled with the marker gene expression being plotted.

```{r}
# read in marker genes table
marker_genes_df <- readr::read_tsv(params$marker_genes_file, show_col_types = FALSE) |>
  # account for genes being from multiple sources
  dplyr::select(cell_type, ensembl_gene_id, gene_symbol) |>
  dplyr::distinct()

# get list of all cell types found
cell_types <- unique(marker_genes_df$cell_type)

# get the sum of expression of all genes for each cell type
gene_exp_df <- cell_types |>
  purrr::map(\(type){
    calculate_sum_markers(marker_genes_df, sce, type)
  }) |>
  purrr::reduce(dplyr::inner_join, by = "barcodes")

# join sum expression columns with clustering results
cluster_exp_df <- all_cluster_results |>
  dplyr::left_join(gene_exp_df, by = "barcodes")
```

```{r, fig.height=10, fig.width=5}
# plot to look at marker gene expression for all cell types across all clusters and all params
k_params <- unique(cluster_exp_df$nn_param) |> as.character()

k_params |>
  purrr::map(\(k) {
    plot_marker_genes(cluster_exp_df,
      k_value = k
    )
  })
```

## Session info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```