### 2024-08-07 analyses/cell-type-wilms-tumor-06/01-clustering_SCPCS000169.Rmd

---
title: "Evaluation of clustering parameters - SCPCS000169"
author: maud-p
date: '`r format(Sys.time(), "%B %d, %Y %H:%M:%S %Z")`'
output:
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: kable
params:
  sample_name: ~
  sample_path: ~
  out_rds_path: ~
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warnings=FALSE)
```

## Introduction

This a template for an analysis notebook using RMarkdown.

In this notebook, we will set up parameters in the seurat workflow [normalization --> reduction --> clustering] for one Wilms tumor sample (SCPCS000169) of the Wilms Tumor dataset (SCPCP000006).

This correspond to the step 2 of the proposed analysis: clustering of cells across a set of parameters for few samples

- https://github.com/AlexsLemonade/OpenScPCA-analysis/discussions/635
- https://github.com/AlexsLemonade/OpenScPCA-analysis/issues/671

### Packages

Load required packages in the following chunk, if needed.
Do not install packages here; only load them with the `library()` function.

```{r packages, message=FALSE, warning=FALSE}
library(SingleCellExperiment)
library(Seurat)      # to use Seurat workflow

library(tidyr)
library(dplyr)
library(DT)          # for table visualization

library(ggplot2)     # for visualization
library(patchwork)   # for visualization
library(ggplotify)
library(SCpubr)      # for visualization
library(viridis)     # for visualization - colors

library(edgeR)       # For pseudobulk based differential expression (DE) analysis
library(DElegate)    # for pseudobulk based DE analysis and identification of marker genes

library(Azimuth)     # For label transfer

library(msigdbr)     # for gene set enrichment analysis or pathway enrichment
library(clusterProfiler)

library(data.table)
```

#### Some config set-up / threshold

We store in a config list names "cfg" parameters used all along the analysis to filter for p-value, log fold change, percentage of expression, etc.

```{r cfg}
cfg = list()
cfg$padj_thershold = 0.05
cfg$lfc_threshold = 1
cfg$rate1_threshold = 0.5
set.seed(12345)
```

#### Base directories

```{r base paths, eval=TRUE, include=TRUE}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "2024-07-08", "SCPCP000006")

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-wilms-tumor-06")
```

#### Input and output files

For this analysis, we worked with the _processed.rds data. 
We builded a Seurat object based on the counts data and re-perform the analysis 
[normalization --> reduction --> clustering] following the Seurat workflow.

We transferred meta.data to keep:

[x] QC data computed by the DataLab

[x] annotation data computed by the DataLab

[x] raw annotation and gene_symbol conversion

```{r paths}
# input
# use download-data.py to download the data

# get all files in the data directory
filelist <- list.files(data_dir, full.names = TRUE)

# select the 40 Wilms tumor single nucleus folders only
filelist <- filelist[! grepl(".tsv", filelist)]

# output
## html report for each of the Wilms tumor sample will be saved in 
path_to_report <- file.path(module_base, "01-clustering")
## some plots from the report will be saved in 
path_to_plot <- file.path(module_base, "/plots")
## after final decision on the clustering strategy, processed RDS files will be saved in 
path_to_result <- file.path(module_base, "/results")
```

```{r paths2}
## to be modified when rendered
i = 2
filedir <- filelist[i] 
sample <- gsub(paste0(data_dir, "/"), "", filedir)

metadata <- read.table(file.path(data_dir, "/single_cell_metadata.tsv"), sep = "\t", header=TRUE)

metadata[metadata$scpca_sample_id == sample,]
```

Here we have a sample of favorable histology, untreated. Few thoughs:

- We thus do not expect anaplasia-associated patterns (TP53 as an example) to be major driver here.

- We do not expect high immune infiltration (induced by chemotherapy).

- We do not expect chemotherapy induced DNA damages.

### Description of the analysis workflow

We perform the following analysis to assess for the quality of clustering:

[1] We perform some quality check to assess any QC-induced clustering (nFeature, nCount, percent.mito).

[2] We add cell cycle information, as we know that in a specific cell cycle state, the transcriptional program is mostly/exclusively related to cell cycle genes and the identity of cells is difficult to determine. 
We expect these cells to cluster together in a cluster of proliferating cells.

[3] We look at specific marker genes that we reported in the table marker.sets/CellType_metadata.csv to check the relevance of the clustering.

[4] We look at specific pathways that we reported in the table marker.sets/Pathways_metadata.csv to check the relevance of the clustering.

[5] We run DElegate::FindAllMarkers2 to find markers of the different clusters and manually check if they do make sense. 
DElegate::FindAllMarkers2 is an improved version of Seurat::FindAllMarkers based on pseudobulk differential expression method.

[6] We perform enrichment analysis of marker genes for each seurat clusters. 
We defined all the genes from the seurat object as the universe and used the MSigDB gene sets.

[7] We plot pca/umap reduction grouping with available annotations (singler_, cellassign_). 
We expect at least immune cells to be correctly label and fall into a few set of clusters.

[8] We run label transfer (Azimuth) to transfer annotation from the fetal kidney atlas human reference. 
We plot pca/umap reduction grouping with latest labels. 
We expect it to be the most representative of the cell types in the sample.

### Selected parameters

Please note: to keep the notebook as straight as possible, we decided to show the analysis for the selected set of parameters:

- SCTransform (default parameters)
- RunPCA (default) 
- RunUMAP (dims = 1:50) 
- FindNeighbors (dims = 1:50) 
- FindClusters (default)

Note: Other parameters have been previously tested, but we would like to show in the following report that the one selected is performing good.

## Analysis content

### Load the processed SingleCellExperiment rds object

```{r load}
file <- list.files(filedir)
file <- file[grepl("_processed.rds", file)]

# open the processed rds object
sce <- readRDS(paste0(filedir, "/", file))
```

### Run Seurat pipeline

```{r seurat_wf, fig.width=8, fig.height=4, out.width='100%'}
# convert to seurat
srat <- CreateSeuratObject(counts = counts(sce),
                                    assay = "RNA",
                                    project = sample
                           )

# convert colData and rowData to data.frame for use in the Seurat object
cell_metadata <- as.data.frame(colData(sce))
row_metadata <- as.data.frame(rowData(sce))

# add cell metadata (colData) from SingleCellExperiment to Seurat
srat@meta.data <- cell_metadata

# add row metadata (rowData) from SingleCellExperiment to Seurat
srat[["RNA"]]@meta.data <- row_metadata

# add metadata from SingleCellExperiment to Seurat
srat@misc <- metadata(sce)

# Normalization
options(future.globals.maxSize= 891289600)
srat <- SCTransform(srat, verbose = F, conserve.memory = TRUE)

# dimensionality reduction
srat <- RunPCA(srat, verbose = F)
srat <- RunUMAP(srat, dims = 1:50, verbose = F)

# clustering
srat    <- FindNeighbors(srat, dims = 1:50, verbose = F)
srat    <- FindClusters(srat, verbose = T)

```

### Visualize seurat clusters

```{r fig.width=20, fig.height=5, out.width='100%'}

d2 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE) + ggtitle("Seurat Cluster - umap")
d1 <- SCpubr::do_DimPlot(srat, reduction="pca", group.by = "seurat_clusters", label = TRUE) + ggtitle("Seurat Cluster - pca")
v <- SCpubr::do_ViolinPlot(srat, features = c( "subsets_mito_percent"), ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(2,2,4))

```

The umap clustering is expected for a triphasic Wilms tumor sample. The 3 main clusters might be:

- blastema cancer cells
- epithelium cancer and/or normal cells
- stroma cancer and/or normal cells

The two small clusters C11 and C12 might be immune and endothelial cells.

To be checked in the following analysis!

### Cell cycle information
```{r cell_cycle, fig.height=20, fig.width=30}
s.genes <- srat@assays$RNA@meta.data$gene_ids[srat@assays$RNA@meta.data$gene_symbol %in% cc.genes$s.genes]
g2m.genes <- srat@assays$RNA@meta.data$gene_ids[srat@assays$RNA@meta.data$gene_symbol %in% cc.genes$g2m.genes]

srat <- CellCycleScoring(srat, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE)
```

```{r fig.width=20, fig.height=5, out.width='100%'}

d2 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "Phase", label = TRUE) + ggtitle("Phase - umap")
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE) + ggtitle("Seurat Cluster - umap")

b <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "Phase",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)
d1 + d2 + b + plot_layout(ncol = 3, widths = c(2,2,4))

```

```{r fig.width=10, fig.height=2.5, out.width='100%'}
 SCpubr::do_ViolinPlot(srat, features = c("S.Score", "G2M.Score"), ncol = 2, group.by = "seurat_clusters", legend.position = "none") 
```

C1 might be composed of cycling cells, in higher proportion than the other clusters.

### Look at specific genes

Here, we open the table of marker genes marker-sets/CellType_metadata.csv.

```{r marker_genes, fig.width=12, fig.height=6, out.width='100%'}

  # open the set of reference marker genes

CellType_metadata <- read.table("marker-sets/CellType_metadata.csv", header = TRUE, sep = ",")

DT::datatable(CellType_metadata, caption = ("CellType_metadata"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

                               buttons = c( 'csv', 'excel')))
```

#### Malignant markers

```{r fig.width=25, fig.height=5, out.width='100%'}
 SCpubr::do_ViolinPlot(srat, features = rownames(srat)[rownames(srat) %in% (CellType_metadata$ENSEMBL_ID[CellType_metadata$cell_class %in% c("malignant")])] , ncol = 9, group.by = "seurat_clusters", legend.position = "none") 
```
The set of clusters 0-1-3-4-7-9-10-13 is NCAM1+ (ENSG00000149294) which is a marker of blastema area.

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000149294", pt.size = 0.2) + ggtitle("NCAM1")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000149294", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

C2-7 is positive for COL6A3 which is a marker of stroma Wilms tumor cells (also normal stroma)

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000163359", pt.size = 0.2) + ggtitle("COL6A3")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000163359", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))

```

#### Immune markers

```{r fig.width=20, fig.height=5, out.width='100%'}
 SCpubr::do_ViolinPlot(srat, features = rownames(srat)[rownames(srat) %in% (CellType_metadata$ENSEMBL_ID[CellType_metadata$cell_class %in% c("immune")])]
                       , ncol = 6, group.by = "seurat_clusters", legend.position = "none") 
```

C11 is PTPRC+ (ENSG00000081237) and must be a cluster of immune cells. Wilms tumor are generally cold tumor, especially untreated Wilms tumor. According to the clinical data, the present sample is untreated (treatment = upfront resection). We do not expect high percentage of immune cells. The size of the cluster would therefore make sense.

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000081237", pt.size = 0.2) + ggtitle("PTPRC = CD45")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000081237", ncol = 1, group.by = "seurat_clusters", legend.position = "none", pt.size = 2)

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

#### Other markers

```{r fig.width=20, fig.height=5, out.width='100%'}
 SCpubr::do_ViolinPlot(srat, features = rownames(srat)[rownames(srat) %in% (CellType_metadata$ENSEMBL_ID[! CellType_metadata$cell_class %in% c("immune", "malignant")])]
                       , ncol = 6, group.by = "seurat_clusters", legend.position = "none") 
```

C12 is VWF+ (ENSG00000110799) and must be a cluster of endothelial cells.

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000110799", pt.size = 0.2) + ggtitle("VWF")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000110799", ncol = 1, group.by = "seurat_clusters", legend.position = "none", pt.size = 2)

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

### Look at specific pathways

Here, we open the table of marker genes marker-sets/Pathway_metadata.csv.

```{r marker_pathways, fig.width=12, fig.height=6, out.width='100%'}
# open the set of reference marker genes

Pathway_metadata <- read.csv("marker-sets/Pathway_metadata.csv", header = TRUE, sep = ",")

DT::datatable(Pathway_metadata, caption = ("CellType_metadata"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

                               buttons = c( 'csv', 'excel')))
```
#### TP53 pathway

here we will calculate a TP53 score using AddMduleScore and the genes of the HALLMARK_P53_PATHWAY gene set.

```{r fig.width=10, fig.height=10, out.width='100%'}

## define genesets
hallmarks <- msigdbr(species = "human", category = "H")

TP53_list = hallmarks %>% 
  filter(gs_name == "HALLMARK_P53_PATHWAY") %>%
  dplyr::distinct(gs_name, gene_symbol, human_ensembl_gene) %>% as.data.frame()

TP53_list <- list(TP53_list$human_ensembl_gene)

srat <- AddModuleScore(srat, features = TP53_list, name = "TP53_score")
```

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "TP53_score1", pt.size = 0.2) + ggtitle("TP53 score")
v <- SCpubr::do_ViolinPlot(srat, features = "TP53_score1", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```
"Normal" cells (immune, endothelial) have a slightly higher TP53 score. I am however not sure how relevant the TP53 score is for Favorable sample as the one we are analysing here. To be tested on an anaplastic sample.

#### DNA repair pathway

here we will calculate a DNA_repair score using AddMduleScore and the genes of the HALLMARK_DNA_REPAIR gene set.

```{r fig.width=10, fig.height=10, out.width='100%'}

DNA_repair = hallmarks %>% 
  filter(gs_name == "HALLMARK_DNA_REPAIR") %>%
  dplyr::distinct(gs_name, gene_symbol, human_ensembl_gene) %>% as.data.frame()

DNA_repair_list <- list(DNA_repair$human_ensembl_gene)

srat <- AddModuleScore(srat, features = DNA_repair_list, name = "DNA_repair_score")
```

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "DNA_repair_score1", pt.size = 0.2) + ggtitle("DNA_repair score")
v <- SCpubr::do_ViolinPlot(srat, features = "DNA_repair_score1", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```
I am not sure how relevant the DNA repair score is for Favorable sample as the one we are analysing here. To be tested on an anaplastic sample.

#### DROSHA target genes

```{r fig.width=10, fig.height=10, out.width='100%'}

## define genesets
c3 <- msigdbr(species = "human", category = "C3", subcategory = "TFT:GTRD")

DROSHA_list = c3 %>% 
  filter(gs_name == "DROSHA_TARGET_GENES") %>%
  dplyr::distinct(gs_name, gene_symbol, human_ensembl_gene) %>% as.data.frame()

DROSHA_list <- list(DROSHA_list$human_ensembl_gene)

srat <- AddModuleScore(srat, features = DROSHA_list, name = "DROSHA_score")
```

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "DROSHA_score1", pt.size = 0.2) + ggtitle("DROSHA score")
v <- SCpubr::do_ViolinPlot(srat, features = "DROSHA_score1", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

#### DICER1 target genes

```{r fig.width=10, fig.height=10, out.width='100%'}
DICER1_list = c3 %>% 
  filter(gs_name == "DICER1_TARGET_GENES") %>%
  dplyr::distinct(gs_name, gene_symbol, human_ensembl_gene) %>% as.data.frame()

DICER1_list <- list(DICER1_list$human_ensembl_gene)

srat <- AddModuleScore(srat, features = DICER1_list, name = "DICER1_score")
```

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "DICER1_score1", pt.size = 0.2) + ggtitle("DICER1 score")
v <- SCpubr::do_ViolinPlot(srat, features = "DICER1_score1", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

C10 might have a slightly altered DROSHA/DICER1 ratio compared to the other cluster, but this might be due to the fact that C10 is mostly in G2M phase (?).

### Find marker genes for each of the seurat clusters

In addition to the list of known marker genes, we used an unbiased approach to find transcripts that characterized the different clusters.
We run DElegate::FindAllMarkers2 to find markers of the different clusters and manually check if they do make sense. 
DElegate::FindAllMarkers2 is an improved version of Seurat::FindAllMarkers based on pseudobulk differential expression method. 
Please check the preprint from Chistoph Hafemeister: https://www.biorxiv.org/content/10.1101/2023.03.28.534443v1
and tool described here: https://github.com/cancerbits/DElegate

```{r fig.width=10, fig.height=10, out.width='100%'}
feature_conversion <- srat@assays$RNA@meta.data
de_results   <- DElegate::FindAllMarkers2(srat, group_column = "seurat_clusters")

#filter the most relevant markers
s.markers <- de_results[de_results$padj < cfg$padj_thershold & de_results$log_fc > cfg$lfc_threshold & de_results$rate1 > cfg$rate1_threshold,]

# add gene symbol for easiest interpretation of the result
s.markers$gene_ids <- s.markers$feature
s.markers <- left_join(s.markers,feature_conversion, by = c( "gene_ids") )
identical(s.markers$feature, s.markers$gene_ids)     # check the quality of the merge, must be true

DT::datatable(s.markers, caption = ("marker genes"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

                               buttons = c( 'csv', 'excel')))

# Select top 5 genes for heatmap plotting
s.markers <- na.omit(s.markers)
s.markers %>%
    group_by(group1) %>%
    top_n(n =  5, wt = log_fc) -> top5

# subset for plotting
cells <- WhichCells(srat, downsample = 100)
ss <- subset(srat, cells = cells)
ss <- ScaleData(ss, features = top5$feature)

p1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
p2 <- DoHeatmap(ss, features = top5$feature,  cells = cells, group.by = "seurat_clusters") + NoLegend() + 
  scale_fill_gradientn(colors =  c("#01665e","#35978f",'darkslategray3', "#f7f7f7", "#fee391","#fec44f","#F9AD03")) 
p3 <- ggplot(srat@meta.data, aes(seurat_clusters, fill = seurat_clusters)) + geom_bar() + NoLegend()

common_title <- sprintf("Unsupervised clustering %s, %d cells", srat@meta.data$orig.ident[1], ncol(srat))
show((((p1 / p3) + plot_layout(heights = c(3,2)) | p2) ) + plot_layout(widths = c(1, 2)) + plot_layout(heights = c(3,1)) + plot_annotation(title = common_title))
DT::datatable(top5[, c(1, 9, 11, 12)], caption = ("top 5 marker genes"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

                               buttons = c( 'csv', 'excel')))

```

The cluster that we found seems to make sense as they can be defined by a few number of specific genes each. I don't know most of the marker genes found, but few point below that could support the analysis and annotation.

UNC5D (C0, dependency receptor), has been described in cancer and shown to regulate p53-dependent apoptosis (https://pubmed.ncbi.nlm.nih.gov/24691657/). We know that especially in blastema cell, p53 signaling play an important role.

PDCH7 (C6, protocadherin 7) is expressed in human kidney renal tubule (normal kidney epithelium, see https://www.proteinatlas.org/ENSG00000169851-PCDH7/tissue/kidney) and has been associated with oncogenic function in lung cancer (https://pubmed.ncbi.nlm.nih.gov/30409919/, https://www.frontiersin.org/journals/pharmacology/articles/10.3389/fphar.2023.1217213/full).

EBF1 (C7, Early B Cell factor 1) has been shown to play a role in kidney development, esp. glomerular, while expressed in stroma progenitor (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6727263/).

ERBB4 (C8, EGF receptor family) is expressed during kidney development of epithelial tubule and uteric bud (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3269924/)

FLT1 (C12, vascular endothelial growth factor receptor 1), PCAM and ELMO1 in the top5 of C13 marker genes confirm the identification of endothelial cells (https://www.ahajournals.org/doi/pdf/10.1161/CIRCRESAHA.109.213983, https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4986701/)

### Enrichment analysis of marker genes

Here we perform enrichment analysis of the marker genes found in the previous section for each Seurat cluster.

We defined as universe/background all the genes expressed in the dataset, meaning the rownames of the Seurat object.

We used three gene sets from MSiGDB:

- the hallmark gene sets  are coherently expressed signatures derived by aggregating many MSigDB gene sets to represent well-defined biological states or processes.
- the C3 : regulatory target gene sets  based on gene target predictions for microRNA seed sequences and predicted transcription factor binding sites.
- the C8 : cell type signature gene sets  curated from cluster markers identified in single-cell sequencing studies of human tissue.

We used enricher function from clusterProfiler to perform enrichment analysis.

```{r fig.width=10, fig.height=10, out.width='100%'}
# define background genes = universe for enrichment
background <-  row_metadata$gene_ids

## define genesets
c8 <- msigdbr(species = "human", category = "C8")

msigdbr_hallmarks = hallmarks %>% dplyr::distinct(gs_name, ensembl_gene) %>% as.data.frame()
msigdbr_c3 = c3 %>% dplyr::distinct(gs_name, ensembl_gene) %>% as.data.frame()
msigdbr_c8 = c8 %>% dplyr::distinct(gs_name, ensembl_gene) %>% as.data.frame()
```

#### Hallmarks MSiGDB gene sets

```{r fig.width=25, fig.height=3, out.width='100%'}
tmp_H <- NULL
for(i in unique(srat$seurat_clusters)){
signature = s.markers$feature[s.markers$group1 == i]
ego_module <- enricher(gene = signature, universe = background, TERM2GENE = msigdbr_hallmarks)
tmp <- NULL

  if(!is.null(ego_module)){
    if(dim(ego_module)[1]>0){
    b_H <- barplot(ego_module, showCategory = 15, font.size = 20, label_format = 40)  
    tmp <- b_H$data
    tmp$set <- "Hallmarks"
    tmp$cluster <- i
    }
  }
tmp_H <- rbind(tmp_H, tmp)
}

ggplot(tmp_H, aes(Count, ID)) +
    geom_bar(stat = "identity", aes(fill=cluster))+

  theme_classic()+ 
    geom_text(
        aes(label = (paste( "Padj = ", round(p.adjust,2)))),
        color = "black",
        size = 3,
        hjust=1,
        position = position_dodge(0.5)
    )+
   theme(text = element_text(size=14))+
  facet_wrap(facets=c("cluster"), ncol = length(unique(tmp_H$cluster)))
```

- C2 and C13 are enriched for EMT signature (Warning count = 3) which makes sense with the hypothesis that they are stroma cluster.

#### C3 MSiGDB regulatory target gene sets

```{r fig.width=25, fig.height=6, out.width='100%'}

tmp_H <- NULL
for(i in unique(srat$seurat_clusters)){
signature = s.markers$feature[s.markers$group1 == i]
ego_module <- enricher(gene = signature, universe = background, TERM2GENE = msigdbr_c3)
tmp <- NULL

  if(!is.null(ego_module)){
    if(dim(ego_module)[1]>0){
    b_H <- barplot(ego_module, showCategory = 15, font.size = 20, label_format = 40)  
    tmp <- b_H$data
    tmp$ID <- factor(tmp$ID, levels = c(tmp$ID[order(tmp$Count)]))
    tmp$set <- "c8"
    tmp$cluster <- i
    }
  }
tmp_H <- rbind(tmp_H, tmp[1:5,])
}

ggplot(tmp_H, aes(Count, ID)) +
    geom_bar(stat = "identity", aes(fill=cluster))+

  theme_classic()+ 
    geom_text(
        aes(label = (paste( "Padj = ", round(p.adjust,2)))),
        color = "black",
        size = 3,
        hjust=1,
        position = position_dodge(0.5)
    )+
   theme(text = element_text(size=14))+
  facet_wrap(facets=c("cluster"), ncol = length(unique(tmp_H$cluster)))
```

- C5, which is though to be more epithelial, has a specific enrichment pattern of PRMD4, F10 and BCL6B targets.
Might be usefull to look more into detail in the lit. if this pattern comes more often accross other samples.

- C13 has a specific enrichement of HMGB2 targets, which is though to be associated with "stemness". 
Would makes sense with the hypothesis tht C13 is blastema.

#### C8 MSiGDB cell type signature gene sets

```{r fig.width=30, fig.height=15, out.width='100%'}

tmp_H <- NULL
for(i in unique(srat$seurat_clusters)){
signature = s.markers$feature[s.markers$group1 == i]
ego_module <- enricher(gene = signature, universe = background, TERM2GENE = msigdbr_c8)
tmp <- NULL

  if(!is.null(ego_module)){
    if(dim(ego_module)[1]>0){
    b_H <- barplot(ego_module, showCategory = 15, font.size = 20, label_format = 40)  
    tmp <- b_H$data
    tmp$ID <- factor(tmp$ID, levels = c(tmp$ID[order(tmp$Count)]))
    tmp$set <- "c8"
    tmp$cluster <- i
    }
  }
tmp_H <- rbind(tmp_H, tmp[1:5,])
}

ggplot(tmp_H, aes(Count, ID)) +
    geom_bar(stat = "identity", aes(fill=cluster))+

  theme_classic()+ 
    geom_text(
        aes(label = (paste( "Padj = ", round(p.adjust,2)))),
        color = "black",
        size = 3,
        hjust=1,
        position = position_dodge(0.5)
    )+
   theme(text = element_text(size=14))+
  facet_wrap(facets=c("cluster"), ncol = length(unique(tmp_H$cluster)))
```

- C0 is enriched in fetal metanephric cells, which makes sense with the hypothesis that C0 is blastema.

- C2, C7 is enriched in adult kidney mesenchymal cells (interstitial, muscle and mesenglial).
As the associated pathways are from "adult" reference, this might suggests that the cells are from normal kidney. 
To be checked with inferCNV profile.

- C5, C6, C8 is enriched in adulte differentiated cell types/pathway of the kidney.
This could suggest that C5 are from normal kidney. 
To be checked with inferCNV profile.

- C12 is enriched in endothelial gene sets, which fits whith the hypothesis that C12 are endothelial cells.

### Annotation from SingleR (no cancer or kidney specific dataset)

Here, we quickly checked annotations that are present in the _processed rds object. 
However, the automated annotation have not been performed using a cancer specific reference or a kidney reference. 
We do not expect a nice labelling of the cells as the overlap of cell types between the reference and the query dataset is poor. 
This support the need to do a proper label transfer from the fetal kidney atlas, which is imho the best reference that can be applied to a Wilms tumor query.

```{r fig.width=15, fig.height=10, out.width='100%'}

d2 <-DimPlot(srat, group.by = "singler_celltype_annotation", reduction = "umap", label = TRUE, repel = TRUE) + NoLegend()

DT::datatable(table(srat$seurat_clusters, srat$singler_celltype_annotation), caption = ("table of SingleR annotation per seurat clusters"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

                               buttons = c( 'csv', 'excel')))

d3 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "singler_celltype_annotation",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

d2|d3
```

### Annotation from cellassign (no cancer or kidney specific dataset)

```{r fig.width=15, fig.height=10, out.width='100%'}

d2 <-DimPlot(srat, group.by = "cellassign_celltype_annotation", reduction = "umap", label = TRUE, repel = TRUE) + NoLegend()

DT::datatable(table(srat$seurat_clusters, srat$cellassign_celltype_annotation)
, caption = ("table of cellassign annotation per seurat clusters"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

                               buttons = c( 'csv', 'excel')))

d3 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "cellassign_celltype_annotation",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

d2|d3
```

Even if the result of SingleR and CellAssign are not specific to a Wilms tumor dataset, it give the first impression that C12 is a cluster of endothelial cells and C11 a cluster of immune cells.

### Azimuth annotation from fetal kidney

Note to the DataLab : where should I save the fetal kidney reference? 
In S3 like results as it is a quite big object or in the folder marker-sets?

For more information related to the reference, please go to https://www.kidneycellatlas.org/
You will find:

- interactive viewer

- h5ad files to download.

Please note that as Wilms tumor have been described to be closer to fetal kidney as mature kidney, we only used the fetal kidney atlas as the reference. 
Also check : https://www.science.org/doi/10.1126/science.aat5031

This part is imho one of the most important step that allow us to have a quick and reliable idea of the composition of the different clusters. 
The predicted compartment are defined into 4 categories:

- endotheliúm
- stroma
- fetal nephron 
- immune

As for SingleR and CellAssign, the annotation of immune cells and endothelial cells is straightforward. 
The stroma compartment should then contain normal and cancer stromal cells.
The fetal nephron compartment contain blastema cancer cells a well as normal and cancer epithelial cells.

Further segregation of cancer versus normal cells will be achieved using a combination of markers/pathways (see above) and inferred CNV (to be done).

```{r fig.height=6, fig.width=6, message=FALSE, warnings=FALSE}

# access to the azimuth kidney reference 
##### to be updated depending on the location of the reference file 
ref_dir = file.path(module_base, "marker-sets/Azimuth_Compatible_Fetal_full/")

reference <- LoadReference(ref_dir)
# Load the query object for mapping

# Preprocess with SCTransform
srat <-  RunAzimuth(srat, ref_dir, assay = 'RNA')
rm(reference)
gc()

DT::datatable(table(srat$seurat_clusters, srat$predicted.compartment), caption = ("table of azimuth compartment annotation per seurat clusters"),               extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',
              buttons = c( 'csv', 'excel')))

DT::datatable(table(srat$seurat_clusters, srat$predicted.celltype), caption = ("table of azimuth annotation per seurat clusters"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',
              buttons = c( 'csv', 'excel')))

```

```{r fig.height=5, fig.width=20, warnings=FALSE}

d1 <- DimPlot(srat, reduction = "umap", group.by = "seurat_clusters", label = TRUE) + NoLegend()
d2 <- DimPlot(srat, reduction = "umap", dims = c(1,2), group.by = "predicted.compartment", label = TRUE, repel = TRUE) + NoLegend()
d3 <- DimPlot(srat, reduction = "umap", dims = c(1,2), group.by = "predicted.celltype", label = TRUE, repel = TRUE) + NoLegend()

d1 | d2 | d3
```

Looking at the predicted compartment, we are satisfied that the immune, endothelium and stroma compartment are specific to a set of clusters. 
It is expected that the fetal__nephron compartment match to both blastema and epithelial cancer and normal kidney. 
It is thus OK to have it labeling two distinct set of clusters.

Looking at the predicted cell types, we observed that the set of cluster [0,1,3,4,9,10,14] is labeled as cap mesenchyme, an early developmental stage of nephron, while the set of clusters [2,7,13] is labeled with more differentiated epithelial units of the kidney. 
This suggest that:

- clusters [0,1,3,4,9,10,14] are more stem, either primitive epithelium or blastem cancer cells. 
- clusters [2,7,13] are more differentiated, either epithelial cancer cells or normal cells.

```{r fig.height=5, fig.width=20, warnings=FALSE}

f1 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "predicted.compartment",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

f2 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "predicted.celltype",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

f1 | f2

```

## Summary and conclusion

Taking together, we can conclude:

- C12 = endothelium (with high confidency)
- C11 = immune cells (with high confidency)

- C2. C7, C13, that form a unit in the umap reduction, are stromal cells, either cancer or normal (with high confidency)
- C[0-1-3-4-9-10-14] that form a unit in the umap reduction, are label as cap mesenchyme (<=> stem) and NCAM+ must be blastema cancer cells.
- C[5-6-8] that form a unit in the umap reduction, are label as epithelial cells (SingleR) and differentiated kidney epithelium subunit (Fetal kidney atlas) must be epithelial (cancer or normal)

The primary aim of this script was to assess the quality of clustering and play with clustering parameters. On this sample, we are happy with the clustering, as we could identify:

- one cluster of immune cells (cell marker PTPRC, SingleR, Azimuth label transfer fro the fetal kidney) 
- one cluster of endothelial cells (cell marker VWF, SingleR, Azimuth label transfer fro the fetal kidney)
- one set of clusters of stromal cells
- one set of cluster of epithelial cells
- one set of cluster of blastema cancer cells.

Additionally, we showed the feasibility to run label transfer using runAzimuth (part of step 3 of the proposed analysis:  https://github.com/maud-p/OpenScPCA-analysis/issues/1).

We also showed the easy and robust identification of normal cells (endothelial and immune cells) that would serve as reference for the CNV inference (inferCNV, step 4 of the analysis).

## Next step

[ ] We will adapt this template to be render on all the 40 Wilms tumor samples of the dataset.

[ ] We will save for each sample the rds file

[ ] We will run inferCNV for each sample to decide the malignant/normal status of some stroma and epithelial cluster, and confirm the blastema annotation.

The next step will provide us a better understanding of the entire cohort. We will then have to set up a strategy to annotate each sample. Open questions are:

[ ] should we annotate single cell                      
or   
[] consider applying similar annotations to all cells in a cluster?

[ ] manual annotation of each cluster / each patient    
or   
[] automated annotation using some threshold?

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```