### 2024-08-07 analyses/cell-type-wilms-tumor-06/01_fetal_reference_kidney.Rmd
add 01-clustering analysis
few lines to get an idea how I build the azimuth compatible fetal kidney reference
Must be improved + add a way to build proper docker container for this.
Or I can maybe directly load somewhere the azimuth compatible reference?

---
title: "Build azimuth compatible fetal kidney reference from the kidney cell atlas"
author: "Maud PLASCHKA"
date: '2024-08-07'
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

The aim is to build an azimuth compatible reference for fetal kidney from the kidney cell atlas.
Download the Fetal_full_v3.h5ad from : https://www.kidneycellatlas.org/fetal-kidney-global
And save it in path_to_resource folder

```{r }
# run with docker image maud_R_Python
    library("reticulate")
    library("scater")
    library("Seurat")
    sc <- import("scanpy")
```
```{r }

file = paste0(path_to_resource, "/Fetal_full_v3.h5ad")

adata <- sc$read_h5ad(file)

counts <- t(adata$layers["counts"])
colnames(counts) <- adata$obs_names$to_list()
rownames(counts) <- adata$var_names$to_list()
counts <- Matrix::Matrix(as.matrix(counts), sparse = T)
seurat <- CreateSeuratObject(counts)
seurat <- AddMetaData(seurat,  adata$obs)
embedding <- adata$obsm["X_umap"]
rownames(embedding) <- adata$obs_names$to_list()
colnames(embedding) <- c("umap_1", "umap_2")
seurat[["umap"]] <- CreateDimReducObject(embedding, key = "umap_")
saveRDS(seurat, file = paste0(path_to_resources,"/Fetal_full.rds"))
```

# Create a azimuth compatible reference

```{r echo=TRUE, fig.height=10, fig.width=14, message=FALSE, warning=FALSE, out.width='100%'}
# run with my standard docker image maud
library(Seurat)
library(sctransform)
library(Azimuth)

seurat <- readRDS( file = paste0(path_to_resources,"/Fetal_full.rds"))
seurat <- subset(seurat, subset = (nCount_RNA > 0) )

DimPlot(seurat, group.by = "compartment")
DimPlot(seurat, group.by = "celltype")

s <- SCTransform(seurat, verbose = FALSE, method = "glmGamPoi", conserve.memory = TRUE)
s <- RunPCA(s, npcs = 50, verbose = FALSE)

s <- RunUMAP(s, dims = 1:50, verbose = FALSE, return.model = TRUE)
DimPlot(s, reduction = "umap", group.by = "compartment") 
DimPlot(s, group.by = "celltype")

Fetal_kidney <- AzimuthReference(
  s,
  refUMAP = "umap",
  refDR = "pca",
  refAssay = "SCT",
  dims = 1:50,
  k.param = 31,
  plotref = "umap",
  plot.metadata = NULL,
  ori.index = NULL,
  colormap = NULL,
  assays = NULL,
  metadata = c("compartment", "celltype"),
  reference.version = "0.0.0",
  verbose = FALSE
)

ref_dir = paste0(path_to_resources)

# save reference in rds file
SaveAnnoyIndex(object = Fetal_kidney[["refdr.annoy.neighbors"]], file = file.path(ref_dir, "idx.annoy"))
saveRDS(object = Fetal_kidney, file = file.path(ref_dir, "ref.Rds"))

```