### 2024-08-09 analyses/cell-type-wilms-tumor-06/00_fetal_reference_kidney.Rmd

---
title: "Build azimuth compatible fetal kidney reference from the kidney cell atlas"
author: "Maud PLASCHKA"
date: '2024-08-07'
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warnings=FALSE)
```

Introduction

The aim is to build an azimuth compatible reference for fetal kidney from the kidney cell atlas.
The rds data can be download using the download link https://datasets.cellxgene.cziscience.com/40ebb8e4-1a25-4a33-b8ff-02d1156e4e9b.rds

## Packages

Load required packages in the following chunk, if needed.
Do not install packages here; only load them with the `library()` function.

```{r packages, message=FALSE, warning=FALSE}
library("Seurat")
library(sctransform)
library(Azimuth)
library(SCpubr)
library(tidyverse)
library(patchwork)
library(SeuratWrappers)
```

## Base directories

```{r base paths, eval=TRUE, include=TRUE}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "2024-07-08", "SCPCP000006")

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-wilms-tumor-06")
```

## Input files

rds data can be downloaded using the URLhttps://datasets.cellxgene.cziscience.com/40ebb8e4-1a25-4a33-b8ff-02d1156e4e9b.rds

Please note that this download link permanently references the current version of the dataset (08/2024). 
If this dataset is updated, a new download link will be created that permanently references the next version of this dataset.

We save it in the marker-sets folder of the module.
Note to the DataLab: should we save it somewhere else? 
I suggest this rds file could be placed here transiently and removed once the reference is build?

```{r path_to_data}
path_to_data <- file.path(module_base, "marker-sets/fetal_full.rds")

url = "https://datasets.cellxgene.cziscience.com/40ebb8e4-1a25-4a33-b8ff-02d1156e4e9b.rds"

download.file(url, path_to_data)
```
## Output file

The azimuth compatible fetal kidney reference will be saved in the marker-sets folder from the module.

```{r path_to_output}
path_to_output <- file.path(module_base, "marker-sets")
```

## Some config set-up / threshold

We store in a config list names "cfg" parameters used all along the analysis to filter for p-value, log fold change, percentage of expression, etc.

```{r cfg}
cfg = list()
cfg$padj_thershold = 0.05
cfg$lfc_threshold = 1
cfg$rate1_threshold = 0.5

set.seed(12345)
```

# Create a azimuth compatible reference

```{r pre_process, echo=TRUE, fig.height=7, fig.width=12, message=FALSE, warning=FALSE, out.width='100%'}
seurat <- readRDS(path_to_data)

d2 <- do_DimPlot(seurat, reduction = "umap", dims = c(1,2), group.by = "compartment", label = TRUE, repel = TRUE) + NoLegend() + ggtitle("umap reduction before SCTransform")
d3 <- do_DimPlot(seurat, reduction = "umap", dims = c(1,2), group.by = "cell_type", label = TRUE, repel = TRUE) + NoLegend()

d2 | d3

options(future.globals.maxSize= 891289600000)
s <- SCTransform(seurat, verbose = FALSE, method = "glmGamPoi", conserve.memory = TRUE)
s <- RunPCA(s, npcs = 50, verbose = FALSE)
s <- RunUMAP(s, dims = 1:50, verbose = FALSE, return.model = TRUE)
```

```{r create_ref, echo=TRUE, fig.height=7, fig.width=12, message=FALSE, warning=FALSE, out.width='100%'}
options(future.globals.maxSize= 891289600000)
Fetal_kidney <- AzimuthReference(
  s,
  refUMAP = "umap",
  refDR = "pca",
  refAssay = "SCT",
  dims = 1:50,
  k.param = 31,
  plotref = "umap",
  plot.metadata = NULL,
  ori.index = NULL,
  colormap = NULL,
  assays = NULL,
  metadata = c("compartment", "cell_type"),
  reference.version = "0.0.0",
  verbose = FALSE
)

d2 <- do_DimPlot(Fetal_kidney, reduction = "refUMAP", dims = c(1,2), group.by = "compartment", label = TRUE, repel = TRUE) + NoLegend() + ggtitle("umap reduction after SCTransform and Seurat workflow")
d3 <- do_DimPlot(Fetal_kidney, reduction = "refUMAP", dims = c(1,2), group.by = "cell_type", label = TRUE, repel = TRUE) + NoLegend()

d2 | d3

# save reference in rds file
SaveAnnoyIndex(object = Fetal_kidney[["refdr.annoy.neighbors"]], file = file.path(path_to_output, "idx.annoy"))
saveRDS(object = Fetal_kidney, file = file.path(path_to_output, "ref.Rds"))

# finally, we can remove the fetal_full.rds file from the marker-sets directory
file.remove(path_to_data)
```
## Characterization of compartment and cell types in the reference

Here, we use an unbiased approach to find transcripts that characterized the different compartments and cell types.

This is just to get markers genes of the different population, in case some could be of interest for the Wilms tumor annotations.

We run DElegate::FindAllMarkers2 to find markers of the different clusters and manually check if they do make sense. 
DElegate::FindAllMarkers2 is an improved version of Seurat::FindAllMarkers based on pseudobulk differential expression method. 
Please check the preprint from Chistoph Hafemeister: https://www.biorxiv.org/content/10.1101/2023.03.28.534443v1
and tool described here: https://github.com/cancerbits/DElegate

### Find marker genes for each of the compartment

```{r markers_compatment, fig.width=8, fig.height=7, out.width='100%'}
de_results   <- DElegate::FindAllMarkers2(s, group_column = "compartment")

#filter the most relevant markers
s.markers <- de_results[de_results$padj < cfg$padj_thershold & de_results$log_fc > cfg$lfc_threshold & de_results$rate1 > cfg$rate1_threshold,]

DT::datatable(s.markers, caption = ("marker genes"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

                               buttons = c( 'csv', 'excel')))

# Select top 5 genes for heatmap plotting
s.markers <- na.omit(s.markers)
s.markers %>%
    group_by(group1) %>%
    top_n(n =  5, wt = log_fc) -> top5

# subset for plotting
Idents(s) <- s$compartment
cells <- WhichCells(s, downsample = 100)
ss <- subset(s, cells = cells)
ss <- ScaleData(ss, features = top5$feature)

p1 <- SCpubr::do_DimPlot(s, reduction="umap", group.by = "compartment", label = TRUE, repel = TRUE) + ggtitle("compartment")
p2 <- DoHeatmap(ss, features = top5$feature,  cells = cells, group.by = "compartment") + NoLegend() + 
  scale_fill_gradientn(colors =  c("#01665e","#35978f",'darkslategray3', "#f7f7f7", "#fee391","#fec44f","#F9AD03")) 
p3 <- ggplot(s@meta.data, aes(compartment, fill = compartment)) + geom_bar() + NoLegend()

common_title <- sprintf("Unsupervised clustering %s, %d cells", s@meta.data$orig.ident[1], ncol(s))
show((((p1 / p3) + plot_layout(heights = c(3,2)) | p2) ) + plot_layout(widths = c(1, 2)) + plot_layout(heights = c(3,1)) + plot_annotation(title = common_title))

```

### Find marker genes for each of the cell types

```{r markers_cell, fig.width=15, fig.height=17, out.width='100%'}
de_results   <- DElegate::FindAllMarkers2(s, group_column = "cell_type")

#filter the most relevant markers
s.markers <- de_results[de_results$padj < cfg$padj_thershold & de_results$log_fc > cfg$lfc_threshold & de_results$rate1 > cfg$rate1_threshold,]

DT::datatable(s.markers, caption = ("marker genes"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

                               buttons = c( 'csv', 'excel')))

# Select top 5 genes for heatmap plotting
s.markers <- na.omit(s.markers)
s.markers %>%
    group_by(group1) %>%
    top_n(n =  5, wt = log_fc) -> top5

# subset for plotting
Idents(s) <- s$cell_type
cells <- WhichCells(s, downsample = 100)
ss <- subset(s, cells = cells)
ss <- ScaleData(ss, features = top5$feature)

p1 <- SCpubr::do_DimPlot(s, reduction="umap", group.by = "cell_type", label = TRUE, repel = TRUE) + ggtitle("cell_type") + NoLegend()
p2 <- DoHeatmap(ss, features = top5$feature,  cells = cells, group.by = "cell_type") + NoLegend() + 
  scale_fill_gradientn(colors =  c("#01665e","#35978f",'darkslategray3', "#f7f7f7", "#fee391","#fec44f","#F9AD03")) 
p3 <- ggplot(s@meta.data, aes(cell_type, fill = cell_type)) + geom_bar() + NoLegend() + scale_x_discrete(guide = guide_axis(angle = 90))

common_title <- sprintf("Unsupervised clustering %s, %d cells", s@meta.data$orig.ident[1], ncol(s))
show((((p1 / p3) + plot_layout(heights = c(3,2)) | p2) ) + plot_layout(widths = c(1, 1)) + plot_layout(heights = c(3,1)) + plot_annotation(title = common_title))

```