### 2024-08-13 analyses/cell-type-wilms-tumor-06/notebook_template/01_fetal_reference_Stewart_Azimuthv5.Rmd

---
title: "Reference label transfer exploration for `r params$sample_id`"
author: "Maud PLASCHKA"
date: '2024-08-13'
params:
  current: '2024-07-08'
  scpca_project_id: "SCPCP000006"
  sample_id: "SCPCS000176"
  ref_url: "https://datasets.cellxgene.cziscience.com/40ebb8e4-1a25-4a33-b8ff-02d1156e4e9b.rds"
  seed: 12345
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warnings=FALSE
                      )
```

## Introduction

The aim is to find an appropriate reference to perform label transfer as a step of the annotation of the Wilms tumor dataset for the sample `param$sample_id` =

```{r}
params$sample_id
```

The use of the right reference is crucial. 
It is recommended that the cell types in the reference is representative to the cell types to be annotated in the query.

Wilms tumors can contain up to three histologies that resemble fetal kidney: blastema, stroma, and epithelia [1-2].
Because of their histological similarity to fetal kidneys, Wilms tumors are thought to arise from developmental derangements in embryonic renal progenitors.

We thus decided to test and compare fetal (kidney) references that could be use in the analysis module.
You can find below the three workflow/reference that we plan to test.

### Human fetal kidney atlas Stewart et al.

We first wanted to try the human fetal kidney atlas to transfer label into the Wilms tumor samples using azimuth. 
You can find more about the human kidney atlas here: https://www.kidneycellatlas.org/ [3]

Using this reference, we test two workflows for label transfer:

- `Azimuth v5`,
- `Seurat v4` as described here  https://satijalab.org/seurat/articles/integration_mapping.html#cell-type-classification-using-an-integrated-reference

### Azimuth Human fetal atlas Cao et al.

Azimuth also provide a human fetal atlas as a reference [4].

The data can be found on Zenodo: 
https://zenodo.org/records/4738021#.YJIW4C2ZNQI

The reference contain cells from 15 organs including kidney from fetal samples. 
Here we will use `Azimuth` to transfer labels from the reference.

##########################################################################################################
In this notebook, we will test the human fetal kidney atlas from Stewart et al. using the Azimuth workflow. 
##########################################################################################################

### Packages

Load required packages in the following chunk, if needed.
Do not install packages here; only load them with the `library()` function.

```{r packages, message=FALSE, warning=FALSE}
library("Seurat")
library(sctransform)
library(Azimuth)
library(SCpubr)
library(tidyverse)
library(patchwork)
library(SingleCellExperiment)

set.seed(params$seed)
```

### Base directories

```{r base paths, eval=TRUE}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", params$current, params$scpca_project_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-wilms-tumor-06")
```

### Input files

#### Reference

rds data can be downloaded using the URL defined in `param$ref_url`.

Please note that this download link permanently references the current version of the dataset (08/2024). 
If this dataset is updated, a new download link will be created that permanently references the next version of this dataset.

We place it in the `scratch` folder of the module because it is a temporary file.

```{r path_to_reference}
path_to_reference <- file.path(module_base, "scratch", "fetal_full.rds")
```

#### Query

In this notebook, we test the label transfer from the fetal kidney atlas (reference) to the Wilms tumor sample defined in `params$sample_id` from the Wilms tumor dataset `param$scpca_project_id`

```{r path_to_query}
filelist <- list.files(data_dir, full.names = TRUE)

# select the 40 Wilms tumor single nucleus folders only
filedir <- filelist[grepl(params$sample_id, filelist)]
```

### Output file

This notebook has two main outputs

#### Azimuth compatible reference

The reference data are downloaded from the `params$ref_url` and transiently saved in the `scratch`folder.
The azimuth compatible fetal kidney reference will be saved in the `results` folder from the module, `reference` subfolder.

Please note that the creation of the Azimuth reference cannot be done within RMarkdown. 
We thus set `include = FALSE` to the corresponding chunk.
Please, before running this code either :

- download the azimuth compatible reference from S3 bucket or
- run manualy chunk by chunk the report to create and save the reference once.

```{r path_to_output}
path_to_output <- file.path(module_base, "results", "references")
```

#### The html notebook saved afetr rendering this notebook (knit).
Will be saved in the `notebook/00-reference` directory.

## Using the Human Kidney Atlas - Fetal as a reference

### Create a azimuth compatible reference

Note: not shown here because the `AzimuthReference` function do not allow knitr.
The reference should be build "manually" prior to knitr the report.

```{r create_ref, eval=FALSE, fig.height=7, fig.width=12, message=FALSE, warning=FALSE, include=FALSE, out.width='100%'}
download.file(params$ref_url, path_to_reference)
seurat <- readRDS(path_to_reference)

options(future.globals.maxSize= 891289600000)
s <- SCTransform(seurat, verbose = FALSE, method = "glmGamPoi", conserve.memory = TRUE)
s <- RunPCA(s, npcs = 50, verbose = FALSE)
s <- RunUMAP(s, dims = 1:50, verbose = FALSE, return.model = TRUE)

Fetal_kidney <- AzimuthReference(
  s,
  refUMAP = "umap",
  refDR = "pca",
  refAssay = "SCT",
  dims = 1:50,
  k.param = 31,
  plotref = "umap",
  plot.metadata = NULL,
  ori.index = NULL,
  colormap = NULL,
  assays = NULL,
  metadata = c("compartment", "cell_type"),
  reference.version = "0.0.0",
  verbose = FALSE
)

d2 <- do_DimPlot(Fetal_kidney, reduction = "refUMAP", dims = c(1,2), group.by = "compartment", label = TRUE, repel = TRUE) + NoLegend() + ggtitle("umap reduction after SCTransform and Seurat workflow")
d3 <- do_DimPlot(Fetal_kidney, reduction = "refUMAP", dims = c(1,2), group.by = "cell_type", label = TRUE, repel = TRUE) + NoLegend()

d2 | d3

if (file.exists(path_to_output)){
   SaveAnnoyIndex(object = Fetal_kidney[["refdr.annoy.neighbors"]], file = file.path(path_to_output, "idx.annoy"))
    saveRDS(object = Fetal_kidney, file = file.path(path_to_output, "ref.Rds"))
} else {
    SaveAnnoyIndex(object = Fetal_kidney[["refdr.annoy.neighbors"]], file = file.path(path_to_output, "idx.annoy"))
    saveRDS(object = Fetal_kidney, file = file.path(path_to_output, "ref.Rds"))

}

# finally, we can remove the fetal_full.rds file from the marker-sets directory
file.remove(path_to_reference)
```

### Load and pre-process the query data

```{r load, message=FALSE, warning=FALSE}
file <- list.files(filedir)
file <- file[grepl("_processed.rds", file)]

# open the processed rds object
sce <- readRDS(paste0(filedir, "/", file))
```

```{r seurat_wf, fig.height=4, fig.width=8, warning=FALSE, out.width='100%'}
# convert to seurat
srat <- CreateSeuratObject(counts = counts(sce),
                                    assay = "RNA",
                                    project = params$sample_id
                           )

# convert colData and rowData to data.frame for use in the Seurat object
cell_metadata <- as.data.frame(colData(sce))
row_metadata <- as.data.frame(rowData(sce))

# add cell metadata (colData) from SingleCellExperiment to Seurat
srat@meta.data <- cell_metadata

# add row metadata (rowData) from SingleCellExperiment to Seurat
srat[["RNA"]]@meta.data <- row_metadata

# add metadata from SingleCellExperiment to Seurat
srat@misc <- metadata(sce)

# Normalization
options(future.globals.maxSize= 8912896000000)
srat <- SCTransform(srat, verbose = F, conserve.memory = TRUE)

# dimensionality reduction
srat <- RunPCA(srat, verbose = F)
srat <- RunUMAP(srat, dims = 1:50, verbose = F)

# clustering
srat    <- FindNeighbors(srat, dims = 1:50, verbose = F)
srat    <- FindClusters(srat, verbose = T)

```

### Azimuth annotation from fetal kidney

```{r fig.height=6, fig.width=6, message=FALSE, warnings=FALSE}

data <-  RunAzimuth(srat, path_to_output, assay = 'RNA')

```

```{r fig.height=8, fig.width=8, warnings=FALSE}

d1 <- DimPlot(data, reduction = "umap", dims = c(1,2), group.by = "predicted.compartment", label = TRUE, repel = TRUE) + NoLegend()
d2 <- DimPlot(data, reduction = "umap", dims = c(1,2), group.by = "predicted.cell_type", label = TRUE, repel = TRUE) + NoLegend()

f1 <- SCpubr::do_BarPlot(sample = data,
                         group.by = "predicted.compartment",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

f2 <- SCpubr::do_BarPlot(sample = data,
                         group.by = "predicted.cell_type",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

(d1/f1) | (d2/f2) 
```

Note:

For some reason, the "cap-mesenchyme cells" has been renamed in cellxgene in "mesenchymal cells".

The cap mesenchyme is a cap of condensed metanephric mesenchyme, comprised of cells which epithelialize and sequentially form the pretubular aggregate (PA), renal vesicle (RV), C-, and S-shaped bodies, and finally the mature nephron. 
The CM contains nephron progenitor cells.

This can be confusing and we just need to pay attention that :

- the fetal nephron / mesenchymal cells are cap-mesenchyme cells. 
In our case, cap-mesenchyme contains blastema and primitive epitheliul cancer cells.

- the stroma / mesenchymal stem are likelly mesenchymal cancer or normal cells.

## References

- [1] https://www.ncbi.nlm.nih.gov/books/NBK373356/

- [2] https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9915828/

- [3] https://www.science.org/doi/10.1126/science.aat5031

- [4] https://www.science.org/doi/10.1126/science.aba7721

## Session info

```{r }
sessionInfo()
```