### 2024-08-16 analyses/cell-type-wilms-tumor-06/notebook_template/01_seurat-processing.Rmd

---
title: "Seurat processing for `r params$sample_id`"
author: "Maud PLASCHKA"
date: '2024-08-07'
params:
  scpca_project_id: "SCPCP000006"
  sample_id: "SCPCS000176"
  seed: 12345
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warnings=FALSE)
```

## Introduction

The aim is to pre-process the sample `param$sample_id` =

```{r}
params$sample_id
```

Here we perform the `Seurat` workflow [normalization --> reduction --> clustering] and explore the clustering results.

This correspond to the step 2 of the proposed analysis: clustering of cells across a set of parameters for few samples

- https://github.com/AlexsLemonade/OpenScPCA-analysis/discussions/635
- https://github.com/AlexsLemonade/OpenScPCA-analysis/issues/671

This notebook will be updated with PR #704 to explore the clustering.

### Packages

Load required packages in the following chunk, if needed.
Do not install packages here; only load them with the `library()` function.

```{r packages, message=FALSE, warning=FALSE}
library(SingleCellExperiment)
library("Seurat")

set.seed(params$seed)
```

### Base directories

```{r base paths, eval=TRUE, include=TRUE}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current", params$scpca_project_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-wilms-tumor-06")
```

### Input files

In this notebook, we pre-processed the Wilms tumor sample defined in `params$sample_id` from the Wilms tumor dataset `param$scpca_project_id`

```{r path_to_query}
filelist <- list.files(data_dir, full.names = TRUE)

# select the 40 Wilms tumor single nucleus folders only
filedir <- filelist[grepl(params$sample_id, filelist)]

```

### Output file

Reports will be saved in the `notebook/01-Seurat` directory. 
The pre-processed `Seurat` object per samples are saved in the `result` folder.
They will be the `input` for the next `notebook-template`.
```{r path_to_output}
output_dir <- file.path(module_base, "results", params$sample_id)
```

### Functions

#### Convert `SingleCellExperiment` object to `Seurat`

```{r sce_to_seurat}
convert_sce_to_seurat <- function(sce, 
                                  sample_id) {
  # Convert a SingleCellExperiment from the ScPCA Portal to a Seurat object
  # 
  # Arguments:
  #	  sce: A SingleCellExperiment from the ScPCA Portal
  #   sample_id: The sample identifier supplied as the project to 
  #              CreateSeuratObject() (character)
  #
  # Returns: A Seurat Object
  # Check class of input
  stopifnot(class(sce) == "SingleCellExperiment")

  srat <- CreateSeuratObject(counts = counts(sce),
                             assay = "RNA",
                             project = sample_id
  )

  # convert colData and rowData to data.frame for use in the Seurat object
  cell_metadata <- as.data.frame(colData(sce))
  row_metadata <- as.data.frame(rowData(sce))

  # add cell metadata (colData) from SingleCellExperiment to Seurat
  srat@meta.data <- cell_metadata

  # add row metadata (rowData) from SingleCellExperiment to Seurat
  srat[["RNA"]]@meta.data <- row_metadata

  # add metadata from SingleCellExperiment to Seurat
  srat@misc <- metadata(sce)

  # return Seurat object
  return(srat)

}

```

#### Perform `Seurat` workflow

`seurat_wf` run the `Seurat` workflow from a `Seurat`object `srat`:

```{r seurat_wf}
perform_norm_dimred_cluster <- function(srat,
                                        transform_method = "glmGamPoi",
                                        conserve_memory = TRUE,
                                        umap_dims = 1:50, 
                                        neighbors_dims = 1:50, 
                                        verbose = FALSE) {

  # Perform normalization, dimensionality reduction, and clustering on a 
  # Seurat object
  #
  # Arguments:
  #   srat: A Seurat object
  #   transform_method: The method passed to SCTransform(); glmGamPoi by default
  #   conserve_memory: Logical specifying whether or not SCTransform() should
  #                    conserve memory
  #   umap_dims: A vector of integers passed to RunUMAP() to specify the number 
  #              of dimensions; 1:50 by default
  #   neighbors_dims: A vector of integers passed to FindNeighbors() to specify
  #                   the dimensions of reduction to use as input
  #   verbose: Logical specifying whether or not functions should be verbose
  #
  # Returns: 
  #   A Seurat object containing all the contents of the normalization, PCA, 
  #   UMAP, neighbors, and clustering steps
  #

  # Check class
  stopifnot(class(srat) == "Seurat")

  # Normalization
  options(future.globals.maxSize= 8912896000000)
  srat <- SCTransform(srat, 
                      method = transform_method,
                      verbose = verbose, 
                      conserve.memory = conserve_memory)

  # dimensionality reduction
  srat <- RunPCA(srat, verbose = verbose)
  srat <- RunUMAP(srat, dims = umap_dims, verbose = verbose)

  # clustering
  srat    <- FindNeighbors(srat, dims = neighbors_dims, verbose = verbose)
  srat    <- FindClusters(srat, verbose = verbose)

  # Return
  return(srat)

}

```

## Analysis

### Load and pre-process the query data

```{r load, message=FALSE, warning=FALSE}
file <- list.files(filedir)
file <- file[grepl("_processed.rds", file)]
# open the processed rds object
sce <- readRDS(paste0(filedir, "/", file))
```

```{r run_seurat_wf, fig.height=4, fig.width=8, warning=FALSE, out.width='100%'}

srat <- convert_sce_to_seurat(sce, 
                                  sample_id = params$sample_id)

srat <- perform_norm_dimred_cluster(srat,
                                        transform_method = "glmGamPoi",
                                        conserve_memory = TRUE,
                                        umap_dims = 1:50, 
                                        neighbors_dims = 1:50, 
                                        verbose = FALSE)
```

## Save the `Seurat`object

```{r save}
saveRDS(object = srat, file = file.path(output_dir, paste0("01-Seurat_", params$sample_id,".Rds")))
```

## Session info

```{r }
sessionInfo()
```