### 2024-08-16 analyses/cell-type-wilms-tumor-06/notebook_template/02b_label-transfer_fetal_kidney_reference_Stewart.Rmd

---
title: "Reference label transfer of the fetal kidney reference (Stewart et al.) for `r params$sample_id`"
author: "Maud PLASCHKA"
date: '2024-08-13'
params:
  scpca_project_id: "SCPCP000006"
  sample_id: "SCPCS000176"
  seed: 12345
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warnings=FALSE
                      )
```

## Introduction

The aim is to perform label transfer as a step of the annotation of the Wilms tumor dataset for the sample `param$sample_id` =

```{r}
params$sample_id
```

The use of the right reference is crucial. 
It is recommended that the cell types in the reference is representative to the cell types to be annotated in the query.

Wilms tumors can contain up to three histologies that resemble fetal kidney: blastema, stroma, and epithelia [1-2].
Because of their histological similarity to fetal kidneys, Wilms tumors are thought to arise from developmental derangements in embryonic renal progenitors.

We thus decided to test and compare fetal (kidney) references that could be use in the analysis module.
You can find below the two references that we plan to use

### Human fetal kidney atlas Stewart et al.

Stewart et al. created and a human fetal kidney atlas.
We will use `Azimuth` to transfer label into the Wilms tumor samples. 
You can find more about the human kidney atlas here: https://www.kidneycellatlas.org/ [3]

### Azimuth Human fetal atlas Cao et al.

Azimuth also provide a human fetal atlas as a reference [4].

The data can be found on Zenodo: 
https://zenodo.org/records/4738021#.YJIW4C2ZNQI

The reference contain cells from 15 organs including kidney from fetal samples. 
Here we will use `Azimuth` to transfer labels from the reference.

In this notebook, we perform label transfer from the human fetal kidney atlas from Stewart et al.

### Packages

Load required packages in the following chunk, if needed.
Do not install packages here; only load them with the `library()` function.

```{r packages, message=FALSE, warning=FALSE}
library("Seurat")
library(sctransform)
library(Azimuth)
library(SCpubr)
library(tidyverse)
library(patchwork)
library(SingleCellExperiment)

set.seed(params$seed)
```

### Base directories

```{r base paths, eval=TRUE}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current", params$scpca_project_id)

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-wilms-tumor-06")
```

### Input files

#### Reference

The reference has been download and pre-process in the `R Script` `download-and-create-fetal-kidney-ref.R`

```{r path_to_reference}
path_to_reference <- file.path(module_base, "results", "references")
```

#### Query

In this notebook, we test the label transfer from the fetal kidney atlas (reference) to the Wilms tumor sample defined in `params$sample_id` from the Wilms tumor dataset `param$scpca_project_id` 
We work with the pre-processed `Seurat`object and labeled with the fetal kidney reference saved in the `results` directory.

```{r path_to_query}
data_dir <- file.path(module_base, "results", params$sample_id)
```

### Output file

Reports will be saved in the `notebook/02-reference` directory. 
The pre-processed and annotated `Seurat` object per samples are saved in the `result` folder.

```{r path_to_output}
output_dir <- file.path(module_base, "results", params$sample_id)
```

## Analysis

### Load the pre-processed `Seurat` object
```{r load, message=FALSE, warning=FALSE}
# open the processed rds object
srat <- readRDS(file.path(data_dir, paste0("00a-fetal_full_label-transfer_",params$sample_id,".Rds")))
```

### Azimuth annotation from fetal kidney

```{r run_azimuth, message=FALSE, warnings=FALSE}

data <-  RunAzimuth(srat, path_to_reference, assay = 'RNA')
# We transfer the annotation to the pre-processed `Seurat`object as we don't want to keep changes done on the query by `RunAzimuth`
metadata_to_trasfer <- data@meta.data[, c("predicted.compartment.score", "predicted.compartment", "predicted.cell_type.score", "predicted.cell_type")]
srat <- AddMetaData(srat, metadata_to_trasfer, col.name = c("fetal_kidney_predicted.compartment.score", "fetal_kidney_predicted.compartment", "fetal_kidney_predicted.cell_type.score", "fetal_kidney_predicted.cell_type"))
```

```{r plot_azimuth, fig.height=8, fig.width=8, warnings=FALSE}

d1 <- DimPlot(srat, reduction = "umap", dims = c(1,2), group.by = "fetal_kidney_predicted.compartment", label = TRUE, repel = TRUE) + NoLegend()
d2 <- DimPlot(srat, reduction = "umap", dims = c(1,2), group.by = "fetal_kidney_predicted.cell_type", label = TRUE, repel = TRUE) + NoLegend()

f1 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "fetal_kidney_predicted.compartment",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(params$sample_id)

f2 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "fetal_kidney_predicted.cell_type",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(params$sample_id)

(d1/f1) | (d2/f2) 
```

Note:

For some reason, the "cap-mesenchyme cells" has been renamed in cellxgene in "mesenchymal cells".

The cap mesenchyme is a cap of condensed metanephric mesenchyme, comprised of cells which epithelialize and sequentially form the pretubular aggregate (PA), renal vesicle (RV), C-, and S-shaped bodies, and finally the mature nephron. 
The CM contains nephron progenitor cells.

This can be confusing and we just need to pay attention that :

- the fetal nephron / mesenchymal cells are cap-mesenchyme cells. 
In our case, cap-mesenchyme contains blastema and primitive epitheliul cancer cells.

- the stroma / mesenchymal stem are likelly mesenchymal cancer or normal cells.

## Save the `Seurat`object

```{r save}
saveRDS(object = srat, file = file.path(output_dir, paste0("00b-fetal_kidey_label-transfer_",params$sample_id,".Rds")))
```

## Session info

```{r }
sessionInfo()
```

## References

- [1] https://www.ncbi.nlm.nih.gov/books/NBK373356/

- [2] https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9915828/

- [3] https://www.science.org/doi/10.1126/science.aat5031

- [4] https://www.science.org/doi/10.1126/science.aba7721