### 2024-10-15 analyses/cell-type-wilms-tumor-14/exploratory_analysis/03_copykat_SCPCL000850.Rmd
add copycat scripts

---
title: "Exploratory analysis using Copykat"
author: "Jingxuan Chen"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

This a template for an analysis notebook using R Markdown.

It is a good idea to start with a brief introduction to the analysis, including the purpose of the analysis, the data used, and the methods applied.
Replace this text with your own introduction, and be sure to update the `title:` and `author:` fields in the YAML header at the top of the document.

Don't forget to rename this file as well!

## Setup

### Packages

```{r packages}
library(dplyr)
library(Seurat)
library(copykat)
library(ggplot2)
```

### Paths

Setting paths to directories and files for input and output at the start of a notebook can be helpful for organization and reproducibility.
You may not know all of the paths you need when you start a notebook, but once you do, keeping them all in one place is helpful, both for yourself in the future and for other users of the notebook.

#### Base directories

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
path_repo <- rprojroot::find_root(rprojroot::is_git_root)

# The path to this module
path_anal <- file.path(path_repo,"analyses","cell-type-wilms-tumor-14") 
```

#### Input and output files

Only run for one sample

```{r paths}
library <- "SCPCL000850"
# Input files

# Output files
```

## Analysis content

Organize the remainder of your content into sections and subsections as appropriate for your analysis.

Add a new code chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I* on Windows or Linux, *Cmd+Option+I* on a Mac.
Be sure to comment your code and explain your steps as you go.

When you save the notebook, an HTML file containing the code and output will be saved alongside it.
To preview the HTML file, click the *Preview* button or press *Ctrl+Shift+K* on Windows or Linux, *Cmd+Shift+K* on a Mac.

The preview shows you a rendered HTML copy of the contents of the editor.
Consequently, *Preview* does not run any R code chunks.
Instead, the output of the chunk when it was last run in the editor is displayed.

It is usually a good idea to rerun the entire notebook before sharing any output, to ensure that the notebook and output files are fully up to date.
You can do this by clicking on the *Run* menu in the toolbar and clicking on *Restart R and Run All Chunks*.

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```