### 2024-11-15 analyses/cell-type-wilms-tumor-06/notebook_template/00b_characterize_fetal_kidney_reference_Stewart.Rmd
fix typos in 00b and 02b, and re-render 00b

---
title: "Characterize the fetal kidney reference from the kidney cell atlas"
author: "Maud PLASCHKA"
date: '2024-08-07'
params:
  padj_threshold: 0.05
  lfc_threshold: 1
  rate1_threshold: 0.5
  seed: 12345
  fetal_kidney_path: "../scratch/fetal_kidney.rds"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
    self_contained: yes
    mode: selfcontained
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warnings = FALSE
)
```

# Introduction

The aim is to characterize the human fetal kidney from the kidney cell atlas.
You can find more about the human kidney atlas here: https://www.kidneycellatlas.org/ [1]
The rds data can be download using the download link: <https://datasets.cellxgene.cziscience.com/40ebb8e4-1a25-4a33-b8ff-02d1156e4e9b.rds>.
The reference was downloaded and created in the script `scripts/prepare-fetal-references.R`.

## Packages

```{r packages, message=FALSE, warning=FALSE}
library(Seurat)
library(SCpubr)
library(tidyverse)
library(patchwork)

set.seed(params$seed)
options(future.globals.maxSize = 891289600000)
```

## Base directories

```{r base paths, eval=TRUE, include=TRUE}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-wilms-tumor-06")
```

## Input files

The input file `r params$fetal_kidney_path` is the output of the script `prepare-fetal-references.R`.

## Output file

We will save the result of the differential expression analysis in `results/references/00b_marker_genes_fetal_kidney_Stewart.csv`.
Notebook is saved in the `notebook/00-reference` directory.

```{r path_to_output}
path_to_output <- file.path(module_base, "results", "references")
```

# Analysis

## Load the reference

```{r pre_process, echo=TRUE, fig.height=7, fig.width=12, message=FALSE, warning=FALSE, out.width='100%'}
fetal_kidney <- readRDS(params$fetal_kidney_path)

d1 <- do_DimPlot(fetal_kidney, reduction = "umap", dims = c(1, 2), group.by = "compartment", label = TRUE, repel = TRUE) + NoLegend()
d2 <- do_DimPlot(fetal_kidney, reduction = "umap", dims = c(1, 2), group.by = "cell_type", label = TRUE, repel = TRUE) + NoLegend()

d1 | d2
```

## Characterization of compartment and cell types in the reference

Here, we use an unbiased approach to find transcripts that characterized the different compartments and cell types.

This is just to get markers genes of the different population, in case some could be of interest for the Wilms tumor annotations.

We run `DElegate::FindAllMarkers2()` to find markers of the different clusters and manually check if they do make sense. 
`DElegate::FindAllMarkers2()` is an improved version of `Seurat::FindAllMarkers()` based on pseudobulk differential expression method. 
Please check the preprint from Hafemeister and Halbritter: https://www.biorxiv.org/content/10.1101/2023.03.28.534443v2
and tool described here: https://github.com/cancerbits/DElegate

### Find marker genes for each of the compartment

```{r markers_compatment, fig.width=8, fig.height=7, out.width='100%'}
de_results <- DElegate::FindAllMarkers2(fetal_kidney, group_column = "compartment", )

# filter the most relevant markers
s.markers <- de_results[de_results$padj < params$padj_threshold & de_results$log_fc > params$lfc_threshold & de_results$rate1 > params$rate1_threshold, ]

DT::datatable(s.markers,
  caption = ("marker genes"),
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("csv", "excel")
  )
)

# Select top 5 genes for heatmap plotting
s.markers <- na.omit(s.markers)
s.markers %>%
  group_by(group1) %>%
  top_n(n = 5, wt = log_fc) -> top5

# subset for plotting
Idents(fetal_kidney) <- fetal_kidney$compartment
cells <- WhichCells(fetal_kidney, downsample = 100)
ss <- subset(fetal_kidney, cells = cells)
ss <- ScaleData(ss, features = top5$feature)

p1 <- SCpubr::do_DimPlot(fetal_kidney, reduction = "umap", group.by = "compartment", label = TRUE, repel = TRUE) + ggtitle("compartment")
p2 <- DoHeatmap(ss, features = top5$feature, cells = cells, group.by = "compartment") + NoLegend() +
  scale_fill_gradientn(colors = c("#01665e", "#35978f", "darkslategray3", "#f7f7f7", "#fee391", "#fec44f", "#F9AD03"))
p3 <- ggplot(fetal_kidney@meta.data, aes(compartment, fill = compartment)) +
  geom_bar() +
  NoLegend()

common_title <- sprintf("Unsupervised clustering %s, %d cells", fetal_kidney@meta.data$orig.ident[1], ncol(fetal_kidney))
show((((p1 / p3) + plot_layout(heights = c(3, 2)) | p2)) + plot_layout(widths = c(1, 2)) + plot_layout(heights = c(3, 1)) + plot_annotation(title = common_title))

write_csv(de_results, file = file.path(path_to_output, "00a_marker_compartment_fetal_kidney_Stewart.csv"))
```

### Find marker genes for each of the cell types

```{r markers_cell, fig.width=17, fig.height=20, out.width='100%'}
de_results <- DElegate::FindAllMarkers2(fetal_kidney, group_column = "cell_type")

# filter the most relevant markers
s.markers <- de_results[de_results$padj < params$padj_threshold & de_results$log_fc > params$lfc_threshold & de_results$rate1 > params$rate1_threshold, ]

DT::datatable(s.markers,
  caption = ("marker genes"),
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("csv", "excel")
  )
)

# Select top 5 genes for heatmap plotting
s.markers <- na.omit(s.markers)
s.markers %>%
  group_by(group1) %>%
  top_n(n = 5, wt = log_fc) -> top5

# subset for plotting
Idents(fetal_kidney) <- fetal_kidney$cell_type
cells <- WhichCells(fetal_kidney, downsample = 100)
ss <- subset(fetal_kidney, cells = cells)
ss <- ScaleData(ss, features = top5$feature)

p1 <- SCpubr::do_DimPlot(fetal_kidney, reduction = "umap", group.by = "cell_type", label = TRUE, repel = TRUE) + ggtitle("cell_type") + NoLegend()
p2 <- DoHeatmap(ss, features = top5$feature, cells = cells, group.by = "cell_type") + NoLegend() +
  scale_fill_gradientn(colors = c("#01665e", "#35978f", "darkslategray3", "#f7f7f7", "#fee391", "#fec44f", "#F9AD03"))
p3 <- ggplot(fetal_kidney@meta.data, aes(cell_type, fill = cell_type)) +
  geom_bar() +
  NoLegend() +
  scale_x_discrete(guide = guide_axis(angle = 90))

common_title <- sprintf("Unsupervised clustering %s, %d cells", fetal_kidney@meta.data$orig.ident[1], ncol(fetal_kidney))
show((((p1 / p3) + plot_layout(heights = c(3, 2)) | p2)) + plot_layout(widths = c(1, 1)) + plot_layout(heights = c(3, 1)) + plot_annotation(title = common_title))

write_csv(de_results, file = file.path(path_to_output, "00a_marker_cell-type_fetal_kidney_Stewart.csv"))
```

## References

- [1] https://www.science.org/doi/10.1126/science.aat5031

## Session info

```{r }
sessionInfo()
```