### 2024-11-20 analyses/cell-type-ewings/template_notebooks/clustering-workflow/01-clustering-metrics.Rmd
use different resolution for cpm

---
title: "Clustering exploration for `r params$library_id`"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
params:
  library_id: "SCPCL000822"
  sce_file: "../../../../data/current/SCPCP000015/SCPCS000490/SCPCL000822_processed.rds"
  cluster_results_file: "SCPCL000822_clusters.tsv"
---

This notebook summarizes clustering metrics for `r params$library_id`.

Louvain clusters are calculated using the following range of parameters:

- Nearest neighbors: 5, 10, 15, 20, 25, 30, 35, and 40
- Resolution: 0.5, 1, 1.5

Leiden clusters are calcualted using the following range of parameters:

- Objective function: CPM and modularity 
- Nearest neighbors: 5, 10, 15, 20, 25, 30, 35, and 40
- Resolution: 0.5, 1, 1.5

- Metrics are then calculated to evaluate clustering results across all parameters: 
  - Silhouette width 
  - Cluster purity 
  - Cluster stability

## Setup

```{r}
# check that sce file exists
stopifnot(
  "sce file does not exist" = file.exists(params$sce_file)
)
```

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)

# set seed
set.seed(2024)
```

```{r base paths}
# The path to this module
module_base <- rprojroot::find_root(rprojroot::is_renv_project)

# source in clustering functions
clustering_functions <- file.path(module_base, "scripts", "utils", "clustering-functions.R")
source(clustering_functions)
```

```{r}
# read in input sce files
sce <- readr::read_rds(params$sce_file)
```

## Calculate clusters

```{r}
# louvain clustering across all parameters 
louvain_results <- rOpenScPCA::sweep_clusters(
  sce,
  algorithm = "louvain",
  nn = seq(5, 40, 5),
  resolution = c(0.5, 1, 1.5)
) |> 
  dplyr::bind_rows() |> 
  # add in objective function so we can combine with leiden results
  dplyr::mutate(objective_function = NA_character_)

# leiden cpm clustering
leiden_cpm_results <- rOpenScPCA::sweep_clusters(
  sce,
  algorithm = "leiden",
  nn = seq(5, 40, 5),
  resolution = c(0.001, .005, .01),
  objective_function = "CPM"
) |> 
  dplyr::bind_rows()

# leiden modularity clustering
leiden_mod_results <- rOpenScPCA::sweep_clusters(
  sce,
  algorithm = "leiden",
  nn = seq(5, 40, 5),
  resolution = c(0.5, 1, 1.5),
  objective_function = "modularity"
) |> 
  dplyr::bind_rows()
```

```{r}
# join all clustering results 
all_cluster_results <- dplyr::bind_rows(list(louvain_results, leiden_cpm_results, leiden_mod_results)) |> 
  # add column for method + objective function 
  # use this for splitting plots throughout report
  dplyr::mutate(cluster_method = glue::glue(
    "{algorithm}-{objective_function}"
  ) |> 
    stringr::str_remove("-NA")
  )
```

## Visualize clusters with UMAP

```{r}
# get umap embeddings and combine into a data frame with cluster assignments
umap_df <- sce |>
  scuttle::makePerCellDF(use.dimred = "UMAP") |>
  # replace UMAP.1 with UMAP1 and get rid of excess columns
  dplyr::select("cell_id" = barcodes, UMAP1 = UMAP.1, UMAP2 = UMAP.2) |>
  dplyr::left_join(all_cluster_results, by = "cell_id")

split_umap_df <- umap_df |> 
  # add a column that shows the combination of nn/res so we get stats for each unique combo of params 
  dplyr::mutate(cluster_params = glue::glue("{nn}nn-{resolution}res")) |> 
  split(umap_df$cluster_method)
```

Below we visualize the cluster assignments using each parameter on a UMAP. 
This can be helpful to see if any parameters lead to any obvious over or under clustering.

```{r, fig.height=15, fig.width=10}
# look at clustering results for each library across all params
split_umap_df |> 
  purrr::imap(\(df, method){

    ggplot(df, aes(x = UMAP1, y = UMAP2, color = cluster)) +
      geom_point(alpha = 0.5, size = 0.1) +
      facet_grid(rows = vars(nn),
                 cols = vars(resolution),
                 labeller = labeller(nn = ~ glue::glue("{.}-nn"),
                                     resolution = ~ glue::glue("{.}-res"))) +
      theme(
        aspect.ratio = 1,
        legend.position = "none", 
        panel.border = element_rect(color = "black", fill = NA)
      ) +
      labs(title = method)

  })
```

## Clustering statistics

Below we calculate a series of statistics:

- Average silhouette width: This metric evaluates cluster separation. 
Values range from -1 to 1.
Cells with large positive silhouette widths are closer to other cells in the same cluster than to cells in different clusters. 
Higher values indicate tighter clusters.
- Average cluster purity: This metric also evaluates cluster separation and tells us the proportion of neighboring cells that are assigned to the same cluster. 
Purity values range from 0-1 with higher purity values indicating clusters that are well separated. 
- Cluster stability: This evaluates how stable the clustering is to input data. 
Stability values range from 0-1 with higher values of cluster stability indicating more reproducible clusters.

```{r, warning=FALSE}
# get a combined stats dataframe with purity and width for all clusters
all_stats <- split_umap_df |> 
  purrr::map(\(df) {get_cluster_stats(sce,df)})
```

### Silhouette width

```{r}
# silhouette width for different params
all_stats |> 
  purrr::imap(\(df, cluster_method){plot_cluster_stats(df, width, cluster_method)})
```

### Cluster purity

```{r}
# cluster purity for different params
all_stats |> 
  purrr::imap(\(df, cluster_method){plot_cluster_stats(df, purity, cluster_method)})
```

### Cluster stability

```{r, warning=FALSE, message=FALSE}
# calculate cluster stability
stability_stats <- split_umap_df |> 
  purrr::map(\(df){ get_cluster_stability(sce, df)})
```

```{r, warning=FALSE, message=FALSE}
# plot stability across all parameters
stability_stats |> 
  purrr::imap(\(df, cluster_method){

    # plot stability across all values of k
    ggplot(df, aes(x = nn, y = ari)) +
      geom_jitter(width = 0.1) +
      facet_wrap(vars(resolution),
                 labeller = labeller(resolution = ~ glue::glue("{.}-res"))) +
      labs(title = "Cluster stability") +
      stat_summary(
        aes(group = nn),
        color = "red",
        # median and quartiles for point range
        fun = "median",
        fun.min = function(x) {
          quantile(x, 0.25)
        },
        fun.max = function(x) {
          quantile(x, 0.75)
        }
      )

  })
```

## Export clusters

```{r}
# export all clustering results
readr::write_tsv(all_cluster_results, params$cluster_results_file)
```

## Session info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```