### 2024-12-20 analyses/hello-clusters/02_compare-clustering-parameters.Rmd
WIP: started sweep notebook

---
title: "Comparing clustering parameters with rOpenScPCA"
date: "`r Sys.Date()`"
author: "Data Lab"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    df_print: paged
---

## Introduction

Clustering algorithms have several parameters which can be varied, leading to different clustering results.
A key question when clustering, therefore, is how to identify a set of parameters that lead to robust and reliable clusters that can be used in downstream analysis.

This notebook provides examples of how to use the `rOpenScPCA` package to:

* Calculate clusters across several different parameterizations
* Calculate QC metrics on across parameterizations

Please refer to the [`01_perform-evaluate-clustering.Rmd`](01_perform-evaluate-clustering.Rmd) notebook for a tutorial on using `rOpenScPCA` functions to:

* Calculate clusters from a single parameterization
* Calculate QC metrics on a single set of clusters

This notebook will use the sample `SCPCS000001` from project `SCPCP000001`, which is assumed present in the `OpenScPCA-analysis/data/current/SCPCP000001` directory, for all examples.
Please [see this documentation](https://openscpca.readthedocs.io/en/latest/getting-started/accessing-resources/getting-access-to-data/) for more information about obtaining ScPCA data.

## Setup

### Packages

```{r packages}
library(rOpenScPCA)

suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(Seurat)
  library(dplyr)
  library(ggplot2)
})

# Set ggplot theme for plots
theme_set(theme_bw())
```

### Paths

```{r base paths}
# The base path for the OpenScPCA repository
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current")

# The path to this module
module_base <- file.path(repository_base, "analyses", "hello-clusters")
```

```{r input file path}
# Path to processed SCE file for sample SCPCS000001
input_sce_file <- file.path(data_dir, "SCPCP000001", "SCPCS000001", "SCPCL000001_processed.rds")
```

### Set the random seed

Because clustering involves random sampling, it is important to set the random seed at the top of your analysis script or notebook to ensure reproducibility.

```{r set seed}
set.seed(2024)
```

## Read in and prepare data

To begin, we'll read in the `SingleCellExperiment` (SCE) object.

```{r read data}
# Read the SCE file
sce <- readRDS(input_sce_file)
```

For the initial cluster calculations and evaluations, we will use the PCA matrix extracted from the SCE object.
As shown in [`01_perform-evaluate-clustering.Rmd`](01_perform-evaluate-clustering.Rmd), it is also possible to use an SCE object or a Seurat object directly.

```{r extract pca data}
# Extract the PCA matrix from an SCE object
pca_matrix <- reducedDim(sce, "PCA")
```

## Perform clustering across a set of parameters

This section will show how to perform clustering across a set of parameters (aka, "sweep" a set of parameters) with `rOpenScPCA::sweep_clusters()`.

This function takes a PCA matrix with rownames representing unique cell ids (e.g., barcodes) as its primary argument, with additional arguments for cluster parameters. 
This function wraps the `rOpenScPCA::calculate_clusters()` function but allows you to provide a vector of parameter values to perform clustering across, as listed below.
Clusters will be calculated for all combinations of parameters values (where applicable); default values that the function will use for any unspecified parameter values are shown in parentheses

* `algorithm`: Which clustering algorithm to use (Louvain)
* `weighting`: Which weighting scheme to use (Jaccard)
* `nn`: The number of nearest neighbors (10)
* `resolution`: The resolution parameter (1; used only with Louvain and Leiden clustering)
* `objective_function`: The objective function to optimize clusters (CPM; used only with Leiden clustering)

`rOpenScPCA::sweep_clusters()` does allow you to specify values for any other parameters.

This function will return a list of data frames of clustering results. 
Each data frame will have the following columns:

* `cell_id`: Unique cell identifiers, obtained from the PCA matrix's row names
* `cluster`: A factor column with the cluster identities
* There will be one column for each clustering parameter used

To demonstrate this function, we'll calculate clusters using the Louvain algorithm while varying the `nn` parameter:

```{r sweep clusters}
# Define nn parameter values of interest
nn <- seq(10, 30, 10)

# Calculate clusters varying nn, but leaving other parameters at their default values
cluster_results_list <- rOpenScPCA::sweep_clusters(
  pca_matrix,
  nn = nn
)
```

The resulting list has a length of three, one data frame for each `nn` parameter tested:

```{r length cluster_results_list}
length(cluster_results_list)
```

We can look at the first few rows of each data frame using [`purrr::map()`](https://purrr.tidyverse.org/reference/map.html) to iterate over the list:

```{r map cluster_results_list}
cluster_results_list |>
  purrr::map(head)
```

Generally speaking, `purrr::map()` can be used to iterate over this list to analyze or visualize each clustering result on its own.
Complementary to this, we can also combine this list into a single data frame to jointly analyze or visualize all clustering results together:

```{r bind cluster_results_list rows}
cluster_results_df <- dplyr::bind_rows(cluster_results_list)
head(cluster_results_df)
```

## Evaluating clustering results

This section will use `purrr::map()` to iterate over each df to calculate silhouette, purity, and stability. 
We'll also plot results.

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```