### 2025-01-16 analyses/cell-type-ewings/exploratory_analysis/07-identify-max-rank-aucell.Rmd
make the report pretty

---
title: "Identify max rank for AUCell"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

This notebook is used to find the appropriate `aucMaxRank` to use with `AUCell` for all samples in `SCPCP00015`.

From the  [`AUCell` vignette](https://bioconductor.org/packages/release/bioc/vignettes/AUCell/inst/doc/AUCell.html):

> It is important to check that most cells have at least the number of expressed/detected genes that are going to be used to calculate the AUC (aucMaxRank in calcAUC()). The histogram provided by AUCell_buildRankings() allows to quickly check this distribution.

Here we will look at a histogram showing the number of genes detected per cell for each sample and use that information to pick a threshold (percentage of genes) to use across all samples.

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)

# set seed
set.seed(2025)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current", "SCPCP000015")

# paths to all processed sce files
sce_files <- list.files(data_dir, pattern = "*processed.rds", full.names = TRUE, recursive = TRUE)
```

```{r}
# read in SCE files 
sce_list <- sce_files |> 
  purrr::map(readr::read_rds)
```

## Filter `SingleCellExperiment` objects

The default `aucMaxRank` is calculated as `ceiling(0.05*nrow(exprMat))` where the `exprMat` is the counts matrix. 
For the processed `SingleCellExperiment` objects in ScPCA we include all genes in the index, regardless of if they are detected in our sample. 
This means the rankings might include a bunch of genes that aren't actually present and the `aucMaxRank` would be calculated based on a set of genes that aren't even there, setting the threshold higher than it should be. 
Because of this, we probably want to filter the `SingleCellExperiment` objects to remove any genes that are not detected in any cells prior to calculating ranks or running `AUCell`.

```{r}
# get a list of filtered sce
# remove any genes that are not detected in any cells 
sce_list <- sce_list |> 
  purrr::map(\(sce){

    genes_to_remove <- rowData(sce)$detected > 0 
    filtered_sce <- sce[genes_to_remove > 0 , ]

  })
```

## Number of genes detected

Below we will look at the distribution of the number of genes detected per cell for each sample.

```{r, fig.height = 10, fig.width=10, warning=FALSE}
# make a list of the `countByCell` as AUCell does
names(sce_list) <- stringr::str_extract(sce_files, "SCPCS\\d+")
count_df_list <- sce_list |>
  purrr::map(counts) |>
  purrr::map(
    \(counts) {
      countByCell <- Matrix::colSums(counts > 0, na.rm = TRUE)
      data.frame(count_by_cell = countByCell)
    })

# first, make the plots
count_df_list |>
  purrr::imap(
    \(count_df, sample_id) {
        count_hist <- ggplot(count_df) + 
          aes(x = count_by_cell) + 
          geom_histogram(color = "grey20", fill = "skyblue", bins = 30) + 
          xlim(0, max(count_df$count_by_cell)) +
          xlab("Number of genes detected per cell")

        count_box <- ggplot(count_df) + 
          aes(x = count_by_cell, y = "") + 
          geom_boxplot(color = "grey20", fill = "skyblue", outlier.size = 0.5) + 
          theme_void() + 
          xlim(0, max(count_df$count_by_cell)) + 
          ggtitle(sample_id)

      count_box / count_hist + patchwork::plot_layout(heights = c(0.2, 1))
    }
  ) |>
  patchwork::wrap_plots(nrow = 5)

# second, print a table of the stats
count_df_list |>
  purrr::map(
     \(count_df) {
         c(
          min=min(count_df$count_by_cell), 
          quantile(count_df$count_by_cell, c(.01,.05, .10, .50, 1))
        )
     }) |>
  dplyr::bind_rows(.id = "sample_id")
```

Now we will look at the potential thresholds for `aucMaxRank` using 5% and 1% of genes. 
The numbers printed here represent the total number of genes that would be used as the `aucMaxRank` for a given dataset using the threshold. 
We can then use the histograms to see how many cells have at least that number of genes detected. 
Remember, we want to pick a number of genes that are detected in most cells.
If we picked a max rank higher than the number of genes detected in most cells, the non-detected genes that are randomly ordered would play an outsized role in our AUC values.

### 5%

```{r}
# get max rank using 1% threshold
threshold_list <- sce_list |> 
  purrr::map_dbl(\(sce){ceiling(nrow(counts(sce))*0.05)})

# make it legible
data.frame(
  max_rank = threshold_list
)
```

This looks too high based on our histograms. 
We would have a lot of cells where the number of genes detected is lower than the max rank, which is not ideal.

### 1%

```{r}
# get max rank using 1% threshold
threshold_list <- sce_list |> 
  purrr::map_dbl(\(sce){ceiling(nrow(counts(sce))*0.01)})

# make it legible
data.frame(
  max_rank = threshold_list
)
```

This looks better to me. 
I also think there's enough variation in some of the samples that we don't want to just pick one max rank for all samples, but instead use a percentage of the total genes detected.

## Conclusions

- Prior to running `AUCell` we should remove any genes that are not detected in any cells.
- Using a cutoff of 1% of the total genes detected seems like a reasonable choice for the `aucMaxRank`.

## Session info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```