### 2025-01-29 analyses/cell-type-ewings/template_notebooks/celltype-exploration.Rmd
adjust plots based on review comments

---
title: "Template notebook for validating cell type assignments for an individual library in SCPCP000015"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
params:
  sample_id: "SCPCS000490"
  library_id: "SCPCL000822" 
  cluster_nn: 20
  cluster_res: 0.5
---

This notebook is meant to be a guide for compiling "final" cell type annotations for an individual library in `SCPCP000015`. 
Results from `aucell-singler-annotation.sh`, `evaulate-clusters.sh`, and `run-aucell-ews-signatures.sh` are all combined and used to assign cell type annotations on a case by case basis.

Instructions for using this guide:

1. Ensure that you have a local copy of the results from `aucell-singler-annotation.sh`, `evaluate-clusters.sh` and `run-aucell-ews-signatures.sh` saved to `results`. 
2. Copy the contents of this notebook to a new notebook titled `<library_id>_celltype-exploration.Rmd` and save in `exploratory_analysis/final_annotation_notebooks`. 
3. Update the `title` in the `yaml` section and replace the `sample_id` and `library_id` with the correct IDs in the `params` list. 
4. Optionally, you may choose to update the choices for clustering based on the results from `evaluate-clusters.sh`. 
All clusters used will be calculated with the Leiden algorithm and the modularity objective function. 
To modify the nearest neighbors (default: 20) and resolution (default: 0.5) chosen use the `cluster_nn` and `cluster_res` params. 
5. Run through the notebook and update any sections of the notebook marked with the `{.manual-exploration}` tag. 
6. Render the completed notebook which will produce the rendered `html` file and a TSV with cell type annotations for that library.

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)

# set seed
set.seed(2024)

# quiet messages
options(readr.show_col_types = FALSE)
ComplexHeatmap::ht_opt(message = FALSE)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current", "SCPCP000015")

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-ewings") 
```

```{r}
# path to sce 
sce_file <- file.path(data_dir, params$sample_id, glue::glue("{params$library_id}_processed.rds"))

# path to workflow results
workflow_results_dir <- file.path(module_base, "results")

singler_results_dir <- file.path(workflow_results_dir, "aucell_singler_annotation", params$sample_id)
singler_results_file <- file.path(singler_results_dir,
                                  glue::glue("{params$library_id}_singler-classifications.tsv"))

cluster_results_dir <- file.path(workflow_results_dir, "clustering", params$sample_id)
cluster_results_file <- file.path(cluster_results_dir,
                                  glue::glue("{params$library_id}_cluster-results.tsv"))

aucell_results_dir <- file.path(workflow_results_dir, "aucell-ews-signatures", params$sample_id)
aucell_results_file <- file.path(aucell_results_dir,
                                 glue::glue("{params$library_id}_auc-ews-gene-signatures.tsv"))

# small gene sets
visser_marker_genes_file <- file.path(module_base, "references", "visser-all-marker-genes.tsv")
cell_state_genes_file <- file.path(module_base, "references", "tumor-cell-state-markers.tsv")
```

```{r}
# output file to save final annotations 
results_dir <- file.path(module_base, "results", "final-annotations")
output_file <- file.path(results_dir, glue::glue("{params$library_id}_celltype-annotations.tsv"))
```

```{r}
# source in setup functions prep_results()
setup_functions <- file.path(module_base, "template_notebooks", "utils", "setup-functions.R")
source(setup_functions)

# source in validation functions 
# calculate_mean_markers(), plot_faceted_umap()
validation_functions <- file.path(module_base, "scripts", "utils", "tumor-validation-helpers.R")
source(validation_functions)

# source in plotting functions 
# expression_umap(), cluster_density_plot(), and annotated_exp_heatmap()
plotting_functions <- file.path(module_base, "template_notebooks", "utils", "plotting-functions.R")
source(plotting_functions)
```

```{r}
stopifnot(
  "sce file does not exist" = file.exists(sce_file),
  "singler results file does not exist" = file.exists(singler_results_file),
  "cluster results file does not exist" = file.exists(cluster_results_file),
  "aucell results file does not exist" = file.exists(aucell_results_file)
)
```

```{r, message=FALSE}
# read in sce
sce <- readr::read_rds(sce_file)

# read in workflow results
singler_df <- readr::read_tsv(singler_results_file)
cluster_df <- readr::read_tsv(cluster_results_file)
aucell_df <- readr::read_tsv(aucell_results_file)

# read in marker genes and combine into one list 
visser_markers_df <- readr::read_tsv(visser_marker_genes_file) |> 
  dplyr::select(cell_type, ensembl_gene_id) |> 
  unique()

cell_state_markers_df <- readr::read_tsv(cell_state_genes_file) |> 
  dplyr::select(cell_type = cell_state, ensembl_gene_id)

all_markers_df <- dplyr::bind_rows(list(visser_markers_df, cell_state_markers_df))
```

## Prepare data for plotting

```{r}
all_results_df <- prep_results(
  sce, 
  singler_df, 
  cluster_df, 
  aucell_df,
  cluster_nn = params$cluster_nn,
  cluster_res = params$cluster_res
  )

cell_types <- unique(all_markers_df$cell_type)

# get the mean expression of all genes for each cell state
gene_exp_df <- cell_types |>
  purrr::map(\(type){
    calculate_mean_markers(all_markers_df, sce, type, cell_type)
  }) |>
  purrr::reduce(dplyr::inner_join, by = "barcodes")

all_info_df <- all_results_df |> 
  dplyr::left_join(gene_exp_df, by = "barcodes") |> 
  dplyr::arrange(cluster)
```

## Summary of workflow results

### `aucell-singler-annotation` assignments

The below UMAP shows the top cell types assigned by `SingleR` in the `aucell-singler-annotation.sh` workflow. 
The top 7 cell types are shown and all other cell types are grouped together as "All remaining cell types".

```{r, fig.height = 5}
plot_faceted_umap(all_info_df, singler_lumped, legend_title = "SingleR cell types") +
  theme(strip.text = element_text(size = 8))
```

### Cluster assignments

The below UMAP shows the cluster assignments output by `evaluate-clusters.sh` using the following parameters:

- Leiden with modularity 
- Nearest neighbors: `r params$cluster_nn`
- Resolution: `r params$cluster_res`

```{r, fig.height = 5}
plot_faceted_umap(all_info_df, cluster, legend_title = "Cluster")
```

### `AUCell` results

The below plots show the AUC values determined by `AUCell` and output from `run-aucell-ews-signatures.sh`. 
The first plot shows the individual AUC values on the UMAP.

```{r}
# get the individual thresholds determined by AUCell for each msigdb geneset
# we want this so we can show the threshold on the density plots 
auc_threshold_df <- all_info_df |> 
  tidyr::pivot_longer(starts_with("threshold_auc_"), names_to = "geneset", values_to = "threshold") |> 
  dplyr::mutate(
    geneset = stringr::str_remove(geneset, "threshold_auc_")
  ) |> 
  dplyr::select(barcodes, geneset, threshold)

# reformat auc data for density plots and UMAPs showing AUC values 
auc_df <- all_info_df |> 
  tidyr::pivot_longer(starts_with("auc_"), names_to = "geneset", values_to = "auc_value") |> 
  dplyr::mutate(
    geneset = stringr::str_remove(geneset, "auc_")
  ) |> 
  dplyr::select(barcodes, UMAP1, UMAP2, geneset, auc_value, cluster) |> 
  dplyr::left_join(auc_threshold_df, by = c("barcodes", "geneset")) |> 
  dplyr::mutate(
    in_geneset =  auc_value > threshold
  )
```

```{r, fig.height=10, fig.width=10}
expression_umap(auc_df, auc_value, geneset)
```

The below plot shows the distribution of AUC values for each gene set colored based on if the AUC value is above the threshold determined by `AUCell`, indicated with a dotted line.

```{r, fig.height=5}

ggplot(auc_df, aes(x = auc_value, color = in_geneset, fill = in_geneset)) +
  geom_density(alpha = 0.5, bw = 0.01) +
  facet_wrap(vars(geneset)) +
  ggplot2::geom_vline(data = auc_df,
                      mapping = aes(xintercept = threshold),
                      lty = 2) +
  theme(
      aspect.ratio = 1,
      strip.background = element_rect(fill = "transparent", linewidth = 0.5),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5)
    ) 
```

Here we look at the AUC values for each gene set across all clusters.

```{r, fig.height=10}
auc_columns <- colnames(all_info_df)[which(startsWith(colnames(all_info_df), "auc_"))]
cluster_density_plot(all_info_df, auc_columns, "AUC")
```

The heatmap below shows the AUC values for all cells and all gene sets. 
The annotations shown are the `SingleR` cell types and clusters.

```{r}
annotated_exp_heatmap(
  all_info_df, 
  exp_columns = auc_columns, 
  cell_type_column = "singler_lumped", 
  cluster_column = "cluster", 
  legend_title = "AUC"
)
```

### Mean expression of custom gene sets

Below we look at the mean expression of all genes in each of the custom gene sets we have defined both on a UMAP and in a density plot.

```{r, fig.height=10, fig.width=10}
# format expression data for UMAP and density plot 
exp_df <- all_info_df |> 
  tidyr::pivot_longer(ends_with("_mean"), names_to = "geneset", values_to = "mean") |>
  dplyr::mutate(tumor_cell_classification = dplyr::if_else(singler_lumped == "tumor", "tumor", "other"))

expression_umap(exp_df, mean, geneset)
```

```{r, fig.height=8}
ggplot(exp_df, aes(x = mean, color = tumor_cell_classification, fill = tumor_cell_classification)) +
  geom_density(alpha = 0.5, bw = 0.2) +
  facet_wrap(vars(geneset)) +
  theme(
    aspect.ratio = 1,
    strip.background = element_rect(fill = "transparent", linewidth = 0.5),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5)
  )
```

Here we look at the mean expression for each gene set across all clusters.

```{r, fig.height=7}
mean_exp_columns <- colnames(all_info_df)[which(endsWith(colnames(all_info_df), "_mean"))]
cluster_density_plot(all_info_df, mean_exp_columns, "mean gene expression")
```

The heatmap below shows the mean expression values for all cells and all gene sets. 
The annotations shown are the `SingleR` cell types and clusters.

```{r}
annotated_exp_heatmap(
  all_info_df, 
  exp_columns = mean_exp_columns, 
  cell_type_column = "singler_lumped", 
  cluster_column = "cluster", 
  legend_title = "Mean\ngene expression"
)
```

## Re-cluster tumor cells {.manual-exploration}

<!-- Use this section to label tumor cells based on the above findings. 
Any cells that are labeled as tumor will then be re-clustered and plots showing only tumor cells
can be created to identify tumor cell states -->

TODO: Functions for re-clustering tumor cells
Show the same plots across the tumor only clusters and assign tumor cell states to each cluster

## Additional exploration {.manual-exploration}

<!--This section is for any additional exploration that may be needed to finalize annotations for this library. 
If not using, please delete this section.
For example, here you may want to dive into the normal cell types and make adjustments as needed
-->

## Validate final tumor and normal annotation {.manual-exploration}

<!-- This section should be used to update the assignments -->

TODO: Insert plots that will be useful for validation (UMAPs, heatmaps, density plots)

## Prepare annotations {.manual-exploration}

<!-- This section should be used to create the final.final table with cell type annotations for export -->

TODO: Code and instructions for exporting such as what columns should be named.

## Session info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```