### 2025-02-04 analyses/cell-type-ewings/exploratory_analysis/08-merged-celltypes.Rmd
incorporate consensus cell types

---
title: "Cell type assignments for SCPCP000015"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
---

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(SingleCellExperiment)
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)

# set seed
set.seed(2024)

# quiet messages
options(readr.show_col_types = FALSE)
ComplexHeatmap::ht_opt(message = FALSE)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current", "results", "merge-sce", "SCPCP000015")

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-ewings") 
```

```{r}
# path to sce 
sce_file <- file.path(data_dir, "SCPCP000015_merged.rds")

# path to workflow results
workflow_results_dir <- file.path(module_base, "results")

singler_results_dir <- file.path(workflow_results_dir, "aucell_singler_annotation")
singler_results_files <- list.files(singler_results_dir, pattern = "*singler-classifications.tsv", full.names = TRUE, recursive = TRUE)
library_ids <- stringr::str_remove(basename(singler_results_files), "_singler-classifications.tsv")

aucell_results_dir <- file.path(workflow_results_dir, "aucell-ews-signatures")
aucell_results_file <- file.path(aucell_results_dir, "SCPCP000015_auc-ews-gene-signatures.tsv")

# small gene sets
visser_marker_genes_file <- file.path(module_base, "references", "visser-all-marker-genes.tsv")
cell_state_genes_file <- file.path(module_base, "references", "tumor-cell-state-markers.tsv")
```

```{r}
# output file to save final annotations 
results_dir <- file.path(module_base, "results", "final-annotations")
output_file <- file.path(results_dir, glue::glue("SCPCP000015_celltype-annotations.tsv"))
```

```{r}
# source in setup functions prep_results()
setup_functions <- file.path(module_base, "template_notebooks", "utils", "setup-functions.R")
source(setup_functions)

# source in validation functions 
# calculate_mean_markers(), plot_faceted_umap()
validation_functions <- file.path(module_base, "scripts", "utils", "tumor-validation-helpers.R")
source(validation_functions)

# source in plotting functions 
# expression_umap(), cluster_density_plot(), and annotated_exp_heatmap()
plotting_functions <- file.path(module_base, "template_notebooks", "utils", "plotting-functions.R")
source(plotting_functions)
```

```{r}
stopifnot(
  "sce file does not exist" = file.exists(sce_file),
  "aucell results file does not exist" = file.exists(aucell_results_file)
)
```

```{r, message=FALSE}
# read in sce
sce <- readr::read_rds(sce_file)

# read in workflow results
singler_df <- singler_results_files |> 
  purrr::set_names(library_ids) |>
  purrr::map(readr::read_tsv) |> 
  dplyr::bind_rows(.id = "library_id")
aucell_df <- readr::read_tsv(aucell_results_file) |> 
  tidyr::separate(barcodes, "-", into = c("library_id", "barcodes"))

# read in marker genes and combine into one list 
visser_markers_df <- readr::read_tsv(visser_marker_genes_file) |> 
  dplyr::select(cell_type, ensembl_gene_id) |> 
  unique()

cell_state_markers_df <- readr::read_tsv(cell_state_genes_file) |> 
  dplyr::select(cell_type = cell_state, ensembl_gene_id)

all_markers_df <- dplyr::bind_rows(list(visser_markers_df, cell_state_markers_df))
```

## Prepare data for plotting

```{r}
all_results_df <- prep_results(
  sce, 
  singler_df = singler_df, 
  cluster_df = NULL, 
  aucell_df = aucell_df,
  consensus_df = consensus_df,
  cluster_nn = params$cluster_nn,
  cluster_res = params$cluster_res,
  join_columns = c("barcodes", "library_id")
  ) |>
  dplyr::mutate(barcodes = glue::glue("{library_id}-{barcodes}"))

cell_types <- unique(all_markers_df$cell_type)

# get the mean expression of all genes for each cell state
gene_exp_df <- cell_types |>
  purrr::map(\(type){
    calculate_mean_markers(all_markers_df, sce, type, cell_type)
  }) |>
  purrr::reduce(dplyr::inner_join, by = "barcodes")

all_info_df <- all_results_df |> 
  dplyr::left_join(gene_exp_df, by = "barcodes")
```

## Summary of workflow results

### `aucell-singler-annotation` assignments

The below UMAP shows the top cell types assigned by `SingleR` in the `aucell-singler-annotation.sh` workflow. 
The top 7 cell types are shown and all other cell types are grouped together as "All remaining cell types".

```{r, fig.height = 5}
plot_faceted_umap(all_info_df, singler_lumped, legend_title = "SingleR cell types") +
  theme(strip.text = element_text(size = 8))
```

### Consensus cell type assignments

```{r, fig.height=5}
plot_faceted_umap(all_info_df, consensus_lumped, legend_title = "Consensus cell types") +
  theme(strip.text = element_text(size = 8))
```

### `AUCell` results

The below plots show the AUC values determined by `AUCell` and output from `run-aucell-ews-signatures.sh`. 
The first plot shows the individual AUC values on the UMAP.

```{r}
# get the individual thresholds determined by AUCell for each msigdb geneset
# we want this so we can show the threshold on the density plots 
auc_threshold_df <- all_info_df |> 
  tidyr::pivot_longer(starts_with("threshold_auc_"), names_to = "geneset", values_to = "threshold") |> 
  dplyr::mutate(
    geneset = stringr::str_remove(geneset, "threshold_auc_")
  ) |> 
  dplyr::select(barcodes, geneset, threshold)

# reformat auc data for density plots and UMAPs showing AUC values 
auc_df <- all_info_df |> 
  tidyr::pivot_longer(starts_with("auc_"), names_to = "geneset", values_to = "auc_value") |> 
  dplyr::mutate(
    geneset = stringr::str_remove(geneset, "auc_")
  ) |> 
  dplyr::select(barcodes, UMAP1, UMAP2, geneset, auc_value) |> 
  dplyr::left_join(auc_threshold_df, by = c("barcodes", "geneset")) |> 
  dplyr::mutate(
    in_geneset =  auc_value > threshold
  )
```

```{r, fig.height=10, fig.width=10}
expression_umap(auc_df, auc_value, geneset)
```

The below plot shows the distribution of AUC values for each gene set colored based on if the AUC value is above the threshold determined by `AUCell`, indicated with a dotted line.

```{r, fig.height=5}

ggplot(auc_df, aes(x = auc_value, color = in_geneset, fill = in_geneset)) +
  geom_density(alpha = 0.5, bw = 0.01) +
  facet_wrap(vars(geneset)) +
  ggplot2::geom_vline(data = auc_df,
                      mapping = aes(xintercept = threshold),
                      lty = 2) +
  theme(
      aspect.ratio = 1,
      strip.background = element_rect(fill = "transparent", linewidth = 0.5),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5)
    ) 
```

Here we look at the AUC values for each gene set across all clusters.

```{r, fig.height=10}
auc_columns <- colnames(all_info_df)[which(startsWith(colnames(all_info_df), "auc_"))]
cluster_density_plot(all_info_df, auc_columns, "consensus_lumped", "AUC")
```

Here we should identify any tumor cells as those that have high expression of `EWS-FLI` targets and low expression of `EWS-FLI` repressed targets. 
I think we want to use the gene sets where we see some separation in AUC values between cell types to do our assignment.

- `aynaud-ews-targets`: > 0.05
- `staege`: > 0.25
- `riggi_down`: < 0.025
- `migayawa_down`: < 0.03

Intriguingly we see high expression of `hallmark_EMT` and the down gene sets (genes that are repressed by `EWS-FLI`) in fibroblasts, which also happen to have high expression of the `EWS-FLI` low targets identifed by Wrenn et al.
That paper identifies `EWS-FLI` low cells as cancer associated fibroblasts.
Let's label these cells accordingly. 
If cells have `in_geneset=TRUE` for the `hallmark_EMT`, `wrenn-nt5e-genes`, and `migayawa_down` they will be classified as `ews_low_cafs`.

- `wrenn-nt5e-genes`: > 0.1
- `hallmark_EMT`: > 0.05
- `migayawa_down`: > 0.03
- `riggi_down`: > 0.025

```{r}
all_info_df <- all_info_df |> 
  dplyr::mutate(
    final_annotation = dplyr::case_when(
      `auc_wrenn-nt5e-genes` > 0.1 & auc_miyagawa_down > 0.03 ~ "ews_low_caf",
      `auc_aynaud-ews-targets` > 0.025 | 
        consensus_annotation %in% c("Unknown", "muscle cell", "smooth muscle cell") ~ "ews_high",
      .default = consensus_annotation
    ),
    final_lumped = final_annotation |>
          forcats::fct_lump_n(8, other_level = "All remaining cell types", ties.method = "first") |>
          forcats::fct_infreq() |>
          forcats::fct_relevel("All remaining cell types", after = Inf)
  )

head(all_info_df)
table(all_info_df$final_annotation)
```

The heatmap below shows the AUC values for all cells and all gene sets. 
The annotations shown are the `SingleR` cell types and clusters.

```{r}
all_info_df <- all_info_df |> 
  dplyr::arrange(final_annotation)

annotated_exp_heatmap(
  all_info_df, 
  exp_columns = auc_columns, 
  cell_type_column = "consensus_lumped", 
  cluster_column = "final_lumped", 
  legend_title = "AUC"
)
```

### Mean expression of custom gene sets

Below we look at the mean expression of all genes in each of the custom gene sets we have defined both on a UMAP and in a density plot.

```{r, fig.height=10, fig.width=10}
# format expression data for UMAP and density plot 
exp_df <- all_info_df |> 
  tidyr::pivot_longer(ends_with("_mean"), names_to = "geneset", values_to = "mean") |>
  dplyr::mutate(tumor_cell_classification = dplyr::if_else(singler_lumped == "tumor", "tumor", "other"))

expression_umap(exp_df, mean, geneset)
```

Here we look at the mean expression for each gene set across all clusters.

```{r, fig.height=7, warning=FALSE}
mean_exp_columns <- colnames(all_info_df)[which(endsWith(colnames(all_info_df), "_mean"))]
cluster_density_plot(all_info_df, mean_exp_columns, annotation_column = "final_lumped", "mean gene expression")
```

What are the other cell types not present here? 
```{r}
lumped_celltypes <- unique(all_info_df$final_lumped)
all_celltypes <- unique(all_info_df$final_annotation)
missing_celltypes <- setdiff(all_celltypes, lumped_celltypes)

missing_only <- all_info_df |> 
  dplyr::filter(final_annotation %in% missing_celltypes)
table(missing_only$final_annotation)
```

I think we can leave these few cells alone and just lump them as "All remaining cell types" for plots as we have been doing.

Now let's make a dot plot! Or a heatmap only looking at key genes! 
In these plots, lets group all T cells together.

```{r}
markers_df <- readr::read_tsv(file.path(module_base, "references", "combined-markers.tsv"))
genes <- markers_df |>
  dplyr::pull(ensembl_gene_id)

gene_cts <- logcounts(sce[genes, ]) |>
  as.matrix() |>
  t() |> 
  as.data.frame()
gene_cts$barcodes <- rownames(gene_cts)

rowData(sce)[genes, ] |> 
  as.data.frame() |> 
  dplyr::select(ends_with(".mean")) |> 
  tibble::rownames_to_column("ensembl_gene_id") |> 
  tidyr::pivot_longer(!ensembl_gene_id, names_to = "library_id", values_to = "mean")

celltypes_df <- endsWith()celltypes_df <- all_info_df |> 
  dplyr::select(barcodes, final_lumped)

genes_df <- gene_cts |> 
  tidyr::pivot_longer(!barcodes, names_to = "ensembl_gene_id", values_to = "gene_exp") |> 
  dplyr::left_join(celltypes_df, by = c("barcodes")) |> 
  dplyr::mutate(
    percent_expression = 
  )

head(genes_df)
```

## Session info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```