### 2025-02-06 analyses/cell-type-consensus/exploratory-notebooks/utils/03-osteosarcoma-consensus-celltypes.Rmd
initiate osteo notebook

---
title: "Consensus cell types across osteosarcoma samples"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    code_folding: show
---

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)
```

## Data setup

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)
module_base <- file.path(repository_base, "analyses", "cell-type-consensus")

# results directory with cell-type-consensus 
results_dir <- file.path(module_base, "results", "cell-type-consensus")

# data directory where project metadata files live
data_dir <- file.path(repository_base, "data", "current")
```

```{r}
# list all results files 
results_files <- list.files(results_dir, pattern = "_consensus-cell-types\\.tsv.\\gz$", full.names = TRUE)

# get project ids from file list  
project_ids <- stringr::str_remove(basename(results_files), "_consensus-cell-types.tsv.gz")
names(results_files) <- project_ids

# remove cell line projects from file list
osteo_project_ids <- c("SCPCP000017", "SCPCP000018", "SCPCP000023")
results_files <- results_files[osteo_project_ids]

# list metadata files 
metadata_files <- list.files(data_dir, pattern = "single_cell_metadata\\.tsv$", full.names = TRUE, recursive = TRUE)
metadata_ids <- stringr::word(dirname(metadata_files), -1, sep = "/")
names(metadata_files) <- metadata_ids
osteo_metadata_files <- metadata_files[osteo_project_ids]

# immune cell types 
consensus_immune_file <- file.path(module_base, "references", "consensus-immune-cell-types.tsv")
```

```{r}
# source summarize_celltypes() function
setup_functions <- file.path(module_base, "exploratory-notebooks", "utils", "setup-functions.R")
source(setup_functions)
```

```{r, message=FALSE}
# read in metadata
all_metadata_df <- osteo_metadata_files |>
  purrr::map(readr::read_tsv) |>
  dplyr::bind_rows() |>
  # select columns that might be useful 
  dplyr::select(
    project_id = scpca_project_id,
    library_id = scpca_library_id,
    disease_timing,
    age,
    sex,
    tissue_location,
    primary_or_metastasis,
    seq_unit
  )

# immune consensus cell types 
consensus_immune_types <- readr::read_tsv(consensus_immune_file) |> 
  dplyr::pull(consensus_annotation)

# read in results and prep data frame for plotting 
all_results_df <- results_files |> 
  purrr::imap(summarize_celltypes) |> 
  dplyr::bind_rows(.id = "project_id") |> 
  # join with sample metadata
  dplyr::left_join(all_metadata_df, by = c("project_id", "library_id")) |>
  dplyr::filter(sample_type != "patient-derived xenograft") |> 
  dplyr::mutate(is_immune = consensus_annotation %in% consensus_immune_types)

head(all_results_df)
```

```{r}
all_results_df |>
  dplyr::count(consensus_annotation) |> 
  dplyr::arrange(desc(n))
```

```{r}
all_results_df <- all_results_df |> 
    dplyr::mutate(
      # get most frequently observed cell types across libraries in that project 
      top_celltypes = forcats::fct_lump_n(consensus_annotation, 10, other_level = "All remaining cell types", ties.method = "first") |> 
        # sort by frequency 
        forcats::fct_infreq() |> 
        # make sure all remaining and unknown are last, use this to assign colors in specific order
        forcats::fct_relevel("All remaining cell types", "Unknown", after = Inf)
    )

# get all unique cell types ordered by frequency 
unique_celltypes <- all_results_df |> 
  dplyr::filter(!top_celltypes %in% c("All remaining cell types", "Unknown")) |> 
  dplyr::pull(top_celltypes) |> 
  unique() |>
  sort() |> 
  as.character()

# get color palette
colors <- c(
  palette.colors(palette = "alphabet")[1:9],
  "grey60", 
  "grey95"
)
names(colors) <- c(unique_celltypes, "All remaining cell types", "Unknown")

ggplot(all_results_df) + 
  aes(
    x = library_id, 
    y = percent_cells_annotation, 
    fill = top_celltypes
  ) +
  geom_col() + 
  facet_wrap(vars(project_id), scales ="free") +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = colors, name = "cell type") +
  theme(axis.text.x = element_blank())
```

```{r}
# t cell types
t_cell_types <- consensus_immune_types[stringr::str_detect(consensus_immune_types, "T")]
b_cell_types <- consensus_immune_types[stringr::str_detect(consensus_immune_types, "B")]
myeloid_cell_types <- c(
  "macrophage",
  "myeloid leukocyte",
  "neutrophil",
  "granulocyte",
  "monocyte",
  "erythroid lineage cell",
  "megakaryocyte",
  "eosinophil",
  "dendritic cell"
)

# look at immune cell types vs. unknown vs. other 
all_results_df <- all_results_df |> 
  dplyr::mutate(
    immune_category = dplyr::case_when(
      is_immune ~ "immune",
      consensus_annotation == "Unknown" ~ "unknown",
      .default = "other"
    ),
    broad_category = dplyr::case_when(
      consensus_annotation %in% t_cell_types ~ "T cell",
      consensus_annotation %in% b_cell_types ~ "B cell",
      consensus_annotation %in% myeloid_cell_types ~ "myeloid",
      consensus_annotation == "Unknown" ~ "unknown",
      .default = "non-immune"
    )
  )
```

```{r}
library_levels <- pct_immune_df <- all_results_df |> 
  dplyr::filter(immune_category == "immune") |> 
  dplyr::group_by(library_id) |> 
  dplyr::summarize(
    percent_immune = sum(total_cells_per_annotation)/total_cells_per_library
  ) |>
  unique() |> 
  dplyr::arrange(desc(percent_immune)) |> 
  dplyr::pull(library_id)

all_results_df <- all_results_df |> 
  dplyr::mutate(
    library_id = forcats::fct_relevel(library_id, library_levels)
  )

ggplot(all_results_df) + 
  aes(
    x = library_id, 
    y = percent_cells_annotation, 
    fill = immune_category
  ) +
  geom_col() + 
  # split samples based on sample type, patient tissue or pdx 
  #facet_wrap(vars(seq_unit), scales ="free") +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("red", "grey60", "grey95"), name = "cell type") +
  theme(axis.text.x = element_blank())
```

```{r}
# define colors for broad categories 
celtype_order <- c("myeloid", "B cell", "T cell", "non-immune", "unknown")

# get color palette
colors <- c(
  palette.colors(palette = "Dark2")[1:3],
  "grey60", 
  "grey95"
)
names(colors) <- celtype_order

ggplot(all_results_df) + 
  aes(
    x = library_id, 
    y = percent_cells_annotation, 
    fill = broad_category
  ) +
  geom_col() + 
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = colors, name = "cell type") +
  theme(axis.text.x = element_blank())
```

```{r}
updated_metadata_df <- all_results_df |> 
  dplyr::mutate(
    primary_or_metastasis = dplyr::if_else(!is.na(primary_or_metastasis), primary_or_metastasis, disease_timing)
  ) |> 
  dplyr::filter(primary_or_metastasis %in% c("Primary", "Metastasis"))

ggplot(updated_metadata_df) + 
  aes(
    x = library_id, 
    y = percent_cells_annotation, 
    fill = immune_category
  ) +
  geom_col() + 
  # split samples based on sample type, patient tissue or pdx 
  facet_wrap(vars(primary_or_metastasis), scales ="free") +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("red", "grey60", "grey95"), name = "cell type") +
  theme(axis.text.x = element_blank())

ggplot(updated_metadata_df, aes(x = primary_or_metastasis, y = percent_immune)) +
  ggforce::geom_sina() +
  stat_summary(fun.y=median, geom="crossbar" , color = "red", linewidth = 0.2) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        plot.margin = margin(10,10,10,10)) +
  labs(
    x = "", 
    y = "Percent of cells annotated as Immune"
  )
```

Do the same thing but with T/B/myeloid percentage? 
Will also need to update code to make sure immune percentage is actually in the df.

Then filter for initial diagnosis vs. recurrence and do the same thing

All samples together sorted by immune population 
facte by project id - what's different about osteo project 23? 
all samples broken up by primary vs. mets, stacked bar chart and sina plot 
all samples broken up by disease timing 
why is project 23 different? 
what about just T cell infiltrate? and macrophage?