### 2025-02-19 analyses/cell-type-consensus/exploratory-notebooks/05-marker-gene-validation.Rmd
initial exploration notebook for validation plots

---
title: "Validation of marker gene expression"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    code_folding: hide
params:
  project_id: "SCPCP000001"
---

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)
```

## Data setup

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)
module_base <- file.path(repository_base, "analyses", "cell-type-consensus")

# results directory with cell-type-consensus 
results_dir <- file.path(module_base, "results", "cell-type-consensus", params$project_id)

# data directory where project metadata files live
data_dir <- file.path(repository_base, "data", "current")
```

```{r}
# get consensus results 
consensus_results_files <- list.files(results_dir, pattern = "_consensus-cell-type-assignments\\.tsv.\\gz$", recursive = TRUE, full.names = TRUE)

# get gene expression files 
gene_exp_files <- list.files(results_dir, pattern = "_marker-gene-expression\\.tsv.\\gz$", recursive = TRUE, full.names = TRUE)

# list of all validation markers
validation_markers_file <- file.path(module_base, "references", "validation-markers.tsv")

# all validation groups mapped to consensus labels 
validation_groups_file <- file.path(module_base, "references", "consensus-validation-groups.tsv")
```

```{r}
# source summarize_celltypes() function
setup_functions <- file.path(module_base, "exploratory-notebooks", "utils", "setup-functions.R")
source(setup_functions)
```

```{r, message=FALSE}
# read in consensus and gene expression results 
consensus_df <- consensus_results_files |> 
  purrr::map(readr::read_tsv) |>
  dplyr::bind_rows()

gene_exp_df <- gene_exp_files |> 
  purrr::map(readr::read_tsv) |>
  dplyr::bind_rows()

# read in validation markers 
markers_df <- readr::read_tsv(validation_markers_file) |> 
  dplyr::mutate(
    plot_label = glue::glue("{gene_symbol}-{validation_group_annotation}")
  )

validation_groups_df <- readr::read_tsv(validation_groups_file) |> 
  # rename final assigned group to avoid conflicts when merging in marker gene expression 
  # we want to separate the marker gene group from the actual cell type annotation
  dplyr::select(consensus_annotation, assigned_group = validation_group_annotation)
```

```{r}
# combine into one data frame 
combined_df <- consensus_df |> 
  dplyr::left_join(validation_groups_df, by = c("consensus_annotation")) |> 
  dplyr::left_join(gene_exp_df, by = c("barcodes", "library_id")) |> 
  # account for the same gene being present in multiple cell types 
  dplyr::left_join(markers_df, by = c("ensembl_gene_id"), relationship = "many-to-many") |> 
  dplyr::rowwise() |> 
  dplyr::mutate(
    detected = logcounts > 0 # column indicating if gene is present or not
  )

# make a data frame that just has the unique genes 
unique_gene_df <- combined_df |> 
  dplyr::filter(gene_observed_count == 1)
```

```{r}
# get total number of cells per final annotation group 
total_cells_df <- unique_gene_df |> 
  dplyr::select(library_id, barcodes, assigned_group) |> 
  unique() |> 
  dplyr::count(assigned_group, name = "total_cells")

# get total number of cells each gene is detected in per group 
# and mean gene expression per group 
group_stats_df <- unique_gene_df |> 
  dplyr::group_by(assigned_group, ensembl_gene_id) |>
  dplyr::summarize(
    detected_count = sum(detected),
    mean_exp = mean(logcounts)
  )

gene_summary_df <- group_stats_df |> 
  # add total cells
  dplyr::left_join(total_cells_df, by = c("assigned_group")) |> 
  dplyr::filter(total_cells > 50) |> 
  # add marker gene groups
  dplyr::left_join(markers_df, by = c("ensembl_gene_id"), relationship = "many-to-many") |> 
  dplyr::rowwise() |> 
  dplyr::mutate(
    # get total percent
    percent_exp = (detected_count/total_cells) * 100,
    # account for NA/unknowns and set axes order
    assigned_group = dplyr::if_else(is.na(assigned_group), "unknown", assigned_group) |> 
      factor(levels =  c(unique(markers_df$validation_group_annotation), "unknown"))
  ) 
```

```{r}
# filter out low expressed genes
dotplot_df <- gene_summary_df |> 
  dplyr::filter(mean_exp > 0, percent_exp > 5) |>
  dplyr::arrange(assigned_group) |> 
  dplyr::mutate(y_label = as.factor(glue::glue("{assigned_group} ({total_cells})")))

# pull out cell types to specify legend and x-axis order
celltype_groups <- dotplot_df |> 
  dplyr::pull(assigned_group) |> 
  unique() |>
  as.character()

# specify x axis order
marker_gene_order <- markers_df |> 
  dplyr::filter(validation_group_annotation %in% celltype_groups,
                gene_symbol %in% dotplot_df$gene_symbol) |> 
  dplyr::pull(gene_symbol)

# remove marker genes that aren't present in final annotations and set x axis order 
dotplot_df <- dotplot_df |> 
  dplyr::filter(validation_group_annotation %in% celltype_groups) |> 
  dplyr::mutate(
    gene_symbol = factor(gene_symbol, levels = marker_gene_order)
  )

dotplot <- ggplot(dotplot_df, aes(y = forcats::fct_rev(y_label), x = gene_symbol, color = mean_exp, size = percent_exp)) +
  geom_point() +
  scale_color_viridis_c(option = "magma") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
  ) +
  labs(
    x = "",
    y = "Cell type annotation"
  )

color_bar <- ggplot(dotplot_df, aes(x = gene_symbol, y = 1, fill = validation_group_annotation)) + 
  geom_tile() + 
  scale_fill_brewer(palette = "Dark2", breaks = celltype_groups) +
  #scale_fill_manual(values = color_palette, breaks = celltype_groups) +
  ggmap::theme_nothing() +
  theme(legend.position = "bottom") +
  labs(fill = "")

dotplot + color_bar +
  patchwork::plot_layout(ncol = 1, heights = c(4, 0.1))

```

```{r}
# note that we can't use the same heatmap function as before since we are looking at cell types as rows rather than gene sets
# we also want to specify the annotation name

# get annotation colors for each of the marker gene types 
filtered_markers_df <- markers_df |> 
  dplyr::filter(validation_group_annotation %in% celltype_groups,
                gene_symbol %in% gene_summary_df$gene_symbol)

# first make a mtx to use for the heatmap with rows as cell types and genes as columns 
heatmap_mtx <- gene_summary_df |> 
  dplyr::select(gene_symbol, assigned_group, mean_exp) |> 
  dplyr::filter(gene_symbol %in% filtered_markers_df$gene_symbol) |> 
  tidyr::pivot_wider(
    names_from = gene_symbol,
    values_from = mean_exp
  ) |> 
  tibble::column_to_rownames("assigned_group") |>
  as.matrix()
# make sure cell types are present in the right order
heatmap_mtx <- heatmap_mtx[celltype_groups, filtered_markers_df$gene_symbol]

marker_gene_categories <- unique(filtered_markers_df$validation_group_annotation)
num_categories <- length(marker_gene_categories)
category_colors <- palette.colors(palette = "Dark2") |>
  head(n = num_categories) |>
  purrr::set_names(marker_gene_categories)

# create annotation for heatmap
annotation <- ComplexHeatmap::columnAnnotation(
  category = filtered_markers_df$validation_group_annotation,
  col = list(
    category = category_colors
  ),
  annotation_legend_param = list(
      title = "Marker gene category"
    )
)

ComplexHeatmap::Heatmap(
    heatmap_mtx,
    # set the color scale based on min and max values
    col = circlize::colorRamp2(seq(min(heatmap_mtx), max(heatmap_mtx), length = 2), colors = c("white", "#00274C")),
    border = TRUE,
    ## Row parameters
    cluster_rows = FALSE,
    row_title = "",
    row_title_side = "left",
    row_names_side = "left",
    row_dend_side = "right",
    row_names_gp = grid::gpar(fontsize = 10),
    ## Column parameters
    cluster_columns = FALSE,
    show_column_names = TRUE,
    column_names_gp = grid::gpar(fontsize = 8),
    top_annotation = annotation,
    heatmap_legend_param = list(
      title = "Mean expression"
    )
  )
```

Repeat the plots above but use mean expression of all marker genes across all cells

```{r}
# get total number of cells each gene is detected in per group 
# and mean gene expression per group 
group_stats_df <- combined_df |> 
  dplyr::group_by(assigned_group, ensembl_gene_id) |>
  dplyr::summarize(
    detected_count = sum(detected),
    mean_exp = mean(logcounts),
    .groups = "drop"
  ) |> 
  # add in validation group for marker genes
  dplyr::left_join(markers_df, by = c("ensembl_gene_id")) |>
  dplyr::group_by(assigned_group, validation_group_annotation) |> 
  dplyr::mutate(
    group_mean_exp = mean(mean_exp),
    group_detected_count = mean(detected_count),
    .groups = "drop"
  ) |> 
  dplyr::select(assigned_group, validation_group_annotation, group_mean_exp, group_detected_count) |>
  unique()

gene_summary_df <- group_stats_df |> 
  # add total cells
  dplyr::left_join(total_cells_df, by = c("assigned_group")) |> 
  dplyr::filter(total_cells > 50) |> 
  # add per gene stats
  #dplyr::left_join(group_stats_df, by = c("ensembl_gene_id", "assigned_group")) |> 
  #dplyr::left_join(markers_df, by = c("ensembl_gene_id"), relationship = "many-to-many") |> 
  dplyr::rowwise() |> 
  dplyr::mutate(
    # get total percent
    percent_exp = (group_detected_count/total_cells) * 100,
    # account for NA/unknowns and set axes order
    assigned_group = dplyr::if_else(is.na(assigned_group), "unknown", assigned_group) |> 
      factor(levels =  c(unique(markers_df$validation_group_annotation), "unknown"))
  ) 
```

```{r}
# filter out low expressed genes
dotplot_df <- gene_summary_df |> 
  dplyr::filter(group_mean_exp > 0, percent_exp > 5) |>
  dplyr::arrange(assigned_group) |> 
  dplyr::mutate(y_label = as.factor(glue::glue("{assigned_group} ({total_cells})")))

# pull out cell types to specify legend and x-axis order
celltype_groups <- dotplot_df |> 
  dplyr::pull(assigned_group) |> 
  unique() |>
  as.character()

# remove marker genes that aren't present in final annotations and set x axis order 
dotplot_df <- dotplot_df |> 
  dplyr::filter(validation_group_annotation %in% celltype_groups) |>
  dplyr::mutate(
    validation_group_annotation = factor(validation_group_annotation, levels = celltype_groups)
  )

dotplot <- ggplot(dotplot_df, aes(y = forcats::fct_rev(y_label), x = validation_group_annotation, color = group_mean_exp, size = percent_exp)) +
  geom_point() +
  scale_color_viridis_c(option = "magma") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
  ) +
  labs(
    x = "Marker gene group",
    y = "Cell type annotation"
  )

color_bar <- ggplot(dotplot_df, aes(x = validation_group_annotation, y = 1, fill = validation_group_annotation)) + 
  geom_tile() + 
  scale_fill_brewer(palette = "Dark2", breaks = celltype_groups) +
  #scale_fill_manual(values = color_palette, breaks = celltype_groups) +
  ggmap::theme_nothing() +
  theme(legend.position = "bottom") +
  labs(fill = "")

dotplot + color_bar +
  patchwork::plot_layout(ncol = 1, heights = c(4, 0.1))

```

```{r}
# get annotation colors for each of the marker gene types 
filtered_markers_df <- markers_df |> 
  dplyr::filter(validation_group_annotation %in% celltype_groups) |> 
  dplyr::select(validation_group_annotation) |> 
  unique()

# first make a mtx to use for the heatmap with rows as cell types and genes as columns 
heatmap_mtx <- gene_summary_df |> 
  dplyr::select(validation_group_annotation, assigned_group, group_mean_exp) |> 
  tidyr::pivot_wider(
    names_from = validation_group_annotation,
    values_from = group_mean_exp
  ) |> 
  tibble::column_to_rownames("assigned_group") |>
  as.matrix()
# make sure cell types are present in the right order
heatmap_mtx <- heatmap_mtx[celltype_groups, filtered_markers_df$validation_group_annotation]

marker_gene_categories <- unique(filtered_markers_df$validation_group_annotation)
num_categories <- length(marker_gene_categories)
category_colors <- palette.colors(palette = "Dark2") |>
  head(n = num_categories) |>
  purrr::set_names(marker_gene_categories)

# create annotation for heatmap
annotation <- ComplexHeatmap::columnAnnotation(
  category = filtered_markers_df$validation_group_annotation,
  col = list(
    category = category_colors
  ),
  annotation_legend_param = list(
      title = "Marker gene category"
    )
)

ComplexHeatmap::Heatmap(
    heatmap_mtx,
    # set the color scale based on min and max values
    col = circlize::colorRamp2(seq(min(heatmap_mtx), max(heatmap_mtx), length = 2), colors = c("white", "#00274C")),
    border = TRUE,
    ## Row parameters
    cluster_rows = FALSE,
    row_title = "",
    row_title_side = "left",
    row_names_side = "left",
    row_dend_side = "right",
    row_names_gp = grid::gpar(fontsize = 10),
    ## Column parameters
    cluster_columns = FALSE,
    show_column_names = TRUE,
    column_names_gp = grid::gpar(fontsize = 8),
    top_annotation = annotation,
    heatmap_legend_param = list(
      title = "Mean expression"
    )
  )
```

## Session info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```