### 2025-02-27 analyses/cell-type-wilms-tumor-06/07_combined_annotation_across_samples_exploration.Rmd
notebook after review

---
title: "Combined annotation exploration for SCPCP000006"
author: "Maud PLASCHKA"
date: "`r Sys.Date()`"
params:
  predicted.celltype.threshold: 0.75
  cnv_threshold_low: 0
  cnv_threshold_high: 2
  testing: 0
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warnings = FALSE,
  knitr.digits.signif = FALSE
)
```

## Introduction

The aim is to combine label transfer and CNV inference to annotate Wilms tumor samples in SCPCP000006.
The proposed annotation will be based on the combination of:

- the label transfer from the fetal kidney reference (Stewart et al.), in particular the `fetal_kidney_predicted.compartment` and `fetal_kidney:predicted.cell_type`, as well as the `prediction.score` for each compartment,

- the predicted CNV calculated using intra-sample endothelial and immune cells (`--reference both`) as normal reference; no reference was used for samples with fewer than 3 predicted normal cells.

In a second time, we will explore and validate the chosen annotation.

We will use some of the [markers genes](https://github.com/AlexsLemonade/OpenScPCA-analysis/tree/main/analyses/cell-type-wilms-tumor-06#the-table-celltype_metadatacsv-contains-the-following-column-and-information) to validate visually the annotations.

The analysis can be summarized as the following:

| first level annotation | second level annotation | selection of the cells                                                                        | marker genes for validation | CNV validation                             |
| ---------------------- | ----------------------- | --------------------------------------------------------------------------------------------- | --------------------------- | ------------------------------------------ |
| normal                 | endothelial             | `compartment == "endothelium" & predicted.score > pred.thr`                                   | `VWF`                       | no CNV                                     |
| normal                 | immune                  | `compartment == "immune" & predicted.score > pred.thr`                                        | `PTPRC`, `CD163`, `CD68`    | no CNV                                     |
| normal                 | kidney                  | `compartment == "fetal_nephron" & predicted.score > pred.thr & cnv_score < cnv.thr.lower`     | `CDH1`, `PODXL`, `LTL`      | no CNV                                     |
| normal                 | stroma                  | `compartment == "stroma" & predicted.score > pred.thr & cnv_score < cnv.thr.upper`            | `VIM`                       | no CNV                                     |
| cancer                 | stroma                  | `compartment == "stroma" & cnv_score > cnv.thr.upper`                                         | `VIM`                       | `proportion_cnv_chr: 1, 4, 11, 16, 17, 18` |
| cancer                 | blastema                | `compartment == "fetal_nephron" & cell_type == "mesenchymal cell" & cnv_score > cnv.thr.upper`| `CITED1`                    | `proportion_cnv_chr: 1, 4, 11, 16, 17, 18` |
| cancer                 | epithelial              | `compartment == "fetal_nephron" & cell_type != "mesenchymal cell" & cnv_score > cnv.thr.upper`| `CDH1`                      | `proportion_cnv_chr: 1, 4, 11, 16, 17, 18` |
| unknown                | -                       | the rest of the cells                                                                         | -                           | -                                          |

with the defined parameters:

- `pred.thr` = `r params$predicted.celltype.threshold`

- `cnv.thr.lower` = `r params$cnv_threshold_low`

- `cnv.thr.upper` = `r params$cnv_threshold_high`

### Packages

```{r packages, message=FALSE, warning=FALSE}
library(Seurat)
library(tidyverse)
library(patchwork)
library(DT)
```

### Base directories

```{r base paths, eval=TRUE}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current", "SCPCP000006")

# The path to this module
module_base <- file.path(repository_base, "analyses", "cell-type-wilms-tumor-06")

result_dir <- file.path(module_base, "results")
```

### Output files

The report will be saved in the `notebook` directory.
The final annotation file `SCPCP000006-annotations.tsv` will be saved in the `results` directory.

```{r}
# cell type annotation table
annotations_tsv <- file.path(result_dir, "SCPCP000006-annotations.tsv")
```

### Input files

In this notebook, we are working with all of the samples in SCPCP000006.

The sample metadata can be found in `sample_metadata_file` in the `data` folder.

```{r }
# sample metadata
sample_metadata_file <- file.path(repository_base, "data", "current", "SCPCP000006", "single_cell_metadata.tsv")
metadata <- read.table(sample_metadata_file, sep = "\t", header = TRUE)
```

We extracted from the pre-processed and labeled `Seurat` object from `results/{sample_id}/06_infercnv_HMM-i3_{sample_id}_reference-{reference}.rds`.
the following information per cell:

- the predicted compartment in `fetal_kidney_predicted.compartment` and `fetal_kidney_predicted.compartment.score`

- the predicted cell type in `fetal_kidney_predicted.cell_type` and `fetal_kidney_predicted.cell_type.score`

- the sample identifier in `sample_id`

- the CNV `{dupli|loss}` estimation per chromosome {i} and arm `{p|q}`: 
`has_{dupli|loss}_chr{i}{p|q}`, in 1 to 22

- the `counts` of [markers genes](https://github.com/AlexsLemonade/OpenScPCA-analysis/tree/main/analyses/cell-type-wilms-tumor-06#the-table-celltype_metadatacsv-contains-the-following-column-and-information)

```{r path_to_query}
# Extract from each Seurat object information that will be used
sample_ids <- metadata |>
  dplyr::filter(seq_unit != "spot") |>
  dplyr::pull(scpca_sample_id) |>
  unique()

# These samples were run with "none" as the reference
none_reference_samples <- c("SCPCS000177", "SCPCS000180", "SCPCS000181", "SCPCS000190", "SCPCS000197")

# Create a data frames of all annotations
cell_type_df <- sample_ids |>
  purrr::map(
    # For each sample_id, do the following:
    \(sample_id) {
      if (sample_id %in% none_reference_samples) {
        reference <- "none"
      } else {
        reference <- "both"
      }

      input_file <- file.path(
        result_dir,
        sample_id,
        glue::glue("06_infercnv_HMM-i3_{sample_id}_reference-{reference}.rds")
      )

      # The file may not be present if this is being run in CI, which is ok.
      # If we are not running in CI and the file doesn't exist, we should error out
      # We should error out if the file does not exist and we are NOT testing
      if (!file.exists(input_file)) {
        if (params$testing) {
          return(NULL)
        } else {
          stop("Input RDS file does not exist.")
        }
      }

      # Read in the Seurat object
      srat <- readRDS(input_file)

      # Create and return a data frame from the Seurat object with relevant annotations
      # this data frame will have four columns: barcode, sample_id, compartment, organ
      data.frame(
        srat@meta.data,

        # cell embedding
        umap = srat@reductions$umap@cell.embeddings,

        # marker genes
        PTPRC = FetchData(object = srat, vars = "ENSG00000081237", layer = "counts"),
        VWF = FetchData(object = srat, vars = "ENSG00000110799", layer = "counts"),
        VIM = FetchData(object = srat, vars = "ENSG00000026025", layer = "counts"),
        CITED1 = FetchData(object = srat, vars = "ENSG00000125931", layer = "counts"),
        CDH1 = FetchData(object = srat, vars = "ENSG00000039068", layer = "counts"),
        PODXL = FetchData(object = srat, vars = "ENSG00000128567", layer = "counts"),
        COL6A3 = FetchData(object = srat, vars = "ENSG00000163359", layer = "counts"),
        SIX2 = FetchData(object = srat, vars = "ENSG00000170577", layer = "counts"),
        NCAM1 = FetchData(object = srat, vars = "ENSG00000149294", layer = "counts"),
        THY1 = FetchData(object = srat, vars = "ENSG00000154096", layer = "counts")

      ) |>
        tibble::rownames_to_column("barcode") |>
        dplyr::mutate(sample_id = sample_id)
    }

  ) |>
  # now combine all dataframes to make one big one
  dplyr::bind_rows()
```

We will also use:

- the samples metadata or clinical data reported in `r file.path(repository_base, "data", "current", "SCPCP000006", "single_cell_metadata.tsv")`

```{r }
# Add clinical metadata
cell_type_df <- cell_type_df |>
  left_join( metadata, by= c("sample_id" = "scpca_sample_id"))

```

- the cytoband file to get the length of each chromosome arm (for plotting only)

```{r input}
# the cytoband file that will be used to extract the size of each chromosome arm for plotting only
arm_order_file <- file.path(module_base, "results", "references", "hg38_cytoBand.txt")
# Load cytoBand file
cytoBand <- read.table(gzfile(arm_order_file), header = FALSE, sep = "\t", stringsAsFactors = FALSE)
colnames(cytoBand) <- c("chrom", "start", "end", "band", "stain")
# Here we define the size of each chromosome arm that will be used for plotting of the mean CNV profile
  # Extract chromosome arm information
  cytoBand <- cytoBand |>
    mutate(arm = substr(band, 1, 1))  # Extract 'p' or 'q' from the band column

  # Define arm boundaries
  chromosome_arms <- cytoBand |>
    group_by(chrom, arm) |>
    summarize(
      Start = min(start),
      End = max(end),
      .groups = "drop"
    ) |>
    mutate(Size = End - Start,
           dupli = paste0("has_dupli_", chrom, arm),
           loss = paste0("has_loss_", chrom, arm),
           chrom_arm = glue::glue("{chrom}{arm}")) |> 
    mutate(Size = Size / sum(Size)) |>
    # Define chromosome arm order
    mutate(chrom_arm = factor(chrom_arm, levels = c(paste0("chr", rep(1:22, each = 2), c("p", "q")),
                                                  "chrXp", "chrXq", "chrYp", "chrYq"))) |>
    # Sort genes by Chromosome arm and Start position
    arrange(chrom_arm, Start)  |> 
    filter(arm != "")

```

## Functions

#### do_Feature_mean

`do_Feature_mean` shows heatmap of mean expression of a feature grouped by a metadata.

- `df` is the name of the table containing metadata and feature expression (counts) per cells
- `group.by` is the name of the metadata to group the cells
- `feature` is the name of the gene to average the expression

```{r fig.width=10, fig.height=4, out.width='100%'}
do_Feature_mean <- function(df, group.by, feature) {
  df <- df |>
    group_by(sample_id, !!sym(group.by)) |>
    summarise(m = mean(!!sym(feature)))

  p <- ggplot(df, aes(x = sample_id, y = !!sym(group.by), fill = m)) +
    geom_tile() +
    scale_fill_viridis_c() +
    theme_bw() +
    theme(text = element_text(size = 20)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) +
    guides(fill = guide_colourbar(title = paste0(feature)))

  return(p)
}
```

#### do_BoxPlot_mean

`do_BoxPlot_mean` shows boxplot of the percentage of cells with CNV in a specific chromosome arm per clinical relevant group (metadata)

- `df` is the name of the table containing metadata and feature expression (counts) per cells
- `group.by` is the name of the metadata to group the cells
- `comparison` is a list of factors in `group.by` for which we will compare the means using an unpaired Wilcoxon test

```{r fig.width=20, fig.height=15, out.width='100%', results='asis'}
do_BoxPlot_mean <- function(df, group.by, comparison){
  tmp <- df |>
    select(!!sym(group.by), sample_id, contains("has_loss_chr"), contains("has_dupli_chr")) |>
    group_by(!!sym(group.by), sample_id) |>
    summarise_all(.funs = mean) |>
    pivot_longer(cols = -c(sample_id,!!sym(group.by)) ) |>
    rename("chromosome_arm" = "name", "percentage" = "value") |>
    mutate(chromosome_arm = factor(chromosome_arm, levels = c(has_dupli_chr, has_loss_chr)))

  p <- ggplot(tmp, aes(x = !!sym(group.by), y = percentage, fill = !!sym(group.by))) +
  geom_boxplot()+
  scale_fill_brewer(palette="PuRd") +
  geom_dotplot(binaxis='y', stackdir='center') +
  facet_wrap(~chromosome_arm, nrow = 5 ) +
  theme_classic()

  return(p)
}

```

## Analysis

### Global CNV score

As done in `06_cnv_infercnv_exploration.Rmd`, we calculate  single CNV score and assess its potential in identifying cells with CNV versus normal cells without CNV.

We simply checked for each chromosome the cell `cnv_chr`.
Would the cell have less than `r params$cnv_threshold_low` chromosome with CNV, the global `cnv_score` will be FALSE.
Would the cell have more than `r params$cnv_threshold_high`  chromosome with CNV, the global `cnv_score` will be TRUE.
The default is set to `unknown`.

The `infercnv` method can lead to false positive CNV.
We introduced a flexibility of +`r params$cnv_threshold_high` over the `cnv_threshold` to reduce the risk of misclassification of normal cells as cancer cells.

Please note that some cancer cell might not have any CNV. 
There is thus a risk of misclassifying cancer cells as normal cells. 
However, we cannot avoid this risk with the data and tools currently available.
This is a limitation of our annotations.

```{r fig.width=10, fig.height=10, out.width='100%', results='asis'}
cell_type_df <- cell_type_df |>
  mutate(cnv_score = rowSums(cell_type_df[, grepl("has_loss_chr|has_dupli_chr", colnames(cell_type_df))], na.rm = TRUE)) |>
  mutate(cnv_score = case_when(
    cnv_score > params$cnv_threshold_high ~ "CNV",
    cnv_score <= params$cnv_threshold_low ~ "no CNV",
    .default = "unknown"
  ))

```

### First level annotation

At first, we like to indicate in the `first.level_annotation` if a cell is normal, cancer or unknown.

- _normal_ cells can be observe in all five compartments (`normal`,`endothelium`, `immune`, `stroma` or `fetal nephron`) and do not have CNV.
We only allow a bit of flexibility in terms of CNV profile for immune and endothelium cells that have a high predicted score.
Indeed, we know that false positive CNV can be observed in a cell type specific manner.
Please note that in `06_infercnv.R`, we renamed all `immune` and `endothelium` cells with a `fetal_kidney_predicted.compartment.score` > `r params$predicted.celltype.threshold` as `normal`.

The threshold used for the `predicted.score` is defined as a parameter of this notebook as `r params$predicted.celltype.threshold`.
The thresholds used for the identification of CNV are also defined as notebook parameters with a minimum of `r params$cnv_threshold_low` and a maximum of `r params$cnv_threshold_high`.

- _cancer_ cells are either from the `stroma` or `fetal nephron` compartments and must have at least few CNV

```{r fig.width=10, fig.height=10, out.width='100%', results='asis'}
# Define normal cells
# We first pick up the immune and endothelial cells annotated via the label transfer compartments under the condition that the predicted score is above the threshold
cell_type_df <- cell_type_df |>
  mutate(first.level_annotation = case_when(
    # assign normal/cancer based on condition
    fetal_kidney_predicted.compartment %in% c("fetal_nephron", "stroma") & 
    cnv_score == "no CNV" ~ "normal",

    fetal_kidney_predicted.compartment %in% c("normal", "immune", "endothelium") &
    fetal_kidney_predicted.compartment.score > params$predicted.celltype.threshold ~ "normal",

    fetal_kidney_predicted.compartment %in% c("fetal_nephron", "stroma") & 
    cnv_score == "CNV" ~ "cancer",

    .default = "unknown"
  ))

```

Using this basic strategy, we identified `r table(cell_type_df$first.level_annotation)["cancer"]` _cancer_ cells, `r table(cell_type_df$first.level_annotation)["normal"]` _normal_ cells.
`r table(cell_type_df$first.level_annotation)["unknown"]` cells remain _unknown_

```{r fig.width=20, fig.height=20, out.width='100%', results='asis'}
ggplot(cell_type_df, aes(x = umap.umap_1, y = umap.umap_2, color = first.level_annotation)) +
  geom_point(size = 0.5, shape = 19, alpha = 0.5) +
  facet_wrap(facets = ~sample_id, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 22))

cell_type_df |>
  dplyr::count(first.level_annotation, sample_id) |>
  DT::datatable()

```

Strikingly, 5 samples show more normal cells than cancer cells:

- `SCPCS000177`
- `SCPCS000180`
- `SCPCS000181`
- `SCPCS000190`
- `SCPCS000197`

These five samples do not have enough immune and endothelium cell to be used as a normal reference in `scripts/06_infercnv.R` and we previously decided to run `scripts/06_infercnv.R` without any reference for these samples.
In that case, the mean expression profile over the sample is taken as the normal reference, biasing the inference of CNV.

For those samples, the annotations based on `infercnv` cannot be trust.
To avoid any confusion in the following steps of the analysis, we decided to force the annotation of the entire samples to `unknown`.

```{r fig.width=10, fig.height=10, out.width='100%', results='asis'}

cell_type_df$first.level_annotation[cell_type_df$sample_id %in% none_reference_samples] <- "unknown"

```

### Second level annotation

#### Normal cells

- Normal cells from the `fetal nephron` compartment must be normal kidney cells.

- Normal cells from the `stroma` compartment must be normal stroma cells.

- Immune and endothelial cells have been already identified by label transfer.

#### Cancer cells

Wilms tumor cancer cells can be:

- _cancer stroma_: We define as _cancer stroma_ all cancer cells from the stroma compartment.

- _blastema_,: we defined as _blastema_ every cancer cell that has a `fetal_kidney_predicted.cell_type == mesenchymal cell`.
We know that these _mesenchymal_ cells are cells from the cap mesenchyme that are not expected to be in a mature kidney.
These blastema cells should express higher _CITED1_ and/or _NCAM1_.

- _cancer epithelium_: we defined as _cancer epithelium_ all cancer cells that are neither stroma nor blastemal cells.
We expect these cells to express epithelial markers.
Their predicted cell type should correspond to more mature kidney epithelial subunits.

```{r fig.width=10, fig.height=10, out.width='100%', results='asis'}
cell_type_df <- cell_type_df |>
  mutate(second.level_annotation = case_when(
    # assign normal cells based on condition
      cell_type_df$fetal_kidney_predicted.compartment   == c("fetal_nephron") &
      cell_type_df$cnv_score                            == "no CNV" ~ "kidney",

      cell_type_df$fetal_kidney_predicted.compartment   == c("stroma") &
      cell_type_df$cnv_score                            == "no CNV" ~ "normal stroma",

      cell_type_df$fetal_kidney_predicted.compartment   == c("normal") &
      cell_type_df$fetal_kidney_predicted.cell_type     == c("endothelial cell") ~ "endothelium",

      cell_type_df$fetal_kidney_predicted.compartment   == c("normal") &
      cell_type_df$fetal_kidney_predicted.cell_type    %in% c("B cell",
                                    "CD4-positive, alpha-beta T cell",
                                    "conventional dendritic cell",
                                    "lymphocyte",
                                    "macrophage",
                                    "mast cell",
                                    "monocyte",
                                    "natural killer cell",
                                    "neutrophil",
                                    "plasmacytoid dendritic cell") ~ "immune",

    # assign cancer cells based on condition
      cell_type_df$fetal_kidney_predicted.compartment    == c("stroma") &
      cell_type_df$cnv_score                             == "CNV" ~ "cancer stroma",

      cell_type_df$fetal_kidney_predicted.compartment    == c("fetal_nephron") &
      cell_type_df$cnv_score                             == "CNV" &
      cell_type_df$fetal_kidney_predicted.cell_type      == "mesenchymal cell" ~ "blastema",

      cell_type_df$fetal_kidney_predicted.compartment    == c("fetal_nephron") &
      cell_type_df$cnv_score                             == "CNV" &
      cell_type_df$fetal_kidney_predicted.cell_type      != "mesenchymal cell" ~ "cancer epithelium",

      sample_id                                         %in% none_reference_samples ~ "unknown",

    .default = "unknown"
  ))

```

#### `Second.level_annotation` of cancer and normal cells - `umap` reduction

```{r fig.width=20, fig.height=20, out.width='100%', results='asis'}
ggplot(cell_type_df, aes(x = umap.umap_1, y = umap.umap_2, color = second.level_annotation)) +
  geom_point(size = 0.5, shape = 19, alpha = 0.5) +
  facet_wrap(facets = ~sample_id, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 22))
```

#### `Second.level_annotation` of cancer and normal cells without `unknown` - `umap` reduction

```{r fig.width=20, fig.height=20, out.width='100%', results='asis'}
cell_type_df_sub <- cell_type_df[cell_type_df$second.level_annotation != "unknown",]
ggplot(cell_type_df_sub, aes(x = umap.umap_1, y = umap.umap_2, color = second.level_annotation)) +
  geom_point(size = 0.5, shape = 19, alpha = 0.5) +
  facet_wrap(facets = ~sample_id, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 22))
```

Per sample, the annotations in the `umap` reduction seem to make sense as we can the three main Wilms tumor components are easily distinguished:

- cancer blastema
- epithelium
- stroma

### Validation cancer versus normal based on the CNV profile

We look for each `second.level_annotation` and `sample_id` at the proportion of cells that has CNV in each of the chromosome.

These heatmaps allow us to:

- validate that mostly tumor cells do present high percentages of cells with CNV (by definition)

- identify chromosome and cell types that might be more sensitive to _FALSE_ positive CNV, such as the _chr6p_ which is a hub for immune genes

- get the repartition per sample of the CNV

#### Loss on long arms

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in 1:22) {
  print(do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = glue::glue("has_loss_chr", i, "q")))
}
```

#### gains on long arms

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in 1:22) {
  print(do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = glue::glue("has_dupli_chr", i, "q")))
}
```

#### loss on short arms

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in c(1:12, 16:20)) {
  print(do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = glue::glue("has_loss_chr", i, "p")))
}
```

#### gains on short arms

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in c(1:12, 16:20)) {
  print(do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = glue::glue("has_dupli_chr", i, "p")))
}
```

### Mean CNV profile

While the above heatmaps can be informative, we have to acknowledge that there are numerous and it is difficult to get a global picture of the CNV profile of the entire dataset.

Here we aim to plot a mean CNV profile across the 35 annotated samples as presented in Figure 1A of [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). This represent the mean SNP Array profile of 64 pediatric kidney cancer (including 60 Wilms tumors).

Here we plot for each chromosome arm (x-axis) the number of samples having a `dupli` (y-axis > 0, in red) or a `loss` (y-axis < 0, in blue).

For a sample, a CNV is defined as `clonal` (opaque) if detected in more than 70% of the cells;
and `subclonal` (transparent) when detected in at least 10% of the cells.

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
# Here we define the vector of ordered chromosomes arms that will be used for plotting
has_dupli_chr <- names(cell_type_df)[stringr::str_starts(names(cell_type_df), "has_dupli_chr")]

# define the vector of ordered chromosomes arms
has_loss_chr <- names(cell_type_df)[stringr::str_starts(names(cell_type_df), "has_loss_chr")]

```

```{r fig.height=10, fig.width=20, message=FALSE, warning=FALSE, out.width='100%', results='asis'}
# prepare the table for the number of sample with a (sub)clonal loss

CNV_profile_loss <- cell_type_df_sub |>
                    select(sample_id, contains("has_loss_chr")) |>
                    group_by(sample_id) |>
                    summarise_all(.funs = mean) |>
                    pivot_longer(
                      cols = -sample_id, 
                      names_to = "chromosome_arm", 
                      values_to = "percentage") |>
                    left_join(chromosome_arms, by = c("chromosome_arm"="loss")) |>
                    mutate(chromosome_arm = factor(chromosome_arm, levels = has_loss_chr)) |>
                    mutate(
                      absolute_70 = ifelse(percentage > 0.7, -1, 0), 
                      absolute_10 = ifelse(percentage > 0.1, -1, 0))

# prepare the table for the number of sample with a (sub)clonal gain

CNV_profile_gain <- cell_type_df_sub |>
                    select(sample_id, contains("has_dupli_chr")) |>
                    group_by(sample_id) |>
                    summarise_all(.funs = mean) |>
                    pivot_longer(
                      cols = -sample_id, 
                      names_to = "chromosome_arm", 
                      values_to = "percentage") |>
                    left_join(chromosome_arms, by = c("chromosome_arm"="dupli")) |>
                    mutate(chromosome_arm = factor(chromosome_arm, levels = has_dupli_chr)) |>
                    mutate(
                      absolute_70 = ifelse(percentage > 0.7, 1, 0), 
                      absolute_10 = ifelse(percentage > 0.1, 1, 0))

# plot
# Here we chose to loop on chromosome arms to then be able to scale each bar to the length of the corresponding genomic region using wrap_plot. 
# Of note, the resulting plot is way better than the one obtained with facet_wrap or facet_grid so far.

loss_profile <- list()
for(chr in unique(CNV_profile_loss$chromosome_arm)){
  loss_profile[[chr]] <- ggplot(CNV_profile_loss[CNV_profile_loss$chromosome_arm == chr,], 
                                aes(x = chromosome_arm, y = absolute_10)) +
                        geom_bar(stat = "identity",  fill = "#41b6c4", alpha = 0.4) +
                        geom_bar(aes(x = chromosome_arm, y = absolute_70), stat = "identity",  fill = "#41b6c4") +
                        scale_x_discrete(labels = gsub("has_loss_", "", chr))  +
                        theme_void() +
                        ylim(c(-40,0)) +
                        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 12), text = element_text(size = 15))
}
loss_profile <- wrap_plots(loss_profile, nrow = 1, widths = chromosome_arms$Size[chromosome_arms$loss %in% unique(CNV_profile_loss$chromosome_arm)], guides = "collect")

gain_profile <- list()
for(chr in unique(CNV_profile_gain$chromosome_arm)){
  gain_profile[[chr]] <- ggplot(CNV_profile_gain[CNV_profile_gain$chromosome_arm == chr,], aes(x = chromosome_arm, y = absolute_10)) +
  geom_bar(stat = "identity",  fill = "#ce1256", alpha = 0.4) +
  geom_bar(aes(x = chromosome_arm, y = absolute_70), stat = "identity",  fill = "#ce1256") +
  theme_void() +
  ylim(c(0,40)) +
  theme(axis.text.x = element_blank())
}
gain_profile <- wrap_plots(gain_profile, nrow = 1, widths = chromosome_arms$Size[chromosome_arms$dupli %in% unique(CNV_profile_gain$chromosome_arm)], guides = "collect")

# combined plot
wrap_plots(gain_profile, loss_profile, ncol = 1) + plot_layout( widths = CNV_profile_gain$Size)

```

Comparing with the profile in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1), we identified important CNV known to play a role in Wilms tumor, including:

- _chr8q_, _chr12_, _chr13_, _chr18q_ gains

- _chr11q_, _chr16_, _chr17q_ loss

However, we spotted some differences:

- we couldn't detect any _chr4q_ loss as described in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1).

- we identified 2 losses in _chr5_ and _chr6_ that do not show up in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). 
These can be `infercnv` induced _FALSE_positive CNV.

Our major concern is the pattern of the _chr1_. 
We do see a tendency of _chr1p_ loss and _chr1q_ gain but it is not as pronounced as in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). 
_chr1q_ gain is however one of the most promising predictive marker of Wilms tumor outcome ([Nelson et al, Curr Opin Pediatr (2020)](https://pmc.ncbi.nlm.nih.gov/articles/PMC7919941/)).
It is however important to notice that the profile shown in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1) derived from SIOP samples that have been pre-treated with chemotherapy, while samples in the dataset are enriched in `upfront resection` samples that haven't been pre-treated, according to the COG protocol. 
So far, we do not know how the pre-operative chemotherapy impact the clonal selection of specific CNV.

### Mean of the percentage of cells with CNV in a specific chromosome arm per clinical relevant group

#### `vital_status`

As described in [Nelson et al, Curr Opin Pediatr (2020)](https://pmc.ncbi.nlm.nih.gov/articles/PMC7919941/), 
[Phelps et al, Children (2018)](https://pmc.ncbi.nlm.nih.gov/articles/PMC6262554/) and 
[Zheng et al, Front. Oncol (2023)](https://www.frontiersin.org/journals/oncology/articles/10.3389/fonc.2023.1137346/full), 
LOH at _chr1p_, _chr11p_ and _chr16q_ as well as _chr1q_ gain are prognostic factors.

Here we aim to see if we identify CNV associated with poorer prognosis.

```{r fig.height=15, fig.width=20, message=FALSE, warning=FALSE, out.width='100%', results='asis'}

do_BoxPlot_mean(df = cell_type_df_sub, group.by = "vital_status", comparison = c("Alive", "Expired") )

```

In our study, we validate the association of _chr1q_ gain, _chr11q_ loss and _chr14q_ loss with poorer prognosis.

##### `treatment`

As we don't know the effect of the treatment on the selection of cells with CNV, it is difficult to compare our CNV profile with the one in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1).
We those check the repartition of CNV in samples that have been pre-treated (i.e `Resection post chemotherapy`) with samples that have been removed surgically before any treatment (i.e `Upfront resection`).

```{r fig.height=15, fig.width=20, message=FALSE, warning=FALSE, out.width='100%', results='asis'}

do_BoxPlot_mean(df = cell_type_df_sub, group.by = "treatment", comparison = c("Upfront resection", "Resection post chemotherapy") )

```

Our analysis suggest that there is indeed an enrichment in cells with CNV after chemotherapy, 
including a tendency of more _chr1q_ gain samples after chemotherapy. 
This could explain some of the differences observed between our mean CNV profile and the one published in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1).

### Validation of second level annotation using marker genes

Finally, we aim to validate the `second level annotation` using marker genes.

#### Immune, `PTPRC` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000081237")
```

Immune cells are well annotated based on _PTPRC_ expression.

#### Endothelium, `VWF` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000110799")
```

Endothelial cells are well annotated based on _VWF_ expression.

#### Stroma

(*) To the best of our knowledges, there is no Wilms tumor specific marker of stroma cells.
They should express in some extend similar markers as normal stromal cells including _VIM_, _THY1_ and _COL6A3_.

##### `Vimentin` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000026025")
```

_VIM_ do not seem to be neither specific nor universal.

##### `COL6A3` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000163359")
```

_COL6A3_ expression is higher in Wilms tumor stroma cells and expressed across a wider range of samples.

##### `THY1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000154096")
```

_THY1_ do not seem to be neither specific nor universal.

#### Blastema

(*) To the best of our knowledges, there is no Wilms tumor specific and universel (i.e. expressed by all cells in every sample) marker of blastema cells.
We expect however blastema cells to express higher levels of stemness markers such as _CITED1_, _SIX1/2_, _NCAM1_.

##### `CITED1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000125931")
```

Blastema cells are enriched in _CITED1_+ cells, but some samples do not express it at all.

##### `NCAM1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000149294")
```

_NCAM1_ seems to be more cancer specific and broadly expressed among samples.

##### `SIX2` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000170577")
```

#### Epithelium

##### `CDH1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000039068")
```

As expected, epithelial cells express higher expression of _CDH1_.

##### `PODXL` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000128567")
```

## Create annotation table for export

This section creates the cell type annotation table for export.

```{r}
annotations_table <- cell_type_df |>
  dplyr::select(
    cell_barcode = barcode,
    scpca_sample_id = sample_id,
    tumor_cell_classification = first.level_annotation,
    cell_type_assignment = second.level_annotation
  ) |>
  mutate(
    # change cancer --> tumor, but keep the other labels
    tumor_cell_classification = ifelse(
      tumor_cell_classification == "cancer", "tumor", tumor_cell_classification
    ),
    cell_type_assignment = str_replace_all(
      cell_type_assignment,
      "cancer ",
      "tumor "
    )
  )

write_tsv(annotations_table, annotations_tsv)
```

Confirm how many samples we have annotations for:
```{r}
length(unique(annotations_table$scpca_sample_id))
```

## Conclusion

- Combining label transfer and CNV inference we have produced draft annotations for 35/40 Wilms tumor samples in SCPCP000006.
We decided not to annotate samples for which the `infercnv` couldn't be run with a normal reference, as we don't trust the results in that case and want to avoid misclassification (`SCPCS000177`, `SCPCS000180`, `SCPCS000181`, `SCPCS000190`, `SCPCS000197`).

- The mean CNV profile of the Wilms tumor cohort seems reasonable in regards to the one published in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). 
We identify some commonalities but also some divergence. 
They can be due to technical limitation of our CNV inference but can also reflect some specificities of this cohort (e.g. pre/post treatment samples).
It would be really useful to compare for some sample the `infercnv` result to a ground truth such as SNP Array (to be requested).

- The heatmaps of CNV proportion and marker genes support our annotations, but signals with some marker genes are very low.
Also, there is no universal marker for each entity of Wilms tumor that cover all tumor cells from all patient.
This makes the validation of the annotations quite difficult.

- However, we could try to take the problem from the other side, and used the current annotation to perform differential expression analysis and try to find marker genes that are consistent across patient and Wilms tumor histologies.

- In each histology (i.e. epithelial and stroma), the distinction between cancer and non cancer cell is difficult (as expected).
In this analysis, we suggested to rely on the CNV score to assess the normality of the cell.
Here again, we could try to run differential expression analysis and compare epithelial (resp. stroma) cancer versus non-cancer cells across patient, aiming to find a share transcriptional program allowing the classification cancer versus normal.

- In our annotation, we haven't taken into account the favorable/anaplastic status of the sample.
However, as anaplasia can occur in every (but do not has to) Wilms tumor histology, I am not sure how to integrate the information into the annotation.

- This notebook could be finally rendered using different parameters, i.e. threshold for the CNV score and predicted score to use.

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```