### 2025-03-20 analyses/infercnv-consensus-cell-type/exploratory-notebooks/01_ewings-consensus-cell-types.Rmd
notebooks -> exploratory-notebooks

---
title: "Explore immune cell types in Ewing sarcoma"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output: html_notebook
---

## Introduction

The goal of this notebook is to explore the distribution of immune consensus cell types in Ewing sarcoma samples (`SCPCP000015`).

Based on these findings, we can determine approaches to evaluate for pooling immune cells to use as a normal reference for `inferCNV`.

## Setup

```{r}
options(readr.show_col_types = FALSE)
```

### Paths

#### Base directories

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current results directory, found within the repository base directory
result_dir <- file.path(repository_base, "data", "current", "results")

# Path to cell-type-ewings results, which also contains consensus annotations
ewings_dir <- file.path(result_dir, "cell-type-ewings", "SCPCP000015")

# Path to immune cell list
immune_url <- "https://raw.githubusercontent.com/AlexsLemonade/OpenScPCA-analysis/refs/heads/main/analyses/cell-type-consensus/references/consensus-immune-cell-types.tsv"
```

#### Input and output files

Set paths to input and output directories and files in the chunk below.

```{r paths}
# Input files
celltype_files <- list.files(
  path = ewings_dir,
  pattern = "_ewing-celltype-assignments\\.tsv$",
  recursive = TRUE,
  full.names = TRUE
)
```

## Analysis

```{r}
immune_celltypes <- readr::read_tsv(immune_url) |>
  dplyr::pull(consensus_annotation)

celltype_df <- celltype_files |>
  purrr::map(readr::read_tsv) |>
  purrr::list_rbind() |>
  dplyr::select(library_id, barcodes, ewing_annotation, consensus_annotation) |>
  dplyr::mutate(
    category = dplyr::case_when(
      consensus_annotation == "Unknown" ~ "Unknown",
      consensus_annotation %in% immune_celltypes ~ "immune",
      .default = "other"
    )
  )
```

```{r}
category_count_df <- celltype_df |>
  dplyr::count(library_id, category) |>
  tidyr::pivot_wider(
    names_from = category,
    values_from = n,
    values_fill = 0
  )

# this way we can print the full table at once; RMarkdown will only show 10 rows at once
knitr::kable(category_count_df)
```
It seems most immune cells are concentrated in 2-3 samples, and most samples don't have any immune cells.

How many immune cells are there total?

```{r}
total_immune <- sum(category_count_df$immune)
total_immune
```

`inferCNV` authors [recommend that at least 100 cells](https://github.com/broadinstitute/infercnv/issues/188) be used for a "normal" reference, so we are above that threshold.

If we were to use all of these, what fraction is that per sample?
We'll just divide the total number of cells by this count, not taking into consideration how many immune cells are actually in that sample, to get a rough sense of how many immune cells there are relative to all.

```{r}
celltype_df |>
  dplyr::count(library_id, name = "total_cells") |>
  dplyr::mutate(fraction_immune = round(total_immune / total_cells, 2)) |>
  knitr::kable()
```

For the low-quality sample `SCPCL001111` this fraction isn't very meaningful since there are fewer cells in the sample than there are pooled immune cells.
For all other samples, the fraction ranges from 0.03 - 0.13.

Let's look a bit more at the breakdown of immune cell types:

```{r}
celltype_df |>
  dplyr::filter(category == "immune") |>
  dplyr::count(library_id, consensus_annotation) |>
  # this will arrange columns (after pivoting) in descending order of total counts
  dplyr::arrange(desc(n)) |>
  tidyr::pivot_wider(
    names_from = consensus_annotation,
    values_from = n,
    values_fill = 0
  )
```

Let's look at the total number of macrophages and all types of T cell, since these are the dominant cell types here.

```{r}
celltype_df |>
  dplyr::mutate(
    consensus_annotation = ifelse(
      stringr::str_detect(consensus_annotation, "T cell"),
      "T cell type",
      consensus_annotation
    )
  ) |>
  dplyr::filter(consensus_annotation %in% c("T cell type", "macrophage")) |>
  dplyr::count(consensus_annotation)
```
These constitute the vast majority of cells, so it would also be an option to use just these groups of cell types.
If we were to want a single immune cell type, then macrophage is the only suitable option to get more than 100 cells.

## Compare to ewings cell typing

We should also check whether any of the immune cells were classified as a tumor cell in the `cell-type-ewings` analysis module.
Ideally this will not have been the case, but it's an important thing to check!

```{r}
celltype_df |>
  dplyr::filter(category == "immune") |>
  dplyr::count(ewing_annotation) |>
  dplyr::filter(stringr::str_detect(ewing_annotation, "tumor"))
```

Well, it appears there are some consensus immune cells which were classified as `tumor`. 
Which samples are these in?

```{r}
celltype_df |>
  dplyr::filter(
    stringr::str_detect(ewing_annotation, "tumor"),
    category == "immune"
  ) |>
  dplyr::count(library_id, ewing_annotation, consensus_annotation)
```

These results show macrophage and T cells, which are the two most prevalent categories immune cell categories.
We might want to investigate these cells further, but at least we should probably exclude them as "normal" cells for `inferCNV` since their true status is bit unclear.

## Conclusions

* There are sufficient immune cells to pool together to run `inferCNV` specifying them as a normal reference
* We could also consider using only `macrophage` cells, or combined `macrophage` and all `T cell` types together, if we wanted to assess any potential effect of limit the types of immune cells feeding into `inferCNV`. 
* There are 17 cells which are annotated as an immune consensus cell type (specifically, macrophage and mature T cell) which the `cell-type-ewings` module annotated as tumor cells.
We would probably want to exclude these cells from any normal reference.

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```