### 2025-06-04 analyses/infercnv-consensus-cell-type/exploratory-notebooks/SCPCP000004/01_nb-consensus-cell-types.Rmd
add exploratory notebook for NB normal cells to determine reference options

---
title: "Explore normal cell types in Neuroblastoma samples"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    code_folding: hide
---

## Introduction

The goal of this notebook is to explore the distribution of normal consensus cell types, specifically immune and endothelial, in the Neuroblastoma tumor (i.e., _not_ the PDX) samples from `SCPCP000004`.

Based on these findings, we can determine approaches to evaluate for pooling  cells to use as a normal reference for `inferCNV`.
We can also determine whether any samples can be run with an internal (non-pooled) reference as a control.

## Setup

```{r}
options(readr.show_col_types = FALSE)
library(ggplot2)
theme_set(theme_bw())
```

### Paths

#### Base directories

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The current data directory, found within the repository base directory
data_dir <- file.path(repository_base, "data", "current", "SCPCP000004")

# The current results directory, found within the repository base directory
result_dir <- file.path(repository_base, "data", "current", "results")

# Path to cell-type-consensus results
consensus_dir <- file.path(result_dir, "cell-type-consensus", "SCPCP000004")
```

#### Input and output files

```{r paths}
# Input consensus cell type files
celltype_files <- list.files(
  path = consensus_dir,
  pattern = "_processed_consensus-cell-types\\.tsv\\.gz$",
  recursive = TRUE,
  full.names = TRUE
) |>
  # set names as library id
  purrr::set_names(
    \(x) {
      stringr::str_split_i(basename(x), "_", 1)
    }
  )

# project sample metadata file
metadata_file <- file.path(data_dir, "single_cell_metadata.tsv")

# Path to all possible immune consensus cell types
immune_url <- "https://raw.githubusercontent.com/AlexsLemonade/OpenScPCA-analysis/refs/heads/main/analyses/cell-type-consensus/references/consensus-immune-cell-types.tsv"
```

We'll also define vectors for the consensus cell types to check for.

```{r}
immune_celltypes <- readr::read_tsv(immune_url) |>
  dplyr::pull(consensus_annotation)

# Vector of all possible endothelial cell types
endo_celltypes <- c(
  "endothelial cell",
  "blood vessel endothelial cell",
  "microvascular endothelial cell",
  "pericyte"
)
```

## Analysis

```{r}
# determine PDX libraries to exclude
tumor_library_ids <- readr::read_tsv(metadata_file) |>
  dplyr::filter(!is_xenograft) |>
  dplyr::pull(scpca_library_id)

# Prepare data frame of all cell types
celltype_df <- celltype_files |>
  # consider only non-PDX
  purrr::keep_at(tumor_library_ids) |>
  purrr::map(readr::read_tsv) |>
  purrr::list_rbind() |>
  dplyr::select(library_id, barcodes, consensus_annotation) |>
  dplyr::mutate(
    category = dplyr::case_when(
      consensus_annotation %in% immune_celltypes ~ "immune",
      consensus_annotation %in% endo_celltypes ~ "endo",
      .default = "other"
    )
  )
```

The table below shows counts of endothelial, immune, and all other cells across libraries:

```{r}
category_count_df <- celltype_df |>
  dplyr::count(library_id, category)

# format for viewing
category_count_df |>
  tidyr::pivot_wider(
    names_from = category,
    values_from = n,
    values_fill = 0
  ) |>
  # order columns
  dplyr::select(library_id, endo, immune, other) |>
  # this way we can print the full table at once; RMarkdown will only show 10 rows at once
  knitr::kable()
```

Looks like we've got a lot of normal cells here, and plenty of libraries we can run with an internal reference. 
Since we have a lot of immune and endothelial cells on their own, we don't need to consider a combined reference for this project.

There are also some libraries with a very small number of cells which are presumably lower-quality and should not be included in any reference.

First, we'll find any libraries that have fewer than 1000 cells and exclude them.

```{r}
discard_libraries <- category_count_df |>
  dplyr::group_by(library_id) |>
  dplyr::summarize(total_cells = sum(n)) |>
  dplyr::filter(total_cells < 1000) |>
  dplyr::pull(library_id)

discard_libraries
```

How many cells would be in each reference for `endo` and `immune` groups?

```{r}
category_count_df <- category_count_df |>
  dplyr::filter(
    category != "other", # we dont need these counts anymore
    !(library_id %in% discard_libraries)
  )

total_ref_cells_df <- category_count_df |>
  dplyr::group_by(category) |>
  dplyr::summarize(total_cells = sum(n))
total_ref_cells_df
```
Well, that's quite a lot of cells!
Let's visualize which libraries these cells are spread across.
The barplot below shows the proportion of cells in the given category from each library.

```{r}
category_prop_df <- category_count_df |>
  dplyr::inner_join(total_ref_cells_df) |>
  dplyr::mutate(prop_total_ref_cells = n / total_cells)

ggplot(category_prop_df) +
  aes(x = library_id, y = prop_total_ref_cells, fill = category) +
  geom_col(position = "dodge") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))
```
The plurality of endothelial cells are from `SCPCL000125`, and the plurality of immune cells are from both `SCPCL000126` and `SCPCL000127`.
These libraries dominate their respective reference cell categories, but cells are still reasonably spread across the other libraries.

Next, let's look more closely at the counts for cell types within these broad categories of `immune` and `endo`, since we may have enough data here to narrow the reference down further:

```{r}
fine_celltype_df <- celltype_df |>
  dplyr::filter(category != "other") |>
  dplyr::count(consensus_annotation, category) |>
  dplyr::arrange(category)

# kable to print the full table at once
knitr::kable(fine_celltype_df)
```

* Endothelial cells are nearly exclusively plain `endothelial cell`
* The immune cells are pretty spread out, and we could have cell-type specific references created from anything with more than 100 cells:

```{r}
fine_celltype_df |>
  dplyr::filter(category == "immune", n > 100) |>
  dplyr::arrange(desc(n))
```
Dominant cell types are T cells (across several categories of T cell) and macrophages.

Finally, let's see which libraries could have an internal reference.
Generally `inferCNV` recommends at least 100 cells in a reference, but since there are so many normal cells here we can increase this threshold since 100 is really a lower bound.
So, for this project, we'll use 250 as a threshold to find which libraries/internal references are possible:

```{r}
category_count_df |>
  dplyr::filter(n >= 250) |>
  dplyr::group_by(library_id) |>
  dplyr::mutate(reference_options = paste(category, collapse = ", ")) |>
  dplyr::select(library_id, reference_options) |>
  unique() |>
  # kable to print the full table at once
  knitr::kable()
```

Notably, we have previously observed with our analysis of `SCPCP000015` in this module that references with immune cells tend to show more total CNV overlap with tumor cells, since immune cells are funky! 
So, it's nice to see that we enough data to be able to use a robust `endo` reference without needing to rely on `immune`, should similar CNV patterns emerge here.

## Conclusions

* We can _definitely_ have `endo` and `immune` pooled references, and the references won't be _strongly_ biased towards coming from a specific library
* Since we have such a high number of cells in each reference cell grouping in the first place, a combined reference (e.g. `endo-immune`) does not seem necessary
* We could also consider a subset of immune cell types for a reference, particularly macrophage and T cell, but a full immune reference seems reasonable to start with
* We have 11 libraries which we could use an internal reference with: 3 of them with `endo`, and _all_ of them with `immune`

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```