### 2025-07-02 analyses/cell-type-neuroblastoma-04/exploratory-notebooks/01_reference-aggregation.Rmd
Add exploratory Rmd notebook and utils functions for jaccard

---
title: "`r glue::glue('Compare aggregated and non-aggregated reference: {params$library_id}')`"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    code_folding: hide
params:
  sample_id: "SCPCS000105"
  library_id: "SCPCL000122"
---

This notebook does some preliminary comparisons of `SingleR` results obtained with the NBAtlas reference, where the reference was either aggregated before training or not.
The goal is to determine whether there are substantial differences between aggregation versions.
Note that the goal of this notebook is _not_ specifically to validate annotations, but to compare whether the aggregated vs non-aggregated reference yield different or similar labels.
There is, however, a brief comparison to consensus annotations.

## Setup

```{r, warning = FALSE}
options(readr.show_col_types = FALSE)
suppressPackageStartupMessages({
  library(ggplot2)
  library(patchwork)
  library(SingleCellExperiment)
})

theme_set(theme_bw())

# Define color ramp for shared use in the heatmaps
heatmap_col_fun <- circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue"))
# Set heatmap padding option
ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(0.6, "in"))
```

### Paths

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

module_dir <- file.path(repository_base, "analyses", "cell-type-neuroblastoma-04")
results_dir <- file.path(module_dir, "results", "singler")
data_dir <- file.path(repository_base, "data", "current", "SCPCP000004")

validation_url <- "https://raw.githubusercontent.com/AlexsLemonade/OpenScPCA-analysis/refs/heads/main/analyses/cell-type-consensus/references/consensus-validation-groups.tsv"
```

### Functions

```{r}
# Source Jaccard and heatmap utilities functions:
source(file.path(module_dir, "exploratory-notebooks", "jaccard-utils.R"))

# This function calculates and returns a confusion matrix
#
# df: data frame with columns of interest
# annotation_col1: first column to compare; arg should be a string
# annotation_col2: second column to compare; arg should be a string
calculate_confusion <- function(
    df,
    annotation_col1,
    annotation_col2) {
  # determine all possible annotation values for caret
  all_levels <- union(
    unique(df[[annotation_col1]]),
    unique(df[[annotation_col2]])
  )

  caret::confusionMatrix(
    factor(df[[annotation_col1]], levels = all_levels),
    factor(df[[annotation_col2]], levels = all_levels),
    mode = "sens_spec"
  )
}

# This function calculates Jaccard similarity between two vectors and plots a heatmap
# Code is based on this scpca-nf code:
# https://github.com/AlexsLemonade/scpca-nf/blob/4bb82aa635b572a62f2028dbec587fcfd2155e26/templates/qc_report/celltypes_supplemental_report.rmd#L50
#
# df: data frame with columns of interest
# annotation_col1: first column to compare; arg should be a string
# annotation_col2: second column to compare; arg should be a string
# label1: label to use in heatmap for annotation_col1 axis
# label2: label to use in heatmap for annotation_col2 axis
make_jaccard_heatmap <- function(
    df,
    annotation_col1,
    annotation_col2,
    label1,
    label2) {
  # Calculate jaccard matrix
  jaccard_matrix <- make_jaccard_matrix(
    df,
    annotation_col1,
    annotation_col2
  )

  create_single_heatmap(
    jaccard_matrix,
    row_title = label2,
    column_title = label1,
    keep_legend_name = label2,
    col_fun = heatmap_col_fun,
    # additional arguments
    column_names_rot = 90
  ) |>
    ComplexHeatmap::draw(heatmap_legend_side = "bottom")
}

# This function calculates and plots delta median for two groups of barcodes: those in agree and those not
# This code is based on this scpca-nf code:
# https://github.com/AlexsLemonade/scpca-nf/blob/4bb82aa635b572a62f2028dbec587fcfd2155e26/templates/qc_report/celltypes_supplemental_report.rmd#L626
#
# df: SingleR data frame with `scores.` columns
# agree_barcodes: Vector of barcodes that have agreeing labels between aggregated and not
# plot_title: Title for plot
plot_delta_median <- function(df, agree_barcodes, plot_title) {
  # extract scores into matrix
  scores_mat <- df |>
    dplyr::select(starts_with("scores.")) |>
    as.matrix()
  rownames(scores_mat) <- df$barcodes

  # Create data frame for plotting with delta median and the full *non-pruned* cell labels
  delta_median_df <- tibble::tibble(
    delta_median = rowMaxs(scores_mat) - rowMedians(scores_mat),
    # Need to grab the non-pruned label for this plot
    full_labels = aggr$labels,
    # if pruned.labels are NA ==> low confidence
    # so, negate for this variable:
    confident = !is.na(df$pruned.labels)
  ) |>
    dplyr::mutate(
      confident = ifelse(confident, "High-quality", "Low-quality"),
      barcodes = df$barcodes,
      annotations_agree = barcodes %in% agree_barcodes
    )

  ggplot(delta_median_df) +
    aes(x = annotations_agree, y = delta_median, color = confident) +
    # don't jitter the height
    geom_jitter(height = 0, alpha = 0.7) +
    theme(legend.position = "bottom") +
    ggtitle(plot_title)
}
```

### Prepare input data

Read SCE object so we can get UMAP and consensus cell types:

```{r}
sce_file <- file.path(
  data_dir,
  params$sample_id,
  glue::glue("{params$library_id}_processed.rds")
)

sce_df <- readRDS(sce_file) |>
  scuttle::makePerCellDF() |>
  tibble::as_tibble() |>
  dplyr::select(
    barcodes,
    consensus_annotation = consensus_celltype_annotation,
    UMAP1 = UMAP.1,
    UMAP2 = UMAP.2
  )
```

Read cell type data frames:
```{r}
validation_df <- readr::read_tsv(validation_url) |>
  dplyr::select(consensus_annotation, validation_group_annotation)

aggr <- readr::read_rds(
  file.path(
    results_dir,
    params$sample_id,
    "aggregated",
    glue::glue("{params$library_id}_singler-annotations.rds")
  )
) |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes") |>
  dplyr::mutate(ref_type = "aggregated")

non_aggr <- readr::read_rds(
  file.path(
    results_dir,
    params$sample_id,
    "non-aggregated",
    glue::glue("{params$library_id}_singler-annotations.rds")
  )
) |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes") |>
  dplyr::mutate(ref_type = "non-aggregated")
```

Join and prepare data for use:

```{r}
singler_df <- data.frame(
  barcodes = aggr$barcodes,
  "aggregated" = aggr$pruned.labels,
  "not_aggregated" = non_aggr$pruned.labels
)

sce_df <- sce_df |>
  dplyr::left_join(singler_df, by = "barcodes") |>
  dplyr::left_join(validation_df, by = "consensus_annotation") |>
  # set validation NA -> Unknown
  dplyr::mutate(
    validation_group_annotation =
      ifelse(is.na(validation_group_annotation), "Unknown", validation_group_annotation)
  )
```

## Aggregated vs. non-aggregated

First a quick and dirty query: What percent of annotations made with either the aggregated and non-aggregated reference versions agree?

```{r}
sum(singler_df$aggregated == singler_df$not_aggregated, na.rm = T) / nrow(singler_df)
```

Where specifically are their disagreements?
This table displays the number of cells that have the given cell label disagreements:

```{r}
singler_df |>
  dplyr::filter(aggregated != not_aggregated) |>
  dplyr::count(aggregated, not_aggregated) |>
  dplyr::arrange(desc(n))
```

### Delta median scores

We'll plot the `delta_median` scores for each `SingleR` inference (aggregated and non-aggregated) for cells where annotations agree vs. differ.
We would like to see that scores will be _lower_ for the cells where inferences between reference versions disagree.
Cells will also be colored by their reliability as determined by `SingleR`.

```{r}
agree_barcodes <- singler_df |>
  dplyr::filter(aggregated == not_aggregated) |>
  dplyr::pull(barcodes)

plot_delta_median(aggr, agree_barcodes, plot_title = "Aggregated delta_median") +
  plot_delta_median(non_aggr, agree_barcodes, plot_title = "Non-aggregated delta_median") &
  theme(legend.position = "none")
```

### Confusion matrix metrics

Here are metrics from a confusion matrix comparing these annotations:

```{r}
calculate_confusion(
  sce_df,
  "aggregated",
  "not_aggregated"
)
```

### Heatmap: Aggregated vs. non-aggregated

This heatmap compares annotations between aggregated and non-aggregated directly:

```{r, fig.height = 8}
make_jaccard_heatmap(
  sce_df,
  "aggregated",
  "not_aggregated",
  "Aggregated labels",
  "Non-aggregated labels"
)
```

## Heatmap: Aggregated vs. consensus annotations

This heatmap compares annotations between `SingleR` inference using the aggregated reference and consensus cell types.
The broader validation grouping for consensus cell types is used to keep the plot manageable.

```{r, fig.height = 10}
make_jaccard_heatmap(
  sce_df,
  "consensus_annotation",
  "aggregated",
  "OpenScPCA Consensus cell type validation group",
  "Aggregated labels"
)
```

## Heatmap: Non-aggregated vs. consensus annotations

This heatmap compares annotations between `SingleR` inference using the non-aggregated reference and consensus cell types.
The broader validation grouping for consensus cell types is used to keep the plot manageable.

```{r, fig.height = 10}
make_jaccard_heatmap(
  sce_df,
  "consensus_annotation",
  "not_aggregated",
  "OpenScPCA Consensus cell type validation group",
  "Non-aggregated labels"
)
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```