### 2025-07-07 analyses/cell-type-neuroblastoma-04/exploratory-notebooks/reference-aggregation.Rmd
wording about valdation groupings and stromal

---
title: "`r glue::glue('Compare aggregated and non-aggregated reference: {params$library_id}')`"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    code_folding: hide
params:
  sample_id: "SCPCS000115"
  library_id: "SCPCL000133"
---

This notebook does some preliminary comparisons of `SingleR` results obtained with the NBAtlas reference, where the reference was either aggregated before training or not.
The goal is to determine whether there are substantial differences between aggregation versions.
Note that the goal of this notebook is _not_ specifically to validate annotations, but to compare whether the aggregated vs non-aggregated reference yield different or similar labels.
There is, however, a brief comparison to consensus annotations.

## Setup

```{r, warning = FALSE}
options(readr.show_col_types = FALSE)
suppressPackageStartupMessages({
  library(ggplot2)
  library(patchwork)
  library(SingleCellExperiment)
})

theme_set(theme_bw())

# Define color ramp for shared use in the heatmaps
heatmap_col_fun <- circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue"))
# Set heatmap padding option
ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(0.6, "in"))
```

### Paths

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

module_dir <- file.path(repository_base, "analyses", "cell-type-neuroblastoma-04")
results_dir <- file.path(module_dir, "results", "singler")
data_dir <- file.path(repository_base, "data", "current", "SCPCP000004")
```

```{r url paths}
# broad consensus cell type groups
validation_url <- "https://raw.githubusercontent.com/AlexsLemonade/OpenScPCA-analysis/refs/heads/main/analyses/cell-type-consensus/references/consensus-validation-groups.tsv"

# umap colors
palette_url <- "https://raw.githubusercontent.com/AlexsLemonade/scpca-paper-figures/refs/heads/main/palettes/validation-group-palette.tsv"
```

```{r file paths}
# SingleR results
aggr_singler_file <- file.path(
  results_dir,
  params$sample_id,
  "aggregated",
  glue::glue("{params$library_id}_singler-results.rds")
)

non_aggr_singler_file <- file.path(
  results_dir,
  params$sample_id,
  "not-aggregated",
  glue::glue("{params$library_id}_singler-results.rds")
)

# SCE file
sce_file <- file.path(
  data_dir,
  params$sample_id,
  glue::glue("{params$library_id}_processed.rds")
)
```

### Functions

```{r}
# Source Jaccard and heatmap utilities functions:
source(file.path(module_dir, "exploratory-notebooks", "jaccard-utils.R"))

# This function calculates and returns a confusion matrix
#
# df: data frame with columns of interest
# annotation_col1: first column to compare; arg should be a string
# annotation_col2: second column to compare; arg should be a string
calculate_confusion <- function(
    df,
    annotation_col1,
    annotation_col2) {
  # determine all possible annotation values for caret
  all_levels <- union(
    unique(df[[annotation_col1]]),
    unique(df[[annotation_col2]])
  )

  caret::confusionMatrix(
    factor(df[[annotation_col1]], levels = all_levels),
    factor(df[[annotation_col2]], levels = all_levels),
    mode = "sens_spec"
  )
}

# This function calculates and plots delta median for two groups of barcodes: those in agree and those not
# This code is based on this scpca-nf code:
# https://github.com/AlexsLemonade/scpca-nf/blob/4bb82aa635b572a62f2028dbec587fcfd2155e26/templates/qc_report/celltypes_supplemental_report.rmd#L626
#
# df: SingleR data frame with `scores.` columns
# agree_barcodes: Vector of barcodes that have agreeing labels between aggregated and not
# plot_title: Title for plot
plot_delta_median <- function(df, agree_barcodes, plot_title) {
  # extract scores into matrix
  scores_mat <- df |>
    dplyr::select(starts_with("scores.")) |>
    as.matrix()
  rownames(scores_mat) <- df$barcodes

  # Create data frame for plotting with delta median and the full *non-pruned* cell labels
  delta_median_df <- tibble::tibble(
    delta_median = rowMaxs(scores_mat) - rowMedians(scores_mat),
    # Need to grab the non-pruned label for this plot
    full_labels = aggr$labels,
    # if pruned.labels are NA ==> low confidence
    # so, negate for this variable:
    confident = !is.na(df$pruned.labels)
  ) |>
    dplyr::mutate(
      confident = ifelse(confident, "High-quality", "Low-quality"),
      barcodes = df$barcodes,
      annotations_agree = barcodes %in% agree_barcodes
    )

  ggplot(delta_median_df) +
    aes(x = annotations_agree, y = delta_median, color = confident) +
    # don't jitter the height
    geom_jitter(height = 0, alpha = 0.7) +
    theme(legend.position = "bottom") +
    ggtitle(plot_title)
}
```

### Prepare input data

Read SCE object so we can get consensus cell types:

```{r}
sce_df <- readRDS(sce_file) |>
  scuttle::makePerCellDF() |>
  dplyr::select(
    barcodes,
    consensus_annotation = consensus_celltype_annotation,
    UMAP1 = UMAP.1,
    UMAP2 = UMAP.2
  )
```

Read cell type data frames:
```{r}
validation_df <- readr::read_tsv(validation_url) |>
  dplyr::select(consensus_annotation, validation_group_annotation)

palette_df <- readr::read_tsv(palette_url) # umap plotting later

aggr <- readr::read_rds(aggr_singler_file) |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes")

non_aggr <- readr::read_rds(non_aggr_singler_file) |>
  as.data.frame() |>
  tibble::rownames_to_column("barcodes")
```

Join and prepare data for use:

```{r}
singler_df <- data.frame(
  barcodes = aggr$barcodes,
  "aggregated" = aggr$pruned.labels,
  "not_aggregated" = non_aggr$pruned.labels
)

celltype_df <- sce_df |>
  dplyr::left_join(singler_df, by = "barcodes") |>
  dplyr::left_join(validation_df, by = "consensus_annotation") |>
  # set validation NA -> unknown
  dplyr::mutate(
    validation_group_annotation =
      ifelse(is.na(validation_group_annotation), "unknown", validation_group_annotation)
  )
```

## Aggregated vs. non-aggregated

First a quick and dirty query: What percent of annotations made with either the aggregated and non-aggregated reference versions agree?

```{r}
sum(singler_df$aggregated == singler_df$not_aggregated, na.rm = T) / nrow(singler_df)
```

Where specifically are their disagreements?
This table displays the number of cells that have the given cell label disagreements:

```{r}
singler_df |>
  dplyr::filter(aggregated != not_aggregated) |>
  dplyr::count(aggregated, not_aggregated) |>
  dplyr::arrange(desc(n))
```

### Delta median scores

We'll plot the `delta_median` scores for each `SingleR` inference (aggregated and non-aggregated) for cells where annotations agree vs. differ.
We would like to see that scores will be _lower_ for the cells where inferences between reference versions disagree.
Cells will also be colored by their reliability as determined by `SingleR`.

```{r}
agree_barcodes <- singler_df |>
  dplyr::filter(aggregated == not_aggregated) |>
  dplyr::pull(barcodes)

plot_delta_median(aggr, agree_barcodes, plot_title = "Aggregated delta_median") +
  plot_delta_median(non_aggr, agree_barcodes, plot_title = "Non-aggregated delta_median") +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

### Confusion matrix metrics

Here are metrics from a confusion matrix comparing these annotations:

```{r}
calculate_confusion(
  celltype_df,
  "aggregated",
  "not_aggregated"
)
```

### Heatmap: Aggregated vs. non-aggregated

This heatmap compares annotations between aggregated and non-aggregated directly:

```{r, fig.height = 8, fig.width = 8}
make_jaccard_heatmap(
  celltype_df,
  "aggregated",
  "not_aggregated",
  "Aggregated labels",
  "Non-aggregated labels"
)
```

## Comparison with consensus cell type annotations

In this section, we compare aggregated & non-aggregated annotations to the consensus cell type labels.
In these comparisons, we consider the broader validation groupings for consensus cell types in order to keep plots manageable and more easily interpreted.

### Heatmap: Aggregated vs. consensus annotations

This heatmap compares annotations between `SingleR` inference using the aggregated reference and consensus cell types.

```{r, fig.height = 8, fig.width = 8}
make_jaccard_heatmap(
  celltype_df,
  "validation_group_annotation",
  "aggregated",
  "OpenScPCA Consensus cell type validation group",
  "Aggregated labels"
)
```

### Heatmap: Non-aggregated vs. consensus annotations

This heatmap compares annotations between `SingleR` inference using the non-aggregated reference and consensus cell types.

```{r,fig.height = 8, fig.width = 8}
make_jaccard_heatmap(
  celltype_df,
  "validation_group_annotation",
  "not_aggregated",
  "OpenScPCA Consensus cell type validation group",
  "Non-aggregated labels"
)
```

## UMAPs

Below, we show the UMAPs colored by cell types for each of aggregated, non-aggregated, and consensus (again, using the validation groupings for plot visibility).

Before plotting cell type labels were harmonized to the extent possible to that `NBAtlas` labels matched the corresponding consensus cell type labels (e.g., `Fibroblast` -> `fibroblast` or `Endothelial` -> `endothelial cell`).
Cell types that were not literally just spelling variants were left alone; note that `stromal` cell types between references were also not harmonized since this group doesn't always have a clear definition.

The UMAP therefore has one overall legend for all possible cell types, so colors are directly comparable across panels.
The specific cell types that are part of either refernce are printed below.

```{r}
# Recode NBAtlas cell types where possible so they match with colors
celltype_recoded_df <- celltype_df |>
  dplyr::select(-consensus_annotation) |>
  # pivot longer for wrangling first
  tidyr::pivot_longer(
    c(validation_group_annotation, aggregated, not_aggregated),
    names_to = "annotation_type",
    values_to = "label"
  ) |>
  dplyr::mutate(
    label_recoded = dplyr::case_when(
      # nbatlas ~ consensus
      label %in% c("Myeloid", "Fibroblast") ~ stringr::str_to_lower(label),
      label == "Endothelial" ~ "endothelial cell",
      label == "Plasma" ~ "plasma cell",
      label == "NK cell" ~ "natural killer cell",
      is.na(label) ~ "unknown",
      .default = label
    )
  )

# Recode some existing colors we don't need
palette_df <- palette_df |>
  dplyr::mutate(
    label_recoded = dplyr::case_when(
      validation_group_annotation == "hematopoietic precursor cell" ~ "Neuroendocrine",
      validation_group_annotation == "muscle cell" ~ "RBCs",
      validation_group_annotation == "adipocyte" ~ "Schwann",
      validation_group_annotation == "pericyte" ~ "pDC",
      validation_group_annotation == "innate lymphoid cell" ~ "Stromal other",
      .default = validation_group_annotation
    )
  ) |>
  dplyr::filter(label_recoded %in% unique(celltype_recoded_df$label_recoded))

# Define palette
pal <- palette_df$color
names(pal) <- palette_df$label_recoded
```

```{r, fig.width = 12}
ggplot(celltype_recoded_df) +
  aes(x = UMAP1, y = UMAP2, color = label_recoded) +
  geom_point(alpha = 0.6, size = 0.5) +
  scale_color_manual(values = pal) +
  facet_wrap(vars(annotation_type)) +
  coord_equal() +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.position = "bottom"
  ) +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1)))
```
Cell types that are part of `NBAtlas` include:

```{r}
celltype_recoded_df |>
  dplyr::filter(annotation_type != "validation_group_annotation") |>
  dplyr::pull(label_recoded) |>
  unique()
```

Cell types that are part of consensus labels include:

```{r}
celltype_recoded_df |>
  dplyr::filter(annotation_type == "validation_group_annotation") |>
  dplyr::pull(label_recoded) |>
  unique()
```

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```