### 2025-07-10 analyses/cell-type-neuroblastoma-04/exploratory-notebooks/explore-singler-results.Rmd
Continued WIP notebook development

---
title: "Explore the SingleR results"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    code_folding: hide
---

## Setup

```{r, warning = FALSE}
options(readr.show_col_types = FALSE)
suppressPackageStartupMessages({
  library(ggplot2)
  library(patchwork)
  library(SingleCellExperiment)
})

theme_set(theme_bw())

# Define color ramp for shared use in the heatmaps
heatmap_col_fun <- circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue"))
# Set heatmap padding option
ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(0.6, "in"))
```

### Paths

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

module_dir <- file.path(repository_base, "analyses", "cell-type-neuroblastoma-04")
results_dir <- file.path(module_dir, "results", "singler") # sample_id/aggregated/NE-tumor-combined/
data_dir <- file.path(repository_base, "data", "current", "SCPCP000004")
```

```{r url paths}
# broad consensus cell type groups
validation_url <- "https://raw.githubusercontent.com/AlexsLemonade/OpenScPCA-analysis/refs/heads/main/analyses/cell-type-consensus/references/consensus-validation-groups.tsv"
```

```{r file paths}
# SingleR files
singler_files <- list.files(
  path = "/Users/sjspielman/ALSF/open-scpca/OpenScPCA-analysis/analyses/cell-type-neuroblastoma-04/results/singler-test_aggregated",
  pattern = "_singler-annotations\\.tsv",
  recursive = TRUE,
  full.names = TRUE
) |>
  # add names as library id
  purrr::set_names(
    \(x) {
      stringr::str_remove(basename(x), "_singler-annotations.tsv")
    }
  )

# merged SCE file
# OBTAINED FROM PORTAL DIRECTLY on 2025-07-09
sce_file <- file.path(
  data_dir,
  "SCPCP000004_merged.rds"
)
```

### Functions

```{r}
# Source Jaccard and heatmap utilities functions:
source(file.path(module_dir, "exploratory-notebooks", "jaccard-utils.R"))

# This function calculates and plots delta median for two groups of barcodes: those in agree and those not
# This code is based on this scpca-nf code:
# https://github.com/AlexsLemonade/scpca-nf/blob/4bb82aa635b572a62f2028dbec587fcfd2155e26/templates/qc_report/celltypes_supplemental_report.rmd#L626
#
# df: SingleR data frame with `scores.` columns
# agree_barcodes: Vector of barcodes that have agreeing labels between aggregated and not
# plot_title: Title for plot
plot_delta_median <- function(df, agree_barcodes, plot_title) {
  # extract scores into matrix
  scores_mat <- df |>
    dplyr::select(starts_with("scores.")) |>
    as.matrix()
  rownames(scores_mat) <- df$barcodes

  # Create data frame for plotting with delta median and the full *non-pruned* cell labels
  delta_median_df <- tibble::tibble(
    delta_median = rowMaxs(scores_mat) - rowMedians(scores_mat),
    # Need to grab the non-pruned label for this plot
    full_labels = aggr$labels,
    # if pruned.labels are NA ==> low confidence
    # so, negate for this variable:
    confident = !is.na(df$pruned.labels)
  ) |>
    dplyr::mutate(
      confident = ifelse(confident, "High-quality", "Low-quality"),
      barcodes = df$barcodes,
      annotations_agree = barcodes %in% agree_barcodes
    )

  ggplot(delta_median_df) +
    aes(x = annotations_agree, y = delta_median, color = confident) +
    # don't jitter the height
    geom_jitter(height = 0, alpha = 0.7) +
    theme(legend.position = "bottom") +
    ggtitle(plot_title)
}
```

### Prepare input data

Read SCE object to get consensus cell types and UMAP coordinates:

```{r}
merged_sce <- readRDS(sce_file)
# immediately remove assays we don't need for space
assay(merged_sce, "spliced") <- NULL
assay(merged_sce, "counts") <- NULL
gc()
```

Read cell type data frames:
```{r}
singler_df <- singler_files |>
  purrr::map(readr::read_tsv) |>
  purrr::list_rbind(names_to = "library_id") |>
  dplyr::select(-delta.next, -labels) |>
  dplyr::mutate(
    # add cell id column for unique rows & joining
    cell_id = glue::glue("{library_id}-{barcodes}"),
    # recode NA -> "unknown_singler"
    pruned.labels = ifelse(
      is.na(pruned.labels),
      "unknown_singler",
      pruned.labels
    )
  )

validation_df <- readr::read_tsv(validation_url) |>
  dplyr::select(consensus_annotation, validation_group_annotation)

####### TEMPORARY!
merged_sce <- merged_sce[, singler_df$cell_id]
merged_sce
```

Join and prepare data for use:

```{r}
celltype_df <- scuttle::makePerCellDF(
  merged_sce,
  use.coldata = c("barcodes", "sample_id", "library_id", "consensus_celltype_annotation"),
  use.dimred = c("UMAP")
) |>
  # there are repeated barcodes so we need to keep cell_id around
  tibble::rownames_to_column("cell_id") |>
  dplyr::rename(
    UMAP1 = UMAP.1,
    UMAP2 = UMAP.2,
    consensus_annotation = consensus_celltype_annotation
  ) |>
  dplyr::left_join(validation_df, by = "consensus_annotation") |>
  # recode NAs to "unknown_consensus" and remove full consensus labels
  dplyr::mutate(validation_group_annotation = ifelse(
    is.na(validation_group_annotation),
    "unknown_consensus",
    validation_group_annotation
  )) |>
  dplyr::select(-consensus_annotation) |>
  dplyr::left_join(singler_df, by = c("cell_id", "barcodes", "library_id"))
```

```{r}
harmonize_celltypes <- function(df, label_column, recoded_column) {
  df |>
    dplyr::mutate(
      {{ recoded_column }} := dplyr::case_when(
        # nbatlas ~ consensus
        {{ label_column }} %in% c("Myeloid", "Fibroblast") ~ stringr::str_to_lower(label),
        {{ label_column }} == "Endothelial" ~ "endothelial cell",
        {{ label_column }} == "Plasma" ~ "plasma cell",
        {{ label_column }} == "NK cell" ~ "natural killer cell",
        # make the difference stromal labels more clearly distinguishable
        {{ label_column }} == "stromal cell" ~ "stromal cell (consensus)",
        {{ label_column }} == "Stromal other" ~ "Stromal other (NBAtlas)",
        .default = {{ label_column }}
      )
    )
}

# Recode NBAtlas cell types where possible so they match with colors
celltype_recoded_df <- celltype_df |>
  # rename to make annotation sources more clear
  dplyr::rename(
    "consensus_validation" = validation_group_annotation,
    "singler" = pruned.labels
  ) |>
  # pivot longer for wrangling
  tidyr::pivot_longer(
    c(consensus_validation, singler),
    names_to = "annotation_type",
    values_to = "label"
  ) |>
  harmonize_celltypes(label, label_recoded) |>
  dplyr::select(-label)

palette_df <- readr::read_tsv("../harmonized-cell-type-palette.tsv")
celltype_pal <- palette_df$hex_color
names(celltype_pal) <- palette_df$harmonized_label
```

```{r, fig.height = 10, fig.width = 10} 
# Pivot wider for heatmap functions
celltype_recoded_wide_df <- celltype_recoded_df |>
  tidyr::pivot_wider(
    names_from = annotation_type,
    values_from = label_recoded
  )

heatmap_col_fun <- circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue"))

make_jaccard_heatmap(
  celltype_recoded_wide_df,
  "consensus_validation",
  "singler",
  "Consensus validation label",
  "SingleR NBAtlas label"
)
```

```{r, fig.width = 12, fig.height = 8}

ggplot(celltype_recoded_df) +
  aes(x = UMAP1, y = UMAP2, color = label_recoded) +
  geom_point(size = 0.25, alpha = 0.5) +
  scale_color_manual(
    values = celltype_pal,
    name = "Harmonized cell types"
  ) +
  facet_wrap(vars(annotation_type)) +
  coord_equal() +
  theme(
    legend.position = "bottom",
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1)))
```

Let's facet just the SingleR annotations:

```{r}
# Adapted from scpca-nf
# https://github.com/AlexsLemonade/scpca-nf/blob/4bb82aa635b572a62f2028dbec587fcfd2155e26/templates/qc_report/celltypes_qc.rmd#L134-L221

#' Create a faceted UMAP panel where each panel has only one cell type colored
#'
#' @param umap_df Data frame with UMAP1 and UMAP2 columns
#' @param annotation_column Column containing broad cell type annotation
#' @param celltype_colors Named vector of colors to use for each broader validation group
#'
#' @return ggplot object containing a faceted UMAP where each cell type is a facet.
#'   In each panel, the cell type of interest is colored red and all other cells are grey.
faceted_umap <- function(umap_df,
                         annotation_column,
                         facet_column,
                         celltype_colors,
                         facet_ncol = 4, 
                         single_celltype = NULL) {

  if (!(is.null(single_celltype))) {
    first_df <- umap_df |>
      dplyr::filter({{annotation_column}} == single_celltype)
    plot_title <- single_celltype
  } else {
    first_df <- umap_df
    plot_title <- ""
  }

  # color by the annotation column but only color one cell type at a time
  faceted_umap <- ggplot(
    first_df,
    aes(x = UMAP1, y = UMAP2, color = {{ annotation_column }})
  ) +
    # set points for all "other" points
    geom_point(
      data = dplyr::select(
        umap_df, -{{ annotation_column }}
      ),
      color = "gray90",
      alpha = 0.5,
      size = 1
    ) +
    scale_color_manual(values = celltype_colors) +
    # set points for desired cell type
    geom_point(size = 1, alpha = 0.5) +
    facet_wrap(
      vars({{ facet_column }}),
      ncol = facet_ncol
    ) +
    coord_equal() +
    theme_bw() +
    theme(
      legend.position = "none",
      axis.ticks = element_blank(), 
      axis.text = element_blank()
    ) +
    ggtitle(plot_title)

  return(faceted_umap)
}
```

```{r fig.width = 14}
celltype_recoded_df |>
  dplyr::filter(annotation_type == "singler") |>
  faceted_umap(
    annotation_column = label_recoded,
    facet_column = label_recoded,
    celltype_colors = celltype_pal
  )
```

For cells that directly match up with consensus:

```{r, fig.width = 30, fig.asp = 1}
direct_celltype_matches <- c(
  "B cell",
  "T cell",
  "myeloid",
  "fibroblast",
  "endothelial cell",
  "plasma cell",
  "natural killer cell",
  # we'll use this label to be able to directly compare unknown_consensus to Neuroendocrine
  "putative-tumor" 
)

celltype_facet_df <- celltype_recoded_df |>
  # recode so Neuroendocrine matches with unknown_consensus in the plot
  dplyr::mutate(
    label_recoded = ifelse(
      label_recoded %in% c("Neuroendocrine", "unknown_consensus"),
      "putative-tumor",
      label_recoded
    )
  ) |>
  dplyr::filter(label_recoded %in% direct_celltype_matches)

unique(celltype_facet_df$label_recoded) |>
  purrr::map(
    \(cell_type) {
      faceted_umap(
        celltype_facet_df,
        annotation_column = label_recoded, 
        celltype_colors = celltype_pal,
        facet_column = annotation_type,
        facet_ncol = 2, 
        single_celltype = cell_type
      )
  }) |>
  patchwork::wrap_plots(ncol = 1)

```

## Dotplot time

This is copied from Ally's code and is NOT YET ADAPTED TO THIS NOTEBOOK.
```{r}
  # prep for plots
  # get total number of cells per final annotation group
  total_cells_df <- consensus_df |>
    dplyr::select(unique_id, barcodes, broad_celltype_group) |>
    dplyr::distinct() |>
    dplyr::count(broad_celltype_group, name = "total_cells")

  # table with one row per unique broad cell type/ marker gene combination
  # first all cells in with the same broad_celltype_group (determined based on consensus_annotation) are grouped together
  # then get the mean gene expression and total percentage of cells that express each marker gene across all cells in that group
  # do this for every possible marker gene/ validation group assignment
  # second we calculate the mean expression and mean percentage of all marker genes in a given validation group (this value is used only in the second section of the report)
  group_stats_df <- consensus_df |>
    # for each assigned cell type/marker gene combo get total detected and mean expression
    # group by both broad group and validation group to account for genes that are expressed in more than one cell type
    dplyr::summarize(
      .by = c("broad_celltype_group", "ensembl_gene_id", "validation_group_annotation"),
      detected_count = sum(detected),
      mean_exp = mean(logcounts)
    ) |>
    # add in validation group for marker genes
    # this includes all possible marker genes and all possible validation group assignments
    dplyr::left_join(markers_df, by = c("ensembl_gene_id", "validation_group_annotation"), relationship = "many-to-many") |> # add total cells
    dplyr::left_join(total_cells_df, by = c("broad_celltype_group")) |>
    # only show cells that have above the minimum total number
    dplyr::filter(total_cells > min_cells) |>
    dplyr::mutate(
      # get total percent expressed
      percent_exp = (detected_count / total_cells) * 100,
      # account for NA/unknowns and set axes order
      broad_celltype_group = factor(broad_celltype_group, levels = c(names(celltype_colors)))
    )

  # get list of celltypes to keep and assign colors
  celltype_groups <- group_stats_df |>
    dplyr::pull(broad_celltype_group) |>
    unique() |>
    as.character()

  # filter markers to those that are actually relevant
  # we will only plot the marker genes for cell types that are part of the assigned broad validation group for this group of samples
  # we don't care about plotting marker genes for cell types that aren't present here
  filtered_markers_df <- markers_df |>
    dplyr::filter(
      validation_group_annotation %in% celltype_groups,
      gene_symbol %in% group_stats_df$gene_symbol
    ) |>
    # ensure order matches the order of the legend
    dplyr::mutate(
      validation_group_annotation = factor(validation_group_annotation, levels = names(celltype_colors))
    ) |>
    dplyr::arrange(validation_group_annotation)

  # specify x axis order for dotplot
  marker_gene_order <- filtered_markers_df |>
    dplyr::pull(gene_symbol)

  # set order for cell types
  celltype_order <- unique(filtered_markers_df$validation_group_annotation)

  # filter out low expressed genes
  dotplot_df <- group_stats_df |>
    dplyr::filter(mean_exp > 0, percent_exp > 10) |>
    dplyr::arrange(broad_celltype_group) |>
    # add a label for the plot
    dplyr::mutate(y_label = glue::glue("{broad_celltype_group} ({total_cells})")) |>
    # remove marker genes that aren't present in final annotations and set x axis order
    dplyr::filter(gene_symbol %in% marker_gene_order) |>
    dplyr::mutate(
      # set orders of gene symbol and validation groups
      y_label = factor(y_label, levels = rev(unique(y_label))),
      gene_symbol = factor(gene_symbol, levels = marker_gene_order),
      validation_group_annotation = factor(validation_group_annotation, levels = celltype_order)
    )

  # make dotplot with marker gene exp
  dotplot <- ggplot(dotplot_df, aes(y = y_label, x = gene_symbol, color = mean_exp, size = percent_exp)) +
    geom_point() +
    scale_color_viridis_c(option = "magma") +
    facet_grid(cols = vars(validation_group_annotation), scales = "free", space = "free") +
    theme_classic() +
    theme(
      strip.background = element_rect(fill = "transparent", color = NA),
      strip.placement = "outside",
      strip.text.x = element_blank(),
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
      axis.ticks.x = element_blank(),
      text = element_text(size = 14),
      panel.spacing = unit(0.5, "lines") # adjust spacing and match with annotation bar
    ) +
    labs(
      x = "",
      y = "Broad cell type annotation",
      color = "Mean gene expression",
      size = "Percent cells expressed"
    )

  # add annotation bar aligning marker genes with validation group
  color_bar <- ggplot(dotplot_df, aes(x = gene_symbol, y = 1, fill = validation_group_annotation)) +
    geom_tile() +
    facet_grid(cols = vars(validation_group_annotation), scales = "free", space = "free") +
    scale_fill_manual(values = celltype_colors, breaks = levels(dotplot_df$validation_group_annotation)) +
    ggmap::theme_nothing() +
    theme(
      strip.background = element_rect(fill = "transparent", color = NA),
      strip.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5, size = 12),
      strip.placement = "outside",
      legend.position = "none",
      panel.spacing = unit(0.5, "lines"),
      strip.clip = "off"
    ) +
    labs(fill = "")

  combined_plot <- color_bar / dotplot +
    patchwork::plot_layout(ncol = 1, heights = c(0.1, 4))

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```