### 2025-08-14 analyses/cell-type-neuroblastoma-04/exploratory-notebooks/scanvi-singler-comparison.Rmd
update section where we look at overall agreement

---
title: "Compare scANVI and SingleR annotations"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    code_folding: hide
---

We have used `SingleR` and `scANVI/scArches` to perform cell type annotation on `SCPCP000004` samples using the `NBAtlas` reference.
The goal of this notebook is to compare these two sets of inferences to determine if they generally agree.

## Setup

```{r, warning = FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(patchwork)
  library(SingleCellExperiment)
})

theme_set(theme_bw())
set.seed(2025) # used for some viz

# Define color ramp for shared use in the heatmaps
heatmap_col_fun <- circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue"))
# Set heatmap padding option
ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(0.6, "in"))

# settings
options(
  dplyr.summarise.inform = FALSE, 
  readr.show_col_types = FALSE
)
```

### Paths

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

merged_dir <- file.path(repository_base, "data", "current", "results", "merge-sce", "SCPCP000004")
module_dir <- file.path(repository_base, "analyses", "cell-type-neuroblastoma-04")
ref_dir <- file.path(module_dir, "references")
results_dir <- file.path(module_dir, "results")
singler_dir <- file.path(results_dir, "singler")
scanvi_dir <- file.path(results_dir, "scanvi")
```

```{r file paths}
# merged SCE file
sce_file <- file.path(
  merged_dir,
  "SCPCP000004_merged.rds"
)

# SingleR files
singler_files <- list.files(
  path = singler_dir,
  pattern = "_singler-annotations\\.tsv",
  recursive = TRUE,
  full.names = TRUE
) |>
  # add names as library id
  purrr::set_names(
    \(x) {
      stringr::str_remove(basename(x), "_singler-annotations.tsv")
    }
  )

# scanvi predictions
scanvi_file <- file.path(
  scanvi_dir, 
  "scanvi-predictions.tsv"
)

# broad consensus cell type groups
validation_url <- "https://raw.githubusercontent.com/AlexsLemonade/OpenScPCA-analysis/refs/heads/main/analyses/cell-type-consensus/references/consensus-validation-groups.tsv"

# palette files
recoded_palette_file <- file.path(
  module_dir,
  "palettes",
  "harmonized-cell-type-palette.tsv"
)
nbatlas_palette_file <- file.path(
  module_dir,
  "palettes",
  "nbatlas-cell-type-palette.tsv"
)

# label mapping
nbatlas_label_map_file <- file.path(ref_dir, "nbatlas-label-map.tsv")
```

### Functions

```{r}
# Source Jaccard and heatmap utilities functions
source(file.path(module_dir, "scripts", "utils", "jaccard-utils.R"))

# Source additional utilities functions:
# - select_nbatlas_markers()
# - harmonize_celltypes()
# - faceted_umap()
# - generate_dotplot()
source(file.path(module_dir, "scripts", "utils", "celltype-utils.R"))
```

### Read and prepare input data

Read merged SCE object:

```{r}
merged_sce <- readRDS(sce_file)
# immediately remove assays we don't need for space
assay(merged_sce, "spliced") <- NULL
assay(merged_sce, "counts") <- NULL
reducedDim(merged_sce, "PCA") <- NULL
```

Read SingleR results:

```{r}
singler_df <- singler_files |>
  purrr::map(readr::read_tsv) |>
  purrr::list_rbind(names_to = "library_id") |>
  dplyr::mutate(
    # add cell id column for unique rows & joining
    cell_id = glue::glue("{library_id}-{barcodes}"),
    # recode NA -> "unknown_singler"
    singler = ifelse(is.na(pruned.labels),"Unknown", pruned.labels)
  ) |>
  dplyr::select(cell_id, singler, singler_delta_next = delta.next)
```

Read and prepare scanvi results:

```{r}
scanvi_full_df <- readr::read_tsv(
  file.path(scanvi_dir, "scanvi_predictions.tsv")
) |>
  dplyr::select(
    cell_id, 
    scanvi = scanvi_prediction, 
    starts_with("pp_")
  ) |>
  tidyr::pivot_longer(
    starts_with("pp_"), 
    names_to = "posterior_celltype", 
    values_to = "posterior"
  ) |>
  dplyr::mutate(posterior_celltype = stringr::str_remove(posterior_celltype, "^pp_"))

# create dedicated column for the posterior associated with the label
scanvi_df <- scanvi_full_df |>
  dplyr::filter(scanvi == posterior_celltype) |>
  dplyr::select(cell_id, scanvi, scanvi_posterior = posterior)
```

Read and prepare additional helper files for viz:

```{r}
# validation groups data frame
validation_df <- readr::read_tsv(validation_url) |>
  dplyr::select(consensus_annotation, validation_group_annotation)

# label mapping
nbatlas_label_map <- readr::read_tsv(nbatlas_label_map_file)

# set up palettes
recoded_palette_df <- readr::read_tsv(recoded_palette_file)
recoded_celltype_pal <- recoded_palette_df$hex_color
names(recoded_celltype_pal) <- recoded_palette_df$harmonized_label
# for this notebook, we want just a single Unknown
recoded_celltype_pal["Unknown"] <- recoded_celltype_pal["unknown_singler"] 
recoded_celltype_pal["unknown_singler"] <- NULL
recoded_celltype_pal["unknown_consensus"] <- NULL

nbatlas_palette_df <- readr::read_tsv(nbatlas_palette_file)
nbatlas_celltype_pal <- nbatlas_palette_df$hex_color
names(nbatlas_celltype_pal) <- nbatlas_palette_df$NBAtlas_label
```

## scANVI posterior probabilities

Before comparing to `SingleR`, we should get a sense of reliability for the `scANVI` results.
`scANVI` returns a cell-specific posterior probability for each label, so we'll begin by looking at these distributions.

First, what is the distribution of posterior probabilities for assigned labels?
We'll make a histogram below using a very visible color to ensure the plot is legible, where each panel is a library, and libraries are ordered by size.

```{r fig.height=12, fig.width=15}
scanvi_df |>
  tidyr::separate(cell_id, into = c("library_id", "barcode")) |>
  dplyr::add_count(library_id) |>
  dplyr::mutate(
    library_id = glue::glue("{library_id} (N={n})"),
    library_id = factor(library_id), 
    library_id = forcats::fct_infreq(library_id), 
    library_id = forcats::fct_rev(library_id)
  ) |>
  ############ into ggplot ################
  ggplot() + 
  aes(x = scanvi_posterior) + 
  # ugly but _very visible_ color
  geom_histogram(bins = 100, fill = "magenta", color = "magenta") + 
  facet_wrap(vars(library_id), scales = "free_y", ncol = 7)
```

Across libraries, the median posterior probability is always exceptionally high at which suggests `scANVI` confidence is roughly the same across libraries.

Let's instead look at these probabilities on a per-cell type basis, this time with a ridge plot including a line for the median value for each distribution:

```{r fig.height=8, fig.width=8}
scanvi_ridgeplot <- scanvi_df |>
  dplyr::group_by(scanvi) |>
  dplyr::mutate(scanvi = glue::glue("{scanvi} (n={dplyr::n()})")) |>
  ########### into ggplot ############
  ggplot() + 
    aes(y = forcats::fct_infreq(scanvi), x = scanvi_posterior) + 
    # show the median
    ggridges::stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.7) +
    labs(x = "posterior probability", y = "scANVI label (number of cells)") 
scanvi_ridgeplot
```

While annotation reliability is about the same on a per-library basis, it seems that certain cell types may be much easier to classify than others which may be a factor of how well represented they are in the `NBAtlas` reference.
There are very few cells in general associated with the cell types that have very low median posterior probabilities (`NKT cell`, `Migratory cDC`, `Circulating NK cell`).

A `0.75` threshold seems like it may be a good middle-ground for preserving labels while removing low-confidence annotations:

```{r fig.height=8, fig.width=8}
scanvi_ridgeplot +
  geom_vline(xintercept = 0.75, color = "firebrick")
```

How many cells will this remove?

```{r}
sum(scanvi_df$scanvi_posterior < 0.75) / nrow(scanvi_df)
```

What are those cell types?

```{r}
scanvi_df |>
  dplyr::filter(scanvi_posterior < 0.75) |>
  dplyr::count(scanvi) |>
  dplyr::arrange(desc(n))
```

We'll therefore recode cells with posterior probabilities below 0.75 to `Unknown`.

```{r}
pp_threshold <- 0.75
scanvi_df <- scanvi_df |>
  dplyr::mutate(scanvi = ifelse(
    scanvi_posterior < pp_threshold, 
    "Unknown",
    scanvi
  ))
```

## Compare labels

We'll create a single data frame with UMAP coordinates and all cell types (consensus, `SingleR`, and `scANVI`). 
We'll do this at two levels of organization:

- One with harmonized labels to be able to compare to consensus
- One with original NBAtlas labels to just compare `SingleR` to `scANVI`

```{r}
celltype_df <- scuttle::makePerCellDF(
  merged_sce,
  use.coldata = c("sample_id", "is_xenograft", "consensus_celltype_annotation"),
  use.dimred = c("UMAP")
) |>
  # there ARE repeated barcodes so we need to keep cell_id around
  tibble::rownames_to_column("cell_id") |>
  dplyr::rename(
    UMAP1 = UMAP.1,
    UMAP2 = UMAP.2,
    consensus_annotation = consensus_celltype_annotation
  ) |>
  dplyr::left_join(validation_df, by = "consensus_annotation") |>
  # recode consensus NAs to "Unknown"
  dplyr::mutate(
    validation_group_annotation = ifelse(
      is.na(validation_group_annotation), "Unknown", validation_group_annotation
    )
  ) |>
  dplyr::select(-consensus_annotation) |>
  # merge in the annotations
  dplyr::left_join(singler_df, by = "cell_id") |>
  dplyr::left_join(scanvi_df, by = "cell_id") |>
  # make it tidy
  tidyr::pivot_longer(
    c(validation_group_annotation, singler, scanvi), 
    names_to = "celltype_method", 
    values_to = "label"
  ) |>
  harmonize_celltypes(label, label_recoded) |>
  dplyr::mutate(label = ifelse(label == "NE", "Neuroendocrine", label)) |>
  tidyr::separate(cell_id, into = c("library_id", "barcode"), remove = FALSE)
```

## UMAP

This section displays the UMAP for the _merged object_ (not integrated) with cell type labels across methods.
We show a panel for each cell type method.

### SingleR vs scANVI

This set of UMAPs shows annotation results from `SingleR` and `scANVI`.

Since there are many cell types in this plot, it has been colored so that cell types in the same "family" each share a color:

* T cells: `T cell`, `Treg`, `CD4+ T cell`, `CD8+ T cell`
* Natural killer cells: `NK cell`, `Circulating NK cell`, `Resident NK cell`, `TOX2+/KIT+ NK cell`
* Conventional dendritic cells: `cDC`, `cDC1`, `cDC2/DC3`, `Migratory cDC`
* Monocytes: `Monocyte`, `Classical monocyte`, `Patrolling monocyte`

```{r, fig.width = 12, fig.height = 8}
nbatlas_t_celltypes <- nbatlas_label_map |>
  dplyr::filter(NBAtlas_label == "T cell") |>
  dplyr::pull(NBAtlas_label_immune_zoom)

nbatlas_nk_celltypes <- nbatlas_label_map |>
  dplyr::filter(NBAtlas_label == "NK cell") |>
  dplyr::pull(NBAtlas_label_immune_zoom)

nbatlas_myeloid_celltypes <- nbatlas_label_map |>
  dplyr::filter(NBAtlas_label == "Myeloid") |>
  dplyr::pull(NBAtlas_label_immune_zoom)

# Note that cDC and Monocyte are not direct labels in the disseminated NBAtlas object,
# but these groupings are used in the paper's marker gene analysis.
nbatlas_cdc_celltypes <- nbatlas_label_map |>
  dplyr::filter(NBAtlas_label == "cDC") |>
  dplyr::pull(NBAtlas_label_immune_zoom)

nbatlas_monocyte_celltypes <- nbatlas_label_map |>
  dplyr::filter(NBAtlas_label == "Monocyte") |>
  dplyr::pull(NBAtlas_label_immune_zoom)

nbatlas_myeloid_celltypes <- nbatlas_label_map |>
  dplyr::filter(NBAtlas_label == "Myeloid") |>
  dplyr::pull(NBAtlas_label_immune_zoom)

umap_direct_df <- celltype_df |>
  dplyr::filter(celltype_method != "validation_group_annotation") |>
  # shuffle points to ensure order they are plotted is random
  dplyr::sample_frac(1) |> 
  # order celltypes so that families are grouped together for a cleaner legend
  # note this will give warnings if any of those cells aren't in the annotation, but that's fine
  dplyr::mutate(
    label = factor(label), 
    label = forcats::fct_relevel(
      label, 
      "T cell", nbatlas_t_celltypes, 
      "NK cell", nbatlas_nk_celltypes,
      "cDC", nbatlas_cdc_celltypes, 
      "Monocyte", nbatlas_monocyte_celltypes
    )
  )

ggplot(umap_direct_df) +
  aes(x = UMAP1, y = UMAP2, color = label) +
  geom_point(alpha = 0.5, size = 0.2) +
  scale_color_manual(values = nbatlas_celltype_pal) +
  facet_wrap(vars(celltype_method)) +
  coord_equal() + 
  theme(
    legend.position = "bottom", 
    axis.text = element_blank(), 
    axis.ticks = element_blank()
  ) +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1)))
```

Cell type annotations look broadly consistent between these methods, although `SingleR` appears to assign more heterogeneity within the "clumps" of cells (which likely correspond to individual libraries in this merged object).

### Harmonized labels for consensus comparison

This set of UMAPs harmonizes cell type labels into broader groupings to enable a comparison with consensus annotations.
Note that the `Unknown` cell type refers to cells uncategorized by `SingleR` or `scANVI`.

```{r, fig.width = 12, fig.height = 8}
ggplot(celltype_df) +
  aes(x = UMAP1, y = UMAP2, color = label_recoded) +
  geom_point(alpha = 0.5, size = 0.2) +
  scale_color_manual(values = recoded_celltype_pal) +
  facet_wrap(vars(celltype_method)) +
  coord_equal() +
  theme(
    legend.position = "bottom", 
    axis.text = element_blank(), 
    axis.ticks = element_blank()
  ) +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1)))
```
We see broad compatibility among all three approaches, and the Unknown consensus indeed appear to be mostly tumor cells (here, Neuroendocrine).
This is reassuring that we have some robust cell type identifications here!

## Heatmap

This section shows heatmaps comparing annotations to one another.

### SingleR vs scANVI

This heatmap directly compares `SingleR` and `scANVI` annotations.

```{r, fig.height = 11, fig.width = 11} 
# make a wide version for the heatmap
celltype_wide_df <- celltype_df |>
  dplyr::filter(celltype_method != "validation_group_annotation") |>
  dplyr::select(-label_recoded) |>
  tidyr::pivot_wider(
    names_from = celltype_method,
    values_from = label
  )

make_jaccard_heatmap(
  celltype_wide_df,
  "scanvi",
  "singler",
  "scANVI label",
  "SingleR label"
)
```

This heatmap shows excellent correspondence between `SingleR` and `scANVI` annotations with most of the weight along the diagonal.

We'll nail this down a bit further with some confusion matrix metrics:

```{r}
# ensure factors are compatible for caret
all_levels <- c(
  celltype_wide_df$scanvi, 
  celltype_wide_df$singler
) |> unique()

caret::confusionMatrix(
  factor(celltype_wide_df$scanvi, levels = all_levels),  
  factor(celltype_wide_df$singler, levels = all_levels)
)$overall

```

### Harmonized labels for consensus comparison

This stacked heatmap compares all three methods to one another, using broader cell type groupings to enable a more direct comparison with consensus annotations.

```{r, fig.height = 13, fig.width = 12} 
# make a wide version for the heatmap
celltype_recoded_wide_df <- celltype_df |>
  dplyr::select(-label) |>
  tidyr::pivot_wider(
    names_from = celltype_method,
    values_from = label_recoded
  )

list(
  "SingleR annotations" = make_jaccard_matrix(celltype_recoded_wide_df, "validation_group_annotation", "singler"),
  "scANVI annotations" = make_jaccard_matrix(celltype_recoded_wide_df, "validation_group_annotation", "scanvi")
) |>
  purrr::imap(
    \(jaccard_mat, name) create_single_heatmap(
        mat = jaccard_mat,
        row_title = name,
        column_title = "Consensus validation groups",
        keep_legend_name = "SingleR annotations" # arbitrary for labeling
    )
  ) |>
  # concatenate vertically into HeatmapList object
  purrr::reduce(ComplexHeatmap::`%v%`) |>
  ComplexHeatmap::draw(heatmap_legend_side = "right")
```

Again, we see excellent correspondence among all three methods.

## Cross-library comparison

Across libraries, what fraction of `SingleR` and `scANVI` labels exactly agree?
Note that this analysis does not consider similar labels (e.g. two types of T cells) but only does a quick look at identical labels.
We'll also highlight the PDX samples here to see if there was any associated bias.

> I think it would be helpful to get a table in here that shows you % of each library that has exact agreement between SingleR and scANVI, % of each library where cells are in the same family (e.g., the same cell is labeled as two different T cell types), and % of each library that disagrees and is not in the same family.

In this case, we consider these families specifically:

* T cells: `T cell`, `Treg`, `CD4+ T cell`, `CD8+ T cell`
* Natural killer cells: `NK cell`, `Circulating NK cell`, `Resident NK cell`, `TOX2+/KIT+ NK cell`
* Conventional dendritic cells (cDC): `cDC`, `cDC1`, `cDC2/DC3`, `Migratory cDC`
* Monocytes: `Monocyte`, `Classical monocyte`, `Patrolling monocyte`
* Myeloid: `Neutrophil` and `Macrophage`

```{r}
# celltype_wide_df uses the original labels
identical_agree_df <- celltype_wide_df |>
  dplyr::group_by(library_id) |>
  dplyr::summarize(
    frac_identical = sum(scanvi == singler)/dplyr::n(), 
    library_size = dplyr::n()
  )

# create a wide data frame with "family" labels where possible
celltype_family_df <- celltype_df |>
  dplyr::filter(celltype_method != "validation_group_annotation") |>
  # create label_family grouping
  dplyr::mutate(
    label_family = dplyr::case_when(
      label %in% nbatlas_t_celltypes ~ "T cell",
      label %in% nbatlas_nk_celltypes ~ "NK cell",
      label %in% nbatlas_cdc_celltypes ~ "cDC",
      label %in% nbatlas_monocyte_celltypes ~ "Monocyte",
      label %in% nbatlas_myeloid_celltypes ~ "Myeloid", 
      .default = label
    )
  ) |>
  dplyr::select(-label_recoded, -label) |>
  tidyr::pivot_wider(
    names_from = celltype_method,
    values_from = label_family
  )

family_agree_df <- celltype_family_df |>
  dplyr::group_by(library_id) |>
  dplyr::summarize(frac_same_family = sum(scanvi == singler)/dplyr::n())

# if families disagree, then there is no agreement at all
disagree_df <- celltype_family_df |>
  dplyr::group_by(library_id) |>
  dplyr::summarize(frac_disagree = sum(scanvi != singler)/dplyr::n())
```

```{r}
# combine into a single table for display & viz
frac_df <- identical_agree_df |>
  dplyr::left_join(family_agree_df) |>
  dplyr::left_join(disagree_df) |>
  # round to 4 decimals
  dplyr::mutate(dplyr::across(contains("frac_"), round, 4)) |>
  dplyr::select(
    library_id, 
    library_size, 
    frac_identical, 
    frac_same_family, 
    frac_disagree
  ) |>
  dplyr::arrange(library_size)

# use DT for improved exploration ability
DT::datatable(frac_df, rownames = FALSE)
```

```{r}
# add PDX column for plotting
frac_df <- celltype_df |> 
  dplyr::select(library_id, is_xenograft) |>
  unique() |>
  dplyr::right_join(frac_df)

ggplot(frac_df) + 
  aes(x = library_size, y = frac_same_family) + 
  geom_point(aes(color = is_xenograft)) + 
  geom_smooth(method = "lm") +
  labs(
    x = "Library size",
    y = "Fraction of SingleR/scANVI labels from same family"
  ) +
  # stop going above 1
  scale_y_continuous(limits = c(0, 1))
```

Most cells have agreeing annotations in the majority of libraries.
There appears to be a slight trend (not signficant) that larger libraries have more agreement between methods, but this also seems driven by the libraries with low correspondence.

### Libraries with high levels of disagreement

There are four libraries (2 PDX and 2 tissue) which show less than 50% agreement, so we should look at these a bit more closely.
Notably, these do _not_ correspond to the four smallest libraries.

## Conclusions

Overall `SingleR` and `scANVI` appear to have made highly similar predictions for the vast majority of libraries.
These cell types are also fairly similar to our existing consensus cell types.
Therefore, we can use agreement between `SingleR` and `scANVI` to derive final annotations for this analysis module.

```{r}
sessionInfo()
```