### 2025-08-21 analyses/infercnv-consensus-cell-type/exploratory-notebooks/reference-query-counts.Rmd
add notebook to look at counts of reference/query

---
title: "Diagnoses and cell types"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output: html_notebook
params:
  n_reference_threshold: 100
---

The goal of this notebook is to look at cell type distributions across samples to determine how many samples could be run through `inferCNV`.
Specifically, we are interested in seeing the relative numbers of cells intended as `inferCNV` reference vs. query.
Cell types were determined to be reference or query based on the sample's diagnosis, as mapped out in `../references/broad-diagnosis-map.tsv`.
We consider a threshold of at least `r params$n_reference_threshold` reference cells needed to run `inferCNV`.

## Setup

```{r, warning = FALSE}
# settings
options(
  dplyr.summarise.inform = FALSE, 
  readr.show_col_types = FALSE
)
```

### Paths

Define directories:

```{r}
module_dir <- rprojroot::find_root(rprojroot::is_renv_project)
repository_base <- file.path(module_dir, "..", "..")

ref_dir <- file.path(module_dir, "references")
data_dir <- file.path(repository_base, "data", "current")
consensus_dir <- file.path(repository_base, "data", "current", "results", "cell-type-consensus")
consensus_module_dir <- file.path(repository_base, "analyses", "cell-type-consensus")
```

Define paths to input files to read in:

```{r}
# map broad to submitted diagnoses
diagnosis_map_file <- file.path(ref_dir, "broad-diagnosis-map.tsv")

# map broad diagnoses to cell type groups
diagnosis_celltype_map_file <- file.path(ref_dir, "diagnosis-celltype-groups.tsv")

# map cell type groups to constituent cell types
validation_file <- file.path(
  consensus_module_dir, 
  "references", 
  "consensus-validation-groups.tsv"
)

# metadata files, for getting sample diagnoses
metadata_files <- list.files(
  path = data_dir,
  pattern = "single_cell_metadata\\.tsv$",
  recursive = TRUE,
  full.names = TRUE
)

# TSVs of consensus cell types
consensus_files <- list.files(
  path = consensus_dir,
  pattern = "_processed_consensus-cell-types\\.tsv\\.gz$",
  recursive = TRUE,
  full.names = TRUE
) |>
  # set names to filter on later so we don't read files we don't need
  purrr::set_names(\(x) {
    stringr::str_split_i(basename(x), "_", 1)
  })
```

### Read input data

```{r}
diagnosis_map <- readr::read_tsv(diagnosis_map_file)
diagnosis_celltype_map <- readr::read_tsv(diagnosis_celltype_map_file)
validation_group_df <- readr::read_tsv(validation_url)
```

```{r}
sample_diagnosis_df <- metadata_files |>
  purrr::map(
    \(file) {
      readr::read_tsv(file) |>
        # don't consider multiplexed or cell line
        dplyr::filter(!is_multiplexed, !is_cell_line) |>
        dplyr::select(contains("scpca"), diagnosis)
  }) |>
  dplyr::bind_rows() |>
  # use inner_join to remove normals
  dplyr::inner_join(diagnosis_map, by = c("diagnosis" = "submitted_diagnosis")) |>
  dplyr::select(-ontology_id, -human_readable_value)
```

```{r}
consensus_df <- consensus_files |>
  # only keep samples in sample_diagnosis_df
  purrr::keep_at(sample_diagnosis_df$scpca_library_id) |>
  purrr::imap(
    \(file, id) {

      # bump guess_max to kill problems() warnings
      readr::read_tsv(file, guess_max = 15000) |>
       dplyr::select(
         scpca_project_id = project_id,
         scpca_sample_id = sample_id,
         scpca_library_id = library_id,
         barcodes,
         consensus_annotation,
         consensus_ontology
       )
  }) |>
  dplyr::bind_rows()
```

### Prepare data for analysis

Here, we map over each diagnosis group to count the number of reference and query cells for each sample in that group.
We then combine into a single data frame of reference/query cell counts per sample.

```{r}
sample_counts_df <- unique(sample_diagnosis_df$diagnosis_group) |>
  purrr::map(
    \(broad_diagnosis) {

      celltype_groups <- diagnosis_celltype_map |>
        dplyr::filter(diagnosis_group == broad_diagnosis) |>
        tidyr::separate_rows(celltype_groups, sep = ", ") |>
        dplyr::pull(celltype_groups)

      reference_celltypes <- validation_group_df |>
        dplyr::filter(validation_group_annotation %in% celltype_groups) |>
        dplyr::pull(consensus_annotation)

      library_ids <- sample_diagnosis_df |>
        dplyr::filter(diagnosis_group == broad_diagnosis) |>
        dplyr::pull(scpca_library_id)

      count_df <- consensus_df |>
        dplyr::filter(scpca_library_id %in% library_ids) |>
        dplyr::mutate(
          cell_class = ifelse(
            consensus_annotation %in% reference_celltypes, 
            "n_reference", 
            "n_query"
        )) |>
        dplyr::count(scpca_library_id, cell_class) |>
        tidyr::pivot_wider(
          names_from = cell_class, 
          values_from = n, 
          values_fill = 0
        ) |>
        dplyr::inner_join(sample_diagnosis_df, by = "scpca_library_id") |>
        dplyr::select(
          scpca_project_id, 
          scpca_library_id, 
          diagnosis_group, 
          diagnosis, 
          # there may not be any n_reference, so this is more flexible
          contains("n_")
        )

      # Create n_reference if it doesn't exist
      if (!("n_reference" %in% names(count_df))) {
        count_df$n_reference <- 0
      }

      return(count_df)
  }) |>
  dplyr::bind_rows()
```

## Analysis

Using a threshold of needing at least `r params$n_reference_threshold` reference cells, how many samples could be run through `inferCNV`?

```{r}
usable_df <- sample_counts_df |>
  dplyr::group_by(diagnosis_group) |>
  dplyr::summarize(
    n_usable = sum(n_reference >= params$n_reference_threshold),
    total_group_samples = dplyr::n(), 
    frac_usable = round(n_usable / total_group_samples, 3) 
  ) |>
  dplyr::arrange(desc(total_group_samples))

usable_df
```
All of our broad diagnosis groups are represented, which is nice to see.

In total, we can run this fraction of ScPCA samples:

```{r}
sum(sample_counts_df$n_reference >= params$n_reference_threshold) / nrow(sample_counts_df)
```

Of the samples with sufficient reference cells, do any have _fewer_ query than reference cells?

```{r}
sample_counts_df |>
  # consider usable only
  dplyr::filter(n_reference >= params$n_reference_threshold) |>
  dplyr::filter(n_query < params$n_reference_threshold)
```

There is one project about 7x as many reference as query cells; let's look more closely:

```{r}
consensus_df |>
  dplyr::filter(scpca_library_id == "SCPCL000021") |>
  dplyr::count(consensus_annotation) |>
  dplyr::arrange(n)
```

This sample indeed is dominated by immune cell types.
This sample may just have very low tumor purity, but I think it's still fine to run through `inferCNV`.

## Conclusions

Currently, _before_ incorporating `SCimilarity` into our consensus cell types, we are able to run about 44% of samples through `inferCNV`. 
All broad diagnosis groups are represented.

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```