### 2025-08-21 analyses/cell-type-consensus/exploratory-notebooks/06-update-consensus-scimilarity.Rmd
figure out how to deal with multiple lcas

---
title: "Summary of cell type ontologies in reference files"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
    df_print: "kable"
---

This notebook explores incorporating annotations from the `SCimilarity` foundational model to the consensus cell type reference. 
The full list of cell type annotations possible for `SCimilarity` can be found in the [`cell-type-scimilarity` module](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-scimilarity/references/scimilarity-mapped-ontologies.tsv).

For the updated consensus cell types, we plan to assign a consensus cell type label if 2 of the 3 cell type annotations agree.

- First, all possible combinations of the three annotations are identified. 
- Then we take all pairs for each combination and find the latest common ancestor (LCA) using the [`ontoProc::findCommonAncestors()` function](https://rdrr.io/bioc/ontoProc/man/findCommonAncestors.html).
- The LCA will then be assigned to the consensus cell type label for that _pair_ if it meets a set of rules that we previously defined (total descendants <= 170 with some exceptions). 
We will look at all cell type labels that are kept and discarded using our previous rules and identify if we need to make updates to the exceptions. 
- Finally we will map the LCA for each possible pair back to the combinations of all three cell types. 
If one possible LCA is identified, that will be used as the consensus cell type label. 
If there are multiple possible LCAs, we will choose the most specific term as the label.

## Setup

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_bw()
)
```

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

# The path to this module
ref_dir <- file.path(repository_base, "analyses", "cell-type-consensus", "references")

# path to ref files
panglao_file <- file.path(ref_dir, "panglao-cell-type-ontologies.tsv")
blueprint_file <- file.path(ref_dir, "blueprint-mapped-ontologies.tsv")
scimilarity_file <- file.path(repository_base, "analyses", "cell-type-scimilarity", "references", "scimilarity-mapped-ontologies.tsv")
```

```{r}
# grab obo file
cl_ont <- ontologyIndex::get_ontology("http://purl.obolibrary.org/obo/cl/releases/2024-09-26/cl-basic.obo")

# set up the graph to use for assigning LCA terms 
parent_terms <- cl_ont$parents
cl_graph <- igraph::make_graph(rbind(unlist(parent_terms), rep(names(parent_terms), lengths(parent_terms))))
```

```{r, message=FALSE}
# read in reference files
panglao_df <- readr::read_tsv(panglao_file) |>
  # rename columns to have panglao in them for easy joining later
  dplyr::select(
    panglao_ontology = "ontology_id",
    panglao_annotation = "human_readable_value",
    original_panglao_name = "panglao_cell_type" # keep original name since some map to the same ontology ID
  ) |> 
  # remove any cell types that don't have ontologies 
  tidyr::drop_na()

# read in blueprint data 
blueprint_df <- readr::read_tsv(blueprint_file)

# read in scimilarity 
scimilarity_df <- readr::read_tsv(scimilarity_file) |> 
  dplyr::select(
    scimilarity_ontology = "scimilarity_celltype_ontology",
    scimilarity_annotation = "human_readable_value",
    original_scimilarity_name = "scimilarity_celltype_annotation"
  )
```

```{r, message=FALSE}
# get total descendants for each term in CL  
# turn cl_ont into data frame with one row per term
cl_df <- data.frame(
  cl_ontology = cl_ont$id,
  cl_annotation = cl_ont$name
) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    descendants = list(ontologyIndex::get_descendants(cl_ont, cl_ontology, exclude_roots = TRUE)),
    total_descendants = length(descendants)
  )
```

## Determine LCA for all possible pairs

In this section we will first create a table of all possible combinations of the three cell type annotations (`SingleR`, `CellAssign`, and `SCimilarity`). 
For each combination, we will get all possible pairs and compute the LCA.

```{r, warning=FALSE, message=FALSE}
# get a data frame with all combinations of panglao, blueprint, and scimilarity terms
# one row for each combination
all_ref_df <- expand.grid(
  panglao_ontology = panglao_df$panglao_ontology,
  blueprint_ontology = blueprint_df$blueprint_ontology,
  scimilarity_ontology = scimilarity_df$scimilarity_ontology
) |> 
  # remove any potentially redundant combinations based on ontology Ids 
  dplyr::distinct() |> 
  # create a column with all possible pairs between the three tools 
  dplyr::mutate(pair_1 = glue::glue("{panglao_ontology};{blueprint_ontology}"),
                pair_2 = glue::glue("{panglao_ontology};{scimilarity_ontology}"),
                pair_3 = glue::glue("{blueprint_ontology};{scimilarity_ontology}")) |> 
  tidyr::pivot_longer(cols = c(pair_1, pair_2, pair_3), names_to = "number", values_to = "pairs")

# get all possible pairs from each combination and get lca for each pair that is unique 
lca_pairs_df <- all_ref_df |> 
  dplyr::select(pairs) |> 
  dplyr::distinct() |> 
  tidyr::separate(pairs, into = c("id1", "id2"), sep = ";") |> 
  dplyr::rowwise() |>
  dplyr::mutate(
    # least common shared ancestor
    lca = list(rownames(ontoProc::findCommonAncestors(id1, id2, g = cl_graph)))
  )
```

```{r, warning=FALSE}
# add lca and total number of descendants to data frame
# expand to have one row per unique combo + unique lca
# remember sometimes there are multiple LCA that will get returned 
lca_df <- lca_pairs_df |>
  dplyr::mutate(
    total_lca = length(lca), # get total number for filtering later 
    lca = paste0(lca, collapse = ",") # make it easier to split the lca terms
  ) |>
  # split each lca term into its own column
  tidyr::separate(lca, into = c("lca_1", "lca_2", "lca_3"), sep = ",") |>
  # transpose so that instead of lca being in a column there is one row per lca 
  tidyr::pivot_longer(
    cols = dplyr::starts_with("lca"),
    names_to = "lca_number",
    values_to = "lca"
  ) |>
  tidyr::drop_na() |>
  dplyr::select(-lca_number) |> # don't need this column 
  # account for any cases where the ontology IDs are exact matches
  # r complains about doing this earlier since the lca column holds lists until now
  dplyr::mutate(lca = dplyr::if_else(id1 == id2, id1, lca)) |> 
  # join in information for each of the lca terms including name and number of descendants
  dplyr::left_join(cl_df, by = c("lca" = "cl_ontology")) |> 
  # add in legible names for id1/id2 pairs 
  dplyr::mutate(
    id1_annotation = cl_ont$name[id1],
    id2_annotation = cl_ont$name[id2]
  )

```

### Retained celltypes using existing rules

We previously excluded any LCA that had > 170 descendants or belonged to one of these cell types: "bone cell", "lining cell", "blood cell", "progenitor cell", "supporting cell".

Let's see what cell types we would keep if we kept these rules and see if any of them seem too "broad".

```{r}
# vector of cell types that we want to exclude 
celltypes_to_exclude <- c("bone cell", "lining cell", "blood cell", "progenitor cell", "supporting cell")

# get only celltypes that remain after implementing the previous rules 
lca_df |>
  dplyr::filter(total_descendants <= 170 & !(cl_annotation %in% celltypes_to_exclude)) |>
  dplyr::select(cl_annotation, total_descendants) |>
  dplyr::arrange(desc(total_descendants)) |>
  unique() |> 
  # print out in a readable table
  DT::datatable(rownames = FALSE)
```

There are a few cell types we might want to explore further that might seem too broad. 
Let's print out a table that look at the possible matches that give rise to those cell types below.

```{r}
celltypes_to_explore <- c(
  "peptide hormone secreting cell",
  "embryonic cell (metazoa)",
  "biogenic amine secreting cell", 
  "protein secreting cell",
  "antibody secreting cell",
  "extracellular matrix secreting cell",
  "surfactant secreting cell"
)

# for all cell types we want to look at print out a table of the pairs that give rise to that cell type
celltypes_to_explore |> 
  purrr::map(
    \(celltype){
      lca_df |> 
        dplyr::filter(cl_annotation == {{celltype}}) |> 
        dplyr::select(id1_annotation, id2_annotation, total_lca, cl_annotation) |> 
        unique()
    }
  )
```

Looking at these tables, I might get rid of "biogenic amine secreting cell", "protein secreting cell", and "extracellular matrix secreting cell". 
These cell types arise because of matches from cells that have similar functions but are very different. 
For example, platelets and mast cells are pretty different but give the LCA term "biogenic amine secreting cell".

I would keep the following:

- Matches that result in "peptide hormone secreting cell" all appear to be from the pancreas. 
- Matches that result in "embryonic cell (metazoa)" are all developmental cell types that are found in the embryo and mirror some pediatric cancer cell types. 
I think having that information could be useful. 
- "Antibody secreting cell" refers to "A lymphocyte of B lineage" according to [Cell Ontology](https://www.ebi.ac.uk/ols4/ontologies/cl/classes/http%253A%252F%252Fpurl.obolibrary.org%252Fobo%252FCL_0000946) and I think knowing that a cell is some sort of B cell is helpful.
- "Surfactant secreting cell" is specific to the lung and both cell types that appear in matches are lung cell types.

One thing to note is that most of these have a total LCA count > 1 meaning there is at least one other match for those cell types. 
Maybe we should consider removing anything with more than 1 LCA which is what we were doing previously when we only were matching Blueprint and Panglao.

### Handling multiple LCA matches

First, let's see what cell types have multiple matches.

```{r}
# updated celltypes to exclude 
celltypes_to_exclude <- c(celltypes_to_exclude, "biogenic amine secreting cell", "protein secreting cell", "extracellular matrix secreting cell")

# create a df with just lca that meet the specified criteria so far 
filtered_lca_df <- lca_df |>
  dplyr::filter(total_descendants <= 170  & !cl_annotation %in% celltypes_to_exclude)

# print out all matches when total lca > 1
filtered_lca_df |> 
  dplyr::filter(total_lca > 1) |> 
  dplyr::select(cl_annotation) |> 
  unique() |> 
  DT::datatable(rownames = FALSE)
```

Previously, when we looked at the LCA for pairs with more than 1 match, we had seen only very broad cell types. 
That doesn't seem to be the case here as we see things like "memory T cell".

#### Re-calculating the LCA

Perhaps instead of just removing the pairs with multiple LCA, we first filter out the terms that don't meet our criteria. 
Then, if multiple LCAs remain, we find the LCA of all the remaining LCAs, and if that new LCA meets our criteria, we use that instead.

```{r, warning=FALSE, message=FALSE}
# get the new LCA number after removing the non-specific terms 
filtered_lca_df <- filtered_lca_df |> 
  dplyr::group_by(id1, id2) |> 
  dplyr::mutate(new_lca_pair = length(lca))
```

```{r, warning=FALSE, message=FALSE}
# now pull out any pairs where only one of the LCA remains 
# find the LCA between the remaining ones and then filter based on total descendants
# are there any that are remaining that still have multiple LCA? 
new_lca_df <- filtered_lca_df |>
  dplyr::filter(new_lca_pair > 1) |>
  dplyr::group_by(id1, id2) |> 
  dplyr::summarize(
    updated_lca = list(unique(rownames(ontoProc::findCommonAncestors(lca, g = cl_graph))))
    ) |> 
  dplyr::rowwise() |> 
  dplyr::mutate(
    new_total_lca = length(updated_lca),
    updated_lca = paste0(updated_lca, collapse = ","),
    id1_annotation = cl_ont$name[id1],
    id2_annotation = cl_ont$name[id2]
  ) |> 
  # split each lca term into its own column
  tidyr::separate(updated_lca, into = c("updated_lca_1", "updated_lca_2", "updated_lca_3"), sep = ",") |>
  # transpose so that instead of lca being in a column there is one row per lca 
  tidyr::pivot_longer(
    cols = dplyr::starts_with("updated_lca"),
    names_to = "lca_number",
    values_to = "updated_lca"
  ) |>
  tidyr::drop_na() |>
  dplyr::select(-lca_number) |>
   dplyr::left_join(cl_df, by = c("updated_lca" = "cl_ontology"))
```

First for any pairs that have LCA > 1, what are they? 
```{r}
new_lca_df |> 
  dplyr::filter(new_total_lca > 1) |> 
  dplyr::select(id1_annotation, id2_annotation, cl_annotation, total_descendants)
```

I think here we just want to keep "kidney tubule cell" if one of the LCA's includes that term. 
In that case, there are three possible LCA but "kidney tubule cell" is the most specific (fewest descendants).

In the other cases, we should keep any possible LCA that meet the total descendants criteria and remove all others.

Now let's look at the remaining matches that give rise to 1 LCA and see what pairs pass the 170 total descendants threshold.

```{r}
new_lca_df |> 
  dplyr::filter(total_descendants <= 170 & new_total_lca == 1) |> 
  dplyr::select(id1_annotation, id2_annotation, cl_annotation)
```

These all seem like reasonable assignments. 
Let's create a dataframe with the pairs and final matches for these so we can update the whole assignment table later.

```{r}
# first get a data frame of the matches we will keep that had multiple LCA and their final annotations
multi_lca_keep_df <- new_lca_df |> 
  dplyr::filter(total_descendants <= 170 & cl_annotation != "epithelial cell of nephron" | cl_annotation == "kidney tubule cell") |> 
  dplyr::select(id1, id2, id1_annotation, id2_annotation, lca = updated_lca, cl_annotation, total_descendants) |> 
  unique()
```

#### Choosing a single LCA

Alternatively, we could identify some systematic approach to choose one of the possible LCAs. 
Let's first look at some of the pairs and see if using the total descendants would be appropriate:

```{r}
filtered_lca_df |> 
  dplyr::filter(new_lca_pair > 1) |> 
  dplyr::ungroup() |> 
  dplyr::select(id1_annotation, id2_annotation, cl_annotation, total_descendants) |> 
  DT::datatable(rownames = FALSE)
```

In looking at these pairs, for the most part we see one term that is obviously "more specific" than the other terms.

- A lot of matches give both "memory T cell" and "mature alpha-beta T cell". 
In this case, "memory T cell" is more specific and has fewer descendants. 
- We see "peptide hormone secreting cell" and another more specific term like "pancreatic endocrine cell" or "enteroendocrine cell" with fewer descendants. 
- "glandular epithelial cell" and "epithelial cell of pancreas" also show up, where "epithelial cell of pancreas" has fewer descendants.

The exception to this is "mononuclear phagocyte" and "myeloid leukocyte" which to me are both equally vague. 
"myeloid leukocyte" does have a fewer number of descendants though so if we use that rule, I do think it would be reasonable to choose "myeloid leukocyte".

Generally, I think we could choose the term with the fewest descendants as the appropriate LCA, which would be consistent with using the number of descendants in the first place to set our original cutoffs. 
This approach seems more straight forward than recalculating the LCA as we did in the previous section.

```{r}
filtered_lca_df <- filtered_lca_df |> 
  dplyr::group_by(id1, id2) |> 
  # get the cl_annotation with the minimum value in total_descendants column 
  # discard all other rows that aren't the minimum value 
  dplyr::slice_min(total_descendants, with_ties = FALSE) |> 
  dplyr::ungroup()
```

### Discarded cell types

Let's look at what possible cell types would be discarded if we used these rules. 
Previously we were keeping "neuron" and "epithelial cell" even though they had a larger number of descendants. 
Are there any other cell types that appear that we should include in that list?

```{r}
lca_df |>
  dplyr::filter(total_descendants > 170) |>
  dplyr::select(cl_annotation, total_descendants) |>
  dplyr::arrange(desc(total_descendants)) |>
  unique() |> 
  DT::datatable(rownames = FALSE)
```

We should probably also include "columnar/cuboidal epithelial cell" and "endo-epithelial cell". 
Let's look at the matches that led to those LCA to see if there is any reason we should not include them.

```{r}
lca_df |> 
  dplyr::filter(cl_annotation == "columnar/cuboidal epithelial cell") |> 
  dplyr::select(id1_annotation, id2_annotation, cl_annotation) |> 
  unique() |> 
  DT::datatable(rownames = FALSE)
```

```{r}
lca_df |> 
  dplyr::filter(cl_annotation == "endo-epithelial cell") |> 
  dplyr::select(id1_annotation, id2_annotation, cl_annotation) |> 
  unique() |> 
  DT::datatable(rownames = FALSE)
```

I think we should add these two terms to our list of cell types to keep alongside "neuron" and "epithelial cell" when the total descendants is higher than 170.

## Updated consensus cell type pairs

Using the above exploration, we should update our rules for labeling the pairs with their LCA to fit one of the following criteria:

- A single LCA term with <= 170 total descendants
- LCA term that is not one of: "bone cell", "lining cell", "blood cell", "progenitor cell", "supporting cell", "biogenic amine secreting cell", "protein secreting cell", or "extracellular matrix secreting cell"
- LCA term that is "neuron" or "epithelial cell"
- The LCA term with the fewest descendants when multiple LCA terms are found that meet the above criteria

```{r}
# create a table with all pairs and their LCA
filtered_lca_df <- lca_df |> 
  # first set criteria to keep cell types, > 170, celltypes of interest 
  dplyr::filter(total_descendants <= 170 & !(cl_annotation %in% celltypes_to_exclude) 
                | cl_annotation %in% c("neuron","columnar/cuboidal epithelial cell", "endo-epithelial cell")
                   # only keep epithelial if not keratinocyte
                | (cl_annotation == "epithelial cell" & id1_annotation != "keratinocyte" & id2_annotation != "keratinocyte")) |> 
  dplyr::group_by(id1, id2) |> 
  # get the cl_annotation with the minimum value in total_descendants column 
  # discard all other rows that aren't the minimum value 
  dplyr::slice_min(total_descendants, with_ties = FALSE) |> 
  dplyr::ungroup()
```

Let's make sure we only have unique combinations:

```{r}
dup <- filtered_lca_df |> 
  dplyr::select(id1, id2) |> 
  duplicated()

dim(filtered_lca_df[dup, ])[1]
```

Now what cell types are we retaining:

```{r}
filtered_lca_df |> 
  dplyr::select(cl_annotation, total_descendants) |> 
  dplyr::arrange(desc(total_descendants)) |> 
  dplyr::distinct() |> 
  DT::datatable(rownames = FALSE)
```

## Obtaining the final consensus cell type

Now that we have the LCA to use for all possible pairs, we need to combine them back with the possible combinations of all three annotation methods to obtain the final consensus cell type. 
For each combination there should be three possible LCA.

- If they are all unknown and no LCA was assigned, then no consensus cell type will be assigned. 
- If there is one unique LCA then that will be used as the consensus cell type. 
- If there are more than one unique LCA, then we will need to decide how to assign the consensus cell type.

```{r}
# combine lca with all combinations 
combined_ref_df <- all_ref_df |> 
  # split pairs to join with final lca assignment
  tidyr::separate(col = "pairs", into = c("id1", "id2"), sep = ";") |> 
  dplyr::left_join(filtered_lca_df, by = c("id1", "id2")) |> 
  # get rid of any combinations that don't have any possible matches 
  tidyr::drop_na()
```

Let's see how often we have a single unique match. 
If there's only one match we'll use that, otherwise we need to figure out how to make the final assignment.

```{r}
consensus_ref_df <- combined_ref_df |> 
  dplyr::group_by(panglao_ontology, blueprint_ontology, scimilarity_ontology) |> 
  dplyr::summarize(
    total_matches = length(unique(cl_annotation)),
    consensus_celltype_annotation = dplyr::case_when(
      total_matches == 1 ~ cl_annotation,
      total_matches > 1 ~ paste0(unique(cl_annotation), collapse = ";")
    ),
    # aggregate total_descendants for the unique annotations only
    total_descendants = paste0(
      unique(total_descendants[match(unique(cl_annotation), cl_annotation)]),
      collapse = ";"
    )

  ) |> 
  dplyr::distinct()

table(consensus_ref_df$total_matches)
```

```{r}
multi_lca_df <- consensus_ref_df |> 
  dplyr::ungroup() |> 
  dplyr::filter(total_matches > 1) |> 
  dplyr::select(consensus_celltype_annotation, total_descendants) |> 
  dplyr::distinct()

multi_lca_df |> 
  DT::datatable(rownames = FALSE)

```

Generally speaking it looks like there are terms that are of the same lineage which makes sense. 
I think we could just stick with our previous guidelines of choosing the label with the fewest descendants or is the most specific term.

The only problem we might have is when terms have an equivalent number of descendants. 
Let's see how many cases that applies too.

```{r}
# first get a dataframe of the minimum descendants, keeping all ties 
final_df <- combined_ref_df |> 
  # first get rid of any cases where more than one pair gives the same lca term 
  dplyr::select(panglao_ontology, blueprint_ontology, scimilarity_ontology, lca, cl_annotation, total_descendants) |> 
  dplyr::distinct() |> 
  # now select the lca term with the minimum total descendants 
  dplyr::group_by(panglao_ontology, blueprint_ontology, scimilarity_ontology) |> 
  dplyr::slice_min(total_descendants)
```

```{r}
# are there any cases where we have ties? 
dup <- final_df |> 
  dplyr::select(panglao_ontology, blueprint_ontology, scimilarity_ontology) |> 
  duplicated()

final_df[dup, ] |> 
  DT::datatable(rownames = FALSE)
```

It looks like there are occassions where the lca term is both `hematopoietic precursor cell` and `stem cell`. 
These terms both have the same number of total descendants. 
Manually looking at the Cell Ontology, they also both only have 2 ancestors. 
I think I would favor choosing `hematopoietic precursor cell` over `stem cell` since that specifically refers to a cell in the blood lineage. 
Also glancing at the terms from blueprint here, they all appear to be from the HSC lineage. 
Let's double check that.

```{r}
final_df[dup, ] |> 
  dplyr::pull(blueprint_ontology) |> 
  as.character() |> 
  unique()
```

These terms correspond to `megakaryocyte-erythroid progenitor cell`,  `hematopoietic stem cell`, `hematopoietic multipotent progenitor cell`, `common lymphoid progenitor`, `granulocyte monoctye progenitor cell`, and `common myeloid progenitor`.

For this case, we should choose `hematopoietic precursor cell` over `stem cell`.

## Conclusions

Combining all the information from this notebook, we should use the following set of rules to create the consensus cell type reference:

Identify the LCA for all possible pairs of annotations and keep the LCA if the following criteria are met:

- LCA has <= 170 total descendants or belongs to `neuron`, `epithelial cell`, `columnar/cuboidal epithelial cell`, or `endo-epithelial cell`
- LCA term that is not one of: "bone cell", "lining cell", "blood cell", "progenitor cell", "supporting cell", "biogenic amine secreting cell", "protein secreting cell", or "extracellular matrix secreting cell"
- If multiple LCAs are found that meet the above criteria, the LCA term with the fewest descendants is retained

For all possible combinations of the three annotations, if at least one pair has an LCA that meet the above criteria, a consensus cell type is assigned using the LCA with the fewest descendants. 
The only exception to this rule is if the annotations return both `hematopoietic precursor cell` and `stem cell` as possible LCAs.
In this case, `hematopoietic precursor cell` is assigned as the consensus cell type.

## Session info

```{r}
sessionInfo()
```