### 2025-08-21 analyses/infercnv-consensus-cell-type/exploratory-notebooks/reference-query-counts.Rmd
fix counting to handle multiplexed samples, and update language throughout to libraries which is actually what we are counting

---
title: "Diagnoses and cell types"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output: html_notebook
params:
  n_reference_threshold: 100
---

The goal of this notebook is to look at cell type distributions across libraries to determine how many could be run through `inferCNV`.
Specifically, we are interested in seeing the relative numbers of cells intended as `inferCNV` reference vs. query.
Cell types were determined to be reference or query based on the sample's diagnosis, as mapped out in `../references/broad-diagnosis-map.tsv`.
We consider a threshold of at least `r params$n_reference_threshold` reference cells needed to run `inferCNV`.

## Setup

```{r, warning = FALSE}
library(ggplot2)
theme_set(theme_bw())

# settings
options(
  dplyr.summarise.inform = FALSE, 
  readr.show_col_types = FALSE
)
```

### Paths

Define directories:

```{r}
module_dir <- rprojroot::find_root(rprojroot::is_renv_project)
repository_base <- file.path(module_dir, "..", "..")

ref_dir <- file.path(module_dir, "references")
data_dir <- file.path(repository_base, "data", "current")
consensus_dir <- file.path(repository_base, "data", "current", "results", "cell-type-consensus")
consensus_module_dir <- file.path(repository_base, "analyses", "cell-type-consensus")
```

Define paths to input files to read in:

```{r}
# map broad to submitted diagnoses
diagnosis_map_file <- file.path(ref_dir, "broad-diagnosis-map.tsv")

# map broad diagnoses to cell type groups
diagnosis_celltype_map_file <- file.path(ref_dir, "diagnosis-celltype-groups.tsv")

# map cell type groups to constituent cell types
validation_file <- file.path(
  consensus_module_dir, 
  "references", 
  "consensus-validation-groups.tsv"
)

# validation palette
palette_url <- "https://raw.githubusercontent.com/AlexsLemonade/scpca-paper-figures/refs/heads/main/palettes/validation-group-palette.tsv"

# metadata files, for getting sample diagnoses
metadata_files <- list.files(
  path = data_dir,
  pattern = "single_cell_metadata\\.tsv$",
  recursive = TRUE,
  full.names = TRUE
)

# TSVs of consensus cell types
consensus_files <- list.files(
  path = consensus_dir,
  pattern = "_processed_consensus-cell-types\\.tsv\\.gz$",
  recursive = TRUE,
  full.names = TRUE
) |>
  # set names to filter on later so we don't read files we don't need
  purrr::set_names(\(x) {
    stringr::str_split_i(basename(x), "_", 1)
  })
```

### Read input data

```{r}
diagnosis_map <- readr::read_tsv(diagnosis_map_file)
diagnosis_celltype_map <- readr::read_tsv(diagnosis_celltype_map_file)
validation_group_df <- readr::read_tsv(validation_file)
palette_df <- readr::read_tsv(palette_url)
celltype_pal <- palette_df$color |>
  purrr::set_names(palette_df$validation_group_annotation)
celltype_pal["query"] <- celltype_pal["unknown"]
```

```{r}
 all_library_diagnosis_df <- metadata_files |>
  purrr::map(
    \(file) {
      df <- readr::read_tsv(file) |>
        dplyr::select(scpca_project_id, scpca_library_id, diagnosis, is_cell_line, is_multiplexed) |>
        # have to join here because of multiplexed; inner_join also removes normals
        dplyr::inner_join(diagnosis_map, by = c("diagnosis" = "submitted_diagnosis")) |>
        dplyr::select(-diagnosis, -ontology_id, -human_readable_value) |>
        dplyr::distinct()
  }) |>
  dplyr::bind_rows()

total_libraries <- nrow(all_library_diagnosis_df) # needed for later math

library_diagnosis_df <- all_library_diagnosis_df |>
  # don't consider cell line (or multiplexed?)
  # for now at least, we are retaining multiplexed
  dplyr::filter(!is_cell_line)
```

```{r}
consensus_df <- consensus_files |>
  # only keep libraries in scpca_project_id_diagnosis_df
  purrr::keep_at(library_diagnosis_df$scpca_library_id) |>
  purrr::imap(
    \(file, id) {

      # bump guess_max to kill problems() warnings
      readr::read_tsv(file, guess_max = 15000) |>
       dplyr::select(
         scpca_project_id = project_id,
         scpca_library_id = library_id,
         barcodes,
         consensus_annotation,
         consensus_ontology
       )
  }) |>
  dplyr::bind_rows()
```

### Prepare data for analysis

Here, we map over each diagnosis group to label each cell as `query` or `reference`.
Then, we count the number of each category.

```{r}
query_reference_df <- unique(library_diagnosis_df$diagnosis_group) |>
  purrr::map(
    \(broad_diagnosis) {

      celltype_groups <- diagnosis_celltype_map |>
        dplyr::filter(diagnosis_group == broad_diagnosis) |>
        tidyr::separate_rows(celltype_groups, sep = ", ") |>
        dplyr::pull(celltype_groups)

      reference_celltypes <- validation_group_df |>
        dplyr::filter(validation_group_annotation %in% celltype_groups) |>
        dplyr::pull(consensus_annotation)

      library_ids <- library_diagnosis_df |>
        dplyr::filter(diagnosis_group == broad_diagnosis) |>
        dplyr::pull(scpca_library_id)

      consensus_df |>
        dplyr::filter(scpca_library_id %in% library_ids) |>
        dplyr::mutate(
          cell_class = ifelse(
            consensus_annotation %in% reference_celltypes, 
            "reference", 
            "query"
        ))
  }) |>
  dplyr::bind_rows()
```

```{r}
library_counts_df <- query_reference_df |>
  dplyr::count(scpca_library_id, cell_class) |>
  tidyr::pivot_wider(
    names_from = cell_class, 
    values_from = n, 
    values_fill = 0
  ) |>
  dplyr::inner_join(library_diagnosis_df, by = "scpca_library_id") |>
  dplyr::select(
    scpca_project_id, 
    scpca_library_id, 
    diagnosis_group, 
    query, 
    reference
  )
```

## Analysis

Using a threshold of needing at least `r params$n_reference_threshold` reference cells, how many libraries could be run through `inferCNV`?

```{r}
usable_df <- library_counts_df |>
  dplyr::group_by(diagnosis_group) |>
  dplyr::summarize(
    n_usable = sum(reference >= params$n_reference_threshold),
    total_group_libs = dplyr::n(), 
    frac_usable = round(n_usable / total_group_libs, 3) 
  ) |>
  dplyr::arrange(desc(total_group_libs)) |>
  dplyr::select(diagnosis_group, total_group_libs, n_usable, frac_usable)

usable_df
```
Except for a handful of "other solid tumors," our broad diagnosis groups are well-represented, which is nice to see.

Which diagnosis groups have _no libraries_ with sufficient reference cells?

```{r}
usable_df |>
  dplyr::filter(frac_usable == 0)
```
It's not too surprising to see `Mixed phenotype acute leukemia` here since most immune cell types can't be included in the reference, and the other groups comprise very small numbers of libraries.

In total, this fraction of ScPCA libraries have sufficient reference cells to run through `inferCNV`:

```{r}
sum(library_counts_df$reference >= params$n_reference_threshold) / total_libraries
```

Of the libraries with sufficient reference cells, do any have _fewer_ query than reference cells?

```{r}
library_counts_df |>
  # consider usable only
  dplyr::filter(reference >= params$n_reference_threshold) |>
  dplyr::filter(query < params$n_reference_threshold)
```

There is one library with about 7x as many reference as query cells; let's look more closely:

```{r}
consensus_df |>
  dplyr::filter(scpca_library_id == "SCPCL000021") |>
  dplyr::count(consensus_annotation) |>
  dplyr::arrange(n)
```

This library indeed is dominated by immune cell types with very low tumor purity, but it still seems to have enough query cells to not be a problem and should be ok to run.

## Plots

In this section, we plot the relative contributions of each cell type group in the libraries we can run. 
We show the relative proportions of consensus cell type groups in the _reference_, and we show all other query cells grouped together in gray.
The x-axis is library ids, and the y-axis is the proportion of each cell type group.

```{r}
usable_libraries <- library_counts_df |>
  dplyr::filter(reference >= params$n_reference_threshold) |>
  dplyr::pull(scpca_library_id)

plot_df <- query_reference_df |>
  dplyr::filter(scpca_library_id %in% usable_libraries) |>
  dplyr::left_join(validation_group_df, by = c("consensus_ontology", "consensus_annotation")) |>
  dplyr::select(
    scpca_project_id, 
    scpca_library_id, 
    barcodes, 
    cell_class, 
    validation_group_annotation
  ) |>
  dplyr::mutate(
    label = ifelse(
      cell_class == "query", 
      "query",
      validation_group_annotation
    ),
    # move "query" first
    label = forcats::fct_relevel(label, "query", after = 0)
  )
```

```{r fig.width = 16, fig.height = 14}
ggplot(plot_df) + 
  aes(x = scpca_library_id, fill = label) + 
  geom_bar(position = "fill") +
  scale_fill_manual(values = celltype_pal) + 
  facet_wrap(
    vars(scpca_project_id), 
    scales = "free_x", 
    ncol = 4
  ) + 
  ylab("proportion") +
  theme(
    axis.text.x = element_blank(), 
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 13),
    legend.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

## Conclusions

Currently, _before_ incorporating `SCimilarity` into our consensus cell types, we are able to run about 44% of libraries through `inferCNV`. 
Most broad diagnosis groups are represented, except `Mixed phenotype acute leukemia` and a handful of other solid tumors.
All eligible projects except `SCPCP000011` (Retinoblastoma samples) are represented as well.

## Session Info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```