### 2025-09-09 analyses/cell-type-scimilarity/exploratory-notebooks/01-scimilarity-results-summary.Rmd

---
title: "Explore SCimilarity cell type annotations"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    code_folding: hide
---

This notebook summarizes the findings from assigning cell type labels to all ScPCA samples using `SCimilarity`.

Running this notebook requires the results from the `cell-type-consensus` and `cell-type-scimilarity` module in `OpenScPCA-nf` to be saved to `OpenScPCA-analysis/data/current/results`.

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)

# Define color ramp for shared use in the heatmaps
heatmap_col_fun <- circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue"))
# Set heatmap padding option
ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(0.6, "in"))

# settings
options(
  dplyr.summarise.inform = FALSE, 
  readr.show_col_types = FALSE
)

```

## Functions

```{r}
# function to read in both scimilarity and consensus results and combine into a single data frame 
prep_data <- function(library_id, scimilarity_file, consensus_file){

  scimilarity_df <- readr::read_tsv(scimilarity_file)
  consensus_df <- readr::read_tsv(consensus_file)

  annotation_df <- consensus_df |> 
    dplyr::left_join(scimilarity_df, by = c("barcodes" = "barcode")) |> 
    dplyr::mutate(
      cell_id = glue::glue("{library_id}-{barcodes}"), 
      total_cells_per_library = dplyr::n()
    )

  return(annotation_df)

}
```

```{r}
# Copied from scpca-nf template report 
#' Format DT datatables
#'
#' @param df Data frame to format
#'
#' @return DT datatable
format_datatable <- function(df, caption) {
  df |>
    DT::datatable(
      rownames = FALSE,
      extensions = "Scroller",
      caption = caption,
      options = list(
        scroller = TRUE,
        deferRender = TRUE,
        scrollX = TRUE,
        scrollY = 400,
        scrollCollapse = TRUE,
        language = list(search = "Filter:"),
        columnDefs = list(list(
          # render missing values as "NA" rather than blanks in the table
          targets = "_all",
          render = DT::JS(
            "function(data, type, row, meta) {",
            "return data === null ? 'NA' : data;",
            "}"
          )
        ))
      )
    )
}
```

## Data setup

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)
# results directory with cell-type-consensus 
results_dir <- file.path(repository_base, "data", "current", "results")
consensus_results_dir <- file.path(results_dir, "cell-type-consensus")
scimilarity_results_dir <- file.path(results_dir, "cell-type-scimilarity")

# diagnoses table used for labeling plots from cell-type-consensus module
diagnoses_file <- file.path(repository_base, "analyses", "cell-type-consensus", "sample-info", "project-diagnoses.tsv")
```

```{r}
# use the jaccard functions from cell-type-neuroblastoma-04 module
jaccard_functions <- file.path(repository_base, "analyses", "cell-type-neuroblastoma-04", "scripts", "utils", "jaccard-utils.R")
source(jaccard_functions)
```

```{r}
# list all results files 
consensus_results_files <- list.files(consensus_results_dir, pattern = "_processed_consensus-cell-types\\.tsv.\\gz$", recursive = TRUE, full.names = TRUE)
consensus_files_df <- data.frame(
  library_id = stringr::word(basename(consensus_results_files), 1, sep = "_"),
  consensus_file = consensus_results_files
)

scimilarity_results_files <- list.files(scimilarity_results_dir, pattern = "_processed_scimilarity-celltype-assignments\\.tsv.\\gz$", recursive = TRUE, full.names = TRUE)
scimilarity_files_df <- data.frame(
  library_id = stringr::word(basename(scimilarity_results_files), 1, sep = "_"),
  scimilarity_file = scimilarity_results_files
)

all_files_df <- scimilarity_files_df |> 
  dplyr::left_join(consensus_files_df, by = "library_id")

# define cell line projects to remove
cell_line_projects <- c("SCPCP000020", "SCPCP000024")
```

```{r}
# check that everything has a matching consensus file 
# we joined the consensus into the scimilarity so only libraries with scimilarity are included 
if(sum(is.na(all_files_df$consensus_file)) > 0){
  stop("Missing matching consensus cell type file for all libraries")
}
```

```{r, message=FALSE, warning=FALSE}
# read in diagnoses
diagnoses_df <- readr::read_tsv(diagnoses_file)

# read in all results as a single dataframe
results_df <- all_files_df |> 
  purrr::pmap(prep_data) |> 
  dplyr::bind_rows() |> 
  # remove cell line projects
  dplyr::filter(!project_id %in% cell_line_projects) |> 
  # add in diagnoses 
  dplyr::left_join(diagnoses_df, by = "project_id") |> 
  dplyr::mutate(
    # create a label for plotting
    project_label = glue::glue("{project_id}:{diagnosis}")
  )
```

## `SCimilarity` results

Below we will summarize the results from `SCimilarity` on all ScPCA projects.

### Table of all cell types in ScPCA

First, we just look at the cell types identified and see if any of the cells are not assigned.

```{r}
sum(is.na(results_df$scimilarity_celltype_annotation))
```

It looks like every cell gets a cell type, so that's something to keep in mind when looking at these results.

Let's see what cell types are presented in our data?

```{r}
results_df |> 
  dplyr::ungroup() |> 
  dplyr::count(scimilarity_celltype_annotation, name = "total_cells") |> 
  dplyr::arrange(desc(total_cells)) |> 
      format_datatable(caption = "All samples")
```

So it looks like we have a lot of "native cell" labels which really just means the cell is a cell and didn't align with anything else in the reference. 
This label is probably similar to our "other" label in `CellAssign`.

Also, most of our cells are progenitor cells or hematopoietic stem cells, which either means a lot of the cells are tumor cells and show stem and progenitor like properties or these are from the blood tumors which make up a large part of our atlas.
Either way, we have cell types!

### Table of cell types per project

This section shows a table of all cell types identified in each project. 
Each project will have its own table.

```{r}
# get all the unique labels
project_labels <- unique(results_df$project_label)

tables <- project_labels |> 
  purrr::map(\(label) {

    results_df |> 
      dplyr::filter(project_label == label) |> 
      dplyr::group_by(scimilarity_celltype_annotation) |> 
      dplyr::summarise(
        total_cells = dplyr::n()
      ) |> 
      dplyr::arrange(desc(total_cells)) |> 
      format_datatable(caption = label)

  })

htmltools::tagList(tables)
```

### Comparison to existing consensus cell types

In this section, we compare the top 15 `SCimilarity` cell type annotations to the consensus cell types determined using only `CellAssign` and `SingleR` annotations for all libraries in a single project. 
The rows correspond to consensus cell types and the columns correspond to `SCimilarity` annotations. 
All other cell types not in the top 10 represented cell types in a project are grouped into the "All remaining cell types" category.

```{r}
results_df <- results_df |> 
    dplyr::group_by(project_id) |> 
    dplyr::mutate(
      # get most frequently observed cell types across libraries in that project 
      scimilarity_celltype_lumped = forcats::fct_lump_n(scimilarity_celltype_annotation, 15, other_level = "All remaining cell types", ties.method = "first") |> 
        # sort by frequency 
        forcats::fct_infreq() |> 
        # make sure all remaining and unknown are last, use this to assign colors in specific order
        forcats::fct_relevel("All remaining cell types", after = Inf) |> 
        as.character()
    )
```

```{r, fig.height = 10, fig.width = 10}
project_labels |> 
  purrr::walk(\(label){

    results_df |> 
      dplyr::filter(project_label == label) |> 
      make_jaccard_heatmap(
        annotation_col1 = "scimilarity_celltype_lumped",
        annotation_col2 = "consensus_annotation",
        label1 = glue::glue("{label} \nSCimilarity"),
        label2 = "Consensus"
      )

  })
```

As expected there is some direct agreement or places where more broad consensus cell types are split up across more granular cell types in `SCimilarity`. 
A few notable examples:

- `SCPCP000001`: Cells labeled as macrophages in consensus are split between glial and microglial cells in `SCimilarity`. 
Natural killer cells match up between consensus and `SCimilarity`. 
- `SCPCP000006`: Endothelial cells match up between consensus and `SCimilarity`. 
The unknown cells in consensus correspond to mesenchymal stem cells and a variety of kidney cells. 
- `SCPCP000007`: Mature T cell in consensus is split between CD4 and CD8 T cells in `SCimilarity`.

We also see places where there is not agreement, which is to be expected.

## Conclusions

Of the 203 possible cell types in `SCimilarity`, 191 of them are represented in ScPCA samples. 
There are quite a few scenarios where consensus cell types "agree" with `SCimilarity`. 
It will be informative to look at the new consensus cell types and how they change with the addition of the `SCimilarity` annotations, but in general the addition of `SCimilarity` seems quite useful!

## Session info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```