### 2025-09-12 analyses/cell-type-consensus/exploratory-notebooks/07-compare-consensus.Rmd
notebook comparing consensus cell type outputs

---
title: "Compare consensus cell types with SCimilarity to consensus cell types without SCimilarity"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    code_folding: hide
---

This notebook compares the results from assigning consensus cell types with `SCimilarity` to the previous consensus cell types without `SCimilarity`. 
All results from the `cell-type-consensus` module in `OpenScPCA-nf` must be saved to `results` prior to rendering this notebook.

Note that any multiplexed libraries have been excluded from this notebook as `SCimilarity` is not run on those libraries when using the `OpenScPCA-nf` module.

```{r packages}
suppressPackageStartupMessages({
  # load required packages
  library(ggplot2)
})

# Set default ggplot theme
theme_set(
  theme_classic()
)

# Define color ramp for shared use in the heatmaps
heatmap_col_fun <- circlize::colorRamp2(c(0, 1), colors = c("white", "darkslateblue"))
# Set heatmap padding option
ComplexHeatmap::ht_opt(TITLE_PADDING = grid::unit(0.6, "in"))

# settings
options(
  dplyr.summarise.inform = FALSE, 
  readr.show_col_types = FALSE
)
```

## Data setup

```{r base paths}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)
module_base <- file.path(repository_base, "analyses", "cell-type-consensus")

# results directory with cell-type-consensus 
results_dir <- file.path(module_base, "results", "cell-type-consensus")

# diagnoses table used for labeling plots 
diagnoses_file <- file.path(module_base, "sample-info", "project-diagnoses.tsv")
```

```{r}
# use the jaccard functions from cell-type-neuroblastoma-04 module
jaccard_functions <- file.path(repository_base, "analyses", "cell-type-neuroblastoma-04", "scripts", "utils", "jaccard-utils.R")
source(jaccard_functions)
```

```{r}
# list all results files 
results_files <- list.files(results_dir, pattern = "_consensus-cell-types\\.tsv.\\gz$", recursive = TRUE, full.names = TRUE)

# define cell line projects to remove
cell_line_projects <- c("SCPCP000020", "SCPCP000024")
```

```{r, message=FALSE, warning=FALSE}
# read in diagnoses
diagnoses_df <- readr::read_tsv(diagnoses_file)

# read in results and prep data frame for plotting 
all_results_df <- results_files |> 
  purrr::map(readr::read_tsv, col_types = "c")|> 
  dplyr::bind_rows() |> 
  # remove cell line projects
  dplyr::filter(!project_id %in% cell_line_projects,
                # remove multiplexed libraries since we didn't run scimilarity on them
                !stringr::str_detect(sample_id, ";")) |> 
  # add in diagnoses 
  dplyr::left_join(diagnoses_df, by = "project_id") |> 
  dplyr::mutate(
    # create a label for plotting
    project_label = glue::glue("{project_id}:{diagnosis}")
  )
```

```{r, message=FALSE, warning=FALSE}
# Turn into a dataframe with one row per library per celltype
# Make a table for each consensus method separately and then stack by consensus cell type old/new 
# one row for old and one row for each cell type

# first add columns for total cells per library, number of new and old cell types
grouped_df <- all_results_df |> 
    dplyr::group_by(library_id) |> 
    dplyr::mutate(
      total_cells_per_library = dplyr::n(),
      num_old_celltypes = length(unique(consensus_annotation)),
      num_new_celltypes = length(unique(existing_consensus_celltype_annotation))
    ) |>
    dplyr::ungroup()

# get stats for old consensus cell types
old_consensus_df <- grouped_df |> 
    dplyr::group_by(project_id, project_label, diagnosis, library_id, sample_type, existing_consensus_celltype_annotation, existing_consensus_celltype_ontology) |> 
    dplyr::summarize(num_celltypes = unique(num_new_celltypes), # constant for each library
                     total_cells_per_annotation = dplyr::n(),
                     total_cells_per_library = unique(total_cells_per_library)) |>
    dplyr::mutate(
      # add percentage 
      percent_cells_annotation = round((total_cells_per_annotation / total_cells_per_library) * 100, 2)
    ) |> 
    dplyr::ungroup()

# stats for new consensus cell types
new_consensus_df <- grouped_df |> 
    dplyr::group_by(project_id, project_label, diagnosis, library_id, sample_type, consensus_annotation, consensus_ontology) |> 
    dplyr::summarize(num_celltypes = unique(num_old_celltypes), # constant for each library
                     total_cells_per_annotation = dplyr::n(),
                     total_cells_per_library = unique(total_cells_per_library)) |>
    dplyr::mutate(
      # add percentage 
      percent_cells_annotation = round((total_cells_per_annotation / total_cells_per_library) * 100, 2)
    ) |> 
    dplyr::ungroup()

# combine into a single df and add a column to indiciate old/new 
df_list <- list(old_consensus_df, new_consensus_df) |> 
  purrr::set_names(c("no_scimilarity", "with_scimilarity"))

all_grouped_df <- dplyr::bind_rows(df_list, .id = "consensus_type")
```

## Number of unknown cells

The plot below shows the number of cells annotated as unknown for each library using old (without `SCimilarity`) vs. new (with `SCimilarity`) consensus annotations. 
The red bar indicates the median percentage of cells.

```{r, fig.height=7}
# only get the unknown cells
unknown_only <- all_grouped_df |> 
  dplyr::filter(consensus_annotation == "Unknown" | existing_consensus_celltype_annotation == "Unknown")

# compare the distribution of unknown with and without scimilarity
ggplot(unknown_only, aes(x = consensus_type, y = percent_cells_annotation)) +
  ggforce::geom_sina(size = 0.1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        plot.margin = margin(10,10,10,10)) +
  labs(
    x = "", 
    y = "Percent of cells annotated as Unknown"
  ) +
  stat_summary(fun=median, geom="crossbar" , color = "red", linewidth = 0.2)

```

As expected, the percentage of cells annotated as unknown is lower when `SCimilarity` is incorporated, suggesting we are able to annotate more cells with a meaningful label that is not unknown.

## Number of cell types observed

Below we look at the number of cell types observed in each project for all samples.

```{r, fig.height=15, fig.width=10}
num_celltypes_df <- all_grouped_df |> 
  # add a new line for facet labels 
  dplyr::mutate(facet_label = glue::glue("{project_id}\n{diagnosis}")) |>
  dplyr::select(facet_label, library_id, num_celltypes, consensus_type) |> 
  unique()

ggplot(num_celltypes_df, 
       aes(x = consensus_type, y = num_celltypes, group = library_id)) +
  geom_point(aes(color = consensus_type)) +
  geom_line(aes(group = library_id), color = "gray60", alpha = 0.7) +
  facet_wrap(vars(facet_label), ncol = 3) +
  labs(
    x = "Consensus cell type",
    y = "Number of cell types"
  ) +
  theme_bw()

```

As expected, we see that the number of cell types identified increases with the addition of `SCimilarity`.

## Heatmaps comparing old and new consensus cell types

In this section, we compare the top 15 cell type annotations for consensus cell types with `SCimilarity` (rows) to consensus cell types without `SCimilarity` (columns). 
All other cell types not in the top 15 represented cell types in a project are grouped into the "All remaining cell types" category.

Note that in some cases, there are fewer than 15 unique cell types.

```{r, warning=FALSE}
all_results_df <- all_results_df |> 
    dplyr::group_by(project_id) |> 
    dplyr::mutate(
      # get most frequently observed cell types across libraries in that project 
      existing_consensus_celltype_lumped = forcats::fct_lump_n(existing_consensus_celltype_annotation, 15, other_level = "All remaining cell types", ties.method = "first") |> 
        # sort by frequency 
        forcats::fct_infreq() |> 
        # make sure all remaining and unknown are last, use this to assign colors in specific order
        forcats::fct_relevel("All remaining cell types", after = Inf) |> 
        as.character(), 
      # do the same thing for the new consensus cell types
      consensus_celltype_lumped = forcats::fct_lump_n(consensus_annotation, 15, other_level = "All remaining cell types", ties.method = "first") |> 
        # sort by frequency 
        forcats::fct_infreq() |> 
        # make sure all remaining and unknown are last, use this to assign colors in specific order
        forcats::fct_relevel("All remaining cell types", after = Inf) |> 
        as.character(), 
      # jaccard functions expect a cell id column
      cell_id = glue::glue("{library_id}-{barcodes}")
    )

project_labels <- unique(all_results_df$project_label)
```

```{r, fig.height = 10, fig.width = 10}
project_labels |> 
  purrr::walk(\(label){

    all_results_df |> 
      dplyr::filter(project_label == label) |> 
      make_jaccard_heatmap(
        annotation_col1 = "existing_consensus_celltype_lumped",
        annotation_col2 = "consensus_celltype_lumped",
        label1 = glue::glue("{label} \nNo Scimilarity"),
        label2 = "With SCimilarity"
      )

  })
```

In general we see a fair amount of agreement in the overall cell types are defined with a bit more granularity in the cell types assigned with `SCimilarity`.

## Distribution of new consensus cell types

Now we look at the distribution of the cell types in each sample. 
For these plots, we will pull out the top 9 cell types for each project. 
All other cells will be labeled with "All remaining cell types".

The top cell types are determined by counting how many libraries each cell type is found in within a project and taking the most frequent types.

```{r, warning=FALSE}
plot_df <- all_grouped_df |> 
  dplyr::group_by(project_id) |> 
  # remove the old results from plotting
  dplyr::select(-c(existing_consensus_celltype_annotation, existing_consensus_celltype_ontology)) |> 
  tidyr::drop_na(consensus_annotation) |> 
  dplyr::mutate(
    # get most frequently observed cell types across libraries in that project 
    top_celltypes = forcats::fct_lump_n(consensus_annotation, 9, other_level = "All remaining cell types", ties.method = "first") |> 
      # sort by frequency 
      forcats::fct_infreq() |> 
      # make sure all remaining and unknown are last, use this to assign colors in specific order
      forcats::fct_relevel("All remaining cell types", "Unknown", after = Inf)
  ) |> 
  dplyr::ungroup()

# get all unique cell types ordered by frequency 
unique_celltypes <- plot_df |> 
  dplyr::filter(!top_celltypes %in% c("All remaining cell types", "Unknown")) |> 
  dplyr::pull(top_celltypes) |> 
  unique() |>
  sort() |> 
  as.character()

# get color palette
colors <- c(
  colorRampPalette(RColorBrewer::brewer.pal(12, "Paired"))(length(unique_celltypes)),
  "grey60", # all remaining
  "grey95" # unknown
)
names(colors) <- c(unique_celltypes, "All remaining cell types", "Unknown")
```

```{r, fig.height=60, fig.width=10}
# stacked bar chart showing the distribution of the top 9 cell types for each project, including Unknown
project_labels |> 
  purrr::map(\(label){

    project_df <- plot_df |> 
      dplyr::filter(project_label == label) |> 
      dplyr::mutate(
        # relevel factors for specific project 
        top_celltypes = forcats::fct_infreq(top_celltypes) |> 
          forcats::fct_relevel("All remaining cell types", "Unknown", after = Inf)
      )

    # make a stacked bar chart with top cell types 
    ggplot(project_df) + 
      aes(
        x = library_id, 
        y = percent_cells_annotation, 
        fill = top_celltypes
      ) +
      geom_col() + 
      # split samples based on sample type, patient tissue or pdx 
      facet_wrap(vars(sample_type), scales ="free") +
      scale_y_continuous(expand = c(0,0)) +
      scale_fill_manual(values = colors, name = "cell type") +
      ggtitle(label) +
      theme(axis.text.x = element_blank())

    }) |>
  patchwork::wrap_plots(ncol = 1)
```

We are definitely labeling more cells than we were previously and in some cases we may be assigning a "normal" cell type to tumor cells. 
This is not totally surprising as many solid tumors have cells that resemble fibroblasts and muscle cells. 
We also see various progenitors and HSCs being labeled in the leukemias which reflect how leukemia cells resemble those cell types. 
Overall, I think adding in `SCimilarity` is giving us a lot more information than without it.

## Session info

```{r session info}
# record the versions of the packages used in this analysis and other environment information
sessionInfo()
```